### 2020-09-24 Adaptation.Rmd
new

---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE}
library(ggplot2)
library(tidyverse)
library(ggthemes)
library(lme4)
library(lmerTest)
library(lsmeans)
library(car)
library(agricolae)
library(multcomp)
```

```{r include=FALSE}
Data=read_csv("Alldata.csv")
```

```{r include=FALSE}
Data <- Data %>% 
  filter(state != "havelock")
```

# Gender

```{r}
Data1 <- Data %>% 
  filter(!is.na(gender)) %>% 
  group_by(state, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally() %>% 
  mutate(Perc=(n/nn)*100) 
```

```{r}
GrantJune = data.frame(FEM = 33,
                     MALE = 38)
chisq.test(GrantJune)

GrantJuly = data.frame(FEM = 27,
                     MALE = 36)
chisq.test(GrantJuly)

ClayJuly = data.frame(FEM = 37,
                     MALE = 34)
chisq.test(ClayJuly)

ClayJune = data.frame(FEM = 33,
                     MALE = 37)
chisq.test(ClayJune)

HavelockJune = data.frame(FEM = 14,
                     MALE = 19)
chisq.test(HavelockJune)

MacombJuly = data.frame(FEM = 13,
                     MALE = 22)
chisq.test(MacombJuly)

MacombJune = data.frame(FEM = 36,
                     MALE = 34)
chisq.test(MacombJune)

ArlingtonJuly = data.frame(FEM = 30,
                     MALE = 39)
chisq.test(ArlingtonJuly)

ArlingtonJune = data.frame(FEM = 33,
                     MALE = 38)
chisq.test(ArlingtonJune)
```

```{r}
Data1$state <- factor(Data1$state, levels=c("grant", "clay center", "havelock", "macomb", "arlington"), labels=c("Grant", "Clay Center", "Lincoln", "Macomb", "Arlington"))
ggplot(Data1, aes(x=gender, y=Perc, fill=gender)) + geom_bar(stat="identity", width=0.3) + facet_wrap(planting~state, ncol=4) + labs(x="gender", y="Percentage (%)") + theme(legend.position = "none") + theme_solarized() + ggsave("gender.png", units="in", width=12, height=6, dpi=600)
```

# Resultado Geral

```{r}
Data$state <- factor(Data$state, levels=c("grant", "clay center", "havelock", "macomb", "arlington"), labels=c("Grant, NE", "Clay Center, NE", "Lincoln, NE", "Macomb, IL", "Arlington, WI"))
```

```{r echo=FALSE, warning=FALSE, width=12, height=6}
ggplot(Data, aes(x=crop, y=weight, fill=crop)) + geom_boxplot() + 
  #geom_violin() + 
  stat_summary(fun.y=mean, fill="red", geom="point", 
                           shape=21, size=3, show_guide = FALSE) + 
  scale_x_discrete(breaks=c(0, 50, 100, 150, 200, 250, 300, 350, 400, 450, 500, 550, 600, 650, 700, 750)) +
  facet_wrap(planting~state, ncol=5) + scale_fill_brewer(palette = "Set3") +
  theme_wsj() + theme(legend.position="none", strip.text = element_text(size=14, face="bold")) + labs(x="", y="Biomass (g)") +
  ggsave("Adaptation.png", units="in", width=12, height=6, dpi=600)
```

```{r}
model1=lmer(weight ~ crop * state * planting + (1|plant), data=Data)
#summary(model1)
plot(ranef(model1))
shapiro.test(Data$weight)
Anova(model1, test.statistic = "F", type = "II")
lsmeans(model1, ~ crop | planting | state, contr="pairwise", adjust="none")

```

# Arlington

```{r echo=FALSE}
Arlington <- Data %>% 
  filter(state == "arlington")

model2=lmer(weight ~ crop * planting + (1|plant), data=Arlington)
#summary(model1)
plot(ranef(model2))
shapiro.test(Arlington$weight)
Anova(model2, test.statistic = "F", type = "II")
lsmeans(model2, "crop", contr="pairwise", adjust="none")
lsmeans(model2, "planting", contr="pairwise", adjust="none")
```

# Macomb

```{r echo=FALSE}
Macomb <- Data %>% 
  filter(state == "macomb")

model3=lmer(weight ~ crop * planting + (1|plant), data=Macomb)
#summary(model1)
plot(ranef(model3))
shapiro.test(Macomb$weight)
Anova(model3, test.statistic = "F", type = "II")
lsmeans(model3, ~ crop | planting, contr="pairwise", adjust="none")
```

# Grant

```{r echo=FALSE}
Grant <- Data %>% 
  filter(state == "grant")

model4=lmer(weight ~ crop * planting + (1|plant), data=Grant)
#summary(model1)
plot(ranef(model4))
shapiro.test(Grant$weight)
A=Anova(model4, test.statistic = "F", type = "II")
Grantlsm<-lsmeans(model4, ~ crop | planting, contr="pairwise", adjust="none")
cld(Grantlsm, by="crop", Letters=letters, adjust="none", reversed=TRUE, sort=TRUE)

library(multcomp)
ph <- glht(A, linfct = mcp(ind = "Tukey"))
cld(ph)
```

# Clay Center

```{r echo=FALSE}
Clay <- Data %>% 
  filter(state == "clay center")

model5=lmer(weight ~ crop * planting + (1|plant), data=Clay)
#summary(model1)
plot(ranef(model5))
shapiro.test(Clay$weight)
Anova(model5, test.statistic = "F", type = "II")
lsmeans(model5, ~ crop | planting, contr="pairwise", adjust="none")
```

# Havelock

```{r echo=FALSE}
Havelock <- Data %>% 
  filter(state == "havelock")

model6=lmer(weight ~ crop * planting + (1|plant), data=Clay)
#summary(model1)
plot(ranef(model6))
shapiro.test(Havelock$weight)
Anova(model6, test.statistic = "F", type = "II")
lsmeans(model6, ~ crop | planting, contr="pairwise", adjust="none")
```

### 2020-09-24 Analysis.Rmd
new

---
title: "Results"
author: "Maxwel Coura Oliveira"
date: "10/3/2019"
output: pdf_document
---

```{r include=FALSE}
library(tidyverse)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(ggpubr)
library(bestNormalize)
library(ggthemes)
library(nortest)
library(pbkrtest)
```

```{r include=FALSE}
data <- read_csv("Data.csv") %>% 
  drop_na(weight)
data$year <- factor(data$year)
glimpse(data)
```

```{r warning=FALSE}
data %>% 
#  filter(location == "clay center") %>% 
ggplot() + aes(x=doyh, y=weight, color=planting, size=height) + geom_jitter(alpha=0.5) +  facet_wrap(location~crop, ncol = 3, scales = "free_y") + theme_bw() + geom_smooth(method = "lm") +
  ggsave("biomass.pdf", height=18, width=9)
#  geom_vline(xintercept = 182, linetype="dotted", 
#                color = "blue", size=1.5)
```

# Testing Normality
```{r}
ggdensity(data$weight)
pearson.test(data$weight)
pearson.test(data$weight)
```

```{r}
ggdensity(log(data$weight))

pearson.test(log(data$weight))
pearson.test(data$weight) 
```

```{r}
databn <- bestNormalize(data$weight)
data$biomass <- databn$x.t
ggdensity(data$biomass)
```

```{r}
pearson.test(data$biomass)
pearson.test(data$biomass)
```

```{r}
ggplot(data, aes(x=crop, y=weight, color=crop)) + geom_jitter(alpha=0.4) +
  geom_boxplot() +
   coord_flip() + theme_bw() +
  facet_grid(year~planting~location) 
```

```{r warning=FALSE}
Model <- lmer(log(weight) ~ crop*planting*location + (1|year:plant), data=data)
summary(Model)
anova(Model)
```

```{r}
Model1 <- lmer(log(weight) ~ crop*planting + (1|year:plant), data=data)
```

```{r}
PBmodcomp(Model1, Model, nsim=500, cl=2) 
```

```{r}
anova(Model)
```

```{r}
lsm <- emmeans(Model, ~ crop * planting | location ,  adjust="none", type="response")
lsm
```

```{r warning=FALSE}
cld <- CLD(lsm, adjust="none", Letters = letters, reversed = TRUE, type="response")
cld

#write_csv(cld, path="cld.csv")
```

```{r}
cld$location <- factor(cld$location, levels = c("grant", "clay center",  "havelock", "macomb",  "arlington"), labels=c("Grant, NE", "Clay Center, NE",  "Lincoln, NE", "Macomb, IL", "Arlington, WI"))

cld$crop <- factor(cld$crop, levels = c("corn", "soybean", "fallow"), labels=c("corn", "soybean", "fallow"))
cld$planting <- factor(cld$planting, levels =c("June", "July"), labels=c("June", "July"))
```

```{r}
ggplot(cld, aes(x=crop, y=response, color=crop, label=.group)) + 
  geom_point(size=3) + theme_bw() + labs(x="", y=expression(paste("Plamer amaranth (g plant"^"-1",")"))) +
  geom_errorbar(aes(ymin=lower.CL, ymax=upper.CL), width=0.1, size=1) +
  geom_text(nudge_y = c(100, 80, 80, 70, 40, 30, 200, 170, 100, 70, 60, 50, 260, 100, 100, 
                        100, 80, 50, 140, 100, 90, 50, 60, 50, 100, 80, 70, 70, 60, 60), size=6) +
  facet_grid(planting~location) + coord_flip() +
  scale_color_manual(values = c("blue", "darkgreen", "black", "orange", "red")) +
  theme(legend.position = "none", strip.text = element_text(size=18, face="bold"), axis.title = element_text(size=18),
        axis.text.x = element_text(size=17, angle = 45, hjust = 1), axis.text.y = element_text(size=17)) +
  ggsave("adaptation.png", height=8, width=20, dpi=600)
```

```{r}
data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) 
```

```{r}
arlington1 <- data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) %>% 
  filter(location == "arlington" & planting == "June")

  chisq.test(arlington1$n)

arlington2 <- data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) %>% 
  filter(location == "arlington" & planting == "July")

  chisq.test(arlington2$n) 
```

```{r}
macomb1 <- data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) %>% 
  filter(location == "macomb" & planting == "June")

chisq.test(macomb1$n)

macomb2 <- data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) %>% 
  filter(location == "macomb" & planting == "July")

chisq.test(macomb2$n) 
```

```{r}
havelock1 <- data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) %>% 
  filter(location == "havelock" & planting == "June")

chisq.test(havelock1$n)

havelock2 <- data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) %>% 
  filter(location == "havelock" & planting == "July")

chisq.test(havelock2$n) 
```

```{r}
clay1 <- data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) %>% 
  filter(location == "clay center" & planting == "June")

chisq.test(clay1$n)

clay2 <- data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) %>% 
  filter(location == "clay center" & planting == "July")

chisq.test(clay2$n) 
```

```{r}
grant1 <- data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) %>% 
  filter(location == "grant" & planting == "June")

chisq.test(grant1$n)

grant2 <- data %>% 
  filter(!is.na(gender)) %>% 
  group_by(location, planting) %>% 
  count(gender, na.rm=TRUE) %>% 
  add_tally(name="nn") %>% 
  mutate(Perc=(n/nn)*100) %>% 
  filter(location == "grant" & planting == "July")

chisq.test(grant2$n)
```

### 2020-11-27 Analysis.Rmd
atualizacao

@@
 glimpse(data)
+```
+
+```{r}
+data %>% 
+  summarise(max(doyh) - min(doyh))
+```
+
+```{r}
+data %>% 
+  group_by(crop, location, planting) %>% 
+  summarise(mean = mean(weight), min = min(weight), max = max(weight)) %>% 
+  ggplot(aes(x=planting, y=mean, fill = planting)) +
+    geom_col() +
+  facet_grid(location ~ crop) +
+  coord_flip()
+```
+
+```{r}
+data %>% 
+  ggplot(aes(x = location, y = weight)) +
+  geom_bar(stat="identity", fun.y = "mean") +
+  facet_grid(planting ~ crop) +
+  coord_flip()
 ```
@@
 ```{r warning=FALSE}
-Model <- lmer(log(weight) ~ crop*planting*location + (1|year:plant), data=data)
-summary(Model)
+new_dt <- data %>% 
+  filter(crop %in% "fallow",
+         planting %in% "June") 
+Model <- lmer(log(weight) ~ location + (1|year:location) + (1|plant:location), data= new_dt)
+#summary(Model)
 anova(Model)
@@
 ```{r}
+
 Model1 <- lmer(log(weight) ~ crop*planting + (1|year:plant), data=data)
@@
 ```{r}
-lsm <- emmeans(Model, ~ crop * planting | location ,  adjust="none", type="response")
+?emmeans
+```
+
+```{r}
+lsm <- emmeans(Model, ~  location ,  adjust="none", type="response")
 lsm

### 2020-11-27 Manuscript/Adaptation.Rmd
atualizacao

---
title: "Adaptation of Palmer amaranth to the upper Midwest United States"
runtitle: "Short Title"
documentclass: frontiersSCNS  # or frontiersHLTH, or frontiersFPHY
author:
  - name: Maxwel Coura Oliveira
    affiliation: '1'
    etal: Oliveira # First author's last name. 
  - name: Amit J Jhala
    affiliation: '2'
  - name: Mark Bernards
    affiliation: '3'
  - name: Strahinja Stephanovic
    affiliation: '4'
  - name: Chris Proctor
    affiliation: '2'
  - name: Rodrigo Werle
    affiliation: '1'
    email: rwerle@wisc.edu # Indicates corresponding Author
    institution: Laboratory X, Institute X, Department of Agronomy and Horticulture, University of Wisconsin-Madison
    street: 1575 Linden Drive
    city: Madison
    state: Wisconsin # only USA, Australia, Canada
    zip: 53706
    country: United States of America
affiliation:
  - id: '1'
    department: Department of Agronomy
    institution: University of Wisconsin-Madison
    city: Madison
    state: Wisconsin # only USA, Australia, Canada
    country: United States of America
  - id: '2' 
    department: Department of Agronomy and Horticulture
    institution: University of Nebraska-Lincoln
    city: Lincoln
    state: Nebraska # only USA, Australia, Canada
    country: United States of America
  - id: '3' 
    department: Department X
    institution: Western Illinois University
    city: Macomb
    state: Illinois # only USA, Australia, Canada
    country: United States of America
  - id: '4' 
    department: Department of Agronomy and Horticulture
    institution: University of Nebraska-Lincoln
    city: Grant
    state: Nebraska # only USA, Australia, Canada
    country: United States of America

output: rticles::frontiers_article
csl: frontiers.csl
bibliography: test.bib
---

\begin{abstract}

Abstract length and content varies depending on article type. Refer to 
\url{http://www.frontiersin.org/about/AuthorGuidelines} for abstract requirement
and length according to article type.

%All article types: you may provide up to 8 keywords; at least 5 are mandatory.
\tiny
 \keyFont{ \section{Keywords:} Text Text Text Text Text Text Text Text }

\end{abstract}

# Introduction {-}
Cite fancy references [@Neuro2013]. As demonstrated in @Gene2012, citations can
also be automatically reference. Multiple references are separated by semicolons
[@Neuro2013; @Neurobot2013].

# Results {-}

<!-- Comments should be HTML-style comments -->

## Subsection 1 {-}

You can use `R` chunks directly to plot graphs.

```{r graph, echo = TRUE, message=FALSE, fig.show = "hide"}
x <- 0:100
set.seed(999)
y <- 2 * (x + rnorm(length(x), sd = 3) + 3)
plot(x, y)
```

## Subsection 2 {-}

Frontiers requires figures to be submitted individually, in the same order as
they are referred to in the manuscript. Figures will then be automatically
embedded at the bottom of the submitted manuscript. Kindly ensure that each
table and figure is mentioned in the text and in numerical order. Permission
must be obtained for use of copyrighted material from other sources (including
the web). Please note that it is compulsory to follow figure instructions.
Figures which are not according to the guidelines will cause substantial delay
during the production process.

# Discussion

# Disclosure/Conflict-of-Interest Statement {-}

<!--  
Frontiers follows the recommendations by the International Committee of Medical
Journal Editors (http://www.icmje.org/ethical_4conflicts.html) which require
that all financial, commercial or other relationships that might be perceived by
the academic community as representing a potential conflict of interest must be
disclosed. If no such relationship exists, authors will be asked to declare that
the research was conducted in the absence of any commercial or financial
relationships that could be construed as a potential conflict of interest. When
disclosing the potential conflict of interest, the authors need to address the
following points:

 - Did you or your institution at any time receive payment or services from a
   third party for any aspect of the submitted work?
 - Please declare financial relationships with entities that could be perceived
   to influence, or that give the appearance of potentially influencing, what
   you wrote in the submitted work.
 - Please declare patents and copyrights, whether pending, issued, licensed
   and/or receiving royalties relevant to the work.
 - Please state other relationships or activities that readers could perceive to
   have influenced, or that give the appearance of potentially influencing, what
   you wrote in the submitted work.
 -->

The authors declare that the research was conducted in the absence of any
commercial or financial relationships that could be construed as a potential
conflict of interest.

# Author Contributions {-}

<!--

When determining authorship the following criteria should be observed:

 - Substantial contributions to the conception or design of the work; or the
   acquisition, analysis, or interpretation of data for the work; AND
 - Drafting the work or revising it critically for important intellectual
   content; AND
 - Final approval of the version to be published ; AND
 - Agreement to be accountable for all aspects of the work in ensuring that
   questions related to the accuracy or integrity of any part of the work are
   appropriately investigated and resolved.

Contributors who meet fewer than all 4 of the above criteria for authorship
should not be listed as authors, but they should be acknowledged.
(http://www.icmje.org/roles_a.html)

-->

The statement about the authors and contributors can be up to several sentences
long, describing the tasks of individual authors referred to by their initials
and should be included at the end of the manuscript before the References
section.

# Acknowledgments {-}

Funding:

# Supplemental Data

Supplementary Material should be uploaded separately on submission, if there are
Supplementary Figures, please include the caption in the same file as the
figure. LaTeX Supplementary Material templates can be found in the Frontiers
LaTeX folder

# References

A reference list should be automatically created here. However it won't. Pandoc 
will place the list of references at the end of the document instead. There are 
no convenient solution for now to force Pandoc to do otherwise. The easiest way 
to get around this problem is to edit the LaTeX file created by Pandoc before 
compiling it again using the traditional LaTeX commands.

# Figures {-}

```{r, Figure-1, ref.label = "graph", results = "hide", echo = FALSE, message = FALSE, fig.height=4, fig.width=4, fig.align='center', fig.cap='Figure caption', out.width = "85mm", out.height = "85mm"}
# You can also refer to code chunks from above to place figures at the bottom.
```

### 2020-11-27 Manuscript/Adaptation.tex
atualizacao

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.
%                                              %
%                                                                                                                                                         %
% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 3.4 Generated 2018/06/15 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%

\documentclass[utf8]{frontiersSCNS}

%\setcitestyle{square} % for Physics and Applied Mathematics and Statistics articles
\usepackage{url,hyperref,lineno,microtype,subcaption}
\usepackage[onehalfspacing]{setspace}

\linenumbers

% BELOW TAKEN FROM rticles plos template
%
% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Below is from frontiers
%
\bibliographystyle{frontiersinSCNS}
% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% Leave a blank line between paragraphs instead of using \\

\def\keyFont{\fontsize{8}{11}\helveticabold }

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

\def\Authors{
  Maxwel Coura Oliveira\,\textsuperscript{1},
  Amit J Jhala\,\textsuperscript{2},
  Mark Bernards\,\textsuperscript{3},
  Strahinja Stephanovic\,\textsuperscript{4},
  Chris Proctor\,\textsuperscript{2},
  Rodrigo Werle\,\textsuperscript{1*}}

\def\Address{

  \textsuperscript{1} Department of Agronomy, University of Wisconsin-Madison,  Madison,  Wisconsin,  United States of America

  \textsuperscript{2} Department of Agronomy and Horticulture, University of Nebraska-Lincoln,  Lincoln,  Nebraska,  United States of America

  \textsuperscript{3} Department X, Western Illinois University,  Macomb,  Illinois,  United States of America

  \textsuperscript{4} Department of Agronomy and Horticulture, University of Nebraska-Lincoln,  Grant,  Nebraska,  United States of America
  }

  \def\firstAuthorLast{Oliveira {et~al.}}

  \def\corrAuthor{Rodrigo Werle}\def\corrAddress{Laboratory X, Institute X, Department of Agronomy and Horticulture,
University of Wisconsin-Madison\\1575 Linden Drive\\Madison, Wisconsin, 53706 United States of America}\def\corrEmail{\href{mailto:rwerle@wisc.edu}{\nolinkurl{rwerle@wisc.edu}}}

\begin{document}
\onecolumn
\firstpage{1}

\title[Short Title]{Adaptation of Palmer amaranth to the upper Midwest United States}
\author[\firstAuthorLast]{\Authors}
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}

\maketitle

\begin{abstract}

Abstract length and content varies depending on article type. Refer to 
\url{http://www.frontiersin.org/about/AuthorGuidelines} for abstract requirement
and length according to article type.

%All article types: you may provide up to 8 keywords; at least 5 are mandatory.
\tiny
 \keyFont{ \section{Keywords:} Text Text Text Text Text Text Text Text }

\end{abstract}

\hypertarget{introduction}{%
\section*{Introduction}\label{introduction}}
\addcontentsline{toc}{section}{Introduction}

Cite fancy references (LastName1 et al., 2013). As demonstrated in
OtherAuthor and Coauthor (2012), citations can also be automatically
reference. Multiple references are separated by semicolons (LastName1 et
al., 2013; Author4 and Author5, 2013).

\hypertarget{results}{%
\section*{Results}\label{results}}
\addcontentsline{toc}{section}{Results}

\hypertarget{subsection-1}{%
\subsection*{Subsection 1}\label{subsection-1}}
\addcontentsline{toc}{subsection}{Subsection 1}

You can use \texttt{R} chunks directly to plot graphs.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{x <-}\StringTok{ }\DecValTok{0}\OperatorTok{:}\DecValTok{100}
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{999}\NormalTok{)}
\NormalTok{y <-}\StringTok{ }\DecValTok{2} \OperatorTok{*}\StringTok{ }\NormalTok{(x }\OperatorTok{+}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\KeywordTok{length}\NormalTok{(x), }\DataTypeTok{sd =} \DecValTok{3}\NormalTok{) }\OperatorTok{+}\StringTok{ }\DecValTok{3}\NormalTok{)}
\KeywordTok{plot}\NormalTok{(x, y)}
\end{Highlighting}
\end{Shaded}

\hypertarget{subsection-2}{%
\subsection*{Subsection 2}\label{subsection-2}}
\addcontentsline{toc}{subsection}{Subsection 2}

Frontiers requires figures to be submitted individually, in the same
order as they are referred to in the manuscript. Figures will then be
automatically embedded at the bottom of the submitted manuscript. Kindly
ensure that each table and figure is mentioned in the text and in
numerical order. Permission must be obtained for use of copyrighted
material from other sources (including the web). Please note that it is
compulsory to follow figure instructions. Figures which are not
according to the guidelines will cause substantial delay during the
production process.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{disclosureconflict-of-interest-statement}{%
\section*{Disclosure/Conflict-of-Interest
Statement}\label{disclosureconflict-of-interest-statement}}
\addcontentsline{toc}{section}{Disclosure/Conflict-of-Interest
Statement}

The authors declare that the research was conducted in the absence of
any commercial or financial relationships that could be construed as a
potential conflict of interest.

\hypertarget{author-contributions}{%
\section*{Author Contributions}\label{author-contributions}}
\addcontentsline{toc}{section}{Author Contributions}

The statement about the authors and contributors can be up to several
sentences long, describing the tasks of individual authors referred to
by their initials and should be included at the end of the manuscript
before the References section.

\hypertarget{acknowledgments}{%
\section*{Acknowledgments}\label{acknowledgments}}
\addcontentsline{toc}{section}{Acknowledgments}

Funding:

\hypertarget{supplemental-data}{%
\section{Supplemental Data}\label{supplemental-data}}

Supplementary Material should be uploaded separately on submission, if
there are Supplementary Figures, please include the caption in the same
file as the figure. LaTeX Supplementary Material templates can be found
in the Frontiers LaTeX folder

\hypertarget{references}{%
\section{References}\label{references}}

A reference list should be automatically created here. However it won't.
Pandoc will place the list of references at the end of the document
instead. There are no convenient solution for now to force Pandoc to do
otherwise. The easiest way to get around this problem is to edit the
LaTeX file created by Pandoc before compiling it again using the
traditional LaTeX commands.

\hypertarget{figures}{%
\section*{Figures}\label{figures}}
\addcontentsline{toc}{section}{Figures}

\begin{figure}

{\centering \includegraphics[width=85mm,height=85mm]{Adaptation_files/figure-latex/Figure-1-1}

}

\caption{Figure caption}\label{fig:Figure-1}
\end{figure}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Neurobot2013}{}%
Author4, N., and Author5, N. (2013). Title of the article.
\emph{Frontiers in Neurorobotics} 7.
doi:\href{https://doi.org/10.3389/fnbot.2013.56789}{10.3389/fnbot.2013.56789}.

\leavevmode\hypertarget{ref-Neuro2013}{}%
LastName1, A., LastName2, A., and LastName2, A. (2013). Article title.
\emph{Frontiers in Neuroscience} 30, 10127--10134.
doi:\href{https://doi.org/10.3389/fnins.2013.12345}{10.3389/fnins.2013.12345}.

\leavevmode\hypertarget{ref-Gene2012}{}%
OtherAuthor, N., and Coauthor, N. S. (2012). Article title.
\emph{Frontiers in Genetics} 30, 16417--16418.
doi:\href{https://doi.org/10.3389/fgene.2012.54321}{10.3389/fgene.2012.54321}.

\end{document}

### 2021-06-07 new_analyses.Rmd
as

---
title: "new_analses"
author: "Maxwel Coura Oliveira"
date: "6/7/2021"
output: html_document
---

```{r}
library(tidyverse)
library(lubridate)
```

```{r}
data <- read_csv("Data.csv")
```

```{r}
data %>% 
  filter(location != "grant") %>% 
  mutate_if(is_character, as_factor) -> data1
```

```{r}
data1 %>% 
  ggplot(aes(x = crop, y = weight, color = crop)) +
  geom_boxplot() +
  geom_point() +
  facet_grid(location ~ planting)
```

### 2021-06-08 new_analyses.Rmd
as

@@
 library(lubridate)
+library(drc)
+library(broom)
 ```
@@
 ```{r}
-data <- read_csv("Data.csv")
+data <- read_csv("data2.csv")
 ```
@@
 data %>% 
-  filter(location != "grant") %>% 
-  mutate_if(is_character, as_factor) -> data1
+  filter(location != "grant") -> data1
 ```
@@
 data1 %>% 
-  ggplot(aes(x = crop, y = weight, color = crop)) +
-  geom_boxplot() +
+  rename(month = planting,
+         planting = planting_1) %>% 
+  mutate(planting = mdy(planting),
+         harvest = mdy(harvest),
+         doyp = yday(planting)) %>% 
+  mutate_if(is_character, as_factor) -> data2
+```
+
+```{r}
+data2 %>% 
+  mutate(harvest = as.character(harvest)) %>% 
+  mutate(harvest = str_replace(harvest, "2018", "2019"),
+         harvest = ymd(harvest)) %>% 
+  filter(!is.na(harvest)) %>% 
+  dplyr::select(location, month, crop, harvest) %>% 
+  group_by(location, month, crop) %>% 
+  count(harvest, name = "harvest_day") -> harvest
+```
+
+```{r}
+data2 %>% 
+  filter(!is.na(harvest)) %>% 
+  dplyr::select(location, month, crop, harvest) %>% 
+  group_by(location, month, crop) %>% 
+  count(location, month, crop, name = "total") -> total
+```
+
+```{r}
+total %>% 
+  left_join(harvest) %>% 
+  mutate(prop = harvest_day / total) %>% 
+  mutate(lag = lag(prop)) %>% 
+  mutate(prob = cumsum(prop)) -> cum_sum
+```
+
+```{r}
+data2 %>% 
+  mutate(harvest = str_replace(harvest, "2018", "2019"),
+         harvest = ymd(harvest)) %>% 
+  filter(!is.na(harvest)) %>% 
+  left_join(cum_sum) %>% 
+  rename(harvest_com = harvest) %>% 
+  dplyr::select(-lag) -> data3
+```
+
+```{r}
+data3 %>% 
+  ggplot(aes(x = dap, y = prob, color = crop)) +
   geom_point() +
-  facet_grid(location ~ planting)
+  facet_grid(location ~ month)
 ```
+
+# Arlington
+
+```{r}
+arlington <- data3 %>% 
+  filter(location == "arlington") %>% 
+  filter(month == "July")
+```
+
+```{r}
+model_arl <- drm(prob*100 ~ doyh, crop, fct = W2.3(), data = arlington)
+
+plot(model_arl, col = c(1,2,3), type = "average", ylab = "Cumulative flowering (%)",
+     xlab = "Day of the year")
+```
+
+```{r}
+#broom::tidy(model_arl)
+summary(model_arl)
+```
+
+```{r}
+ED(model_arl, c(10, 50, 90), interval = "delta", type = "absolute")
+```

### 2021-06-10 new_analyses.Rmd
as

@@
 library(broom)
+library(tidytext)
+library(ggthemes)
 ```
@@
 ```
-
-# Arlington
-
 ```{r}
-arlington <- data3 %>% 
-  filter(location == "arlington") %>% 
-  filter(month == "July")
+car::leveneTest(prob ~ location, data = data3)
 ```
@@
 ```{r}
-model_arl <- drm(prob*100 ~ doyh, crop, fct = W2.3(), data = arlington)
+data3 %>% 
+  unite("trt", c("location", "crop"), sep = "-") -> data4
+```
 
-plot(model_arl, col = c(1,2,3), type = "average", ylab = "Cumulative flowering (%)",
+# June
+
+```{r}
+june <-  data4 %>% 
+  filter(month == "June")
+model <- drm(prob*100 ~ doyh, trt, fct = W2.3(fixed = c(NA, 100, NA)), data = june)
+
+plot(model, col = c(1,2,3), type = "average", ylab = "Cumulative flowering (%)",
+     xlab = "Day of the year")
+```
+```{r}
+#broom::tidy(model_arl)
+summary(model)
+```
+
+```{r}
+ed_values <- ED(model, c(10, 50, 90), interval = "delta", type = "absolute")
+```
+
+```{r}
+ed_values %>% 
+  as.data.frame() %>% 
+  rownames_to_column(., var = "name") %>% 
+  as_tibble() %>% 
+  separate(name, c("param", "trt", "ED"), sep = ":") %>% 
+  janitor::clean_names() -> ed_values1
+```
+
+```{r}
+ed_values1 %>% 
+  mutate_if(is_character, as_factor) %>% 
+  mutate(trt = reorder_within(trt, by = estimate, within = ed)) %>% 
+  ggplot(aes(x = trt, y = estimate)) +
+  scale_x_reordered() +
+  geom_point() +
+  geom_linerange(aes(ymin = lower, ymax = upper)) +
+  labs(y = NULL, x = NULL,
+       title = "DOY of 10, 50, and 90% Palmer amaranth flowering") +
+  coord_flip() +
+#  scale_color_brewer(palette = "Paired") +
+  facet_wrap(~ ed, scales = "free") +
+  theme_bw() +
+  theme(legend.position = "none") +
+  ggsave("june.png", width = 12, height = 5)
+```
+
+# July
+
+```{r}
+july <-  data4 %>% 
+  filter(month == "July")
+model1 <- drm(prob*100 ~ doyh, trt, fct = W2.3(fixed = c(NA, 100, NA)), data = july)
+
+plot(model1, col = c(1,2,3), type = "average", ylab = "Cumulative flowering (%)",
      xlab = "Day of the year")
@@
 #broom::tidy(model_arl)
-summary(model_arl)
+summary(model1)
 ```
@@
 ```{r}
-ED(model_arl, c(10, 50, 90), interval = "delta", type = "absolute")
+ed_values <- ED(model1, c(10, 50, 90), interval = "delta", type = "absolute")
 ```

### 2021-06-16 new_analyses.Rmd
as

@@
 library(ggthemes)
+library(ggtext)
 ```
@@
   separate(name, c("param", "trt", "ED"), sep = ":") %>% 
-  janitor::clean_names() -> ed_values1
+  janitor::clean_names() %>% 
+  separate("trt", c("location", "crop"), sep = "-") %>% 
+  mutate(location = str_to_title(location)) %>% 
+  mutate_if(is_character, as_factor) -> ed_values1
+```
+
+  annotate("text", x = 0.65, y = 168, label = "June", 
+           color = "#8B8000", size = 2) +
+  annotate("text", x = 0.65, y = 197, label = "July", 
+           color = "#80461b", size = 2) +
+  annotate("text", x = 0.65, y = 228.5, label = "August", 
+           color = "darkgreen", size = 2) +
+  annotate("text", x = 0.65, y = 259, label = "September", 
+           color = "blue", size = 2) +
+
+```{r}
+label1 <- tibble(ed = "50% flowering", crop = "corn", 
+                   x = 3.5, y = 185, label = "Transplanting \nday")
+rect1 <- tibble(ed = "50% flowering", crop = "corn", 
+               x = 3.5, y = 153, xend = 3.5, yend = 165)
 ```
@@
 ```{r}
+month <- tibble(ed = "90% flowering",
+                crop = "corn",
+                x = c(0.65, 0.65, 0.65, 0.65),
+                y = c(168, 197, 228.5, 259), 
+                label = c("June", "July", "August", "September"))
+
 ed_values1 %>% 
-  mutate_if(is_character, as_factor) %>% 
-  mutate(trt = reorder_within(trt, by = estimate, within = ed)) %>% 
-  ggplot(aes(x = trt, y = estimate)) +
+  filter(crop == "corn") %>% 
+  mutate(ed = case_when(
+    ed == "10" ~ "10% flowering",
+    ed == "50" ~ "50% flowering",
+    ed == "90" ~ "90% flowering",
+    TRUE ~ "ed")) %>% 
+  group_by(crop,  ed) %>% 
+  mutate(location = reorder_within(location, 
+                                   by = estimate, 
+                                   within = ed)) %>% 
+  ggplot(aes(x = location, y = estimate)) +
   scale_x_reordered() +
-  geom_point() +
-  geom_linerange(aes(ymin = lower, ymax = upper)) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 153, ymax= 183), 
+            fill = "#FFFDD0", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 183, ymax= 214), 
+            fill = "#80461b", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 214, ymax= 244), 
+            fill = "lightgreen", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 244, ymax= 274), 
+            fill = "blue", alpha=0.05) +
+  geom_hline(yintercept = 153, size = 1.1, 
+             color = "red",
+             alpha = 1) +
+  geom_point(size = 2) +
+  geom_linerange(aes(ymin = lower, ymax = upper), size = 1.2) +
+  geom_text(data = month, aes(x = x, y = y), label = month$label,
+            size = 2, color = c("#8B8000", "#80461b", 
+                                  "darkgreen", "blue")) +
   labs(y = NULL, x = NULL,
-       title = "DOY of 10, 50, and 90% Palmer amaranth flowering") +
+       title = NULL) +
+  facet_grid(ed ~ crop, scales = "free_y") +
+  geom_curve(data = rect1, aes(x = x, y = y, 
+                               xend = xend, yend = yend),
+             arrow = arrow(length = unit(0.07, "inch")), 
+               size = 0.4, curvature = 0.3) + 
+  geom_text(data = label1, 
+            aes(x = x, y = y), 
+            size = 2,
+            label = label2$label, color = "red") +
   coord_flip() +
-#  scale_color_brewer(palette = "Paired") +
-  facet_wrap(~ ed, scales = "free") +
-  theme_bw() +
-  theme(legend.position = "none") +
-  ggsave("june.png", width = 12, height = 5)
+  theme_test() +
+  theme(legend.position = "none",
+        plot.title = element_markdown(),
+        strip.text = element_markdown(face = "bold"),
+        plot.title.position = "plot") -> p1c
+```
+
+```{r}
+month <- tibble(ed = "90% flowering",
+                crop = "fallow",
+                x = c(0.65, 0.65, 0.65, 0.65),
+                y = c(168, 197, 228.5, 259), 
+                label = c("June", "July", "August", "September"))
+
+ed_values1 %>% 
+  filter(crop == "fallow") %>% 
+  mutate(ed = case_when(
+    ed == "10" ~ "10% flowering",
+    ed == "50" ~ "50% flowering",
+    ed == "90" ~ "90% flowering",
+    TRUE ~ "ed")) %>% 
+  group_by(crop,  ed) %>% 
+  mutate(location = reorder_within(location, 
+                                   by = estimate, 
+                                   within = ed)) %>% 
+  ggplot(aes(x = location, y = estimate)) +
+  scale_x_reordered() +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 153, ymax= 183), 
+            fill = "#FFFDD0", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 183, ymax= 214), 
+            fill = "#80461b", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 214, ymax= 244), 
+            fill = "lightgreen", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 244, ymax= 274), 
+            fill = "blue", alpha=0.05) +
+  geom_hline(yintercept = 153, size = 1.1, 
+             color = "red",
+             alpha = 1) +
+  geom_point(size = 2) +
+  geom_linerange(aes(ymin = lower, ymax = upper), size = 1.2) +
+  geom_text(data = month, aes(x = x, y = y), label = month$label,
+            size = 2, color = c("#8B8000", "#80461b", 
+                                  "darkgreen", "blue")) +
+  labs(y = "Day of year", x = NULL,
+       title = NULL) +
+  facet_grid(ed ~ crop, scales = "free_y") +
+  coord_flip() +
+  theme_test() +
+  theme(legend.position = "none",
+        plot.title = element_markdown(),
+        strip.text = element_markdown(face = "bold"),
+        plot.title.position = "plot") -> p1f
+```
+
+```{r}
+month <- tibble(ed = "90% flowering",
+                crop = "corn",
+                x = c(0.65, 0.65, 0.65, 0.65),
+                y = c(168, 197, 228.5, 259), 
+                label = c("June", "July", "August", "September"))
+
+ed_values1 %>% 
+  filter(crop == "corn") %>% 
+  mutate(ed = case_when(
+    ed == "10" ~ "10% flowering",
+    ed == "50" ~ "50% flowering",
+    ed == "90" ~ "90% flowering",
+    TRUE ~ "ed")) %>% 
+  group_by(crop,  ed) %>% 
+  mutate(location = reorder_within(location, 
+                                   by = estimate, 
+                                   within = ed)) %>% 
+  ggplot(aes(x = location, y = estimate)) +
+  scale_x_reordered() +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 153, ymax= 183), 
+            fill = "#FFFDD0", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 183, ymax= 214), 
+            fill = "#80461b", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 214, ymax= 244), 
+            fill = "lightgreen", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 244, ymax= 274), 
+            fill = "blue", alpha=0.05) +
+  geom_hline(yintercept = 153, size = 1.1, 
+             color = "red",
+             alpha = 1) +
+  geom_point(size = 2) +
+  geom_linerange(aes(ymin = lower, ymax = upper), size = 1.2) +
+  geom_text(data = month, aes(x = x, y = y), label = month$label,
+            size = 2, color = c("#8B8000", "#80461b", 
+                                  "darkgreen", "blue")) +
+  labs(y = NULL, x = NULL,
+       title = NULL) +
+  facet_grid(ed ~ crop, scales = "free_y") +
+  coord_flip() +
+  theme_test() +
+  theme(legend.position = "none",
+        plot.title = element_markdown(),
+        strip.text = element_markdown(face = "bold"),
+        plot.title.position = "plot") -> p1s
+```
+
+```{r}
+library(patchwork)
+f1 <- p1c | p1f | p1s +
+  plot_annotation(
+  title = "DOY of 10, 50, and 90% Palmer amaranth flowering after first transplanting") 
+  ggsave("June.png", width = 9, height = 5, dpi = 600, f1)
 ```
@@
   filter(month == "July")
-model1 <- drm(prob*100 ~ doyh, trt, fct = W2.3(fixed = c(NA, 100, NA)), data = july)
+model2 <- drm(prob*100 ~ doyh, trt, fct = W2.3(fixed = c(NA, 100, NA)), data = july)
 
-plot(model1, col = c(1,2,3), type = "average", ylab = "Cumulative flowering (%)",
+plot(model2, col = c(1,2,3), type = "average", ylab = "Cumulative flowering (%)",
      xlab = "Day of the year")
@@
 #broom::tidy(model_arl)
-summary(model1)
+summary(model2)
 ```
@@
 ```{r}
-ed_values <- ED(model1, c(10, 50, 90), interval = "delta", type = "absolute")
+ed_values2 <- ED(model2, c(10, 50, 90), interval = "delta", type = "absolute")
 ```
+
+```{r}
+ed_values2 %>% 
+  as.data.frame() %>% 
+  rownames_to_column(., var = "name") %>% 
+  as_tibble() %>% 
+  separate(name, c("param", "trt", "ED"), sep = ":") %>% 
+  janitor::clean_names() %>% 
+  separate("trt", c("location", "crop"), sep = "-") %>% 
+  mutate(location = str_to_title(location)) %>% 
+  mutate_if(is_character, as_factor) -> ed_values3
+```
+
+```{r}
+label2 <- tibble(ed = "50% flowering", crop = "corn", 
+                   x = 3.5, y = 211, label = "Transplanting \nday")
+rect2 <- tibble(ed = "50% flowering", crop = "corn", 
+               x = 3.5, y = 183, xend = 3.5, yend = 195)
+```
+
+```{r}
+month <- tibble(ed = "90% flowering",
+                crop = "corn",
+                x = c( 0.65, 0.65, 0.65),
+                y = c(197, 228.5, 259), 
+                label = c("July", "August", "September"))
+```
+
+```{r}
+ed_values3 %>% 
+  filter(crop == "corn") %>% 
+  mutate(ed = case_when(
+    ed == "10" ~ "10% flowering",
+    ed == "50" ~ "50% flowering",
+    ed == "90" ~ "90% flowering",
+    TRUE ~ "ed")) %>% 
+  group_by(crop,  ed) %>% 
+  mutate(location = reorder_within(location, 
+                                   by = estimate, 
+                                   within = ed)) %>%
+  ggplot(aes(x = location,
+             y = estimate)) +
+  scale_x_reordered() +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 183, ymax= 214), 
+            fill = "#80461b", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 214, ymax= 244), 
+            fill = "lightgreen", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 244, ymax= 274), 
+            fill = "blue", alpha=0.05) +
+  geom_hline(yintercept = 183, size = 1.1, 
+             color = "red",
+             alpha = 1) +
+  geom_point(size = 2) +
+  geom_linerange(aes(ymin = lower, ymax = upper), size = 1.2) +
+  geom_text(data = month, aes(x = x, y = y), label = month$label,
+            size = 2.5, color = c("#80461b", "darkgreen", "blue")) +
+  labs(y = NULL, x = NULL) +
+  facet_grid(ed ~ crop, scales = "free_y") +
+  geom_curve(data = rect2, aes(x = x, y = y, 
+                               xend = xend, yend = yend),
+             arrow = arrow(length = unit(0.07, "inch")), 
+               size = 0.4, curvature = 0.3) + 
+  geom_text(data = label2, 
+            aes(x = x, y = y), 
+            size = 2,
+            label = label2$label, color = "red") +
+  coord_flip() +
+  theme_test() +
+  theme(legend.position = "none",
+        strip.text = element_markdown(face = "bold"),
+        plot.title = element_markdown(),
+        plot.title.position = "plot") -> p2c
+```
+
+```{r}
+month <- tibble(ed = "90% flowering",
+                crop = "fallow",
+                x = c( 0.65, 0.65, 0.65),
+                y = c(197, 228.5, 259), 
+                label = c("July", "August", "September"))
+```
+
+```{r}
+ed_values3 %>% 
+  filter(crop == "fallow") %>% 
+  mutate(ed = case_when(
+    ed == "10" ~ "10% flowering",
+    ed == "50" ~ "50% flowering",
+    ed == "90" ~ "90% flowering",
+    TRUE ~ "ed")) %>% 
+  group_by(crop,  ed) %>% 
+  mutate(location = reorder_within(location, 
+                                   by = estimate, 
+                                   within = ed)) %>%
+  ggplot(aes(x = location,
+             y = estimate)) +
+  scale_x_reordered() +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 183, ymax= 214), 
+            fill = "#80461b", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 214, ymax= 244), 
+            fill = "lightgreen", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 244, ymax= 274), 
+            fill = "blue", alpha=0.05) +
+  geom_hline(yintercept = 183, size = 1.1, 
+             color = "red",
+             alpha = 1) +
+  geom_point(size = 2) +
+  geom_linerange(aes(ymin = lower, ymax = upper), size = 1.2) +
+  geom_text(data = month, aes(x = x, y = y), label = month$label,
+            size = 2.5, color = c("#80461b", "darkgreen", "blue")) +
+  labs(y = "Day of year", x = NULL) +
+  facet_grid(ed ~ crop, scales = "free_y") +
+  coord_flip() +
+  theme_test() +
+  theme(legend.position = "none",
+        strip.text = element_markdown(face = "bold"),
+        plot.title = element_markdown(),
+        plot.title.position = "plot") -> p2f
+```
+
+```{r}
+month <- tibble(ed = "90% flowering",
+                crop = "soybean",
+                x = c( 0.65, 0.65, 0.65),
+                y = c(197, 228.5, 259), 
+                label = c("July", "August", "September"))
+
+ed_values3 %>% 
+  filter(crop == "soybean") %>% 
+  mutate(ed = case_when(
+    ed == "10" ~ "10% flowering",
+    ed == "50" ~ "50% flowering",
+    ed == "90" ~ "90% flowering",
+    TRUE ~ "ed")) %>% 
+  group_by(crop,  ed) %>% 
+  mutate(location = reorder_within(location, 
+                                   by = estimate, 
+                                   within = ed)) %>%
+  ggplot(aes(x = location,
+             y = estimate)) +
+  scale_x_reordered() +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 183, ymax= 214), 
+            fill = "#80461b", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 214, ymax= 244), 
+            fill = "lightgreen", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
+                        ymin= 244, ymax= 274), 
+            fill = "blue", alpha=0.05) +
+  geom_hline(yintercept = 183, size = 1.1, 
+             color = "red",
+             alpha = 1) +
+  geom_point(size = 2) +
+  geom_linerange(aes(ymin = lower, ymax = upper), size = 1.2) +
+  geom_text(data = month, aes(x = x, y = y), label = month$label,
+            size = 2.5, color = c("#80461b", "darkgreen", "blue")) +
+  labs(y = NULL, x = NULL) +
+  facet_grid(ed ~ crop, scales = "free_y") +
+  coord_flip() +
+  theme_test() +
+  theme(legend.position = "none",
+        strip.text = element_markdown(face = "bold"),
+        plot.title = element_markdown(),
+        plot.title.position = "plot") -> p2s
+p2s
+```
+
+```{r}
+library(patchwork)
+f3 <- p2c | p2f | p2s +
+  plot_annotation(
+  title = "DOY of 10, 50, and 90% Palmer amaranth flowering after second transplanting") 
+  ggsave("July.png", width = 9, height = 5, dpi = 600, f3)
+```

### 2021-06-23 Frontiers/Frontiers.Rmd
gh

--- 2020-11-27 Manuscript/Adaptation.Rmd
@@
 ---
-title: "Adaptation of Palmer amaranth to the upper Midwest United States"
+title: "Adaptation of Palmer amaranth to the upper Midwest"
 runtitle: "Short Title"
@@
 author:
-  - name: Maxwel Coura Oliveira
+  - name: Maxwel C Oliveira
     affiliation: '1'
@@
     affiliation: '3'
-  - name: Strahinja Stephanovic
-    affiliation: '4'
   - name: Chris Proctor
+    affiliation: '2'
+  - name: Strahinja Stepanovic
     affiliation: '2'
@@
     affiliation: '1'
-    email: rwerle@wisc.edu # Indicates corresponding Author
-    institution: Laboratory X, Institute X, Department of Agronomy and Horticulture, University of Wisconsin-Madison
-    street: 1575 Linden Drive
+    email: rwerle@uwisc.edu # Indicates corresponding Author
+    institution: Department of Agronomy, University of Wisconsin-Madison, United States
+    street: 1575 Linden Dr
     city: Madison
     state: Wisconsin # only USA, Australia, Canada
-    zip: 53706
-    country: United States of America
+    zip: 53705
+    country: United States
 affiliation:
@@
     state: Wisconsin # only USA, Australia, Canada
-    country: United States of America
+    country: United States
   - id: '2' 
@@
     state: Nebraska # only USA, Australia, Canada
-    country: United States of America
+    country: United States
   - id: '3' 
-    department: Department X
+    department: Department of Agronomy
     institution: Western Illinois University
     city: Macomb
-    state: Illinois # only USA, Australia, Canada
-    country: United States of America
-  - id: '4' 
-    department: Department of Agronomy and Horticulture
-    institution: University of Nebraska-Lincoln
-    city: Grant
-    state: Nebraska # only USA, Australia, Canada
-    country: United States of America
-
+    state: Illnois # only USA, Australia, Canada
+    country: United States
 output: rticles::frontiers_article
@@
 \tiny
- \keyFont{ \section{Keywords:} Text Text Text Text Text Text Text Text }
+ \keyFont{ \section{Keywords:} Text Text Text Text Text Text Evolution Weed }
 
@@
 [@Neuro2013; @Neurobot2013].
+
+# Material and Methods {-}
+
+## Plant material {-}
+
+The study was performed with a *A. palmeri* accession (Per1) from Perkins County, Nebraska. Per1 accession collection is documented in [], with no  confirmed herbicide resistance.
 

### 2021-06-23 Frontiers/Frontiers.tex
gh

--- 2020-11-27 Manuscript/Adaptation.tex
@@
 
+% Pandoc citation processing
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc} % for calculating minipage widths
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+
 \usepackage{color}
@@
 \def\Authors{
-  Maxwel Coura Oliveira\,\textsuperscript{1},
+  Maxwel C Oliveira\,\textsuperscript{1},
   Amit J Jhala\,\textsuperscript{2},
   Mark Bernards\,\textsuperscript{3},
-  Strahinja Stephanovic\,\textsuperscript{4},
   Chris Proctor\,\textsuperscript{2},
+  Strahinja Stepanovic\,\textsuperscript{2},
   Rodrigo Werle\,\textsuperscript{1*}}
@@
 
-  \textsuperscript{1} Department of Agronomy, University of Wisconsin-Madison,  Madison,  Wisconsin,  United States of America
+  \textsuperscript{1} Department of Agronomy, University of
+Wisconsin-Madison,  Madison,  Wisconsin,  United States
 
-  \textsuperscript{2} Department of Agronomy and Horticulture, University of Nebraska-Lincoln,  Lincoln,  Nebraska,  United States of America
+  \textsuperscript{2} Department of Agronomy and
+Horticulture, University of
+Nebraska-Lincoln,  Lincoln,  Nebraska,  United States
 
-  \textsuperscript{3} Department X, Western Illinois University,  Macomb,  Illinois,  United States of America
-
-  \textsuperscript{4} Department of Agronomy and Horticulture, University of Nebraska-Lincoln,  Grant,  Nebraska,  United States of America
+  \textsuperscript{3} Department of Agronomy, Western Illinois
+University,  Macomb,  Illnois,  United States
   }
@@
 
-  \def\corrAuthor{Rodrigo Werle}\def\corrAddress{Laboratory X, Institute X, Department of Agronomy and Horticulture,
-University of Wisconsin-Madison\\1575 Linden Drive\\Madison, Wisconsin, 53706 United States of America}\def\corrEmail{\href{mailto:rwerle@wisc.edu}{\nolinkurl{rwerle@wisc.edu}}}
+  \def\corrAuthor{Rodrigo Werle}\def\corrAddress{Department of Agronomy,
+University of Wisconsin-Madison, United States\\1575 Linden
+Dr\\Madison, Wisconsin, 53705 United
+States}\def\corrEmail{\href{mailto:rwerle@uwisc.edu}{\nolinkurl{rwerle@uwisc.edu}}}
 
@@
 
-\title[Short Title]{Adaptation of Palmer amaranth to the upper Midwest United States}
+\title[Short Title]{Adaptation of Palmer amaranth to the upper Midwest}
 \author[\firstAuthorLast]{\Authors}
@@
 al., 2013; Author4 and Author5, 2013).
+
+\hypertarget{material-and-methods}{%
+\section*{Material and Methods}\label{material-and-methods}}
+\addcontentsline{toc}{section}{Material and Methods}
+
+\hypertarget{plant-material}{%
+\subsection*{Plant material}\label{plant-material}}
+\addcontentsline{toc}{subsection}{Plant material}
+
+The study was performed with a \emph{A. palmeri} accession (Per1) from
+Perkins County, Nebraska. The accession Per1 was documented as
+glyphosate-susceptible {[}{]}.
 
@@
 \begin{Highlighting}[]
-\NormalTok{x <-}\StringTok{ }\DecValTok{0}\OperatorTok{:}\DecValTok{100}
-\KeywordTok{set.seed}\NormalTok{(}\DecValTok{999}\NormalTok{)}
-\NormalTok{y <-}\StringTok{ }\DecValTok{2} \OperatorTok{*}\StringTok{ }\NormalTok{(x }\OperatorTok{+}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\KeywordTok{length}\NormalTok{(x), }\DataTypeTok{sd =} \DecValTok{3}\NormalTok{) }\OperatorTok{+}\StringTok{ }\DecValTok{3}\NormalTok{)}
-\KeywordTok{plot}\NormalTok{(x, y)}
+\NormalTok{x }\OtherTok{\textless{}{-}} \DecValTok{0}\SpecialCharTok{:}\DecValTok{100}
+\FunctionTok{set.seed}\NormalTok{(}\DecValTok{999}\NormalTok{)}
+\NormalTok{y }\OtherTok{\textless{}{-}} \DecValTok{2} \SpecialCharTok{*}\NormalTok{ (x }\SpecialCharTok{+} \FunctionTok{rnorm}\NormalTok{(}\FunctionTok{length}\NormalTok{(x), }\AttributeTok{sd =} \DecValTok{3}\NormalTok{) }\SpecialCharTok{+} \DecValTok{3}\NormalTok{)}
+\FunctionTok{plot}\NormalTok{(x, y)}
 \end{Highlighting}
@@
 
-{\centering \includegraphics[width=85mm,height=85mm]{Adaptation_files/figure-latex/Figure-1-1}
+{\centering \includegraphics[width=85mm,height=85mm]{Frontiers_files/figure-latex/Figure-1-1}
 
@@
 \hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
 \leavevmode\hypertarget{ref-Neurobot2013}{}%
@@
 
+\end{CSLReferences}
+
 \end{document}

### 2021-06-23 new_analyses.Rmd
s
gh

@@
 ```
+
+#### -----
+
+```{r}
+library(ggdist)
+library(PupillometryR)
+```
+
+# Biomass
+
+```{r}
+data <- read_csv("data2.csv")
+```
+
+```{r}
+data %>% 
+  rename(month = planting,
+         planting = planting_1) %>% 
+  mutate(planting = mdy(planting),
+         harvest = mdy(harvest),
+         doyp = yday(planting)) %>% 
+  mutate_if(is_character, as_factor) -> data1
+```
+
+```{r}
+data1 %>% 
+  mutate(month = fct_recode(month,
+                            "first planting" = "June",
+                            "second planting" = "July")) -> data1
+```
+
+```{r}
+data1 %>% 
+  mutate(location = str_to_title(location)) %>% 
+  filter(crop == "corn") %>% 
+  filter(!is.na(weight)) %>% 
+  ggplot(aes(x = fct_reorder(location, weight), y = weight,
+             color = location, 
+             fill = location)) + 
+  geom_jitter(position = position_nudge(x = -0.2, y = 0), alpha = 0.1) +
+  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
+  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), 
+                   stat = "ydensity",
+                   alpha = 0.8, scale = "count") +
+  stat_summary(fun = "mean", colour = "red", size = 1, geom = "point") +
+  scale_color_viridis_d(option = "H") + 
+  scale_fill_viridis_d(option = "H") + 
+  theme_test() +
+  facet_grid(month ~ crop, scales = "free") +
+  coord_flip() +
+  labs(x = NULL, y = NULL) +
+  theme(legend.position = "none",
+        strip.text = element_markdown(face = "bold")) -> fg1
+```
+
+```{r}
+data1 %>% 
+  mutate(location = str_to_title(location)) %>% 
+  filter(crop == "fallow") %>% 
+  filter(!is.na(weight)) %>% 
+  ggplot(aes(x = fct_reorder(location, weight), y = weight,
+             color = location, 
+             fill = location)) + 
+  geom_jitter(position = position_nudge(x = -0.2, y = 0), alpha = 0.1) +
+  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
+  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), 
+                   stat = "ydensity",
+                   alpha = 0.8, scale = "count") +
+  stat_summary(fun = "mean", colour = "red", size = 1, geom = "point") +
+  scale_color_viridis_d(option = "H") + 
+  scale_fill_viridis_d(option = "H") + 
+  theme_test() +
+  facet_grid(month ~ crop, scales = "free") +
+  coord_flip() +
+  labs(x = NULL, y = "Biomass (g)") +
+  theme(legend.position = "none",
+        strip.text = element_markdown(face = "bold")) -> fg2
+```
+
+```{r}
+data1 %>% 
+  mutate(location = str_to_title(location)) %>% 
+  filter(crop == "soybean") %>% 
+  filter(!is.na(weight)) %>% 
+  ggplot(aes(x = fct_reorder(location, weight), y = weight,
+             color = location, 
+             fill = location)) + 
+  geom_jitter(position = position_nudge(x = -0.2, y = 0), alpha = 0.1) +
+  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
+  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), 
+                   stat = "ydensity",
+                   alpha = 0.8, scale = "count") +
+  stat_summary(fun = "mean", colour = "red", size = 1, geom = "point") +
+  scale_color_viridis_d(option = "H") + 
+  scale_fill_viridis_d(option = "H") + 
+  theme_test() +
+  facet_grid(month ~ crop, scales = "free") +
+  coord_flip() +
+  labs(x = NULL, y = NULL) +
+  theme(legend.position = "none",
+        strip.text = element_markdown(face = "bold")) -> fg3
+```
+
+```{r}
+f4 <- fg1 | fg2 | fg3 +
+  plot_annotation(
+  title = NULL) 
+  ggsave("biomass.png", width = 9, height = 5, dpi = 600, f4)
+```
+
+# Height
+
+```{r}
+data1 %>% 
+  mutate(location = str_to_title(location)) %>% 
+  filter(crop == "corn") %>% 
+  filter(!is.na(height)) %>% 
+  ggplot(aes(x = fct_reorder(location, height), y = height,
+             color = location, 
+             fill = location)) + 
+  geom_jitter(position = position_nudge(x = -0.2, y = 0), alpha = 0.1) +
+  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
+  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), 
+                   stat = "ydensity",
+                   alpha = 0.8, scale = "count") +
+  stat_summary(fun = "mean", colour = "red", size = 1, geom = "point") +
+  scale_color_viridis_d(option = "H") + 
+  scale_fill_viridis_d(option = "H") + 
+  theme_test() +
+  facet_grid(month ~ crop, scales = "free") +
+  coord_flip() +
+  labs(x = NULL, y = NULL) +
+  theme(legend.position = "none",
+        strip.text = element_markdown(face = "bold")) -> fg4
+```
+
+```{r}
+data1 %>% 
+  mutate(location = str_to_title(location)) %>% 
+  filter(crop == "fallow") %>% 
+  filter(!is.na(height)) %>% 
+  ggplot(aes(x = fct_reorder(location, height), y = height,
+             color = location, 
+             fill = location)) + 
+  geom_jitter(position = position_nudge(x = -0.2, y = 0), alpha = 0.1) +
+  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
+  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), 
+                   stat = "ydensity",
+                   alpha = 0.8, scale = "count") +
+  stat_summary(fun = "mean", colour = "red", size = 1, geom = "point") +
+  scale_color_viridis_d(option = "H") + 
+  scale_fill_viridis_d(option = "H") + 
+  theme_test() +
+  facet_grid(month ~ crop, scales = "free") +
+  coord_flip() +
+  labs(x = NULL, y = "Height (cm)") +
+  theme(legend.position = "none",
+        strip.text = element_markdown(face = "bold")) -> fg5
+```
+
+```{r}
+data1 %>% 
+  mutate(location = str_to_title(location)) %>% 
+  filter(crop == "soybean") %>% 
+  filter(!is.na(height)) %>% 
+  ggplot(aes(x = fct_reorder(location, height), y = height,
+             color = location, 
+             fill = location)) + 
+  geom_jitter(position = position_nudge(x = -0.2, y = 0), alpha = 0.1) +
+  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
+  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), 
+                   stat = "ydensity",
+                   alpha = 0.8, scale = "count") +
+  stat_summary(fun = "mean", colour = "red", size = 1, geom = "point") +
+  scale_color_viridis_d(option = "H") + 
+  scale_fill_viridis_d(option = "H") + 
+  theme_test() +
+  facet_grid(month ~ crop, scales = "free") +
+  coord_flip() +
+  labs(x = NULL, y = NULL) +
+  theme(legend.position = "none",
+        strip.text = element_markdown(face = "bold")) -> fg6
+```
+
+```{r}
+f5 <- fg4 | fg5 | fg6 +
+  plot_annotation(
+  title = NULL) 
+  ggsave("height.png", width = 9, height = 5, dpi = 600, f5)
+```

### 2021-07-05 Frontiers/Frontiers.Rmd
as

@@
 csl: frontiers.csl
-bibliography: test.bib
+bibliography: adaptation.bib
 ---
@@
 # Introduction {-}
+
 Cite fancy references [@Neuro2013]. As demonstrated in @Gene2012, citations can
@@
 
-## Plant material {-}
+## Plant material and growing conditions {-}
 
-The study was performed with a *A. palmeri* accession (Per1) from Perkins County, Nebraska. Per1 accession collection is documented in [], with no  confirmed herbicide resistance.
+The study was performed with a *A.
+palmeri* accession (Per1) from Perkins County, Nebraska.
+Per1 accession collection is documented in [@oliveira2021a], with no reported herbicide resistance.
+Seeds were planted in plastic trays containing potting-mix.
+Emerged seedlings (1 cm) were transplanted into 200 cm<sup>-3</sup> plastic pots (a plant pot<sup>-1</sup>).
+Palmer amaranth seedlings were supplied with adequate water and kept under greenhouse conditions at Arlington, Clay Center, Lincoln, and Macomb; and kept outdoors in Grant.
+At 7 to 10 cm height, Palmer amaranth seedlings were transplanted to the field for the onset of the experiment.
+
+## Field study {-}
+
+The experiment was conducted under field conditions at five locations: Arlington, WI (Washington County), Clay Center, NE (Clay County), Grant, NE (Perkins County), Lincoln, NE (Lancaster County), and Macomb, IL (McDonough County)
 

### 2021-07-05 Frontiers/Frontiers.tex
as

@@
 
-Cite fancy references (LastName1 et al., 2013). As demonstrated in
-OtherAuthor and Coauthor (2012), citations can also be automatically
-reference. Multiple references are separated by semicolons (LastName1 et
-al., 2013; Author4 and Author5, 2013).
+Cite fancy references (\textbf{Neuro2013?}). As demonstrated in
+(\textbf{Gene2012?}), citations can also be automatically reference.
+Multiple references are separated by semicolons (\textbf{Neuro2013?};
+\textbf{Neurobot2013?}).
 
@@
 The study was performed with a \emph{A. palmeri} accession (Per1) from
-Perkins County, Nebraska. The accession Per1 was documented as
-glyphosate-susceptible {[}{]}.
+Perkins County, Nebraska. Per1 accession collection is documented in
+(Oliveira et al., 2021), with no reported herbicide resistance. Seeds
+were planted in plastic trays containing potting-mix. Emerged seedlings
+(1 cm) were transplated into 200 cm-3 plastic pots. Palmer amaranth
+seedlings supplied with adequate water and kept under greenhouse
+conditions at 28/20 C day/night temperature with 80\$ relative humidity.
+Artificial lighting was provided using metal halide lamps (600 mmol m-2
+s-1) to ensure 15-h photoperiod.
+
+The experiment was conducted under field conditions at five locations:
+Arlington, WI (Washington County), Clay Center, NE (Clay County), Grant,
+NE (Perkins County), Lincoln, NE (Lancaster County), and Macomb, IL
+(McDonough County)
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-Neurobot2013}{}%
-Author4, N., and Author5, N. (2013). Title of the article.
-\emph{Frontiers in Neurorobotics} 7.
-doi:\href{https://doi.org/10.3389/fnbot.2013.56789}{10.3389/fnbot.2013.56789}.
-
-\leavevmode\hypertarget{ref-Neuro2013}{}%
-LastName1, A., LastName2, A., and LastName2, A. (2013). Article title.
-\emph{Frontiers in Neuroscience} 30, 10127--10134.
-doi:\href{https://doi.org/10.3389/fnins.2013.12345}{10.3389/fnins.2013.12345}.
-
-\leavevmode\hypertarget{ref-Gene2012}{}%
-OtherAuthor, N., and Coauthor, N. S. (2012). Article title.
-\emph{Frontiers in Genetics} 30, 16417--16418.
-doi:\href{https://doi.org/10.3389/fgene.2012.54321}{10.3389/fgene.2012.54321}.
+\leavevmode\hypertarget{ref-oliveira2021a}{}%
+Oliveira, M. C., Giacomini, D. A., Arsenijevic, N., Vieira, G., Tranel,
+P. J., and Werle, R. (2021). Distribution and validation of genotypic
+and phenotypic glyphosate and {PPO}-inhibitor resistance in {Palmer}
+amaranth ({Amaranthus} palmeri) from southwestern {Nebraska}. \emph{Weed
+Technology} 35, 65--76.
+doi:\href{https://doi.org/10.1017/wet.2020.74}{10.1017/wet.2020.74}.
 

### 2021-07-06 Frontiers/Frontiers.Rmd
tr

@@
 
-Cite fancy references [@Neuro2013]. As demonstrated in @Gene2012, citations can
-also be automatically reference. Multiple references are separated by semicolons
-[@Neuro2013; @Neurobot2013].
+kgghkg
 
@@
 
-The experiment was conducted under field conditions at five locations: Arlington, WI (Washington County), Clay Center, NE (Clay County), Grant, NE (Perkins County), Lincoln, NE (Lancaster County), and Macomb, IL (McDonough County)
+The experiment was conducted under field conditions at five locations: Arlington (Washington County, Wisconsin), Clay Center (Clay County, Nebraska), Grant (Perkins County, Nebraska), Lincoln (Lancaster County, Nebraska), and Macomb (McDonough County, Illinois)
+
+Palmer amaranth flowering was monitored every other day throughout the study.
+When a plant started flowering, it was identified as male or female.
+Plant height was measured from soil surface to the plant apical meristem.
+Then, aboveground plant biomass was harvest near soil surface and oven dried at 65 C until reaching constant weight before the weight of biomass (g plant<sup>-1</sup>) was recorded.
+
+## Statistical analyses {-}
+
+The statistical analyses were performed using R statistical software version 4.0.1.
+
+The cumulative Palmer amaranth flowering estimation was determined using a asymetrical three parameter log logistic Weibull function of the drc package .
+
+$$Y(x) = 0 + (d-0) exp (-exp(b(log(x)-e)))$$
+In this model, *Y* is the Palmer amaranth cumulative flowering, *d* is the upper limit (set to 100), and *e* is the XXX, and *x* day of year.
 

### 2021-07-06 Frontiers/Frontiers.tex
tr

@@
 
-Cite fancy references (\textbf{Neuro2013?}). As demonstrated in
-(\textbf{Gene2012?}), citations can also be automatically reference.
-Multiple references are separated by semicolons (\textbf{Neuro2013?};
-\textbf{Neurobot2013?}).
+kgghkg
 
@@
 
-\hypertarget{plant-material}{%
-\subsection*{Plant material}\label{plant-material}}
-\addcontentsline{toc}{subsection}{Plant material}
+\hypertarget{plant-material-and-growing-conditions}{%
+\subsection*{Plant material and growing
+conditions}\label{plant-material-and-growing-conditions}}
+\addcontentsline{toc}{subsection}{Plant material and growing conditions}
 
@@
 were planted in plastic trays containing potting-mix. Emerged seedlings
-(1 cm) were transplated into 200 cm-3 plastic pots. Palmer amaranth
-seedlings supplied with adequate water and kept under greenhouse
-conditions at 28/20 C day/night temperature with 80\$ relative humidity.
-Artificial lighting was provided using metal halide lamps (600 mmol m-2
-s-1) to ensure 15-h photoperiod.
+(1 cm) were transplanted into 200 cm-3 plastic pots (a plant pot-1).
+Palmer amaranth seedlings were supplied with adequate water and kept
+under greenhouse conditions at Arlington, Clay Center, Lincoln, and
+Macomb; and kept outdoors in Grant. At 7 to 10 cm height, Palmer
+amaranth seedlings were transplanted to the field for the onset of the
+experiment.
+
+\hypertarget{field-study}{%
+\subsection*{Field study}\label{field-study}}
+\addcontentsline{toc}{subsection}{Field study}
 
 The experiment was conducted under field conditions at five locations:
-Arlington, WI (Washington County), Clay Center, NE (Clay County), Grant,
-NE (Perkins County), Lincoln, NE (Lancaster County), and Macomb, IL
-(McDonough County)
+Arlington (Washington County, Wisconsin), Clay Center (Clay County,
+Nebraska), Grant (Perkins County, Nebraska), Lincoln (Lancaster County,
+Nebraska), and Macomb (McDonough County, Illinois)
+
+Palmer amaranth flowering was monitored every other day throughout the
+study. When a plant started flowering, it was identified as male or
+female. Plant height was measured from soil surface to the plant apical
+meristem. Then, aboveground plant biomass was harvest near soil surface
+and oven dried at 65 C until reaching constant weight before the weight
+of biomass (g plant-1) was recorded.
+
+\hypertarget{statistical-analyses}{%
+\subsection*{Statistical analyses}\label{statistical-analyses}}
+\addcontentsline{toc}{subsection}{Statistical analyses}
+
+The statistical analyses were performed using R statistical software
+version 4.0.1.
+
+The cumulative Palmer amaranth flowering estimation was determined using
+a three parameter log logistic Weibull function of the drc package .
+
+\[Y(x) = 0 + (d-0) exp (-exp(b(log(x)-e)))\] In this model, \emph{Y} is
+the Palmer amaranth cumulative flowering, \emph{d} is the upper limit
+(set to 100), and \emph{e} is the XXX, and \emph{x} day of year.
 

### 2021-07-06 new_analyses.Rmd
tr

@@
   ggplot(aes(x = location, y = estimate)) +
-  scale_x_reordered() +
+  scale_x_reordered(expand = c(.01, .01)) +
   geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
@@
             size = 2,
-            label = label2$label, color = "red") +
+            label = label1$label, color = "red") +
   coord_flip() +
@@
              y = estimate)) +
-  scale_x_reordered() +
+  scale_x_reordered(expand = c(.01, .01)) +
   geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,
@@
              y = estimate)) +
-  scale_x_reordered() +
+    scale_x_reordered(expand = c(.01, .01)) +
   geom_rect(mapping=aes(xmin= 0.5, xmax= 4.5,

### 2021-07-06 Frontiers/Frontiers.Rmd
gf

--- 2021-07-05 Frontiers/Frontiers.Rmd
@@
 Per1 accession collection is documented in [@oliveira2021a], with no reported herbicide resistance.
-Seeds were planted in plastic trays containing potting-mix.
+Three weeks prior to the field experiment, seeds were planted in plastic trays containing potting-mix.
 Emerged seedlings (1 cm) were transplanted into 200 cm<sup>-3</sup> plastic pots (a plant pot<sup>-1</sup>).
 Palmer amaranth seedlings were supplied with adequate water and kept under greenhouse conditions at Arlington, Clay Center, Lincoln, and Macomb; and kept outdoors in Grant.
-At 7 to 10 cm height, Palmer amaranth seedlings were transplanted to the field for the onset of the experiment.
+Palmer amaranth seedlings were kept under greenhouse/outdoors until the onset of the experiment (7 to 10 cm height).
 
@@
 
-The experiment was conducted under field conditions at five locations: Arlington, WI (Washington County), Clay Center, NE (Clay County), Grant, NE (Perkins County), Lincoln, NE (Lancaster County), and Macomb, IL (McDonough County)
+The experiment was conducted in 2018 and 2019 under field conditions at five locations: Arlington (Washington County, Wisconsin), Clay Center (Clay County, Nebraska), Grant (Perkins County, Nebraska), Lincoln (Lancaster County, Nebraska), and Macomb (McDonough County, Illinois).
+
+The experimental unit were adjacent 9.1 m wide (12 rows at 72.2 cm row spacing) by 10.7 m long.
+Each experimental unit was planted with corn or soybean, or left fallow.
+Palmer amaranth seedlings were  transplanted to the field experiment by making a whole in the soil (6 cm deep and 8 cm wide); and gently transferring in the ground (potting mix + two seedlings).
+After a week, if both plants were alive, one was eliminated.
+There were two transplant timing: early (June 1^st^) and late (July 1^st^).
+There were 24 Palmer amaranth plants in each crop/fallow and timing, with a total of 144 plants.
+The study was repeated twice.
+
+After transplanting, Palmer amaranth flowering was monitored until the end of the study.
+When a plant started flowering, the day was recorded, plant sex was identified as male or female, and plant height was measured from soil surface to the plant top.
+Then, aboveground plant biomass was harvest near soil surface and oven dried at 65 C until reaching constant weight before the weight of biomass (g plant^-1^) was recorded.
+
+## Statistical analyses {-}
+
+The statistical analyses were performed using R statistical software version 4.0.1.
+
+The cumulative Palmer amaranth flowering estimation was determined using a asymmetrical three parameter log logistic Weibull model of the drc package [@ritz2015].
+
+$$Y(x) = 0 + (d-0) exp (-exp(b(log(x)-e)))$$
+In this model, *Y* is the Palmer amaranth cumulative flowering, *d* is the upper limit (set to 100), and *e* is the XXX, and *x* day of year (doy).
+
+The doy for 10, 50, and 90% Palmer amaranth cumulative flowering were determined using the *ED* function of drc package.
+Also, the 10, 50, and 90% Palmer amaranth cumulative flowering were compared among crop/fallow and timings using the *EDcomp* function of drc package.
+The EDcomp function compares the ratio of cumulative flowering using t-statistics, where P-value < 0.05 indicates that we fail to reject the null hypothesis.
 
@@
 
-```{r graph, echo = TRUE, message=FALSE, fig.show = "hide"}
-x <- 0:100
-set.seed(999)
-y <- 2 * (x + rnorm(length(x), sd = 3) + 3)
-plot(x, y)
+```{r fig1, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../june.png")
 ```
@@
 
-```{r, Figure-1, ref.label = "graph", results = "hide", echo = FALSE, message = FALSE, fig.height=4, fig.width=4, fig.align='center', fig.cap='Figure caption', out.width = "85mm", out.height = "85mm"}
+```{r, Figure-2, ref.label = "fig1", echo = FALSE, message = FALSE, fig.height=3, fig.width=5, fig.align='center', fig.cap='Figure caption', out.width = "150mm", out.height = "85mm"}
 # You can also refer to code chunks from above to place figures at the bottom.

### 2021-07-06 Frontiers/Frontiers.tex
gf

@@
 
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-
 \def\Authors{
@@
 Perkins County, Nebraska. Per1 accession collection is documented in
-(Oliveira et al., 2021), with no reported herbicide resistance. Seeds
-were planted in plastic trays containing potting-mix. Emerged seedlings
-(1 cm) were transplanted into 200 cm-3 plastic pots (a plant pot-1).
-Palmer amaranth seedlings were supplied with adequate water and kept
-under greenhouse conditions at Arlington, Clay Center, Lincoln, and
-Macomb; and kept outdoors in Grant. At 7 to 10 cm height, Palmer
-amaranth seedlings were transplanted to the field for the onset of the
-experiment.
+(Oliveira et al., 2021), with no reported herbicide resistance. Three
+weeks prior to the field experiment, seeds were planted in plastic trays
+containing potting-mix. Emerged seedlings (1 cm) were transplanted into
+200 cm-3 plastic pots (a plant pot-1). Palmer amaranth seedlings were
+supplied with adequate water and kept under greenhouse conditions at
+Arlington, Clay Center, Lincoln, and Macomb; and kept outdoors in Grant.
+Palmer amaranth seedlings were kept under greenhouse/outdoors until the
+onset of the experiment (7 to 10 cm height).
 
@@
 
-The experiment was conducted under field conditions at five locations:
-Arlington (Washington County, Wisconsin), Clay Center (Clay County,
-Nebraska), Grant (Perkins County, Nebraska), Lincoln (Lancaster County,
-Nebraska), and Macomb (McDonough County, Illinois)
+The experiment was conducted in 2018 and 2019 under field conditions at
+five locations: Arlington (Washington County, Wisconsin), Clay Center
+(Clay County, Nebraska), Grant (Perkins County, Nebraska), Lincoln
+(Lancaster County, Nebraska), and Macomb (McDonough County, Illinois).
 
-Palmer amaranth flowering was monitored every other day throughout the
-study. When a plant started flowering, it was identified as male or
-female. Plant height was measured from soil surface to the plant apical
-meristem. Then, aboveground plant biomass was harvest near soil surface
-and oven dried at 65 C until reaching constant weight before the weight
-of biomass (g plant-1) was recorded.
+The experimental unit were adjacent 9.1 m wide (12 rows at 72.2 cm row
+spacing) by 10.7 m long. Each experimental unit was planted with corn or
+soybean, or left fallow. Palmer amaranth seedlings were transplanted to
+the field experiment by making a whole in the soil (6 cm deep and 8 cm
+wide); and gently transferring in the ground (potting mix + two
+seedlings). After a week, if both plants were alive, one was eliminated.
+There were two transplant timing: early (June 1\textsuperscript{st}) and
+late (July 1\textsuperscript{st}). There were 24 Palmer amaranth plants
+in each crop/fallow and timing, with a total of 144 plants. The study
+was repeated twice.
+
+After transplanting, Palmer amaranth flowering was monitored until the
+end of the study. When a plant started flowering, the day was recorded,
+plant sex was identified as male or female, and plant height was
+measured from soil surface to the plant top. Then, aboveground plant
+biomass was harvest near soil surface and oven dried at 65 C until
+reaching constant weight before the weight of biomass (g
+plant\textsuperscript{-1}) was recorded.
 
@@
 The cumulative Palmer amaranth flowering estimation was determined using
-a three parameter log logistic Weibull function of the drc package .
+a asymmetrical three parameter log logistic Weibull model of the drc
+package (Ritz et al., 2015).
 
@@
 the Palmer amaranth cumulative flowering, \emph{d} is the upper limit
-(set to 100), and \emph{e} is the XXX, and \emph{x} day of year.
+(set to 100), and \emph{e} is the XXX, and \emph{x} day of year (doy).
+
+The doy for 10, 50, and 90\% Palmer amaranth cumulative flowering were
+determined using the \emph{ED} function of drc package. Also, the 10,
+50, and 90\% Palmer amaranth cumulative flowering were compared among
+crop/fallow and timings using the \emph{EDcomp} function of drc package.
+The EDcomp function compares the ratio of flowering timing using
+t-statistics, where P-value \textless{} 0.05 indicates that we fail to
+reject the null hypothesis.
 
@@
 You can use \texttt{R} chunks directly to plot graphs.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{x }\OtherTok{\textless{}{-}} \DecValTok{0}\SpecialCharTok{:}\DecValTok{100}
-\FunctionTok{set.seed}\NormalTok{(}\DecValTok{999}\NormalTok{)}
-\NormalTok{y }\OtherTok{\textless{}{-}} \DecValTok{2} \SpecialCharTok{*}\NormalTok{ (x }\SpecialCharTok{+} \FunctionTok{rnorm}\NormalTok{(}\FunctionTok{length}\NormalTok{(x), }\AttributeTok{sd =} \DecValTok{3}\NormalTok{) }\SpecialCharTok{+} \DecValTok{3}\NormalTok{)}
-\FunctionTok{plot}\NormalTok{(x, y)}
-\end{Highlighting}
-\end{Shaded}
 
@@
 
-{\centering \includegraphics[width=85mm,height=85mm]{Frontiers_files/figure-latex/Figure-1-1}
+{\centering \includegraphics[width=150mm,height=85mm]{../june}
 
@@
 
-\caption{Figure caption}\label{fig:Figure-1}
+\caption{Figure caption}\label{fig:Figure-2}
 \end{figure}
@@
 
+\leavevmode\hypertarget{ref-ritz2015}{}%
+Ritz, C., Baty, F., Streibig, J. C., and Gerhard, D. (2015).
+Dose-{Response Analysis Using R}. \emph{PLOS ONE} 10, e0146021.
+doi:\href{https://doi.org/10.1371/journal.pone.0146021}{10.1371/journal.pone.0146021}.
+
 \end{CSLReferences}

### 2021-07-09 Frontiers/Frontiers.Rmd
LP

@@
 ```{r fig1, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../weather/Figure 1.png")
+```
+
+```{r fig2, echo = FALSE, message=FALSE, fig.show = "hide"}
 knitr::include_graphics("../june.png")
@@
 
-```{r, Figure-2, ref.label = "fig1", echo = FALSE, message = FALSE, fig.height=3, fig.width=5, fig.align='center', fig.cap='Figure caption', out.width = "150mm", out.height = "85mm"}
+```{r, Figure-1, ref.label = "fig1", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Mean average temperature (C) and montly sum precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL', out.width = "160mm", out.height = "100mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
 ```
+
+```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Figure caption', out.width = "160mm", out.height = "90mm"}
+# You can also refer to code chunks from above to place figures at the bottom.
+```

### 2021-07-09 Frontiers/Frontiers.tex
LP

@@
 crop/fallow and timings using the \emph{EDcomp} function of drc package.
-The EDcomp function compares the ratio of flowering timing using
+The EDcomp function compares the ratio of cumulative flowering using
 t-statistics, where P-value \textless{} 0.05 indicates that we fail to
@@
 
-{\centering \includegraphics[width=150mm,height=85mm]{../june}
+{\centering \includegraphics[width=160mm,height=100mm]{../weather/Figure 1}
+
+}
+
+\caption{Mean average temperature (C) and montly sum precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-1}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=160mm,height=90mm]{../june}
 

### 2021-07-12 new_analyses.Rmd
,jkj

@@
 data %>% 
-  filter(location != "grant") -> data1
+  filter(location != "grant") %>% 
+  mutate(location = str_replace(location, "havelock", "lincon"))-> data1
 ```
@@
 ```
-
-  annotate("text", x = 0.65, y = 168, label = "June", 
-           color = "#8B8000", size = 2) +
-  annotate("text", x = 0.65, y = 197, label = "July", 
-           color = "#80461b", size = 2) +
-  annotate("text", x = 0.65, y = 228.5, label = "August", 
-           color = "darkgreen", size = 2) +
-  annotate("text", x = 0.65, y = 259, label = "September", 
-           color = "blue", size = 2) +
 
@@
   title = NULL) 
-  ggsave("height.png", width = 9, height = 5, dpi = 600, f5)
+  ggsave("height.png", width = 9, height = 5, dpi = 300, f5)
 ```
+
+###--------------------------
+
+```{r}
+data %>% 
+  filter(location != "grant") %>% 
+  mutate(location = str_replace(location, "havelock", "lincon"))-> data1
+```
+
+```{r}
+data1 %>% 
+  rename(month = planting,
+         planting = planting_1) %>% 
+  mutate(planting = mdy(planting),
+         harvest = mdy(harvest),
+         doyp = yday(planting)) %>% 
+  mutate_if(is_character, as_factor) -> data2
+```
+
+```{r}
+data2 %>% 
+  mutate(harvest = as.character(harvest)) %>% 
+  mutate(harvest = str_replace(harvest, "2018", "2019"),
+         harvest = ymd(harvest)) %>% 
+  filter(!is.na(harvest)) %>% 
+  dplyr::select(location, month, crop, harvest) %>% 
+  group_by(month, crop) %>% 
+  count(harvest, name = "harvest_day") -> harvest
+```
+
+```{r}
+data2 %>% 
+  filter(!is.na(harvest)) %>% 
+  dplyr::select(location, month, crop, harvest) %>% 
+  group_by(month, crop) %>% 
+  count(month, crop, name = "total") -> total
+```
+
+```{r}
+total %>% 
+  left_join(harvest) %>% 
+  mutate(prop = harvest_day / total) %>% 
+  mutate(lag = lag(prop)) %>% 
+  mutate(prob = cumsum(prop)) %>% 
+  mutate(doyh = yday(harvest)) -> cum_sum
+```
+
+```{r}
+data2 %>% 
+  mutate(harvest = str_replace(harvest, "2018", "2019"),
+         harvest = ymd(harvest)) %>% 
+  filter(!is.na(harvest)) %>% 
+  ungroup() %>% 
+  left_join(cum_sum) %>% 
+  rename(harvest_com = harvest) %>% 
+  dplyr::select(-lag) -> data3
+```
+
+# June
+
+```{r}
+data3 %>% 
+  filter(month == "June") -> month_jun 
+model3 <- drm(prob*100 ~ doyh, crop, fct = W1.3(), data = month_jun)
+
+plot(model3, col = c(1,2,3), type = "average", ylab = "Cumulative flowering (%)",
+     xlab = "Day of the year")
+```
+
+```{r}
+#broom::tidy(model_arl)
+summary(model3)
+```
+
+```{r}
+ed_values3 <- ED(model3, c(10, 50, 90), interval = "delta", type = "absolute")
+```
+
+```{r}
+ed_values3 %>% 
+  as.data.frame() %>% 
+  rownames_to_column(., var = "name") %>% 
+  as_tibble() %>% 
+  separate(name, c("param", "crop", "ED"), sep = ":") %>% 
+  janitor::clean_names() %>% 
+  mutate_if(is_character, as_factor) -> ed_values3
+```
+
+```{r}
+label1 <- tibble(ed = "90% flowering", 
+                   x = 2.5, y = 185, label = "Transplanting \nday")
+rect1 <- tibble(ed = "90% flowering", 
+               x = 2.5, y = 153, xend = 2.5, yend = 165)
+```
+
+```{r}
+month <- tibble(ed = "90% flowering",
+                x = c(0.65, 0.65, 0.65, 0.65),
+                y = c(168, 197, 228.5, 259), 
+                label = c("June", "July", "August", "September"))
+```
+
+```{r}
+ed_values3 %>% 
+  mutate(month = "First transplanting") %>% 
+  mutate(ed = case_when(
+    ed == "10" ~ "10% flowering",
+    ed == "50" ~ "50% flowering",
+    ed == "90" ~ "90% flowering",
+    TRUE ~ "ed")) %>% 
+  group_by(month,  ed) %>% 
+  mutate(crop = reorder_within(crop, 
+                                   by = estimate, 
+                                   within = ed)) %>% 
+  ggplot(aes(x = crop, y = estimate)) +
+  scale_x_reordered(expand = c(.01, .01)) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 3.5,
+                        ymin= 153, ymax= 183), 
+            fill = "#FFFDD0", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 3.5,
+                        ymin= 183, ymax= 214), 
+            fill = "#80461b", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 3.5,
+                        ymin= 214, ymax= 244), 
+            fill = "lightgreen", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 3.5,
+                        ymin= 244, ymax= 274), 
+            fill = "blue", alpha=0.05) +
+  geom_hline(yintercept = 153, size = 1.1, 
+             color = "#CCCC00",
+             alpha = 1) +
+  geom_point(size = 2, aes(color = crop, 
+                           shape = crop)) +
+  geom_linerange(aes(ymin = lower, ymax = upper, color = crop, 
+                     shape = crop), size = 1.3) +
+  scale_color_manual(values = c("#412234","#EE4B2B","#32CD32",
+                                "#412234","#EE4B2B","#32CD32",
+                                "#412234","#EE4B2B","#32CD32")) +
+  scale_shape_manual(values = c(15,16,17,15,16,17,15,16,17)) +
+  facet_grid(ed ~ month, scales = "free_y") +
+  coord_flip() +
+  geom_text(data = month, aes(x = x, y = y), label = month$label,
+            size = 2, color = c("#8B8000", "#80461b", 
+                                  "darkgreen", "blue")) +
+  labs(y = NULL, x = NULL,
+       title = NULL) +
+  geom_curve(data = rect1, aes(x = x, y = y, 
+                               xend = xend, yend = yend),
+             arrow = arrow(length = unit(0.07, "inch")), 
+               size = 0.4, curvature = 0.3) + 
+  geom_text(data = label1, 
+            aes(x = x, y = y), 
+            size = 2,
+            label = label1$label, color = "#CCCC00") +
+  coord_flip() +
+  theme_test() +
+  theme(legend.position = "none",
+        plot.title = element_markdown(),
+        strip.text = element_markdown(face = "bold"),
+        plot.title.position = "plot") -> p1c
+p1c
+```
+
+# July
+
+```{r}
+data3 %>% 
+  filter(month == "July") -> month_jul 
+model4 <- drm(prob*100 ~ doyh, crop, fct = W1.3(), data = month_jul)
+
+plot(model4, col = c(1,2,3), type = "average", ylab = "Cumulative flowering (%)",
+     xlab = "Day of the year")
+```
+
+```{r}
+#broom::tidy(model_arl)
+summary(model4)
+```
+
+```{r}
+ed_values4 <- ED(model4, c(10, 50, 90), interval = "delta", type = "absolute")
+```
+
+```{r}
+ed_values4 %>% 
+  as.data.frame() %>% 
+  rownames_to_column(., var = "name") %>% 
+  as_tibble() %>% 
+  separate(name, c("param", "crop", "ED"), sep = ":") %>% 
+  janitor::clean_names() %>% 
+  mutate_if(is_character, as_factor) -> ed_values4
+```
+
+```{r}
+label2 <- tibble(ed = "90% flowering", 
+                   x = 2.5, y = 211, label = "Transplanting \nday")
+rect2 <- tibble(ed = "90% flowering",  
+               x = 2.5, y = 183, xend = 2.5, yend = 195)
+```
+
+```{r}
+month1 <- tibble(ed = "90% flowering",
+                x = c(0.65, 0.65, 0.65),
+                y = c(197, 228.5, 259), 
+                label = c("July", "August", "September"))
+```
+
+```{r}
+ed_values4 %>% 
+  mutate(month = "Second transplanting") %>% 
+  mutate(ed = case_when(
+    ed == "10" ~ "10% flowering",
+    ed == "50" ~ "50% flowering",
+    ed == "90" ~ "90% flowering",
+    TRUE ~ "ed")) %>% 
+  group_by(month,  ed) %>% 
+  mutate(crop = reorder_within(crop, 
+                                   by = estimate, 
+                                   within = ed)) %>% 
+  ggplot(aes(x = crop, y = estimate)) +
+  scale_x_reordered(expand = c(.01, .01)) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 3.5,
+                        ymin= 183, ymax= 214), 
+            fill = "#80461b", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 3.5,
+                        ymin= 214, ymax= 244), 
+            fill = "lightgreen", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.5, xmax= 3.5,
+                        ymin= 244, ymax= 274), 
+            fill = "blue", alpha=0.05) +
+  geom_hline(yintercept = 183, size = 1.1, 
+             color = "#D2691E",
+             alpha = 1) +
+  geom_point(size = 2, aes(color = crop, shape = crop)) +
+  geom_linerange(aes(ymin = lower, ymax = upper, color = crop), size = 1.3) +
+  scale_color_manual(values = c("#412234","#EE4B2B","#32CD32",
+                                "#412234","#EE4B2B","#32CD32",
+                                "#412234","#32CD32","#EE4B2B")) +
+  scale_shape_manual(values = c(15,16,17,15,16,17,15,17,16)) +
+  geom_text(data = month1, aes(x = x, y = y), label = month1$label,
+            size = 2, color = c("#80461b", "darkgreen", "blue")) +
+  labs(y = NULL, x = NULL) +
+  facet_grid(ed ~ month, scales = "free_y") +
+  geom_curve(data = rect2, aes(x = x, y = y, 
+                               xend = xend, yend = yend),
+             arrow = arrow(length = unit(0.07, "inch")), 
+               size = 0.4, curvature = 0.3) + 
+  geom_text(data = label2, 
+            aes(x = x, y = y), 
+            size = 2,
+            label = label2$label, color = "#D2691E") +
+  coord_flip() +
+  theme_test() +
+  theme(legend.position = "none",
+        strip.text = element_markdown(face = "bold"),
+        plot.title = element_markdown(),
+        plot.title.position = "plot") -> p2c
+```
+
+```{r}
+library(patchwork)
+f3 <- p1c | p2c +
+  plot_annotation(
+  title = "DOY of 10, 50, and 90% Palmer amaranth flowering") 
+  ggsave("Combined.png", width = 6, height = 5, dpi = 600, f3)
+```
+
+# Biomass
+
+```{r}
+data <- read_csv("data2.csv")
+```
+
+```{r}
+data %>% 
+  rename(month = planting,
+         planting = planting_1) %>% 
+  mutate(planting = mdy(planting),
+         harvest = mdy(harvest),
+         doyp = yday(planting)) -> data1
+```
+
+```{r}
+data1 %>% 
+  mutate(month = fct_recode(month,
+                            "First transplanting" = "June",
+                            "Second transplanting" = "July")) -> data1
+```
+
+```{r}
+data1 %>% 
+  filter(!is.na(weight)) %>% 
+  group_by(location, month, crop, doyh) %>% 
+  summarise(biomass = sum(weight))  %>% 
+  ggplot(aes(x = biomass, y = crop)) +
+  geom_density_ridges() +
+  facet_grid(location ~ month)
+```
+
+```{r}
+options(scipen = 999)
+library(ggstream)
+library(ggridges)
+
+data1 %>% 
+  filter(!is.na(weight)) %>% 
+  dplyr::select(month, crop, weight, doyh) %>% 
+  group_by(month, crop) %>% 
+  mutate(n = n()) %>% 
+  mutate(rate = weight / n) %>% 
+  group_by(month, crop, doyh) %>% 
+  summarise(biomass = sum(rate)) %>% 
+  mutate(doyh = as.integer(doyh)) -> biomass
+```
+
+```{r}
+biomass %>% 
+  ggplot(aes(x = doyh, y = biomass, fill = crop, label = crop)) +
+  geom_stream(bw = 0.75, geom = "polygon", type = "ridge") +
+#  geom_stream_label(size = 4, type = "proportional", n_grid = 1000) +
+  facet_grid(~ month, scales = "free_x") +
+  scale_fill_manual(values = c("#412234","#EE4B2B","#32CD32")) +
+  labs(y = "Biomass (g plant)") +
+  theme_test()
+
+ggsave("biomass.png")
+```
+
+```{r}
+options(scipen = 999)
+library(ggstream)
+library(ggridges)
+
+data1 %>% 
+  filter(!is.na(height)) %>% 
+  dplyr::select(month, crop, height, doyh) %>% 
+  group_by(month, crop) %>% 
+  mutate(n = n()) %>% 
+  mutate(rate = height / n) %>% 
+  group_by(month, crop, doyh) %>% 
+  summarise(height = sum(rate)) %>% 
+  mutate(doyh = as.integer(doyh)) -> height
+```
+
+```{r}
+height %>% 
+  ggplot(aes(x = doyh, y = height, fill = crop, label = crop)) +
+  geom_stream(bw = 0.75, geom = "polygon", type = "proportional") +
+  facet_grid(~ month, scales = "free_x") +
+  geom_stream_label(size = 4, type = "proportional", n_grid = 1000) +
+  scale_fill_manual(values = c("#412234","#EE4B2B","#32CD32")) +
+  labs(y = "Height (g plant)") +
+  theme_test()
+```

### 2021-07-13 new_analyses.Rmd
lkl

@@
 library(ggtext)
+library(patchwork)
+library(extrafont)
+library(showtext)
+showtext_auto()
+font_add_google("Roboto", "roboto") 
 ```
@@
 ```
+```{r}
+data1 %>% 
+  filter(!is.na(weight)) %>% 
+  group_by(year, location, month, crop) %>% 
+  mutate(n = n()) %>% 
+  mutate(biomass = weight / n) -> biom
+```
+```{r}
+biom %>% 
+  ggplot(aes(x = log(biomass))) + 
+  geom_histogram()
+```
+
+```{r}
+library(lme4)
+library(lmerTest)
+library(emmeans)
+```
+
+```{r}
+biom %>%
+  group_by(month) %>% 
+  nest(-month) %>% 
+  mutate(model = map(data, ~ lmer(log(biomass) ~ crop * location + 
+                                 (1|year), data = .x)),
+         anova = map(model, anova),
+         emmeans = map(model, ~ emmeans(., ~ location | crop, 
+                                        type = "response"))) -> analysis
+```
+
+```{r}
+#anova
+analysis$anova[[2]]
+```
+
+```{r}
+analysis$emmeans[[2]] %>% 
+  as_tibble() %>% 
+  mutate(month = "Second transplanting") -> second
+
+analysis$emmeans[[1]] %>% 
+  as_tibble() %>% 
+  mutate(month = "First transplanting") %>% 
+  bind_rows(second) %>% 
+  mutate(location = str_to_title(location)) %>% 
+  mutate(location = fct_recode(location,
+                              "Lincoln" = "Havelock")) -> result
+```
+
+```{r}
+result %>% 
+  filter(month == "First transplanting") %>% 
+  group_by(month, crop) %>% 
+  mutate(location = reorder_within(location, 
+                                   by = response, 
+                                   within = crop)) %>% 
+  ggplot(aes(x = location, y = response,
+             color = location)) +
+  geom_point(size = 2) +
+  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL)) +
+  scale_color_manual(values = c(1,2,4,3,5, 1,3,2,4,5, 1,2,4,3,5)) +
+  facet_grid(crop ~ month, scales = "free") +
+  coord_flip() +
+  ylim(0, 30) +
+  scale_x_reordered() +
+  labs(y = expression(paste("Biomass (g plant "^"-1",")")), x = NULL) +
+  theme_test() +
+  theme(legend.position = "none") -> first
+
+ggsave("first.png", width = 3)
+```
+
+```{r}
+theme_set(theme_test(base_family = "roboto"))
+result %>% 
+  filter(month == "Second transplanting") %>% 
+  group_by(month, crop) %>% 
+  mutate(location = reorder_within(location, 
+                                   by = response, 
+                                   within = crop)) %>% 
+  ggplot(aes(x = location, y = response,
+             color = location)) +
+  geom_point(size = 2) +
+  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL)) +
+  scale_color_manual(values = c(1,2,3,4,5, 1,4,2,3,5, 1,2,3,4,5)) +
+  facet_grid(crop ~ month, scales = "free") +
+  coord_flip() +
+  ylim(0, 80) +
+  scale_x_reordered() +
+  labs(y = expression(paste("Biomass (g plant "^"-1",")")), x = NULL) +
+  theme_test() +
+  theme(legend.position = "none") -> second
+
+ggsave("second.png", width = 3)
+```
+
+```{r}
+first | second
+
+ggsave("biomass.png", width = 6)
+```
+
+```{r}
+biom %>% 
+  ggplot(aes(x = crop, y = biomass, label = n)) +
+  geom_jitter(alpha = 0.2) +
+  geom_boxplot() +
+  facet_grid(~month, scales = "free") +
+  coord_flip()
+```
+
+```{r}
+data1 %>% 
+  ggplot(aes(x = crop, y = weight)) +
+  geom_jitter(alpha = 0.2) +
+  facet_grid(month ~ location) +
+  coord_flip()
+```
 

### 2021-07-14 new_analyses.Rmd
k
jh

@@
   title = "DOY of 10, 50, and 90% Palmer amaranth flowering after first transplanting") 
-  ggsave("June.png", width = 9, height = 5, dpi = 600, f1)
+  ggsave("figures/June.png", width = 9, height = 5, dpi = 600, f1)
 ```
@@
   title = "DOY of 10, 50, and 90% Palmer amaranth flowering after second transplanting") 
-  ggsave("July.png", width = 9, height = 5, dpi = 600, f3)
+  ggsave("figures/July.png", width = 9, height = 5, dpi = 600, f3)
 ```
@@
   title = NULL) 
-  ggsave("biomass.png", width = 9, height = 5, dpi = 600, f4)
+  ggsave("figures/biomass.png", width = 9, height = 5, dpi = 600, f4)
 ```
@@
   title = NULL) 
-  ggsave("height.png", width = 9, height = 5, dpi = 300, f5)
+  ggsave("figures/height.png", width = 9, height = 5, dpi = 300, f5)
 ```
@@
   title = "DOY of 10, 50, and 90% Palmer amaranth flowering") 
-  ggsave("Combined.png", width = 6, height = 5, dpi = 600, f3)
+  ggsave("figures/Combined.png", width = 6, height = 5, dpi = 600, f3)
 ```
@@
 
-ggsave("first.png", width = 3)
+ggsave("figures/first.png", width = 3)
 ```
@@
 
-ggsave("second.png", width = 3)
+ggsave("figures/second.png", width = 3)
 ```
@@
 
-ggsave("biomass.png", width = 6)
+ggsave("figures/biomass.png", width = 6)
+```
+
+```{r}
+biom %>%
+  group_by(month) %>% 
+  nest(-month) %>% 
+  mutate(model = map(data, ~ lmer(log(biomass) ~ crop + 
+                                 (1|year/location), data = .x)),
+         anova = map(model, anova),
+         emmeans = map(model, ~ emmeans(., ~ crop, 
+                                        type = "response"))) -> analysis2
+```
+
+```{r}
+#anova
+analysis2$anova[[2]]
+```
+
+```{r}
+analysis2$emmeans[[2]] %>% 
+  as_tibble() %>% 
+  mutate(month = "Second transplanting") -> second2
+
+analysis2$emmeans[[1]] %>% 
+  as_tibble() %>% 
+  mutate(month = "First transplanting") %>% 
+  bind_rows(second2) -> result2
+```
+
+```{r}
+result2 %>% 
+  filter(month == "First transplanting") %>% 
+  ggplot(aes(x = crop, y = response,
+             color = crop)) +
+  geom_point(size = 2) +
+  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL)) +
+  scale_color_manual(values = c(1,2,4,3,5, 1,3,2,4,5, 1,2,4,3,5)) +
+  facet_grid(~ month, scales = "free") +
+  coord_flip() +
+  ylim(0, 30) +
+  labs(y = expression(paste("Biomass (g plant "^"-1",")")), x = NULL) +
+  theme_test() +
+  theme(legend.position = "none") -> first
+
+ggsave("figures/first_crop.png", width = 3)
+```
+
+```{r}
+theme_set(theme_test(base_family = "roboto"))
+result2 %>% 
+  filter(month == "Second transplanting") %>% 
+  ggplot(aes(x = crop, y = response,
+             color = crop)) +
+  geom_point(size = 2) +
+  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL)) +
+  scale_color_manual(values = c(1,2,3,4,5, 1,4,2,3,5, 1,2,3,4,5)) +
+  facet_grid(crop ~ month, scales = "free") +
+  coord_flip() +
+  labs(y = expression(paste("Biomass (g plant "^"-1",")")), x = NULL) +
+  theme_test() +
+  theme(legend.position = "none") -> second
+
+ggsave("figures/second_crop.png", width = 3)
+```
+
+```{r}
+first | second
+
+ggsave("figures/biomass.png", width = 6)
 ```
@@
   coord_flip()
-```
-
-```{r}
-data1 %>% 
-  filter(!is.na(weight)) %>% 
-  group_by(location, month, crop, doyh) %>% 
-  summarise(biomass = sum(weight))  %>% 
-  ggplot(aes(x = biomass, y = crop)) +
-  geom_density_ridges() +
-  facet_grid(location ~ month)
 ```
@@
 
-ggsave("biomass.png")
+ggsave("figures/biomass.png")
 ```

### 2021-07-14 new_analyses.Rmd
MKN

@@
 ```{r}
+theme_set(theme_test(base_family = "roboto"))
+result %>% 
+  ggplot(aes(x = location, y = response,
+             color = month)) +
+  geom_point(size = 2, position = position_dodge2(width = 0.5)) +
+  geom_linerange(aes(ymin = lower.CL, ymax = upper.CL),
+                 position = position_dodge2(width = 0.5)) +
+#  scale_color_manual(values = c(1,2,3,4,5, 1,4,2,3,5, 1,2,3,4,5)) +
+  facet_grid(~ crop, scales = "free") +
+  coord_flip() +
+  labs(y = expression(paste("Biomass (g plant "^"-1",")")), x = NULL) +
+  theme_test() +
+  theme(legend.position = "bottom")
+
+ggsave("figures/combined2.png", width = 6)
+```
+
+```{r}
 result %>% 
@@
 biom %>%
+  unite("crop_yr", c("crop", "year"), sep = "-") %>% 
+  filter(!is.na(weight)) %>% 
   group_by(month) %>% 
   nest(-month) %>% 
-  mutate(model = map(data, ~ lmer(log(biomass) ~ crop + 
-                                 (1|year/location), data = .x)),
+  mutate(model = map(data, ~ lmer(log(biomass) ~ crop_yr + 
+                                 (1|location), data = .x)),
          anova = map(model, anova),
-         emmeans = map(model, ~ emmeans(., ~ crop, 
+         emmeans = map(model, ~ emmeans(., ~ crop_yr, 
                                         type = "response"))) -> analysis2
@@
   filter(month == "First transplanting") %>% 
-  ggplot(aes(x = crop, y = response,
-             color = crop)) +
+  ggplot(aes(x = crop_yr, y = response,
+             color = crop_yr)) +
   geom_point(size = 2) +
@@
 
-ggsave("figures/first_crop.png", width = 3)
+ggsave("figures/first_crop.png")
 ```
@@
   filter(month == "Second transplanting") %>% 
-  ggplot(aes(x = crop, y = response,
-             color = crop)) +
+  ggplot(aes(x = crop_yr, y = response,
+             color = crop_yr)) +
   geom_point(size = 2) +
@@
   scale_color_manual(values = c(1,2,3,4,5, 1,4,2,3,5, 1,2,3,4,5)) +
-  facet_grid(crop ~ month, scales = "free") +
-  coord_flip() +
+  facet_grid(~ month, scales = "free") +
+  coord_flip() +
+  ylim(0, 20) +
   labs(y = expression(paste("Biomass (g plant "^"-1",")")), x = NULL) +
@@
 
-ggsave("figures/second_crop.png", width = 3)
+ggsave("figures/second_crop.png")
 ```

### 2021-07-14 new_analyses.Rmd
hjgh

@@
 biom %>%
-  unite("crop_yr", c("crop", "year"), sep = "-") %>% 
+#  unite("crop_yr", c("crop", "year"), sep = "-") %>% 
   filter(!is.na(weight)) %>% 
-  group_by(month) %>% 
-  nest(-month) %>% 
-  mutate(model = map(data, ~ lmer(log(biomass) ~ crop_yr + 
+  group_by(year) %>% 
+  nest(-year) %>% 
+  mutate(model = map(data, ~ lmer(log(biomass) ~ location * crop * month + 
                                  (1|location), data = .x)),
          anova = map(model, anova),
-         emmeans = map(model, ~ emmeans(., ~ crop_yr, 
+         emmeans = map(model, ~ emmeans(., ~ location * crop * month, 
                                         type = "response"))) -> analysis2
@@
   as_tibble() %>% 
-  mutate(month = "Second transplanting") -> second2
+  mutate(year = "2019") -> second2
 
@@
   as_tibble() %>% 
-  mutate(month = "First transplanting") %>% 
+  mutate(year = "2018") %>% 
   bind_rows(second2) -> result2

### 2021-07-20 Frontiers/Frontiers.Rmd
jhjh

@@
 ---
-title: "Adaptation of Palmer amaranth to the upper Midwest"
+title: "Adaptation of Palmer amaranth to croppins systems"
 runtitle: "Short Title"
@@
 
-kgghkg
+Palmer amaranth (*Amaranthus palmeri* S.
+Watson) is an annual forb native to southwestern United States and northern Mexico.
+The species is a fast growing C4 species, XXX reported that Palmer amaranth can grow a inch per day.
+Palmer amaranth reproduce by dioecious plants outcrossing, a single plant can produce up to a million seeds..
+Palmer amaranth as well as other Amaranthaceae species have been reported as an edible plant but it has become a concern to agriculture.
+The species is currently considered one of the most problematic weed species to US cropping systems.
+
+ family. It is a dioecious species that can reach up to 3 m tall and produce a million seed per plant.
+
+The species biology in combination to human intervention in agriculture made Palmer amaranth one of the most difficult species to control in the United States.
+Weed control in US agriculture is heavily based on herbicides.
+Palmer amaranth has showed a remarkable capacity to evolve resistance to herbicides.
+Thus far, Palmer amaranth has evolved resistance to X herbicide sites of action, which increases control complexity.
+Palmer amaranth is widespread to southern United States and it is currently moving to northern states.
+The increase movement of Palmer amaranth to US Midwest warrants investigation on species adaptation to that environment.
+Most studies are based on reactive but not in the proactive management.
+
+The Palmer amaranth biology combined to human intervention in agriculture
+
+ Palmer amaranth troublesomeness
+
+The objective of this study was to investigate the flowering pattern, biomass, and height of Palmer amaranth growing under different environments and timing across five locations in the US Midwest.
 
@@
 ```{r fig1, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../weather/Figure 1.png")
+knitr::include_graphics("../data analysis/weather/Figure 1.png")
 ```
@@
 ```{r fig2, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../june.png")
+knitr::include_graphics("../data analysis/figures/combine 3.png")
 ```
@@
 
-```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Figure caption', out.width = "160mm", out.height = "90mm"}
+```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Figure caption', out.width = "150mm", out.height = "260mm"}
 # You can also refer to code chunks from above to place figures at the bottom.

### 2021-07-20 Frontiers/Frontiers.tex
jhjh

@@
 
-\title[Short Title]{Adaptation of Palmer amaranth to the upper Midwest}
+\title[Short Title]{Adaptation of Palmer amaranth to croppins systems}
 \author[\firstAuthorLast]{\Authors}
@@
 
-kgghkg
+Palmer amaranth (\emph{Amaranthus palmeri}) is an indigenous species
+from southwestern United States and northern Mexico. Palmer amaranth is
+a C4 annual broadleaf forb within the \textbf{Amarantacea} family.
+Palmer amaranth is currently considered one of the most troublesome weed
+species in the United States.
 
@@
 
-{\centering \includegraphics[width=160mm,height=100mm]{../weather/Figure 1}
+{\centering \includegraphics[width=160mm,height=100mm]{../data analysis/weather/Figure 1}
 
@@
 
-{\centering \includegraphics[width=160mm,height=90mm]{../june}
+{\centering \includegraphics[width=150mm,height=260mm]{../data analysis/figures/combine 3}
 

### 2021-07-20 data analysis/map.Rmd
jhjh

---
title: "map"
author: "Maxwel Coura Oliveira"
date: "2/1/2019"
output: html_document
---

```{r include=FALSE}
library(tidyverse)
library(ggplot2)
library(knitr)
library(kableExtra)
library(scales)
library(ggthemes)
library(usmap)
library(ggmap)
library(ggrepel)
library(maps)
library(grid)
library(mapdata)
library(maptools)
library(stringr)
library(dplyr)
library(ggsn)
library(choroplethr)
```

```{r include=FALSE, warning=FALSE}
region <-c("Arlington, WI", "Macomb, IL", "Grant, NE", "Lincoln, NE", "Clay Center, NE")
Latitude <-c(43.335148, 40.45921,  40.83999664, 40.806862, 40.51999792)
Longitude <-c(-89.372523, -90.6718, -101.722330444, -96.681679, -98.052999788)

Map<-data.frame(region, Latitude, Longitude)
```

```{r include=FALSE}
states = map_data("state")
df <- subset(states, region %in% c("wisconsin", "nebraska", "illinois", "iowa"))
```

```{r include=FALSE}
counties <- map_data("county")
midwest <- subset(counties, region %in% c("wisconsin", "nebraska", "illinois", "iowa"))
```

```{r map, echo=FALSE}
ggplot(data = df, mapping = aes(x = long, y = lat, group = group, color=region)) + 
  geom_point(data=Map, mapping=aes(x=Longitude, y=Latitude, group=region, color=region), size=5) +
  geom_polygon(data = midwest, fill = "NA", color = "gray") +
  geom_polygon(color = "black", fill = NA) + scale_color_brewer(palette=c("Set1")) +
  labs(fill="", y="", x="", caption="") +
  north(df, location = 'topright', symbol = 1, anchor = c(x = -86, y = 47.3)) +
  theme_map() + 
  scalebar(data=df, dist = 100, dist_unit="km", model = 'International', location = "bottomleft", transform = TRUE) + 
  annotate("text", x = -101, y = 38, label = "Figure: Maxwel C Oliveira", size=4) +
      theme(axis.text=element_blank(), 
        axis.title=element_blank(), 
        panel.background = element_rect(fill="white", color = "black"), 
        panel.grid.major = element_line(color = "white"), 
  panel.grid.minor = element_blank(),
  legend.position=c(0.05,0.8), legend.text = element_text(size = 20, colour = "black"), legend.key = element_rect(fill=NA), panel.border = element_rect(fill=NA,color="black", size=0.5, linetype="solid"),
legend.background = element_rect(fill =NA), legend.key.width = unit(2.2, "line"), legend.title=element_blank()) +
  ggsave("map.png", dpi=600, width=18, height=12)
```

### 2021-07-20 data analysis/new_analyses.Rmd
jhjh

--- 2021-07-14 new_analyses.Rmd
@@
 ```
+
 ```{r}

### 2021-07-20 data analysis/weather/weather.Rmd
jhjh

---
title: "weather"
author: "Maxwel Coura Oliveira"
date: "7/8/2021"
output: html_document
---

```{r}
library(tidyverse) # Data wrangling
library(lubridate) # Dates operations
library(daymetr)
library(chirps)
library(patchwork)
library(ggtext)
library(nasapower)
library(vegan)
```

# Load data

You can load your data with read_csv if you have it in a spreadsheet.

```{r}
# your data
df_time <- tibble(ID = c('1','2','3','4','5'), 
                  Site = c('Arlington','Macomb','Lincoln',
                               'Clay Center', 'Grant'),
                  # Both coordinates in decimal format
                  latitude = c(43.18, 40.45, 40.85, 40.57, 40.8430),
                  longitude = c(-89.20, -90.67, -96.62, -98.13, -101.72),
                  Start = c('1988-01-01','1988-01-01','1988-01-01',
                                '1988-01-01', '1988-01-01'),
                  Early_planting = c("2018-06-01", "2018-06-01",
                                     "2018-06-01",
                                     "2018-06-01", "2018-06-01"),
                  Late_planting = c("2018-06-01", "2018-06-01", "2018-06-01",
                                    "2018-06-01", "2018-06-01"),
                      End = c('2020-12-31','2020-12-31','2020-12-31',
                              '2020-12-31', '2020-12-31'))

df_time <- df_time %>% 
  mutate_at(5:8, ymd) # as date
```

## Daymet

```{r function}
source("weather_daymet.R")
```

```{r get_data}
#dpp default = 0 - dpp = days prior start date

df_daymet <- weather_daymet(input = df_time, dpp = 0) 
```

```{r}
df_daymet %>% 
  janitor::clean_names() %>% 
  mutate(month_lab = month(date, label = TRUE)) -> adaptation_weather
```

```{r}
adaptation_weather %>% 
  filter(month %in% c(6, 7, 8, 9)) %>% 
  mutate(year = as.character(year),
         highlight = ifelse(year %in% c("2018", "2019"), TRUE, FALSE),
         variable = if_else(highlight == TRUE, year, "NA")) 
```

```{r}
text1 <- tibble(site = "Arlington")
text2 <- tibble(site = "Clay Center")
```

```{r}
adaptation_weather %>% 
  filter(month %in% c(6, 7, 8, 9)) %>% 
  mutate(year = as.character(year),
         highlight = ifelse(year %in% c("2018", "2019"), TRUE, FALSE),
         variable = if_else(highlight == TRUE, year, "NA")) %>% 
  ggplot(aes(x = month_lab, y = tmean)) +
  geom_jitter(aes(color = variable), alpha = 0.05, shape = 0) +
  scale_color_manual(values = c("red", "blue", "gray90")) +
  geom_line(adaptation_weather %>% 
               filter(month %in% c(6, 7, 8, 9) & year %in% c(2018)) %>% 
               group_by(site, month_lab) %>% 
               summarise(tmean = mean(tmean)),
             mapping = aes(x = month_lab, y = tmean, group = 1), 
             color = "red") +
  geom_line(adaptation_weather %>% 
               filter(month %in% c(6, 7, 8, 9) & year %in% c(2019)) %>% 
               group_by(site, month_lab) %>% 
               summarise(tmean = mean(tmean)),
             mapping = aes(x = month_lab, y = tmean, group = 1), 
             color = "blue") +
  geom_point(adaptation_weather %>% 
               filter(month %in% c(6, 7, 8, 9) & year %in% c(2018)) %>% 
               group_by(site, month_lab) %>% 
               summarise(tmean = mean(tmean)),
             mapping = aes(x = month_lab, y = tmean), 
             fill = "red", shape = 22, size = 3) +
  geom_point(adaptation_weather %>% 
               filter(month %in% c(6, 7, 8, 9) & year %in% c(2019)) %>% 
               group_by(site, month_lab) %>% 
               summarise(tmean = mean(tmean)),
             mapping = aes(x = month_lab, y = tmean), 
             fill = "blue", shape = 22, size = 3) +
  geom_richtext(data = text1, aes(x = 2, y = 7), 
            label = "Mean air temperature <br>in <b style='color:red;'>2018</b> and <b style='color:blue;'>2019</b>", size = 2, 
            label.color = NA, fill = NA) +
  geom_richtext(data = text2, aes(x = 2.2, y = 10), 
            label = "30 years <br>daily mean <br>air temperature", size = 2, 
             label.color = NA, fill = NA, color = "gray90") +
  geom_curve(data = text2, 
             x = 2, y = 13, 
             xend = 2.5, yend = 16,
             arrow = arrow(length = unit(0.07, "inch")), 
             size = 0.4, curvature = -0.3, color = "gray90") + 
  facet_grid(~ site) +
  labs(x = NULL, y = "Temperature (C)") +
  theme_test() +
  theme(legend.position = "none") -> temp

ggsave("weather.png", height = 3, width = 6)
```

```{r}
text3 <- tibble(site = "Grant")
text4 <- tibble(site = "Lincoln")
```

```{r}
adaptation_weather %>% 
  filter(month %in% c(6, 7, 8, 9)) %>% 
  group_by(month_lab, site, year) %>% 
  summarise(pp = sum(pp)) %>% 
  mutate(year = as.character(year),
         highlight = ifelse(year %in% c("2018", "2019"), TRUE, FALSE),
         variable = if_else(highlight == TRUE, year, "NA")) %>% 
  ggplot(aes(x = month_lab, y = pp)) +
  geom_jitter(aes(color = variable), alpha = 0.2) +
  scale_color_manual(values = c("red", "blue", "gray90")) +
  geom_line(adaptation_weather %>% 
               filter(month %in% c(6, 7, 8, 9) & year %in% c(2018)) %>% 
               group_by(site, month_lab) %>% 
               summarise(pp = sum(pp)),
             mapping = aes(x = month_lab, y = pp, group = 1), 
             color = "red") +
  geom_line(adaptation_weather %>% 
               filter(month %in% c(6, 7, 8, 9) & year %in% c(2019)) %>% 
               group_by(site, month_lab) %>% 
               summarise(pp = sum(pp)),
             mapping = aes(x = month_lab, y = pp, group = 1), 
             color = "blue") +
  geom_point(adaptation_weather %>% 
               filter(month %in% c(6, 7, 8, 9) & year %in% c(2018)) %>% 
               group_by(site, month_lab) %>% 
               summarise(pp = sum(pp)),
             mapping = aes(x = month_lab, y = pp), 
             fill = "red", shape = 21, size = 3) +
  geom_point(adaptation_weather %>% 
               filter(month %in% c(6, 7, 8, 9) & year %in% c(2019)) %>% 
               group_by(site, month_lab) %>% 
               summarise(pp = sum(pp)),
             mapping = aes(x = month_lab, y = pp), 
             fill = "blue", shape = 21, size = 3) +
  geom_richtext(data = text3, aes(x = 2.6, y = 330), 
            label = "Monthly sum precipitation <br>in <b style='color:red;'>2018</b> and <b style='color:blue;'>2019</b>", size = 2, 
            label.color = NA, fill = NA) +
  geom_richtext(data = text4, aes(x = 3.6, y = 320), 
            label = "30 years <br>monthly sum <br>precipitation", size = 2, 
             label.color = NA, fill = NA, color = "gray90") +
  geom_curve(data = text4, 
             x = 2.6, y = 310, 
             xend = 2, yend = 320,
             arrow = arrow(length = unit(0.07, "inch")), 
             size = 0.4, curvature = -0.3, color = "gray90") + 
  facet_grid(~ site) +
  labs(x = NULL, y = "Precipitation (mm)") +
  theme_test() +
  theme(legend.position = "none") -> ppp

ggsave("ppp.png", height = 3, width = 6)
```

```{r}
temp / ppp

ggsave("Figure 1.png")
```

### 2021-07-21 Frontiers/Frontiers.Rmd
cv

--- 2021-07-09 Frontiers/Frontiers.Rmd
@@
 ---
-title: "Adaptation of Palmer amaranth to the upper Midwest"
+title: "Adaptation of Palmer amaranth to cropping systems"
 runtitle: "Short Title"
@@
 
-kgghkg
+Palmer amaranth (*Amaranthus palmeri* S.
+Watson) is currently considered one of the most economically damaged weed species to cropping systems in the United States [@ward2013].
+The species has showed a remarkable capacity to evolve resistance to herbicides.
+Thus far, Palmer amaranth has evolved resistance to eight herbicide sites of action, increasing the weed management complexity [@lindsay2017].
+Uncontrolled Palmer amaranth in competition for water, light and nutrients can drastically reduce crop yields [@berger2015].
+Although currently considered a problematic weed, Palmer amaranth would emerged as a threat to sustainable agriculture only in the 1990s.
+
+Palmer amaranth is a fast growing summer annual forb indigenous to northern Mexico and southern US [@sauer1957], Palmer amaranth weediness is likely a result of human intervention in combination with species biology.
+Use and movement of farm equipment across locations, conservation agriculture, and reliance on herbicides for weed management are the main human driver selection of Palmer amaranth to cropping systems.
+On the other hand, Palmer amaranth is a prolific seed producer with a C4 photosynthetic apparatus.
+With a dioecy nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging herbicide resistant alleles among plants.
+Also, Palmer amaranth small seed size tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure[@yuundefined/ed], and animals [@farmer2017].
+These combinations make Palmer amaranth one of the most successful cases of weed adaption to cropping systems.
+
+Palmer amaranth is commonly found at crop [@garetson2019] and non-crop land [@bagavathiannan2016] in the southern United States, and its range is expanding northward.
+For example, herbicide resistant Palmer amaranth is widespread in Nebraska  [@oliveira2021a] and Michigan [@kohrt2017].
+Successful cases of Palmer eradication is documented in Minnesota [@yuundefined/ed] and Iowa.
+Palmer amaranth was also found in Argentina [@larran2017], Brazil [@kupper2017], and Italy [@milani2021].
+Climate change can increase Palmer amaranth invasion to Australia and Sub-Sahara Africa [@kistner2018].
+Palmer amaranth can survive under continuous water stress and produce a significant amount of seeds @chahal2018].
+Produced seeds from Palmer amaranth growing under drier conditions are less dormant and prompt for germination [@matzrafi2021].
+Therefore, Palmer amaranth has showed a plasticity to adapt and potential to invade new environments.
+
+talk about the ecological approach
+
+The increase movement of Palmer amaranth to US Midwest warrants investigation on species adaptation to that environment. Most studies are based on reactive but not in the proactive management.
+
+Understanding Palmer amaranth biology can enhance our knowledge on species adaptation to new environments and aid on design proactive and ecological weed management strategies.
+Therefore, the objective of this study was to investigate the flowering pattern, biomass, and height of Palmer amaranth growing under different environments and timings across five locations.
 
@@
 # Author Contributions {-}
+
+MCO design, wrote,
 

### 2021-07-21 Frontiers/Frontiers.tex
cv

@@
 
-\title[Short Title]{Adaptation of Palmer amaranth to croppins systems}
+\title[Short Title]{Adaptation of Palmer amaranth to cropping systems}
 \author[\firstAuthorLast]{\Authors}
@@
 
-Palmer amaranth (\emph{Amaranthus palmeri}) is an indigenous species
-from southwestern United States and northern Mexico. Palmer amaranth is
-a C4 annual broadleaf forb within the \textbf{Amarantacea} family.
-Palmer amaranth is currently considered one of the most troublesome weed
-species in the United States.
+Palmer amaranth (\emph{Amaranthus palmeri} S. Watson) is currently
+considered one of the most economically damaged weed species to cropping
+systems in the United States (Ward et al., 2013). The species has showed
+a remarkable capacity to evolve resistance to herbicides. Thus far,
+Palmer amaranth has evolved resistance to eight herbicide sites of
+action, increasing the weed management complexity (Lindsay et al.,
+2017). Uncontrolled Palmer amaranth in competition for water, light and
+nutrients can drastically reduce crop yields (Berger et al., 2015).
+Although currently considered a problematic weed, Palmer amaranth would
+emerged as a threat to sustainable agriculture only in the 1990s.
+
+Palmer amaranth is a fast growing summer annual forb indigenous to
+northern Mexico and southern US (Sauer, 1957), Palmer amaranth weediness
+is likely a result of human intervention in combination with species
+biology. Use and movement of farm equipment across locations,
+conservation agriculture, and reliance on herbicides for weed management
+are the main human driver selection of Palmer amaranth to cropping
+systems. On the other hand, Palmer amaranth is a prolific seed producer
+with a C4 photosynthetic apparatus. With a dioecy nature, Palmer
+amaranth male and female plants are obligate outcrosser species,
+increasing the chances of exchanging herbicide resistant alleles among
+plants. Also, Palmer amaranth small seed size tend to thrive in
+no-tillage systems (Price et al., 2011), and spread across locations
+through farm equipment (Sauer, 1972), manure(Yu et al., undefined/ed),
+and animals (Farmer et al., 2017). These combinations make Palmer
+amaranth one of the most successful cases of weed adaption to cropping
+systems.
+
+Palmer amaranth is commonly found at crop (Garetson et al., 2019) and
+non-crop land (Bagavathiannan and Norsworthy, 2016) in the southern
+United States, and its range is expanding northward. For example,
+herbicide resistant Palmer amaranth is widespread in Nebraska (Oliveira
+et al., 2021) and Michigan (Kohrt et al., 2017). Successful cases of
+Palmer eradication is documented in Minnesota (Yu et al., undefined/ed)
+and Iowa. Palmer amaranth was also found in Argentina (Larran et al.,
+2017), Brazil (Küpper et al., 2017), and Italy (Milani et al., 2021).
+Climate change can increase Palmer amaranth invasion to Australia and
+Sub-Sahara Africa (Kistner and Hatfield, 2018). Palmer amaranth can
+survive under continuous water stress and produce a significant amount
+of seeds Chahal et al. (2018){]}. Produced seeds from Palmer amaranth
+growing under drier conditions are less dormant and prompt for
+germination (Matzrafi et al., 2021). Therefore, Palmer amaranth has
+showed a plasticity to adapt and potential to invade new environment.
+
+The increase movement of Palmer amaranth to US Midwest warrants
+investigation on species adaptation to that environment. Most studies
+are based on reactive but not in the proactive management.
+
+Understanding Palmer amaranth biology can enhance our knowledge on
+species adaptation to new environments and aid on design proactive and
+ecological weed management strategies. Therefore, the objective of this
+study was to investigate the flowering pattern, biomass, and height of
+Palmer amaranth growing under different environments and timings across
+five locations.
 
@@
 
+MCO design, wrote,
+
 The statement about the authors and contributors can be up to several
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-bagavathiannan2016}{}%
+Bagavathiannan, M. V., and Norsworthy, J. K. (2016). Multiple-{Herbicide
+Resistance Is Widespread} in {Roadside Palmer Amaranth Populations}.
+\emph{PLOS ONE} 11, e0148748.
+doi:\href{https://doi.org/10.1371/journal.pone.0148748}{10.1371/journal.pone.0148748}.
+
+\leavevmode\hypertarget{ref-berger2015}{}%
+Berger, S. T., Ferrell, J. A., Rowland, D. L., and Webster, T. M.
+(2015). Palmer {Amaranth} ({Amaranthus} palmeri) {Competition} for
+{Water} in {Cotton}. \emph{Weed Science} 63, 928--935.
+doi:\href{https://doi.org/10.1614/WS-D-15-00062.1}{10.1614/WS-D-15-00062.1}.
+
+\leavevmode\hypertarget{ref-chahal2018}{}%
+Chahal, P. S., Irmak, S., Jugulam, M., and Jhala, A. J. (2018).
+Evaluating {Effect} of {Degree} of {Water Stress} on {Growth} and
+{Fecundity} of {Palmer} amaranth ({Amaranthus} palmeri) {Using Soil
+Moisture Sensors}. \emph{Weed Science} 66, 738--745.
+doi:\href{https://doi.org/10.1017/wsc.2018.47}{10.1017/wsc.2018.47}.
+
+\leavevmode\hypertarget{ref-farmer2017}{}%
+Farmer, J. A., Webb, E. B., Pierce, R. A., and Bradley, K. W. (2017).
+Evaluating the potential for weed seed dispersal based on waterfowl
+consumption and seed viability. \emph{Pest Management Science} 73,
+2592--2603. doi:\href{https://doi.org/10.1002/ps.4710}{10.1002/ps.4710}.
+
+\leavevmode\hypertarget{ref-garetson2019}{}%
+Garetson, R., Singh, V., Singh, S., Dotray, P., and Bagavathiannan, M.
+(2019). Distribution of herbicide-resistant {Palmer} amaranth
+({Amaranthus} palmeri) in row crop production systems in {Texas}.
+\emph{Weed Technology} 33, 355--365.
+doi:\href{https://doi.org/10.1017/wet.2019.14}{10.1017/wet.2019.14}.
+
+\leavevmode\hypertarget{ref-kistner2018}{}%
+Kistner, E. J., and Hatfield, J. L. (2018). Potential {Geographic
+Distribution} of {Palmer Amaranth} under {Current} and {Future
+Climates}. \emph{Agricultural \& Environmental Letters} 3, 170044.
+doi:\href{https://doi.org/10.2134/ael2017.12.0044}{10.2134/ael2017.12.0044}.
+
+\leavevmode\hypertarget{ref-kohrt2017}{}%
+Kohrt, J. R., Sprague, C. L., Nadakuduti, S. S., and Douches, D. (2017).
+Confirmation of a {Three}-{Way} ({Glyphosate}, {ALS}, and {Atrazine})
+{Herbicide}-{Resistant Population} of {Palmer Amaranth} ({Amaranthus}
+palmeri) in {Michigan}. \emph{Weed Science} 65, 327--338.
+doi:\href{https://doi.org/10.1017/wsc.2017.2}{10.1017/wsc.2017.2}.
+
+\leavevmode\hypertarget{ref-kupper2017}{}%
+Küpper, A., Borgato, E. A., Patterson, E. L., Netto, A. G., Nicolai, M.,
+Carvalho, S. J. P. de, Nissen, S. J., Gaines, T. A., and Christoffoleti,
+P. J. (2017). Multiple {Resistance} to {Glyphosate} and {Acetolactate
+Synthase Inhibitors} in {Palmer Amaranth} ({Amaranthus} palmeri)
+{Identified} in {Brazil}. \emph{Weed Science} 65, 317--326.
+doi:\href{https://doi.org/10.1017/wsc.2017.1}{10.1017/wsc.2017.1}.
+
+\leavevmode\hypertarget{ref-larran2017}{}%
+Larran, A. S., Palmieri, V. E., Perotti, V. E., Lieber, L., Tuesca, D.,
+and Permingeat, H. R. (2017). Target-site resistance to acetolactate
+synthase ({ALS})-inhibiting herbicides in {Amaranthus} palmeri from
+{Argentina}. \emph{Pest Management Science} 73, 2578--2584.
+doi:\href{https://doi.org/10.1002/ps.4662}{10.1002/ps.4662}.
+
+\leavevmode\hypertarget{ref-lindsay2017}{}%
+Lindsay, K., Popp, M., Norsworthy, J., Bagavathiannan, M., Powles, S.,
+and Lacoste, M. (2017). {PAM}: {Decision Support} for {Long}-{Term
+Palmer Amaranth} ({Amaranthus} palmeri) {Control}. \emph{Weed
+Technology} 31, 915--927.
+doi:\href{https://doi.org/10.1017/wet.2017.69}{10.1017/wet.2017.69}.
+
+\leavevmode\hypertarget{ref-matzrafi2021}{}%
+Matzrafi, M., Osipitan, O. A., Ohadi, S., and Mesgaran, M. B. (2021).
+Under pressure: Maternal effects promote drought tolerance in progeny
+seed of {Palmer} amaranth ({Amaranthus} palmeri). \emph{Weed Science}
+69, 31--38.
+doi:\href{https://doi.org/10.1017/wsc.2020.75}{10.1017/wsc.2020.75}.
+
+\leavevmode\hypertarget{ref-milani2021}{}%
+Milani, A., Panozzo, S., Farinati, S., Iamonico, D., Sattin, M., Loddo,
+D., and Scarabel, L. (2021). Recent {Discovery} of {Amaranthus} palmeri
+{S}. {Watson} in {Italy}: {Characterization} of {ALS}-{Resistant
+Populations} and {Sensitivity} to {Alternative Herbicides}.
+\emph{Sustainability} 13, 7003.
+doi:\href{https://doi.org/10.3390/su13137003}{10.3390/su13137003}.
+
 \leavevmode\hypertarget{ref-oliveira2021a}{}%
@@
 
+\leavevmode\hypertarget{ref-price2011}{}%
+Price, A. J., Balkcom, K. S., Culpepper, S. A., Kelton, J. A., Nichols,
+R. L., and Schomberg, H. (2011). Glyphosate-resistant {Palmer} amaranth:
+{A} threat to conservation tillage. \emph{Journal of Soil and Water
+Conservation} 66, 265--275.
+doi:\href{https://doi.org/10.2489/jswc.66.4.265}{10.2489/jswc.66.4.265}.
+
 \leavevmode\hypertarget{ref-ritz2015}{}%
@@
 
+\leavevmode\hypertarget{ref-sauer1957}{}%
+Sauer, J. (1957). Recent {Migration} and {Evolution} of the {Dioecious
+Amaranths}. \emph{Evolution} 11, 11--31.
+doi:\href{https://doi.org/10.2307/2405808}{10.2307/2405808}.
+
+\leavevmode\hypertarget{ref-sauer1972}{}%
+Sauer, J. D. (1972). The dioecious amaranths: A new species name and
+major range extensions. \emph{Madroño} 21, 426--434. Available at:
+\url{http://www.jstor.org/stable/41423815}.
+
+\leavevmode\hypertarget{ref-ward2013}{}%
+Ward, S. M., Webster, T. M., and Steckel, L. E. (2013). Palmer
+{Amaranth} ({Amaranthus} palmeri): {A Review}. \emph{Weed Technology}
+27, 12--27.
+doi:\href{https://doi.org/10.1614/WT-D-12-00113.1}{10.1614/WT-D-12-00113.1}.
+
+\leavevmode\hypertarget{ref-yuundefinedux2fed}{}%
+Yu, E., Blair, S., Hardel, M., Chandler, M., Thiede, D., Cortilet, A.,
+Gunsolus, J., and Becker, R. (undefined/ed). Timeline of {Palmer}
+amaranth ({Amaranthus} palmeri) invasion and eradication in {Minnesota}.
+\emph{Weed Technology}, 1--31.
+doi:\href{https://doi.org/10.1017/wet.2021.32}{10.1017/wet.2021.32}.
+
 \end{CSLReferences}

### 2021-07-23 Frontiers/Frontiers.Rmd
as

--- 2021-07-09 Frontiers/Frontiers.Rmd
@@
 
-kgghkg
+Palmer amaranth (*Amaranthus palmeri* S.
+Watson) is currently considered one of the most economically damaged weed species to cropping systems in the United States [@ward2013].
+The species has showed a remarkable capacity to evolve resistance to herbicides.
+Palmer amaranth has evolved resistance to eight herbicide sites of action, increasing the weed management complexity [@lindsay2017].
+Uncontrolled Palmer amaranth in competition for water, light and nutrients can drastically reduce crop yields [@berger2015].
+Palmer amaranth is documented with potential to reduce 91%, 68%, and X% of corn [@massinga2001], soybean [@klingaman1994], and cotton yields.
+
+Palmer amaranth is a fast growing summer annual forb indigenous to Sonoran Desert [@sauer1957].
+The species would eventually emerge as a threat to US agriculture in the 1990s.
+Palmer amaranth weediness is likely a result of human-assisted selection in combination with species biology.
+The increased use and movement of farm equipment across locations, conservation agriculture (e.g., no-till), and reliance on herbicides for weed management are the main human mediated selection of Palmer amaranth to cropping systems.
+On the other hand, Palmer amaranth is a prolific seed producer with a C4 photosynthetic apparatus [@ward2013].
+With a dioecy nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging herbicide resistant alleles among plants [@oliveira2018].
+Also, Palmer amaranth small seed size (1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure [@yu2021], animals [@farmer2017], and plant propagules [@hartzler2016].
+Therefore, Palmer amaranth dispersal capacity make the species one of the most successful cases of weed adaption to cropping systems.
+
+Light and temperature are likely the main environment requirements for Palmer amaranth successful adaptation.
+Palmer amaranth is reported with an extended germination period [@jha2010].
+Germination of Palmer amaranth is triggered by 18 C soil temperature [@keeley1987], and optimal germination and biomass production were at 35/30 C day and night temperatures [@guo2003].
+Water has not shown to limit Palmer amaranth fitness.
+Under continuous water stress, Palmer amaranth survived and produced at least 14000 seeds plant-1 [@chahal2018].
+Also, seeds from Palmer amaranth growing under water stress conditions were heavier, less dormant, and prompt for germination [@matzrafi2021].
+The continuous global temperature warming can impact agriculture and promote niches for Palmer amaranth invasion/adaptation into new environments.
+Currently, the greatest climatic risk of Palmer amaranth establishment is agronomic crops in Australia and Sub-Sahara Africa [@kistner2018].
+Temperature is a key factor limiting Palmer amaranth northward range expansion [@briscoerunquist2019]; however, under future climate change Palmer amaranth is likely to expand northward into Canada and Northern Europe [@kistner2018].
+
+Palmer amaranth is already found in agronomic crops of South America [@larran2017; @kupper2017] and Southern Europe [@milani2021].
+In the US, Palmer amaranth is commonly found at crop [@garetson2019] and non-crop land [@bagavathiannan2016] in the warm southern United States but its range is expanding to cool temperatures northward.
+For example, herbicide resistant Palmer amaranth is widespread in Nebraska  [@oliveira2021a],  Michigan [@kohrt2017], and Connecticut [@aulakh2021].
+Successful cases of Palmer amaranth invasion and near to eradication is well documented in Minnesota [@yu2021] and Iowa.
+No Palmer amaranth actively growing was found in Canada; however, Palmer amaranth seeds was detected in sweet potato slips [@page2021].
+Nonetheless, it seems fated to manage Palmer amaranth in agronomic crops throughout multiple environments in the near future.
+Therefore, strategies on Palmer amaranth management should encompass the agroecosystem level but not attempts to eradicate the weed.
+The continuous Palmer amaranth dispersal and potential establishment into northern US/Canada warrant investigation on species growing morphology in such environments.
+
+Most Palmer amaranth studies are based on reactive (e,g, herbicide and tillage) rather than proactive management.
+Understanding Palmer amaranth biology and growing strategies under different environments can enhance our knowledge on species adaptation.
+Also, it can aid on designing proactive and ecological tactics to limit Palmer amaranth range expansion and reduce its negative impact while conserving diversity.
+Therefore, the objective of this study was to investigate the flowering pattern, biomass production, and height of Palmer amaranth growing under different environments and timings across five locations.
 

### 2021-07-23 Frontiers/Frontiers.tex
as

@@
 systems in the United States (Ward et al., 2013). The species has showed
-a remarkable capacity to evolve resistance to herbicides. Thus far,
-Palmer amaranth has evolved resistance to eight herbicide sites of
-action, increasing the weed management complexity (Lindsay et al.,
-2017). Uncontrolled Palmer amaranth in competition for water, light and
+a remarkable capacity to evolve resistance to herbicides. Palmer
+amaranth has evolved resistance to eight herbicide sites of action,
+increasing the weed management complexity (Lindsay et al., 2017).
+Uncontrolled Palmer amaranth in competition for water, light and
 nutrients can drastically reduce crop yields (Berger et al., 2015).
-Although currently considered a problematic weed, Palmer amaranth would
-emerged as a threat to sustainable agriculture only in the 1990s.
+Palmer amaranth is documented with potential to reduce 91\%, 68\%, and
+X\% of corn (Massinga et al., 2001), soybean (Klingaman and Oliver,
+1994), and cotton yields.
 
 Palmer amaranth is a fast growing summer annual forb indigenous to
-northern Mexico and southern US (Sauer, 1957), Palmer amaranth weediness
-is likely a result of human intervention in combination with species
-biology. Use and movement of farm equipment across locations,
-conservation agriculture, and reliance on herbicides for weed management
-are the main human driver selection of Palmer amaranth to cropping
-systems. On the other hand, Palmer amaranth is a prolific seed producer
-with a C4 photosynthetic apparatus. With a dioecy nature, Palmer
-amaranth male and female plants are obligate outcrosser species,
-increasing the chances of exchanging herbicide resistant alleles among
-plants. Also, Palmer amaranth small seed size tend to thrive in
-no-tillage systems (Price et al., 2011), and spread across locations
-through farm equipment (Sauer, 1972), manure(Yu et al., undefined/ed),
-and animals (Farmer et al., 2017). These combinations make Palmer
-amaranth one of the most successful cases of weed adaption to cropping
-systems.
+Sonoran Desert (Sauer, 1957). The species would eventually emerge as a
+threat to US agriculture in the 1990s. Palmer amaranth weediness is
+likely a result of human-assisted selection in combination with species
+biology. The increased use and movement of farm equipment across
+locations, conservation agriculture (e.g., no-till), and reliance on
+herbicides for weed management are the main human mediated selection of
+Palmer amaranth to cropping systems. On the other hand, Palmer amaranth
+is a prolific seed producer with a C4 photosynthetic apparatus (Ward et
+al., 2013). With a dioecy nature, Palmer amaranth male and female plants
+are obligate outcrosser species, increasing the chances of exchanging
+herbicide resistant alleles among plants (Oliveira et al., 2018). Also,
+Palmer amaranth small seed size (1 mm) tend to thrive in no-tillage
+systems (Price et al., 2011), and spread across locations through farm
+equipment (Sauer, 1972), manure (Yu et al., 2021), animals (Farmer et
+al., 2017), and plant propagules (Hartzler and Anderson, 2016).
+Therefore, Palmer amaranth dispersal capacity make the species one of
+the most successful cases of weed adaption to cropping systems.
 
-Palmer amaranth is commonly found at crop (Garetson et al., 2019) and
-non-crop land (Bagavathiannan and Norsworthy, 2016) in the southern
-United States, and its range is expanding northward. For example,
-herbicide resistant Palmer amaranth is widespread in Nebraska (Oliveira
-et al., 2021) and Michigan (Kohrt et al., 2017). Successful cases of
-Palmer eradication is documented in Minnesota (Yu et al., undefined/ed)
-and Iowa. Palmer amaranth was also found in Argentina (Larran et al.,
-2017), Brazil (Küpper et al., 2017), and Italy (Milani et al., 2021).
-Climate change can increase Palmer amaranth invasion to Australia and
-Sub-Sahara Africa (Kistner and Hatfield, 2018). Palmer amaranth can
-survive under continuous water stress and produce a significant amount
-of seeds Chahal et al. (2018){]}. Produced seeds from Palmer amaranth
-growing under drier conditions are less dormant and prompt for
-germination (Matzrafi et al., 2021). Therefore, Palmer amaranth has
-showed a plasticity to adapt and potential to invade new environment.
+Light and temperature are likely the main environment requirements for
+Palmer amaranth successful adaptation. Palmer amaranth is reported with
+an extended germination period (Jha et al., 2010). Germination of Palmer
+amaranth is triggered by 18 C soil temperature (Keeley et al., 1987),
+and optimal germination and biomass production were at 35/30 C day and
+night temperatures (Guo and Al-Khatib, 2003). Water has not shown to
+limit Palmer amaranth fitness. Under continuous water stress, Palmer
+amaranth survived and produced at least 14000 seeds plant-1 (Chahal et
+al., 2018). Also, seeds from Palmer amaranth growing under water stress
+conditions were heavier, less dormant, and prompt for germination
+(Matzrafi et al., 2021). The continuous global temperature warming can
+impact agriculture and promote niches for Palmer amaranth
+invasion/adaptation into new environments. Currently, the greatest
+climatic risk of Palmer amaranth establishment is agronomic crops in
+Australia and Sub-Sahara Africa (Kistner and Hatfield, 2018).
+Temperature is a key factor limiting Palmer amaranth northward range
+expansion (Briscoe Runquist et al., 2019); however, under future climate
+change Palmer amaranth is likely to expand northward into Canada and
+Northern Europe (Kistner and Hatfield, 2018).
 
-The increase movement of Palmer amaranth to US Midwest warrants
-investigation on species adaptation to that environment. Most studies
-are based on reactive but not in the proactive management.
+Palmer amaranth is already found in agronomic crops of South America
+(Larran et al., 2017; Küpper et al., 2017) and Southern Europe (Milani
+et al., 2021). In the US, Palmer amaranth is commonly found at crop
+(Garetson et al., 2019) and non-crop land (Bagavathiannan and
+Norsworthy, 2016) in the warm southern United States but its range is
+expanding to cool temperatures northward. For example, herbicide
+resistant Palmer amaranth is widespread in Nebraska (Oliveira et al.,
+2021), Michigan (Kohrt et al., 2017), and Connecticut (Aulakh et al.,
+2021). Successful cases of Palmer amaranth invasion and near to
+eradication is well documented in Minnesota (Yu et al., 2021) and Iowa.
+No Palmer amaranth actively growing was found in Canada; however, Palmer
+amaranth seeds was detected in sweet potato slips (Page et al., 2021).
+Nonetheless, it seems fated to manage Palmer amaranth in agronomic crops
+throughout multiple environments in the near future. Therefore,
+strategies on Palmer amaranth management should encompass the
+agroecosystem level but not attempts to eradicate the weed. The
+continuous Palmer amaranth dispersal and potential establishment into
+northern US/Canada warrant investigation on species growing morphology
+in such environments.
 
-Understanding Palmer amaranth biology can enhance our knowledge on
-species adaptation to new environments and aid on design proactive and
-ecological weed management strategies. Therefore, the objective of this
-study was to investigate the flowering pattern, biomass, and height of
-Palmer amaranth growing under different environments and timings across
-five locations.
+Most Palmer amaranth studies are based on reactive (e,g, herbicide and
+tillage) rather than proactive management. Understanding Palmer amaranth
+biology and growing strategies under different environments can enhance
+our knowledge on species adaptation. Also, it can aid on designing
+proactive and ecological tactics to limit Palmer amaranth range
+expansion and reduce its negative impact while conserving diversity.
+Therefore, the objective of this study was to investigate the flowering
+pattern, biomass production, and height of Palmer amaranth growing under
+different environments and timings across five locations.
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-aulakh2021}{}%
+Aulakh, J. S., Chahal, P. S., Kumar, V., Price, A. J., and Guillard, K.
+(2021). Multiple herbicide-resistant {Palmer} amaranth ({Amaranthus}
+palmeri) in {Connecticut}: Confirmation and response to {POST}
+herbicides. \emph{Weed Technology} 35, 457--463.
+doi:\href{https://doi.org/10.1017/wet.2021.6}{10.1017/wet.2021.6}.
+
 \leavevmode\hypertarget{ref-bagavathiannan2016}{}%
@@
 doi:\href{https://doi.org/10.1614/WS-D-15-00062.1}{10.1614/WS-D-15-00062.1}.
+
+\leavevmode\hypertarget{ref-briscoerunquist2019}{}%
+Briscoe Runquist, R. D., Lake, T., Tiffin, P., and Moeller, D. A.
+(2019). Species distribution models throughout the invasion history of
+{Palmer} amaranth predict regions at risk of future invasion and reveal
+challenges with modeling rapidly shifting geographic ranges. \emph{Sci
+Rep} 9, 2426.
+doi:\href{https://doi.org/10.1038/s41598-018-38054-9}{10.1038/s41598-018-38054-9}.
 
@@
 
+\leavevmode\hypertarget{ref-guo2003}{}%
+Guo, P., and Al-Khatib, K. (2003). Temperature effects on germination
+and growth of redroot pigweed ({Amaranthus} retroflexus), {Palmer}
+amaranth ({A}. Palmeri), and common waterhemp ({A}. rudis). \emph{Weed
+Science} 51, 869--875.
+doi:\href{https://doi.org/10.1614/P2002-127}{10.1614/P2002-127}.
+
+\leavevmode\hypertarget{ref-hartzler2016}{}%
+Hartzler, B., and Anderson, M. (2016). Palmer amaranth: {It}'s here, now
+what? 10.
+
+\leavevmode\hypertarget{ref-jha2010}{}%
+Jha, P., Norsworthy, J. K., Riley, M. B., and Bridges, W. (2010). Annual
+{Changes} in {Temperature} and {Light Requirements} for {Germination} of
+{Palmer Amaranth} ({Amaranthus} palmeri) {Seeds Retrieved} from {Soil}.
+\emph{Weed Science} 58, 426--432.
+doi:\href{https://doi.org/10.1614/WS-D-09-00038.1}{10.1614/WS-D-09-00038.1}.
+
+\leavevmode\hypertarget{ref-keeley1987}{}%
+Keeley, P. E., Carter, C. H., and Thullen, R. J. (1987). Influence of
+{Planting Date} on {Growth} of {Palmer Amaranth} ({Amaranthus} palmeri).
+\emph{Weed Science} 35, 199--204.
+doi:\href{https://doi.org/10.1017/S0043174500079054}{10.1017/S0043174500079054}.
+
 \leavevmode\hypertarget{ref-kistner2018}{}%
@@
 doi:\href{https://doi.org/10.2134/ael2017.12.0044}{10.2134/ael2017.12.0044}.
+
+\leavevmode\hypertarget{ref-klingaman1994}{}%
+Klingaman, T. E., and Oliver, L. R. (1994). Palmer {Amaranth}
+({Amaranthus} palmeri) {Interference} in {Soybeans} ({Glycine} max).
+\emph{Weed Science} 42, 523--527.
+doi:\href{https://doi.org/10.1017/S0043174500076888}{10.1017/S0043174500076888}.
 
@@
 
+\leavevmode\hypertarget{ref-massinga2001}{}%
+Massinga, R. A., Currie, R. S., Horak, M. J., and Boyer, J. (2001).
+Interference of {Palmer} amaranth in corn. \emph{Weed Science} 49,
+202--208.
+doi:\href{https://doi.org/10.1614/0043-1745(2001)049\%5B0202:IOPAIC\%5D2.0.CO;2}{10.1614/0043-1745(2001)049{[}0202:IOPAIC{]}2.0.CO;2}.
+
 \leavevmode\hypertarget{ref-matzrafi2021}{}%
@@
 
+\leavevmode\hypertarget{ref-oliveira2018}{}%
+Oliveira, M. C., Gaines, T. A., Patterson, E. L., Jhala, A. J., Irmak,
+S., Amundsen, K., and Knezevic, S. Z. (2018). Interspecific and
+intraspecific transference of metabolism-based mesotrione resistance in
+dioecious weedy {Amaranthus}. \emph{The Plant Journal} 96, 1051--1063.
+doi:\href{https://doi.org/10.1111/tpj.14089}{10.1111/tpj.14089}.
+
 \leavevmode\hypertarget{ref-oliveira2021a}{}%
@@
 doi:\href{https://doi.org/10.1017/wet.2020.74}{10.1017/wet.2020.74}.
+
+\leavevmode\hypertarget{ref-page2021}{}%
+Page, E. R., Nurse, R. E., Meloche, S., Bosveld, K., Grainger, C.,
+Obeid, K., Filotas, M., Simard, M.-J., and Laforest, M. (2021). Import
+of {Palmer} amaranth ({Amaranthus} palmeri {S}. {Wats}.) Seed with sweet
+potato ({Ipomea} batatas ({L}.) {Lam}) slips. \emph{Can. J. Plant Sci.},
+CJPS-2020-0321.
+doi:\href{https://doi.org/10.1139/CJPS-2020-0321}{10.1139/CJPS-2020-0321}.
 
@@
 
-\leavevmode\hypertarget{ref-yuundefinedux2fed}{}%
+\leavevmode\hypertarget{ref-yu2021}{}%
 Yu, E., Blair, S., Hardel, M., Chandler, M., Thiede, D., Cortilet, A.,
-Gunsolus, J., and Becker, R. (undefined/ed). Timeline of {Palmer}
-amaranth ({Amaranthus} palmeri) invasion and eradication in {Minnesota}.
+Gunsolus, J., and Becker, R. (2021). Timeline of {Palmer} amaranth
+({Amaranthus} palmeri) invasion and eradication in {Minnesota}.
 \emph{Weed Technology}, 1--31.

### 2021-07-26 Frontiers/Frontiers.Rmd
as

@@
 Uncontrolled Palmer amaranth in competition for water, light and nutrients can drastically reduce crop yields [@berger2015].
-Palmer amaranth is documented with potential to reduce 91%, 68%, and X% of corn [@massinga2001], soybean [@klingaman1994], and cotton yields.
+Palmer amaranth is documented with potential to reduce 91%, 68%, and 54% of corn [@massinga2001], soybean [@klingaman1994], and cotton [@morgan2001] yields.
 
@@
 Palmer amaranth weediness is likely a result of human-assisted selection in combination with species biology.
-The increased use and movement of farm equipment across locations, conservation agriculture (e.g., no-till), and reliance on herbicides for weed management are the main human mediated selection of Palmer amaranth to cropping systems.
+Farm mechanization, conservation agriculture (e.g., no-till), and reliance on herbicides for weed management are the main human mediated selection of Palmer amaranth to cropping systems.
 On the other hand, Palmer amaranth is a prolific seed producer with a C4 photosynthetic apparatus [@ward2013].
@@
 Palmer amaranth is reported with an extended germination period [@jha2010].
-Germination of Palmer amaranth is triggered by 18 C soil temperature [@keeley1987], and optimal germination and biomass production were at 35/30 C day and night temperatures [@guo2003].
+Germination of Palmer amaranth is triggered by 18 C soil temperature [@keeley1987], and optimal germination and biomass production occur at 35/30 C day and night temperatures [@guo2003].
 Water has not shown to limit Palmer amaranth fitness.
@@
 The continuous global temperature warming can impact agriculture and promote niches for Palmer amaranth invasion/adaptation into new environments.
-Currently, the greatest climatic risk of Palmer amaranth establishment is agronomic crops in Australia and Sub-Sahara Africa [@kistner2018].
-Temperature is a key factor limiting Palmer amaranth northward range expansion [@briscoerunquist2019]; however, under future climate change Palmer amaranth is likely to expand northward into Canada and Northern Europe [@kistner2018].
+Currently, it is estimated that the greatest climatic risk of Palmer amaranth establishment is agronomic crops in Australia and Sub-Sahara Africa [@kistner2018].
+Temperature is a key factor limiting Palmer amaranth expansion to cooler geographies [@briscoerunquist2019]; however, under future climate change Palmer amaranth is likely to expand northward into Canada and Northern Europe [@kistner2018].
 
-Palmer amaranth is already found in agronomic crops of South America [@larran2017; @kupper2017] and Southern Europe [@milani2021].
-In the US, Palmer amaranth is commonly found at crop [@garetson2019] and non-crop land [@bagavathiannan2016] in the warm southern United States but its range is expanding to cool temperatures northward.
+Palmer amaranth is already documented in agronomic crops of South America [@larran2017; @kupper2017] and Southern Europe [@milani2021].
+In the US, Palmer amaranth is established at crop [@garetson2019] and non-crop land [@bagavathiannan2016] in the warm southern United States but its range is expanding to cool temperatures northward.
 For example, herbicide resistant Palmer amaranth is widespread in Nebraska  [@oliveira2021a],  Michigan [@kohrt2017], and Connecticut [@aulakh2021].
-Successful cases of Palmer amaranth invasion and near to eradication is well documented in Minnesota [@yu2021] and Iowa.
+Successful cases of Palmer amaranth invasion and near to eradication is reported in Minnesota [@yu2021] and Iowa.
 No Palmer amaranth actively growing was found in Canada; however, Palmer amaranth seeds was detected in sweet potato slips [@page2021].
@@
 Therefore, strategies on Palmer amaranth management should encompass the agroecosystem level but not attempts to eradicate the weed.
-The continuous Palmer amaranth dispersal and potential establishment into northern US/Canada warrant investigation on species growing morphology in such environments.
+Most tactics to manage Palmer amaranth are based technology fixes [@scott2011], which are short-term (e.g., herbicide and/or tillage) rather than long-term ecological management.
 
-Most Palmer amaranth studies are based on reactive (e,g, herbicide and tillage) rather than proactive management.
-Understanding Palmer amaranth biology and growing strategies under different environments can enhance our knowledge on species adaptation.
-Also, it can aid on designing proactive and ecological tactics to limit Palmer amaranth range expansion and reduce its negative impact while conserving diversity.
-Therefore, the objective of this study was to investigate the flowering pattern, biomass production, and height of Palmer amaranth growing under different environments and timings across five locations.
+The continuous Palmer amaranth dispersal and potential establishment into northern United States warrant investigations on species morphology growing in such environments.
+Understanding Palmer amaranth biology and growing strategies under different agroecossystems can enhance our knowledge on species adaptation.
+It can also aid on designing proactive and ecological tactics to limit the species range expansion, reduce its negative impact, and design resilient and sustainable farming systems [@maclaren2020].
+Therefore, the objective of this study was to investigate the flowering pattern, biomass production, and height of Palmer amaranth growing under in corn, soybean and fallow at two timings across five locations in the mid/upper United States Midwest.
 

### 2021-07-26 Frontiers/Frontiers.tex
as

@@
 Palmer amaranth is documented with potential to reduce 91\%, 68\%, and
-X\% of corn (Massinga et al., 2001), soybean (Klingaman and Oliver,
-1994), and cotton yields.
+54\% of corn (Massinga et al., 2001), soybean (Klingaman and Oliver,
+1994), and cotton (Morgan et al., 2001) yields.
 
@@
 likely a result of human-assisted selection in combination with species
-biology. The increased use and movement of farm equipment across
-locations, conservation agriculture (e.g., no-till), and reliance on
-herbicides for weed management are the main human mediated selection of
-Palmer amaranth to cropping systems. On the other hand, Palmer amaranth
-is a prolific seed producer with a C4 photosynthetic apparatus (Ward et
-al., 2013). With a dioecy nature, Palmer amaranth male and female plants
-are obligate outcrosser species, increasing the chances of exchanging
-herbicide resistant alleles among plants (Oliveira et al., 2018). Also,
-Palmer amaranth small seed size (1 mm) tend to thrive in no-tillage
-systems (Price et al., 2011), and spread across locations through farm
-equipment (Sauer, 1972), manure (Yu et al., 2021), animals (Farmer et
-al., 2017), and plant propagules (Hartzler and Anderson, 2016).
-Therefore, Palmer amaranth dispersal capacity make the species one of
-the most successful cases of weed adaption to cropping systems.
+biology. Farm mechanization, conservation agriculture (e.g., no-till),
+and reliance on herbicides for weed management are the main human
+mediated selection of Palmer amaranth to cropping systems. On the other
+hand, Palmer amaranth is a prolific seed producer with a C4
+photosynthetic apparatus (Ward et al., 2013). With a dioecy nature,
+Palmer amaranth male and female plants are obligate outcrosser species,
+increasing the chances of exchanging herbicide resistant alleles among
+plants (Oliveira et al., 2018). Also, Palmer amaranth small seed size (1
+mm) tend to thrive in no-tillage systems (Price et al., 2011), and
+spread across locations through farm equipment (Sauer, 1972), manure (Yu
+et al., 2021), animals (Farmer et al., 2017), and plant propagules
+(Hartzler and Anderson, 2016). Therefore, Palmer amaranth dispersal
+capacity make the species one of the most successful cases of weed
+adaption to cropping systems.
 
@@
 amaranth is triggered by 18 C soil temperature (Keeley et al., 1987),
-and optimal germination and biomass production were at 35/30 C day and
+and optimal germination and biomass production occur at 35/30 C day and
 night temperatures (Guo and Al-Khatib, 2003). Water has not shown to
@@
 impact agriculture and promote niches for Palmer amaranth
-invasion/adaptation into new environments. Currently, the greatest
-climatic risk of Palmer amaranth establishment is agronomic crops in
-Australia and Sub-Sahara Africa (Kistner and Hatfield, 2018).
-Temperature is a key factor limiting Palmer amaranth northward range
-expansion (Briscoe Runquist et al., 2019); however, under future climate
-change Palmer amaranth is likely to expand northward into Canada and
-Northern Europe (Kistner and Hatfield, 2018).
+invasion/adaptation into new environments. Currently, it is estimated
+that the greatest climatic risk of Palmer amaranth establishment is
+agronomic crops in Australia and Sub-Sahara Africa (Kistner and
+Hatfield, 2018). Temperature is a key factor limiting Palmer amaranth
+expansion to cooler geographies (Briscoe Runquist et al., 2019);
+however, under future climate change Palmer amaranth is likely to expand
+northward into Canada and Northern Europe (Kistner and Hatfield, 2018).
 
@@
 strategies on Palmer amaranth management should encompass the
-agroecosystem level but not attempts to eradicate the weed. The
-continuous Palmer amaranth dispersal and potential establishment into
-northern US/Canada warrant investigation on species growing morphology
-in such environments.
+agroecosystem level but not attempts to eradicate the weed. Most tactics
+to manage Palmer amaranth are based technology fixes (Scott, 2011),
+which are short-term (e.g., herbicide and/or tillage) rather than
+long-term Palmer amaranth management.
 
-Most Palmer amaranth studies are based on reactive (e,g, herbicide and
-tillage) rather than proactive management. Understanding Palmer amaranth
-biology and growing strategies under different environments can enhance
-our knowledge on species adaptation. Also, it can aid on designing
-proactive and ecological tactics to limit Palmer amaranth range
-expansion and reduce its negative impact while conserving diversity.
-Therefore, the objective of this study was to investigate the flowering
-pattern, biomass production, and height of Palmer amaranth growing under
-different environments and timings across five locations.
+The continuous Palmer amaranth dispersal and potential establishment
+into northern United States warrant investigations on species morphology
+growing in such environments. Understanding Palmer amaranth biology and
+growing strategies under different agroecossystems can enhance our
+knowledge on species adaptation. Also, it can aid on designing proactive
+and ecological tactics to limit the species range expansion, reduce its
+negative impact, and design resilient and sustainable farming systems
+(MacLaren et al., 2020). Therefore, the objective of this study was to
+investigate the flowering pattern, biomass production, and height of
+Palmer amaranth growing under in corn, soybean and fallow at two timings
+across five locations in the mid/upper United States Midwest.
 
@@
 
+\leavevmode\hypertarget{ref-maclaren2020}{}%
+MacLaren, C., Storkey, J., Menegat, A., Metcalfe, H., and
+Dehnen-Schmutz, K. (2020). An ecological future for weed science to
+sustain crop production and the environment. {A} review. \emph{Agron.
+Sustain. Dev.} 40, 24.
+doi:\href{https://doi.org/10.1007/s13593-020-00631-6}{10.1007/s13593-020-00631-6}.
+
 \leavevmode\hypertarget{ref-massinga2001}{}%
@@
 doi:\href{https://doi.org/10.3390/su13137003}{10.3390/su13137003}.
+
+\leavevmode\hypertarget{ref-morgan2001}{}%
+Morgan, G. D., Baumann, P. A., and Chandler, J. M. (2001). Competitive
+{Impact} of {Palmer Amaranth} ({Amaranthus} palmeri) on {Cotton}
+({Gossypium} hirsutum) {Development} and {Yield}. \emph{Weed Technology}
+15, 408--412.
+doi:\href{https://doi.org/10.1614/0890-037X(2001)015\%5B0408:CIOPAA\%5D2.0.CO;2}{10.1614/0890-037X(2001)015{[}0408:CIOPAA{]}2.0.CO;2}.
 
@@
 
+\leavevmode\hypertarget{ref-scott2011}{}%
+Scott, D. (2011). The {Technological Fix Criticisms} and the
+{Agricultural Biotechnology Debate}. \emph{J Agric Environ Ethics} 24,
+207--226.
+doi:\href{https://doi.org/10.1007/s10806-010-9253-7}{10.1007/s10806-010-9253-7}.
+
 \leavevmode\hypertarget{ref-ward2013}{}%

### 2021-07-26 Frontiers/Frontiers.Rmd
as

@@
 ---
-title: "Adaptation of Palmer amaranth to cropping systems"
+title: "Adaptation of Palmer amaranth (*Amaranthus palmeri*) to agroecossystems"
 runtitle: "Short Title"
@@
 The species has showed a remarkable capacity to evolve resistance to herbicides.
-Palmer amaranth has evolved resistance to eight herbicide sites of action, increasing the weed management complexity [@lindsay2017].
+Palmer amaranth has evolved resistance to eight herbicide sites of action [@heap2021], increasing the weed management complexity [@lindsay2017].
 Uncontrolled Palmer amaranth in competition for water, light and nutrients can drastically reduce crop yields [@berger2015].
-Palmer amaranth is documented with potential to reduce 91%, 68%, and 54% of corn [@massinga2001], soybean [@klingaman1994], and cotton [@morgan2001] yields.
+Palmer amaranth is documented with potential to reduce 91%, 68%, and 54% of corn [@massinga2001], soybean [@klingaman1994], and cotton [@morgan2001] yields, respectively.
 
@@
 With a dioecy nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging herbicide resistant alleles among plants [@oliveira2018].
-Also, Palmer amaranth small seed size (1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure [@yu2021], animals [@farmer2017], and plant propagules [@hartzler2016].
+Also, Palmer amaranth small seed size (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure [@hartzler2016], animals [@farmer2017], and plant propagules [@yu2021].
 Therefore, Palmer amaranth dispersal capacity make the species one of the most successful cases of weed adaption to cropping systems.
@@
 The continuous global temperature warming can impact agriculture and promote niches for Palmer amaranth invasion/adaptation into new environments.
-Currently, it is estimated that the greatest climatic risk of Palmer amaranth establishment is agronomic crops in Australia and Sub-Sahara Africa [@kistner2018].
-Temperature is a key factor limiting Palmer amaranth expansion to cooler geographies [@briscoerunquist2019]; however, under future climate change Palmer amaranth is likely to expand northward into Canada and Northern Europe [@kistner2018].
+Currently, it is estimated that the greatest climatic risk of Palmer amaranth establishment are agronomic crops in Australia and Sub-Sahara Africa [@kistner2018].
+Temperature is a key factor limiting Palmer amaranth expansion to cooler geographies [@briscoerunquist2019]; however, under future climate change Palmer amaranth is likely to expand northward into Canada and Northern Europe [@kistner2018; @briscoerunquist2019].
 
@@
 For example, herbicide resistant Palmer amaranth is widespread in Nebraska  [@oliveira2021a],  Michigan [@kohrt2017], and Connecticut [@aulakh2021].
-Successful cases of Palmer amaranth invasion and near to eradication is reported in Minnesota [@yu2021] and Iowa.
+Successful cases of Palmer amaranth invasion and near to eradication is reported in Minnesota [@yu2021].
 No Palmer amaranth actively growing was found in Canada; however, Palmer amaranth seeds was detected in sweet potato slips [@page2021].
-Nonetheless, it seems fated to manage Palmer amaranth in agronomic crops throughout multiple environments in the near future.
-Therefore, strategies on Palmer amaranth management should encompass the agroecosystem level but not attempts to eradicate the weed.
-Most tactics to manage Palmer amaranth are based technology fixes [@scott2011], which are short-term (e.g., herbicide and/or tillage) rather than long-term ecological management.
+Nonetheless, it seems fated the need to manage Palmer amaranth in agronomic crops throughout multiple environments in the near future.
+Therefore, strategies on Palmer amaranth management should encompass the agroecosystem level but not only attempts to eradicate the weed.
+Most tactics to manage Palmer amaranth are based on technology fixes [@scott2011], which are short-term (e.g., herbicide and/or tillage) rather than long-term weed management.
 
@@
 palmeri* accession (Per1) from Perkins County, Nebraska.
-Per1 accession collection is documented in [@oliveira2021a], with no reported herbicide resistance.
+Per1 accession collection is documented with no reported herbicide resistance [@oliveira2021a].
 Three weeks prior to the field experiment, seeds were planted in plastic trays containing potting-mix.
@@
 
-The experimental unit were adjacent 9.1 m wide (12 rows at 72.2 cm row spacing) by 10.7 m long.
-Each experimental unit was planted with corn or soybean, or left fallow.
-Palmer amaranth seedlings were  transplanted to the field experiment by making a whole in the soil (6 cm deep and 8 cm wide); and gently transferring in the ground (potting mix + two seedlings).
-After a week, if both plants were alive, one was eliminated.
-There were two transplant timing: early (June 1^st^) and late (July 1^st^).
-There were 24 Palmer amaranth plants in each crop/fallow and timing, with a total of 144 plants.
+The field experimental unit were six adjacent 9.1 m wide (12 rows at 72.2 cm row spacing) by 10.7 m long.
+Each experimental unit was planted with corn or soybean, or under fallow condition.
+Palmer amaranth seedlings were  transplanted to the field experiment by making a whole in the soil (6 cm deep and 8 cm wide); and gently transferring to the ground (potting mix + two seedlings).
+After a week, one was eliminated and one was kept.
+There were two transplant timing: first (June 1^st^) and second (July 1^st^).
+There were 24 Palmer amaranth plants in each experimental unit, with a total of 144 plants.
 The study was repeated twice.
@@
 
-The statistical analyses were performed using R statistical software version 4.0.1.
+The statistical analyses were performed using R statistical software version 4.0.1. Data across locations and year were combined.
 
@@
 The EDcomp function compares the ratio of cumulative flowering using t-statistics, where P-value < 0.05 indicates that we fail to reject the null hypothesis.
+
+Palmer amaranth height and biomass were performed with a linear mixed model using *lmer* function from "lme4" package [@bates2015].
+Plant height and biomass were transformed to meet model assumption of normality .
+In the model, agroecosystem (crop, soybean, fallow) was the fixed effect and year nested with location the random effects.
+Analysis of variance was performed with *anova* function from "car" package [@fox2018].
+Marginal means and compact letter display were estimated with *emmeans* and *cld* from packages "emmeans" and multcomp [@hothorn2008]
 

### 2021-07-26 Frontiers/Frontiers.tex
as

@@
 
-\title[Short Title]{Adaptation of Palmer amaranth to cropping systems}
+\title[Short Title]{Adaptation of Palmer amaranth (\emph{Amaranthus
+palmeri}) to agroecossystems}
 \author[\firstAuthorLast]{\Authors}
@@
 a remarkable capacity to evolve resistance to herbicides. Palmer
-amaranth has evolved resistance to eight herbicide sites of action,
-increasing the weed management complexity (Lindsay et al., 2017).
-Uncontrolled Palmer amaranth in competition for water, light and
+amaranth has evolved resistance to eight herbicide sites of action
+(Heap, 2021), increasing the weed management complexity (Lindsay et al.,
+2017). Uncontrolled Palmer amaranth in competition for water, light and
 nutrients can drastically reduce crop yields (Berger et al., 2015).
@@
 54\% of corn (Massinga et al., 2001), soybean (Klingaman and Oliver,
-1994), and cotton (Morgan et al., 2001) yields.
+1994), and cotton (Morgan et al., 2001) yields, respectively.
 
@@
 increasing the chances of exchanging herbicide resistant alleles among
-plants (Oliveira et al., 2018). Also, Palmer amaranth small seed size (1
-mm) tend to thrive in no-tillage systems (Price et al., 2011), and
-spread across locations through farm equipment (Sauer, 1972), manure (Yu
-et al., 2021), animals (Farmer et al., 2017), and plant propagules
-(Hartzler and Anderson, 2016). Therefore, Palmer amaranth dispersal
+plants (Oliveira et al., 2018). Also, Palmer amaranth small seed size
+(e.g, 1 mm) tend to thrive in no-tillage systems (Price et al., 2011),
+and spread across locations through farm equipment (Sauer, 1972), manure
+(Hartzler and Anderson, 2016), animals (Farmer et al., 2017), and plant
+propagules (Yu et al., 2021). Therefore, Palmer amaranth dispersal
 capacity make the species one of the most successful cases of weed
@@
 however, under future climate change Palmer amaranth is likely to expand
-northward into Canada and Northern Europe (Kistner and Hatfield, 2018).
+northward into Canada and Northern Europe (Kistner and Hatfield, 2018;
+Briscoe Runquist et al., 2019).
 
-Palmer amaranth is already found in agronomic crops of South America
-(Larran et al., 2017; Küpper et al., 2017) and Southern Europe (Milani
-et al., 2021). In the US, Palmer amaranth is commonly found at crop
+Palmer amaranth is already documented in agronomic crops of South
+America (Larran et al., 2017; Küpper et al., 2017) and Southern Europe
+(Milani et al., 2021). In the US, Palmer amaranth is established at crop
 (Garetson et al., 2019) and non-crop land (Bagavathiannan and
@@
 2021). Successful cases of Palmer amaranth invasion and near to
-eradication is well documented in Minnesota (Yu et al., 2021) and Iowa.
-No Palmer amaranth actively growing was found in Canada; however, Palmer
-amaranth seeds was detected in sweet potato slips (Page et al., 2021).
+eradication is reported in Minnesota (Yu et al., 2021). No Palmer
+amaranth actively growing was found in Canada; however, Palmer amaranth
+seeds was detected in sweet potato slips (Page et al., 2021).
 Nonetheless, it seems fated to manage Palmer amaranth in agronomic crops
@@
 which are short-term (e.g., herbicide and/or tillage) rather than
-long-term Palmer amaranth management.
+long-term ecological management.
 
@@
 growing strategies under different agroecossystems can enhance our
-knowledge on species adaptation. Also, it can aid on designing proactive
+knowledge on species adaptation. It can also aid on designing proactive
 and ecological tactics to limit the species range expansion, reduce its
@@
 The study was performed with a \emph{A. palmeri} accession (Per1) from
-Perkins County, Nebraska. Per1 accession collection is documented in
-(Oliveira et al., 2021), with no reported herbicide resistance. Three
-weeks prior to the field experiment, seeds were planted in plastic trays
+Perkins County, Nebraska. Per1 accession collection is documented with
+no reported herbicide resistance (Oliveira et al., 2021). Three weeks
+prior to the field experiment, seeds were planted in plastic trays
 containing potting-mix. Emerged seedlings (1 cm) were transplanted into
@@
 
-The experimental unit were adjacent 9.1 m wide (12 rows at 72.2 cm row
-spacing) by 10.7 m long. Each experimental unit was planted with corn or
-soybean, or left fallow. Palmer amaranth seedlings were transplanted to
-the field experiment by making a whole in the soil (6 cm deep and 8 cm
-wide); and gently transferring in the ground (potting mix + two
-seedlings). After a week, if both plants were alive, one was eliminated.
-There were two transplant timing: early (June 1\textsuperscript{st}) and
-late (July 1\textsuperscript{st}). There were 24 Palmer amaranth plants
-in each crop/fallow and timing, with a total of 144 plants. The study
-was repeated twice.
+The field experimental unit were six adjacent 9.1 m wide (12 rows at
+72.2 cm row spacing) by 10.7 m long. Each experimental unit was planted
+with corn or soybean, or under fallow condition. Palmer amaranth
+seedlings were transplanted to the field experiment by making a whole in
+the soil (6 cm deep and 8 cm wide); and gently transferring to the
+ground (potting mix + two seedlings). After a week, one was eliminated
+and one was kept. There were two transplant timing: first (June
+1\textsuperscript{st}) and second (July 1\textsuperscript{st}). There
+were 24 Palmer amaranth plants in each experimental unit, with a total
+of 144 plants. The study was repeated twice.
 
@@
 The statistical analyses were performed using R statistical software
-version 4.0.1.
+version 4.0.1. Data across locations and year were combined.
 
@@
 reject the null hypothesis.
+
+Palmer amaranth height and biomass were performed with a linear mixed
+model using \emph{lmer} function from ``lme4'' package (Bates et al.,
+2015). Plant height and biomass were transformed to meet model
+assumption of normality . In the model, agroecosystem (crop, soybean,
+fallow) was the fixed effect and year nested with location the random
+effects. Analysis of variance was performed with \emph{anova} function
+from ``car'' package (Fox and Weisberg, 2018). Marginal means and
+compact letter display were estimated with \emph{emmeans} and \emph{cld}
+from packages ``emmeans'' and multcomp (Hothorn et al., 2008)
 
@@
 
+\leavevmode\hypertarget{ref-bates2015}{}%
+Bates, D., Mächler, M., Bolker, B., and Walker, S. (2015). Fitting
+{Linear Mixed}-{Effects Models Using} Lme4. \emph{Journal of Statistical
+Software} 67, 1--48.
+doi:\href{https://doi.org/10.18637/jss.v067.i01}{10.18637/jss.v067.i01}.
+
 \leavevmode\hypertarget{ref-berger2015}{}%
@@
 
+\leavevmode\hypertarget{ref-fox2018}{}%
+Fox, J., and Weisberg, S. (2018). \emph{An {R Companion} to {Applied
+Regression}}. {SAGE Publications} Available at:
+\url{http://books.google.com?id=uPNrDwAAQBAJ}.
+
 \leavevmode\hypertarget{ref-garetson2019}{}%
@@
 what? 10.
+
+\leavevmode\hypertarget{ref-heap2021}{}%
+Heap, I. (2021). Internation {Herbicide}-{Resistant Weed Database}.
+Available at: \url{http://www.weedscience.org/Home.aspx} {[}Accessed
+July 26, 2021{]}.
+
+\leavevmode\hypertarget{ref-hothorn2008}{}%
+Hothorn, T., Bretz, F., and Westfall, P. (2008). Simultaneous
+{Inference} in {General Parametric Models}. \emph{Biometrical Journal}
+50, 346--363.
+doi:\href{https://doi.org/10.1002/bimj.200810425}{10.1002/bimj.200810425}.
 

### 2021-07-26 Frontiers/Frontiers.Rmd
as

@@
 ---
-title: "Adaptation of Palmer amaranth (*Amaranthus palmeri*) to agroecossystems"
-runtitle: "Short Title"
+title: "Palmer amaranth (*Amaranthus palmeri*) adaptation to agroecossystems"
+runtitle: "Palmer amaranth adaptation"
 documentclass: frontiersSCNS  # or frontiersHLTH, or frontiersFPHY
@@
 \tiny
- \keyFont{ \section{Keywords:} Text Text Text Text Text Text Evolution Weed }
+ \keyFont{ \section{Keywords:} Evolution Flowering Management Pigweed Weed}
 
@@
 
+```{r fig1, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../data analysis/weather/Figure 1.png")
+```
+
 The field experimental unit were six adjacent 9.1 m wide (12 rows at 72.2 cm row spacing) by 10.7 m long.
@@
 
-## Subsection 1 {-}
+## Cumulative flowering {-}
 
 You can use `R` chunks directly to plot graphs.
-
-```{r fig1, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/weather/Figure 1.png")
-```
 
@@
 
-## Subsection 2 {-}
+## Palmer amaranth height and biomass {-}
 
@@
 
-MCO design, wrote,
-
-<!--
-
-When determining authorship the following criteria should be observed:
-
- - Substantial contributions to the conception or design of the work; or the
-   acquisition, analysis, or interpretation of data for the work; AND
- - Drafting the work or revising it critically for important intellectual
-   content; AND
- - Final approval of the version to be published ; AND
- - Agreement to be accountable for all aspects of the work in ensuring that
-   questions related to the accuracy or integrity of any part of the work are
-   appropriately investigated and resolved.
-
-Contributors who meet fewer than all 4 of the above criteria for authorship
-should not be listed as authors, but they should be acknowledged.
-(http://www.icmje.org/roles_a.html)
-
--->
-
-The statement about the authors and contributors can be up to several sentences
-long, describing the tasks of individual authors referred to by their initials
-and should be included at the end of the manuscript before the References
-section.
+RW: designed the experiments; AJ, CP, MB, MO, and SS: conducted the experiments; MO: analyzed the data and wrote the manuscript; AJ, CP, MB, MO, SS, and RW: conceptualized the research.
+All authors reviewed the manuscript.
 
@@
 
-Funding:
+Funding: This work received no specific grant from any funding agency, commercial, or not-for-profit sectors
 
@@
 
-```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Figure caption', out.width = "150mm", out.height = "260mm"}
+```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Figure caption', out.width = "150mm", out.height = "200mm"}
 # You can also refer to code chunks from above to place figures at the bottom.

### 2021-07-26 Frontiers/Frontiers.tex
as

@@
 
-\title[Short Title]{Adaptation of Palmer amaranth (\emph{Amaranthus
-palmeri}) to agroecossystems}
+\title[Palmer amaranth adaptation]{Palmer amaranth (\emph{Amaranthus
+palmeri}) adaptation to agroecossystems}
 \author[\firstAuthorLast]{\Authors}
@@
 invasion/adaptation into new environments. Currently, it is estimated
-that the greatest climatic risk of Palmer amaranth establishment is
+that the greatest climatic risk of Palmer amaranth establishment are
 agronomic crops in Australia and Sub-Sahara Africa (Kistner and
@@
 seeds was detected in sweet potato slips (Page et al., 2021).
-Nonetheless, it seems fated to manage Palmer amaranth in agronomic crops
-throughout multiple environments in the near future. Therefore,
-strategies on Palmer amaranth management should encompass the
-agroecosystem level but not attempts to eradicate the weed. Most tactics
-to manage Palmer amaranth are based technology fixes (Scott, 2011),
-which are short-term (e.g., herbicide and/or tillage) rather than
-long-term ecological management.
+Nonetheless, it seems fated the need to manage Palmer amaranth in
+agronomic crops throughout multiple environments in the near future.
+Therefore, strategies on Palmer amaranth management should encompass the
+agroecosystem level but not only attempts to eradicate the weed. Most
+tactics to manage Palmer amaranth are based on technology fixes (Scott,
+2011), which are short-term (e.g., herbicide and/or tillage) rather than
+long-term weed management.
 
@@
 
-\hypertarget{subsection-1}{%
-\subsection*{Subsection 1}\label{subsection-1}}
-\addcontentsline{toc}{subsection}{Subsection 1}
+\hypertarget{cumulative-flowering}{%
+\subsection*{Cumulative flowering}\label{cumulative-flowering}}
+\addcontentsline{toc}{subsection}{Cumulative flowering}
 
@@
 
-\hypertarget{subsection-2}{%
-\subsection*{Subsection 2}\label{subsection-2}}
-\addcontentsline{toc}{subsection}{Subsection 2}
+\hypertarget{palmer-amaranth-height-and-biomass}{%
+\subsection*{Palmer amaranth height and
+biomass}\label{palmer-amaranth-height-and-biomass}}
+\addcontentsline{toc}{subsection}{Palmer amaranth height and biomass}
 
@@
 
-MCO design, wrote,
-
-The statement about the authors and contributors can be up to several
-sentences long, describing the tasks of individual authors referred to
-by their initials and should be included at the end of the manuscript
-before the References section.
+RW: designed the experiments; AJ, CP, MB, MO, and SS: conducted the
+experiments; MO: analyzed the data and wrote the manuscript; AJ, CP, MB,
+MO, SS, and RW: conceptualized the research. All authors reviewed the
+manuscript.
 
@@
 
-Funding:
+Funding: This work received no specific grant from any funding agency,
+commercial, or not-for-profit sectors
 
@@
 
-{\centering \includegraphics[width=150mm,height=260mm]{../data analysis/figures/combine 3}
+{\centering \includegraphics[width=150mm,height=200mm]{../data analysis/figures/combine 3}
 

### 2021-07-27 Manuscript/Manuscript.Rmd
as

--- 2021-07-26 Frontiers/Frontiers.Rmd
@@
 The field experimental unit were six adjacent 9.1 m wide (12 rows at 72.2 cm row spacing) by 10.7 m long.
-Each experimental unit was planted with corn or soybean, or under fallow condition.
+Each experimental unit was planted with corn or soybean (DSR-1950 R2Y at 296,400 seeds ha 1), or under fallow condition.
 Palmer amaranth seedlings were  transplanted to the field experiment by making a whole in the soil (6 cm deep and 8 cm wide); and gently transferring to the ground (potting mix + two seedlings).
 After a week, one was eliminated and one was kept.
-There were two transplant timing: first (June 1^st^) and second (July 1^st^).
+There were two transplant timing: first (June 1) and second (July 1).
 There were 24 Palmer amaranth plants in each experimental unit, with a total of 144 plants.
@@
 When a plant started flowering, the day was recorded, plant sex was identified as male or female, and plant height was measured from soil surface to the plant top.
-Then, aboveground plant biomass was harvest near soil surface and oven dried at 65 C until reaching constant weight before the weight of biomass (g plant^-1^) was recorded.
+Then, aboveground plant biomass was harvest near soil surface and oven dried at 65 C until reaching constant weight before the weight of biomass (g plant 1) was recorded.
 
@@
 The doy for 10, 50, and 90% Palmer amaranth cumulative flowering were determined using the *ED* function of drc package.
-Also, the 10, 50, and 90% Palmer amaranth cumulative flowering were compared among crop/fallow and timings using the *EDcomp* function of drc package.
+Also, the 10, 50, and 90% Palmer amaranth cumulative flowering were compared among agroecossystems and timings using the *EDcomp* function of drc package.
 The EDcomp function compares the ratio of cumulative flowering using t-statistics, where P-value < 0.05 indicates that we fail to reject the null hypothesis.
@@
 Palmer amaranth height and biomass were performed with a linear mixed model using *lmer* function from "lme4" package [@bates2015].
-Plant height and biomass were transformed to meet model assumption of normality .
+Plant height and biomass were transformed to meet model assumption of normality.
 In the model, agroecosystem (crop, soybean, fallow) was the fixed effect and year nested with location the random effects.
@@
 ```{r fig2, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/combine 3.png")
+knitr::include_graphics("../data analysis/figures/Combined2.png")
 ```
 
-## Palmer amaranth height and biomass {-}
+## Height and biomass {-}
+
+```{r fig3, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../data analysis/figures/figure 3.png")
+```
+
+```{r fig4, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../data analysis/figures/image.jpeg")
+```
 
@@
 
-```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Figure caption', out.width = "150mm", out.height = "200mm"}
+```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "160mm", out.height = "130mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
 ```
+
+```{r, Figure-3, ref.label = "fig3", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Palmer amaranth biomass (A) and height (B) growing in corn, fallow, and soybean across Arlington, Clay Center, Grant, Lincoln, and Macomb', out.width = "130mm", out.height = "70mm"}
+# You can also refer to code chunks from above to place figures at the bottom.
+```
+
+```{r, Figure-4, ref.label = "fig4", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Harvest Palmer amaranth plants at 40 days after first transplant timing.
+From left to right, Palmer amaranth growing in fallow, soybean and corn in Arlington, Wisconsin', out.width = "70mm", out.height = "70mm"}
+# You can also refer to code chunks from above to place figures at the bottom.
+```

### 2021-07-27 Manuscript/Manuscript.tex
as

--- 2021-07-26 Frontiers/Frontiers.tex
@@
 72.2 cm row spacing) by 10.7 m long. Each experimental unit was planted
-with corn or soybean, or under fallow condition. Palmer amaranth
-seedlings were transplanted to the field experiment by making a whole in
-the soil (6 cm deep and 8 cm wide); and gently transferring to the
-ground (potting mix + two seedlings). After a week, one was eliminated
-and one was kept. There were two transplant timing: first (June
-1\textsuperscript{st}) and second (July 1\textsuperscript{st}). There
+with corn or soybean (DSR-1950 R2Y at 296,400 seeds ha 1), or under
+fallow condition. Palmer amaranth seedlings were transplanted to the
+field experiment by making a whole in the soil (6 cm deep and 8 cm
+wide); and gently transferring to the ground (potting mix + two
+seedlings). After a week, one was eliminated and one was kept. There
+were two transplant timing: first (June 1) and second (July 1). There
 were 24 Palmer amaranth plants in each experimental unit, with a total
@@
 biomass was harvest near soil surface and oven dried at 65 C until
-reaching constant weight before the weight of biomass (g
-plant\textsuperscript{-1}) was recorded.
+reaching constant weight before the weight of biomass (g plant 1) was
+recorded.
 
@@
 
-\hypertarget{palmer-amaranth-height-and-biomass}{%
-\subsection*{Palmer amaranth height and
-biomass}\label{palmer-amaranth-height-and-biomass}}
-\addcontentsline{toc}{subsection}{Palmer amaranth height and biomass}
+\hypertarget{height-and-biomass}{%
+\subsection*{Height and biomass}\label{height-and-biomass}}
+\addcontentsline{toc}{subsection}{Height and biomass}
 
@@
 
-{\centering \includegraphics[width=150mm,height=200mm]{../data analysis/figures/combine 3}
+{\centering \includegraphics[width=160mm,height=130mm]{../data analysis/figures/Combined2}
 
@@
 
-\caption{Figure caption}\label{fig:Figure-2}
+\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)}\label{fig:Figure-2}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=130mm,height=70mm]{../data analysis/figures/figure 3}
+
+}
+
+\caption{Palmer amaranth biomass (A) and height (B) growing in corn, fallow, and soybean across Arlington, Clay Center, Grant, Lincoln, and Macomb}\label{fig:Figure-3}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=70mm,height=70mm]{../data analysis/figures/image}
+
+}
+
+\caption{Harvest Palmer amaranth plants at 40 days after first transplant timing. From left to right, Palmer amaranth growing in fallow, soybean and corn in Arlington, Wisconsin}\label{fig:Figure-4}
 \end{figure}

### 2021-07-29 Manuscript/Manuscript.Rmd
jhj

@@
 Palmer amaranth is documented with potential to reduce 91%, 68%, and 54% of corn [@massinga2001], soybean [@klingaman1994], and cotton [@morgan2001] yields, respectively.
+Thus, Palmer amaranth poses a risk to sustainable agriculture.
 
@@
 Palmer amaranth weediness is likely a result of human-assisted selection in combination with species biology.
-Farm mechanization, conservation agriculture (e.g., no-till), and reliance on herbicides for weed management are the main human mediated selection of Palmer amaranth to cropping systems.
+Farm mechanization, conservation agriculture (e.g., no-till), and reliance on herbicides for weed management are the main human-mediated selection of Palmer amaranth into cropping systems.
 On the other hand, Palmer amaranth is a prolific seed producer with a C4 photosynthetic apparatus [@ward2013].
-With a dioecy nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging herbicide resistant alleles among plants [@oliveira2018].
+With a dioecy nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging adaptive traits among plants [@oliveira2018].
 Also, Palmer amaranth small seed size (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure [@hartzler2016], animals [@farmer2017], and plant propagules [@yu2021].
@@
 Palmer amaranth seedlings were supplied with adequate water and kept under greenhouse conditions at Arlington, Clay Center, Lincoln, and Macomb; and kept outdoors in Grant.
-Palmer amaranth seedlings were kept under greenhouse/outdoors until the onset of the experiment (7 to 10 cm height).
+Palmer amaranth seedlings were kept under greenhouse/outdoors until the onset of the experiment (2-3 leaf stage/5 to 8 cm height).
 
@@
 The experiment was conducted in 2018 and 2019 under field conditions at five locations: Arlington (Washington County, Wisconsin), Clay Center (Clay County, Nebraska), Grant (Perkins County, Nebraska), Lincoln (Lancaster County, Nebraska), and Macomb (McDonough County, Illinois).
+
+A glyphosate-resistant soybean cultivar (DSR-1950 R2Y at 296,400 seeds ha 1), and a corn hybrid were planted at
+
+Monthly mean air temperature and sum precipitation were collected from June through September across the five locations in 2018 and 2019 (Figure 1)
 
@@
 Each experimental unit was planted with corn or soybean (DSR-1950 R2Y at 296,400 seeds ha 1), or under fallow condition.
-Palmer amaranth seedlings were  transplanted to the field experiment by making a whole in the soil (6 cm deep and 8 cm wide); and gently transferring to the ground (potting mix + two seedlings).
+Palmer amaranth seedlings (potting mix + two seedlings) were and gently transferring to the ground (6 cm deep and 8 cm wide).
+Twenty-four plants were equidistantly placed (0.76 m apart) between rows within each agroecossystems.
 After a week, one was eliminated and one was kept.
 There were two transplant timing: first (June 1) and second (July 1).
-There were 24 Palmer amaranth plants in each experimental unit, with a total of 144 plants.
+There were 24 Palmer amaranth plants in each experimental unit, with a total of 144 plants for each location.
 The study was repeated twice.
@@
 Analysis of variance was performed with *anova* function from "car" package [@fox2018].
-Marginal means and compact letter display were estimated with *emmeans* and *cld* from packages "emmeans" and multcomp [@hothorn2008]
+Marginal means and compact letter display were estimated with *emmeans* and *cld* from packages "emmeans" and multcomp [@hothorn2008].
 
@@
 
-You can use `R` chunks directly to plot graphs.
+Palmer amaranth cumulative flowering from first transplant timing ranged from late June to early September (Figure 2).
+Palmer amaranth growing in corn resulted in a longer flowering pattern compared to fallow and soybean.
+Palmer amaranth reached 10% flowering in soybean at doy 180, which was slightly different from fallow (doy 180.9; *P* = 0.01) and corn (doy 181.7; *P* = 0.00).
+Nonetheless, the 10% cumulative Palmer amaranth flowering in soybean, fallow and corn occurred at the end of June.
+The 50% Palmer amaranth cumulative flowering occurred in July.
+For example, Palmer amaranth reached 50% in fallow at doy 193.4, followed by soybean (doy 194.8), corn (doy 206.6).
+Similar trend was observed at 90% Palmer amaranth cumulative flowering.
+Palmer amaranth growing in corn reached 90% flowering at doy 252.6 (early September), which was 37.8 and 32.2 days after Palmer amaranth 90% flowering in fallow and soybean, respectively.
 
@@
 
+Palmer amaranth cumulative flowering from second transplanting timing ranged from mid July to mid September (Figure 2).
+All cumulative flowering comparisons among agroecossystems were significant (*P* < 0.00) Palmer amaranth growing in fallow resulted in earlier flowering time compared to soybean and corn.
+Palmer amaranth growing in fallow reached 10%, 50%, and 90% flowering time at day 203.8, 214.4, and 232.2, respectively.
+Palmer amaranth growing in soybean reached 10% flowering at doy 210.9, which was 6 days prior to corn (*P*-value = 0.00).
+Similar trend was observed at 50% flowering, whereas Palmer amaranth reached 50% flowering in corn (doy 233.0) 4 days after soybeans (doy 228.9; *P* = 0.00).
+The 90% Palmer amaranth cumulative flowering occurred at same day in corn (260.9) and soybean (260.5; *P* = 0.66).
+
+```{r eval = FALSE, include = FALSE}
+options(scipen = 7)
+EDcomp(model4, c(90, 90))
+
+result3 %>% 
+  filter(ed == "90" & month == "Second transplanting")
+```
+
 ## Height and biomass {-}
+
+Palmer amaranth accumulated more biomass when growing in fallow compared to Palmer amaranth growing in soybean and corn.
+At first transplanting time, Palmer amaranth biomass was 78.3 g plant-1, 26.4 g plant-1 and 14.9 g plant-1 in fallow, soybean and corn, respectively.
+Palmer amaranth produce less biomass at second transplanting time.
+For example, Palmer amaranth growing in fallow resulted in 49.2 g plant-1, 3.4 g plant-1 in soybean and 1.4 g plant when growing in corn.
+
+```{r eval = FALSE, include = FALSE}
+multcomp::cld(analysis$emmeans[[2]], 
+              alpha=0.05, Letters=letters, 
+              adjust="none", reversed = TRUE)
+```
 
@@
 
-Frontiers requires figures to be submitted individually, in the same order as
-they are referred to in the manuscript. Figures will then be automatically
-embedded at the bottom of the submitted manuscript. Kindly ensure that each
-table and figure is mentioned in the text and in numerical order. Permission
-must be obtained for use of copyrighted material from other sources (including
-the web). Please note that it is compulsory to follow figure instructions.
-Figures which are not according to the guidelines will cause substantial delay
-during the production process.
-
-# Discussion
+# Discussion {-}
 

### 2021-07-29 Manuscript/Manuscript.tex
jhj

@@
 biology. Farm mechanization, conservation agriculture (e.g., no-till),
-and reliance on herbicides for weed management are the main human
-mediated selection of Palmer amaranth to cropping systems. On the other
-hand, Palmer amaranth is a prolific seed producer with a C4
+and reliance on herbicides for weed management are the main
+human-mediated selection of Palmer amaranth into cropping systems. On
+the other hand, Palmer amaranth is a prolific seed producer with a C4
 photosynthetic apparatus (Ward et al., 2013). With a dioecy nature,
 Palmer amaranth male and female plants are obligate outcrosser species,
-increasing the chances of exchanging herbicide resistant alleles among
-plants (Oliveira et al., 2018). Also, Palmer amaranth small seed size
-(e.g, 1 mm) tend to thrive in no-tillage systems (Price et al., 2011),
-and spread across locations through farm equipment (Sauer, 1972), manure
+increasing the chances of exchanging adaptive traits among plants
+(Oliveira et al., 2018). Also, Palmer amaranth small seed size (e.g, 1
+mm) tend to thrive in no-tillage systems (Price et al., 2011), and
+spread across locations through farm equipment (Sauer, 1972), manure
 (Hartzler and Anderson, 2016), animals (Farmer et al., 2017), and plant
@@
 Palmer amaranth seedlings were kept under greenhouse/outdoors until the
-onset of the experiment (7 to 10 cm height).
+onset of the experiment (2-3 leaf stage/5 to 8 cm height).
 
@@
 
+A glyphosate-resistant soybean cultivar (DSR-1950 R2Y at 296,400 seeds
+ha 1), and a corn hybrid were planted at
+
 The field experimental unit were six adjacent 9.1 m wide (12 rows at
@@
 with corn or soybean (DSR-1950 R2Y at 296,400 seeds ha 1), or under
-fallow condition. Palmer amaranth seedlings were transplanted to the
-field experiment by making a whole in the soil (6 cm deep and 8 cm
-wide); and gently transferring to the ground (potting mix + two
-seedlings). After a week, one was eliminated and one was kept. There
-were two transplant timing: first (June 1) and second (July 1). There
-were 24 Palmer amaranth plants in each experimental unit, with a total
-of 144 plants. The study was repeated twice.
+fallow condition. Palmer amaranth seedlings (potting mix + two
+seedlings) were and gently transferring to the ground (6 cm deep and 8
+cm wide). Twenty-four plants were equidistantly placed (0.76 m apart)
+between rows within each agroecossystems. After a week, one was
+eliminated and one was kept. There were two transplant timing: first
+(June 1) and second (July 1). There were 24 Palmer amaranth plants in
+each experimental unit, with a total of 144 plants for each location.
+The study was repeated twice.
 
@@
 50, and 90\% Palmer amaranth cumulative flowering were compared among
-crop/fallow and timings using the \emph{EDcomp} function of drc package.
-The EDcomp function compares the ratio of cumulative flowering using
-t-statistics, where P-value \textless{} 0.05 indicates that we fail to
-reject the null hypothesis.
+agroecossystems and timings using the \emph{EDcomp} function of drc
+package. The EDcomp function compares the ratio of cumulative flowering
+using t-statistics, where P-value \textless{} 0.05 indicates that we
+fail to reject the null hypothesis.
 
@@
 2015). Plant height and biomass were transformed to meet model
-assumption of normality . In the model, agroecosystem (crop, soybean,
+assumption of normality. In the model, agroecosystem (crop, soybean,
 fallow) was the fixed effect and year nested with location the random
@@
 compact letter display were estimated with \emph{emmeans} and \emph{cld}
-from packages ``emmeans'' and multcomp (Hothorn et al., 2008)
+from packages ``emmeans'' and multcomp (Hothorn et al., 2008).
 
@@
 
-You can use \texttt{R} chunks directly to plot graphs.
+Palmer amaranth cumulative flowering from first transplant timing ranged
+from late June to early September. Palmer amaranth growing in corn
+resulted in a longer flowering pattern compared to fallow and soybean.
+Palmer amaranth reached 10\% flowering in soybean at doy 180, which was
+slightly different from fallow (doy 180.9; \emph{P} = 0.01) and corn
+(doy 181.7; \emph{P} = 0.00). Nonetheless, the 10\% cumulative Palmer
+amaranth flowering in soybean, fallow and corn occurred at the end of
+June. The 50\% Palmer amaranth cumulative flowering occurred in July.
+For example, Palmer amaranth reached 50\% in fallow at doy 193.4,
+followed by soybean (doy 194.8), corn (doy 206.6). Similar trend was
+observed at 90\% Palmer amaranth cumulative flowering. Palmer amaranth
+growing in corn reached 90\% flowering at doy 252.6 (early September),
+which was 37.8 and 32.2 days after Palmer amaranth 90\% flowering in
+fallow and soybean, respectively.
+
+Palmer amaranth cumulative flowering from second transplanting timing
+ranged from mid July to mid September. All cumulative flowering
+comparisons among agroecossystems were significant (\emph{P} \textless{}
+0.00) Palmer amaranth growing in fallow resulted in earlier flowering
+time compared to soybean and corn. Palmer amaranth growing in fallow
+reached 10\%, 50\%, and 90\% flowering time at day 203.8, 214.4, and
+232.2, respectively. Palmer amaranth growing in soybean reached 10\%
+flowering at doy 210.9, which was 6 days prior to corn (\emph{P}-value =
+0.00). Similar trend was observed at 50\% flowering, whereas Palmer
+amaranth reached 50\% flowering in corn (doy 233.0) 4 days after
+soybeans (doy 228.9; \emph{P} = 0.00). The 90\% Palmer amaranth
+cumulative flowering occurred at same day in corn (260.9) and soybean
+(260.5; \emph{P} = 0.66).
 
@@
 
-Frontiers requires figures to be submitted individually, in the same
-order as they are referred to in the manuscript. Figures will then be
-automatically embedded at the bottom of the submitted manuscript. Kindly
-ensure that each table and figure is mentioned in the text and in
-numerical order. Permission must be obtained for use of copyrighted
-material from other sources (including the web). Please note that it is
-compulsory to follow figure instructions. Figures which are not
-according to the guidelines will cause substantial delay during the
-production process.
+Palmer amaranth accumulated more biomass when growing in fallow compared
+to Palmer amaranth growing in soybean and corn. At first transplanting
+time, Palmer amaranth biomass was 78.3 g plant-1, 26.4 g plant-1 and
+14.9 g plant-1 in fallow, soybean and corn, respectively. Palmer
+amaranth produce less biomass at second transplanting time. For example,
+Palmer amaranth growing in fallow resulted in 49.2 g plant-1, 3.4 g
+plant-1 in soybean and 1.4 g plant when growing in corn.
 
 \hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\section*{Discussion}\label{discussion}}
+\addcontentsline{toc}{section}{Discussion}
 

### 2021-07-29 Manuscript/Manuscript.Rmd
gfg

@@
 
-Monthly mean air temperature and sum precipitation were collected from June through September across the five locations in 2018 and 2019 (Figure 1)
+Monthly mean air temperature and sum precipitation were obtained using Daymet weather data from June through September across the five locations in 2018 and 2019 [@correndo2021] (Figure 1)
 

### 2021-07-30 Manuscript/Manuscript.Rmd
,M,

@@
 ```{r fig2, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/Combined2.png")
+knitr::include_graphics("../data analysis/figures/Figure 2.png")
 ```
@@
 ```{r fig3, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/figure 3.png")
+knitr::include_graphics("../data analysis/figures/Figure 3.png")
 ```
@@
 
-```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "160mm", out.height = "130mm"}
+```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "130mm", out.height = "130mm"}
 # You can also refer to code chunks from above to place figures at the bottom.

### 2021-07-30 Manuscript/Manuscript.tex
,M,

@@
 54\% of corn (Massinga et al., 2001), soybean (Klingaman and Oliver,
-1994), and cotton (Morgan et al., 2001) yields, respectively.
+1994), and cotton (Morgan et al., 2001) yields, respectively. Thus,
+Palmer amaranth poses a risk to sustainable agriculture.
 
@@
 
+Monthly mean air temperature and sum precipitation were obtained using
+Daymet weather data from June through September across the five
+locations in 2018 and 2019 (\textbf{correndo2021?}) (Figure 1)
+
 The field experimental unit were six adjacent 9.1 m wide (12 rows at
@@
 Palmer amaranth cumulative flowering from first transplant timing ranged
-from late June to early September. Palmer amaranth growing in corn
-resulted in a longer flowering pattern compared to fallow and soybean.
-Palmer amaranth reached 10\% flowering in soybean at doy 180, which was
-slightly different from fallow (doy 180.9; \emph{P} = 0.01) and corn
-(doy 181.7; \emph{P} = 0.00). Nonetheless, the 10\% cumulative Palmer
-amaranth flowering in soybean, fallow and corn occurred at the end of
-June. The 50\% Palmer amaranth cumulative flowering occurred in July.
-For example, Palmer amaranth reached 50\% in fallow at doy 193.4,
+from late June to early September (Figure 2). Palmer amaranth growing in
+corn resulted in a longer flowering pattern compared to fallow and
+soybean. Palmer amaranth reached 10\% flowering in soybean at doy 180,
+which was slightly different from fallow (doy 180.9; \emph{P} = 0.01)
+and corn (doy 181.7; \emph{P} = 0.00). Nonetheless, the 10\% cumulative
+Palmer amaranth flowering in soybean, fallow and corn occurred at the
+end of June. The 50\% Palmer amaranth cumulative flowering occurred in
+July. For example, Palmer amaranth reached 50\% in fallow at doy 193.4,
 followed by soybean (doy 194.8), corn (doy 206.6). Similar trend was
@@
 Palmer amaranth cumulative flowering from second transplanting timing
-ranged from mid July to mid September. All cumulative flowering
-comparisons among agroecossystems were significant (\emph{P} \textless{}
-0.00) Palmer amaranth growing in fallow resulted in earlier flowering
-time compared to soybean and corn. Palmer amaranth growing in fallow
-reached 10\%, 50\%, and 90\% flowering time at day 203.8, 214.4, and
-232.2, respectively. Palmer amaranth growing in soybean reached 10\%
+ranged from mid July to mid September (Figure 2). All cumulative
+flowering comparisons among agroecossystems were significant (\emph{P}
+\textless{} 0.00) Palmer amaranth growing in fallow resulted in earlier
+flowering time compared to soybean and corn. Palmer amaranth growing in
+fallow reached 10\%, 50\%, and 90\% flowering time at day 203.8, 214.4,
+and 232.2, respectively. Palmer amaranth growing in soybean reached 10\%
 flowering at doy 210.9, which was 6 days prior to corn (\emph{P}-value =
@@
 
-{\centering \includegraphics[width=160mm,height=130mm]{../data analysis/figures/Combined2}
+{\centering \includegraphics[width=130mm,height=130mm]{../data analysis/figures/Figure 2}
 
@@
 
-{\centering \includegraphics[width=130mm,height=70mm]{../data analysis/figures/figure 3}
+{\centering \includegraphics[width=130mm,height=70mm]{../data analysis/figures/Figure 3}
 

### 2021-07-30 Manuscript/Manuscript.Rmd
fgfgf

@@
 Palmer amaranth has evolved resistance to eight herbicide sites of action [@heap2021], increasing the weed management complexity [@lindsay2017].
-Uncontrolled Palmer amaranth in competition for water, light and nutrients can drastically reduce crop yields [@berger2015].
+Uncontrolled Palmer amaranth in competition for water, light and nutrients can drastically impact on crop yields [@berger2015].
 Palmer amaranth is documented with potential to reduce 91%, 68%, and 54% of corn [@massinga2001], soybean [@klingaman1994], and cotton [@morgan2001] yields, respectively.
-Thus, Palmer amaranth poses a risk to sustainable agriculture.
+Thus, unmanaged Palmer amaranth poses an economical risk to sustainable agriculture.
 
@@
 Also, Palmer amaranth small seed size (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure [@hartzler2016], animals [@farmer2017], and plant propagules [@yu2021].
-Therefore, Palmer amaranth dispersal capacity make the species one of the most successful cases of weed adaption to cropping systems.
+Palmer amaranth dispersal capacity make the species one of the most successful cases of weed adaption to cropping systems.
 
@@
 Palmer amaranth is already documented in agronomic crops of South America [@larran2017; @kupper2017] and Southern Europe [@milani2021].
-In the US, Palmer amaranth is established at crop [@garetson2019] and non-crop land [@bagavathiannan2016] in the warm southern United States but its range is expanding to cool temperatures northward.
+In the US, Palmer amaranth is well established at crop [@garetson2019] and non-crop land [@bagavathiannan2016] in the warm southern United States but its range is expanding to cool temperatures northward.
 For example, herbicide resistant Palmer amaranth is widespread in Nebraska  [@oliveira2021a],  Michigan [@kohrt2017], and Connecticut [@aulakh2021].
@@
 Most tactics to manage Palmer amaranth are based on technology fixes [@scott2011], which are short-term (e.g., herbicide and/or tillage) rather than long-term weed management.
+Long-term tactics to minimize the negative Palmer amaranth impact in the agroecossystem should include species strategies to adapt and grow into different environments.
 
@@
 
-Palmer amaranth cumulative flowering from first transplant timing ranged from late June to early September (Figure 2).
-Palmer amaranth growing in corn resulted in a longer flowering pattern compared to fallow and soybean.
+Palmer amaranth growing in corn resulted in a longer flowering pattern compared to fallow and soybean (Figure 2A).
 Palmer amaranth reached 10% flowering in soybean at doy 180, which was slightly different from fallow (doy 180.9; *P* = 0.01) and corn (doy 181.7; *P* = 0.00).
@@
 
-Palmer amaranth accumulated more biomass when growing in fallow compared to Palmer amaranth growing in soybean and corn.
+Palmer amaranth accumulated more biomass when growing in fallow compared to Palmer amaranth growing in soybean and corn (Figure 3).
 At first transplanting time, Palmer amaranth biomass was 78.3 g plant-1, 26.4 g plant-1 and 14.9 g plant-1 in fallow, soybean and corn, respectively.
@@
 # Discussion {-}
+
+Our study showed the Palmer amaranth adaptation plasticity when growing into different environments. In general,  These result highlight the plant response to grow degree days.
+
+Palmer amaranth growing plasticity varied within time and space.
+In general, Palmer amaranth produced more biomass and taller plants when transplanted in early June rather than July.
+In addition, at first cohort, when resources (e.g., soil nutrients) and conditions (e.g., light) were not limited for both species, Palmer amaranth growing strategy was likely to mimic the cropping systems (Figure 4).
+In contrast, at second cohort, Palmer amaranth grow and development was limited due to the competitive advantages from the crop, including closed canopy.
+For example, when growing in corn at first cohort, Palmer amaranth plants accumulated lowest biomass, but produce tallest plants.
+Higher Palmer amaranth plants in corn is likely the species strategy to compete for light in between corn rows.
+At first cohort, corn and Palmer amaranth had abundant light over the canopy to grow and develop; however, at second cohort, Palmer amaranth was transplanted when corn canopy was nearly closed, which reduced Palmer amaranth competition.
+As a result, Palmer amaranth height and biomass was nearly 50% lower compared to its first cohort.
+
+Similar trend was observed to Palmer amaranth growing in soybeans. The difference in Palmer amaranth height and biomass was lower due to soybean growing strategy.
+
+Palmer amaranth produced the highest biomass when growing without crop competition (fallow).
+When growing in fallow, cohort time had low impact on Palmer amaranth growing and development.
+Moreover, Palmer amaranth biomass was more variable when growing in fallow.
+
+Palmer amaranth cumulative flowering from first transplant timing ranged from late June to early September.
 

### 2021-07-30 Manuscript/Manuscript.tex
fgfgf

@@
 2017). Uncontrolled Palmer amaranth in competition for water, light and
-nutrients can drastically reduce crop yields (Berger et al., 2015).
+nutrients can drastically impact on crop yields (Berger et al., 2015).
 Palmer amaranth is documented with potential to reduce 91\%, 68\%, and
@@
 1994), and cotton (Morgan et al., 2001) yields, respectively. Thus,
-Palmer amaranth poses a risk to sustainable agriculture.
+unmanaged Palmer amaranth poses an economical risk to sustainable
+agriculture.
 
@@
 (Hartzler and Anderson, 2016), animals (Farmer et al., 2017), and plant
-propagules (Yu et al., 2021). Therefore, Palmer amaranth dispersal
-capacity make the species one of the most successful cases of weed
-adaption to cropping systems.
+propagules (Yu et al., 2021). Palmer amaranth dispersal capacity make
+the species one of the most successful cases of weed adaption to
+cropping systems.
 
@@
 America (Larran et al., 2017; Küpper et al., 2017) and Southern Europe
-(Milani et al., 2021). In the US, Palmer amaranth is established at crop
-(Garetson et al., 2019) and non-crop land (Bagavathiannan and
+(Milani et al., 2021). In the US, Palmer amaranth is well established at
+crop (Garetson et al., 2019) and non-crop land (Bagavathiannan and
 Norsworthy, 2016) in the warm southern United States but its range is
@@
 2011), which are short-term (e.g., herbicide and/or tillage) rather than
-long-term weed management.
+long-term weed management. Long-term tactics to minimize the negative
+Palmer amaranth impact in the agroecossystem should include species
+strategies to adapt and grow into different environments.
 
@@
 Daymet weather data from June through September across the five
-locations in 2018 and 2019 (\textbf{correndo2021?}) (Figure 1)
+locations in 2018 and 2019 (Correndo et al., 2021) (Figure 1)
 
@@
 
-Palmer amaranth cumulative flowering from first transplant timing ranged
-from late June to early September (Figure 2). Palmer amaranth growing in
-corn resulted in a longer flowering pattern compared to fallow and
-soybean. Palmer amaranth reached 10\% flowering in soybean at doy 180,
-which was slightly different from fallow (doy 180.9; \emph{P} = 0.01)
-and corn (doy 181.7; \emph{P} = 0.00). Nonetheless, the 10\% cumulative
-Palmer amaranth flowering in soybean, fallow and corn occurred at the
-end of June. The 50\% Palmer amaranth cumulative flowering occurred in
-July. For example, Palmer amaranth reached 50\% in fallow at doy 193.4,
-followed by soybean (doy 194.8), corn (doy 206.6). Similar trend was
-observed at 90\% Palmer amaranth cumulative flowering. Palmer amaranth
-growing in corn reached 90\% flowering at doy 252.6 (early September),
-which was 37.8 and 32.2 days after Palmer amaranth 90\% flowering in
-fallow and soybean, respectively.
+Palmer amaranth growing in corn resulted in a longer flowering pattern
+compared to fallow and soybean (Figure 2A). Palmer amaranth reached 10\%
+flowering in soybean at doy 180, which was slightly different from
+fallow (doy 180.9; \emph{P} = 0.01) and corn (doy 181.7; \emph{P} =
+0.00). Nonetheless, the 10\% cumulative Palmer amaranth flowering in
+soybean, fallow and corn occurred at the end of June. The 50\% Palmer
+amaranth cumulative flowering occurred in July. For example, Palmer
+amaranth reached 50\% in fallow at doy 193.4, followed by soybean (doy
+194.8), corn (doy 206.6). Similar trend was observed at 90\% Palmer
+amaranth cumulative flowering. Palmer amaranth growing in corn reached
+90\% flowering at doy 252.6 (early September), which was 37.8 and 32.2
+days after Palmer amaranth 90\% flowering in fallow and soybean,
+respectively.
 
@@
 Palmer amaranth accumulated more biomass when growing in fallow compared
-to Palmer amaranth growing in soybean and corn. At first transplanting
-time, Palmer amaranth biomass was 78.3 g plant-1, 26.4 g plant-1 and
-14.9 g plant-1 in fallow, soybean and corn, respectively. Palmer
-amaranth produce less biomass at second transplanting time. For example,
-Palmer amaranth growing in fallow resulted in 49.2 g plant-1, 3.4 g
-plant-1 in soybean and 1.4 g plant when growing in corn.
+to Palmer amaranth growing in soybean and corn (Figure 3). At first
+transplanting time, Palmer amaranth biomass was 78.3 g plant-1, 26.4 g
+plant-1 and 14.9 g plant-1 in fallow, soybean and corn, respectively.
+Palmer amaranth produce less biomass at second transplanting time. For
+example, Palmer amaranth growing in fallow resulted in 49.2 g plant-1,
+3.4 g plant-1 in soybean and 1.4 g plant when growing in corn.
 
@@
 \addcontentsline{toc}{section}{Discussion}
+
+Palmer amaranth cumulative flowering from first transplant timing ranged
+from late June to early September.
 
@@
 doi:\href{https://doi.org/10.1017/wsc.2018.47}{10.1017/wsc.2018.47}.
+
+\leavevmode\hypertarget{ref-correndo2021}{}%
+Correndo, A. A., Moro Rosso, L. H., and Ciampitti, I. A. (2021).
+Retrieving and processing agro-meteorological data from {API}-client
+sources using {R} software. \emph{BMC Research Notes} 14, 205.
+doi:\href{https://doi.org/10.1186/s13104-021-05622-8}{10.1186/s13104-021-05622-8}.
 

### 2021-08-03 Manuscript/Manuscript.Rmd
llkl

--- 2021-07-30 Manuscript/Manuscript.Rmd
@@
 Palmer amaranth (*Amaranthus palmeri* S.
-Watson) is currently considered one of the most economically damaged weed species to cropping systems in the United States [@ward2013].
+Watson) is currently considered one of the most economically damaged weed species to cropping systems in the United States .
 The species has showed a remarkable capacity to evolve resistance to herbicides.
-Palmer amaranth has evolved resistance to eight herbicide sites of action [@heap2021], increasing the weed management complexity [@lindsay2017].
-Uncontrolled Palmer amaranth in competition for water, light and nutrients can drastically reduce crop yields [@berger2015].
-Palmer amaranth is documented with potential to reduce 91%, 68%, and 54% of corn [@massinga2001], soybean [@klingaman1994], and cotton [@morgan2001] yields, respectively.
-Thus, Palmer amaranth poses a risk to sustainable agriculture.
+To date, Palmer amaranth has evolved resistance to eight herbicide sites of action [@heap2021], increasing the weed management complexity [@lindsay2017].
+Unmanaged Palmer amaranth in competition for water, light and nutrients can drastically impact on crop yields [@berger2015].
+For example, Palmer amaranth is documented with potential to reduce 91%, 68%, and 54% of corn [@massinga2001], soybean [@klingaman1994], and cotton [@morgan2001] yields, respectively.
+Thus, Palmer amaranth poses an economical risk to sustainable agriculture.
 
@@
 Palmer amaranth weediness is likely a result of human-assisted selection in combination with species biology.
-Farm mechanization, conservation agriculture (e.g., no-till), and reliance on herbicides for weed management are the main human-mediated selection of Palmer amaranth into cropping systems.
-On the other hand, Palmer amaranth is a prolific seed producer with a C4 photosynthetic apparatus [@ward2013].
+Farm mechanization, conservation agriculture (e.g., no-till), and reliance on herbicides for weed management are the main human-mediated selection of Palmer amaranth into cropping systems [@ward2013].
+On the other hand, Palmer amaranth is a prolific seed producer with a C4 photosynthetic apparatus [@wang1992].
 With a dioecy nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging adaptive traits among plants [@oliveira2018].
 Also, Palmer amaranth small seed size (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure [@hartzler2016], animals [@farmer2017], and plant propagules [@yu2021].
-Therefore, Palmer amaranth dispersal capacity make the species one of the most successful cases of weed adaption to cropping systems.
+The dispersal capacity of Palmer amaranth make the species one of the most successful cases of weed adaption to cropping systems.
 
@@
 Most tactics to manage Palmer amaranth are based on technology fixes [@scott2011], which are short-term (e.g., herbicide and/or tillage) rather than long-term weed management.
+Strategies to manage Palmer amaranth should be built on minimizing the species ability to adapt, grow and develop into agroecossystems.
 
-The continuous Palmer amaranth dispersal and potential establishment into northern United States warrant investigations on species morphology growing in such environments.
+The continuous Palmer amaranth dispersal and potential establishment into northern United States warrant investigations on species morphology in such environments.
 Understanding Palmer amaranth biology and growing strategies under different agroecossystems can enhance our knowledge on species adaptation.
 It can also aid on designing proactive and ecological tactics to limit the species range expansion, reduce its negative impact, and design resilient and sustainable farming systems [@maclaren2020].
-Therefore, the objective of this study was to investigate the flowering pattern, biomass production, and height of Palmer amaranth growing under in corn, soybean and fallow at two timings across five locations in the mid/upper United States Midwest.
+Therefore, the objective of this study was to investigate the flowering pattern, biomass production, and height of Palmer amaranth growing under corn, soybean and fallow at two timings across five locations in the United States Midwest.
 
@@
 
+For example, Palmer amaranth growing at first cohort accumulated 90%, 85% and 31% more biomass when compared to second cohort at corn, soybean and fallow, respectively
+
+In addition, Palmer amaranth growing in corn and soybean at second cohort had a similar end-season flowering pattern.
+
 # Discussion {-}
+
+Our study showed the Palmer amaranth flowering pattern, grow and development varied within agroecossystems and timing.
+In general, Palmer amaranth produced more biomass and taller plants when growing at first cohort rather than second cohort.
+At first cohort, resources (e.g., soil nutrients) and conditions (e.g., light) were more timely available for both species, crop and weed.
+Thus, Palmer amaranth growing strategy was to mimic the crop species (Figure 4).
+In contrast, at second cohort, Palmer amaranth grow and development was limited due to the crop competitive ability at advanced development stages.
+
+Taller Palmer amaranth plants is a possible strategy to compete for light in between crop rows in absence of canopy.
+For example, at first cohort, corn and Palmer amaranth had abundant light to grow and develop, which also aid its extended the flowering pattern.
+However, at second cohort, Palmer amaranth was transplanted when corn canopy was nearly closed, which reduced Palmer amaranth competition.
+As a result, Palmer amaranth height and biomass was nearly 50% and 90% lower compared to its first cohort, respectively.
+Similar trend was observed to Palmer amaranth growing in soybeans.
+Palmer amaranth growing without crop competition produced highest amount of biomass.
+The Palmer amaranth strategy in fallow was to invest biomass in growing plant width and height.
+Nonetheless, Palmer amaranth produced 31% less biomass in second cohort compared to first cohort timing.
+Therefore, crop competition was not the only factor limiting Palmer amaranth growing in the second cohort.
+
+It was not evaluate seed production due to Palmer amaranth harvest at early flowering.
+Nonetheless, it is reported a strong positive correlation between Palmer amaranth biomass and seed production [@schwartz2016; @spaunhorst2018].
+I our study, Palmer amaranth growing at first cohort accumulated nearly 50% more overall biomass when compared to second cohort.
+Therefore, Palmer amaranth plants growing in the second cohort is likely to produce less seeds regardless the agroecossystem.
+Seed production and deposition in the seedbank is a key factor for species perpetuation [@menges1987].
+Palmer amaranth germination is influenced by soil depth and seed viable seeds were fount after 36 months buried in the seedbank [@sosnoskie2013].
+Therefore, tactics that promote crop advantage against Palmer amaranth, including early crop planting, preemergence applied herbicide, and crop residue (e.g.
+cover crops) can minimize the negative impact of Palmer amaranth in agroecossystems.
+
+Palmer amaranth has shown to be responsive to to thermal units (e.g, growing degree days).
+A study in North Carolina demonstrated that Palmer amaranth emergence in response to soil temperature, with seedlings emerging before July prior to crop canopy closure [@piskackova2021].
+In addition, in a bareground study, Palmer emergence emergence pattern in Nebraska was influenced by soil management.
+Early tillage resulted in 90% Palmer amaranth emergence in June
+
+The higher amount of biomass in the first cohort can also be related to the longer flowering period in the first cohort.
+In this study, biomass is related to flowering as plant was harvest at early flowering.
+
+can cause significant soybean yield loss [@korres2019]
 

### 2021-08-03 Manuscript/Manuscript.tex
llkl

@@
 considered one of the most economically damaged weed species to cropping
-systems in the United States (Ward et al., 2013). The species has showed
-a remarkable capacity to evolve resistance to herbicides. Palmer
-amaranth has evolved resistance to eight herbicide sites of action
-(Heap, 2021), increasing the weed management complexity (Lindsay et al.,
-2017). Uncontrolled Palmer amaranth in competition for water, light and
-nutrients can drastically impact on crop yields (Berger et al., 2015).
-Palmer amaranth is documented with potential to reduce 91\%, 68\%, and
-54\% of corn (Massinga et al., 2001), soybean (Klingaman and Oliver,
-1994), and cotton (Morgan et al., 2001) yields, respectively. Thus,
-unmanaged Palmer amaranth poses an economical risk to sustainable
+systems in the United States . The species has showed a remarkable
+capacity to evolve resistance to herbicides. To date, Palmer amaranth
+has evolved resistance to eight herbicide sites of action (Heap, 2021),
+increasing the weed management complexity (Lindsay et al., 2017).
+Unmanaged Palmer amaranth in competition for water, light and nutrients
+can drastically impact on crop yields (Berger et al., 2015). For
+example, Palmer amaranth is documented with potential to reduce 91\%,
+68\%, and 54\% of corn (Massinga et al., 2001), soybean (Klingaman and
+Oliver, 1994), and cotton (Morgan et al., 2001) yields, respectively.
+Thus, Palmer amaranth poses an economical risk to sustainable
 agriculture.
@@
 and reliance on herbicides for weed management are the main
-human-mediated selection of Palmer amaranth into cropping systems. On
-the other hand, Palmer amaranth is a prolific seed producer with a C4
-photosynthetic apparatus (Ward et al., 2013). With a dioecy nature,
-Palmer amaranth male and female plants are obligate outcrosser species,
-increasing the chances of exchanging adaptive traits among plants
-(Oliveira et al., 2018). Also, Palmer amaranth small seed size (e.g, 1
-mm) tend to thrive in no-tillage systems (Price et al., 2011), and
-spread across locations through farm equipment (Sauer, 1972), manure
-(Hartzler and Anderson, 2016), animals (Farmer et al., 2017), and plant
-propagules (Yu et al., 2021). Palmer amaranth dispersal capacity make
-the species one of the most successful cases of weed adaption to
-cropping systems.
+human-mediated selection of Palmer amaranth into cropping systems (Ward
+et al., 2013). On the other hand, Palmer amaranth is a prolific seed
+producer with a C4 photosynthetic apparatus (Wang et al., 1992). With a
+dioecy nature, Palmer amaranth male and female plants are obligate
+outcrosser species, increasing the chances of exchanging adaptive traits
+among plants (Oliveira et al., 2018). Also, Palmer amaranth small seed
+size (e.g, 1 mm) tend to thrive in no-tillage systems (Price et al.,
+2011), and spread across locations through farm equipment (Sauer, 1972),
+manure (Hartzler and Anderson, 2016), animals (Farmer et al., 2017), and
+plant propagules (Yu et al., 2021). The dispersal capacity of Palmer
+amaranth make the species one of the most successful cases of weed
+adaption to cropping systems.
 
@@
 2011), which are short-term (e.g., herbicide and/or tillage) rather than
-long-term weed management. Long-term tactics to minimize the negative
-Palmer amaranth impact in the agroecossystem should include species
-strategies to adapt and grow into different environments.
+long-term weed management. Strategies to manage Palmer amaranth should
+be built on minimizing the species ability to adapt, grow and develop
+into agroecossystems.
 
@@
 into northern United States warrant investigations on species morphology
-growing in such environments. Understanding Palmer amaranth biology and
-growing strategies under different agroecossystems can enhance our
-knowledge on species adaptation. It can also aid on designing proactive
-and ecological tactics to limit the species range expansion, reduce its
+in such environments. Understanding Palmer amaranth biology and growing
+strategies under different agroecossystems can enhance our knowledge on
+species adaptation. It can also aid on designing proactive and
+ecological tactics to limit the species range expansion, reduce its
 negative impact, and design resilient and sustainable farming systems
@@
 investigate the flowering pattern, biomass production, and height of
-Palmer amaranth growing under in corn, soybean and fallow at two timings
-across five locations in the mid/upper United States Midwest.
+Palmer amaranth growing under corn, soybean and fallow at two timings
+across five locations in the United States Midwest.
 
@@
 
+For example, Palmer amaranth growing at first cohort accumulated 90\%,
+85\% and 31\% more biomass when compared to second cohort at corn,
+soybean and fallow, respectively
+
+In addition, Palmer amaranth growing in corn and soybean at second
+cohort had a similar end-season flowering pattern.
+
 \hypertarget{discussion}{%
@@
 
-Palmer amaranth cumulative flowering from first transplant timing ranged
-from late June to early September.
+Our study showed the Palmer amaranth flowering pattern, grow and
+development varied within agroecossystems and timing. In general, Palmer
+amaranth produced more biomass and taller plants when growing at first
+cohort rather than second cohort. At first cohort, resources (e.g., soil
+nutrients) and conditions (e.g., light) were more timely available for
+both species, crop and weed. Thus, Palmer amaranth growing strategy was
+to mimic the crop species (Figure 4). In contrast, at second cohort,
+Palmer amaranth grow and development was limited due to the crop
+competitive ability at advanced development stages.
+
+Taller Palmer amaranth plants is a strategy to compete for light in
+between crop rows in absence of canopy. For example, at first cohort,
+corn and Palmer amaranth had abundant light to grow and develop, which
+likely aid its extended the flowering pattern. However, at second
+cohort, Palmer amaranth was transplanted when corn canopy was nearly
+closed, which reduced Palmer amaranth competition. As a result, Palmer
+amaranth height and biomass was nearly 50\% and 90\% lower compared to
+its first cohort, respectively. Similar trend was observed to Palmer
+amaranth growing in soybeans. Palmer amaranth growing without crop
+competition produced highest amount of biomass. The Palmer amaranth
+strategy in fallow was to invest biomass in growing plant width and
+height. Nonetheless, Palmer amaranth produced 31\% less biomass in
+second cohort compared to first cohort timing. Therefore, crop
+competition was not the only factor limiting Palmer amaranth growing in
+the second cohort.
+
+It was not evaluate seed production due to Palmer amaranth harvest at
+early flowering. Nonetheless, it is reported a strong positive
+correlation between Palmer amaranth biomass and seed production
+(Schwartz et al., 2016; Spaunhorst et al., 2018). I our study, Palmer
+amaranth growing at first cohort accumulated nearly 50\% more overall
+biomass when compared to second cohort. Therefore, Palmer amaranth
+plants growing in the second cohort is likely to produce less seeds
+regardless the agroecossystem. Seed production and deposition in the
+seedbank is a key factor for species perpetuation (Menges, 1987). Palmer
+amaranth germination is influenced by soil depth and seed viable seeds
+were fount after 36 months buried in the seedbank (Sosnoskie et al.,
+2013). Therefore, tactics that promote crop advantage against Palmer
+amaranth, including early crop planting, preemergence applied herbicide,
+and crop residue (e.g.~cover crops) can minimize the negative impact of
+Palmer amaranth in agroecossystems.
+
+Palmer amaranth has shown to be responsive to to thermal units (e.g,
+growing degree days). A study in North Carolina demonstrated that Palmer
+amaranth emergence in response to soil temperature, with seedlings
+emerging before July prior to crop canopy closure (Piskackova et al.,
+2021). In addition, in a bareground study, Palmer emergence emergence
+pattern in Nebraska was influenced by soil management. Early tillage
+resulted in 90\% Palmer amaranth emergence in June
+
+can cause significant soybean yield loss (Korres et al., 2019)
 
@@
 
+\leavevmode\hypertarget{ref-korres2019}{}%
+Korres, N. E., Norsworthy, J. K., and Mauromoustakos, A. (2019). Effects
+of {Palmer Amaranth} ({Amaranthus} palmeri) {Establishment Time} and
+{Distance} from the {Crop Row} on {Biological} and {Phenological
+Characteristics} of the {Weed}: {Implications} on {Soybean Yield}.
+\emph{Weed Science} 67, 126--135.
+doi:\href{https://doi.org/10.1017/wsc.2018.84}{10.1017/wsc.2018.84}.
+
 \leavevmode\hypertarget{ref-kupper2017}{}%
@@
 
+\leavevmode\hypertarget{ref-menges1987}{}%
+Menges, R. M. (1987). Weed {Seed Population Dynamics} during {Six Years}
+of {Weed Management Systems} in {Crop Rotations} on {Irrigated Soil}.
+\emph{Weed Science} 35, 328--332. Available at:
+\url{http://www.jstor.org/stable/4044593}.
+
 \leavevmode\hypertarget{ref-milani2021}{}%
@@
 
+\leavevmode\hypertarget{ref-piskackova2021}{}%
+Piskackova, T. A. R., Reberg-Horton, S. C., Richardson, R. J., Jennings,
+K. M., Franca, L., Young, B. G., and Leon, R. G. (2021). Windows of
+action for controlling palmer amaranth ({Amaranthus} palmeri) using
+emergence and phenology models. \emph{Weed Research} 61, 188--198.
+doi:\href{https://doi.org/10.1111/wre.12470}{10.1111/wre.12470}.
+
 \leavevmode\hypertarget{ref-price2011}{}%
@@
 
+\leavevmode\hypertarget{ref-schwartz2016}{}%
+Schwartz, L. M., Norsworthy, J. K., Young, B. G., Bradley, K. W.,
+Kruger, G. R., Davis, V. M., Steckel, L. E., and Walsh, M. J. (2016).
+Tall {Waterhemp} ({Amaranthus} tuberculatus) and {Palmer} amaranth
+({Amaranthus} palmeri) {Seed Production} and {Retention} at {Soybean
+Maturity}. \emph{Weed Technology} 30, 284--290.
+doi:\href{https://doi.org/10.1614/WT-D-15-00130.1}{10.1614/WT-D-15-00130.1}.
+
 \leavevmode\hypertarget{ref-scott2011}{}%
@@
 doi:\href{https://doi.org/10.1007/s10806-010-9253-7}{10.1007/s10806-010-9253-7}.
+
+\leavevmode\hypertarget{ref-sosnoskie2013}{}%
+Sosnoskie, L. M., Webster, T. M., and Culpepper, A. S. (2013).
+Glyphosate {Resistance Does Not Affect Palmer Amaranth} ({Amaranthus}
+palmeri) {Seedbank Longevity}. \emph{Weed Science} 61, 283--288.
+doi:\href{https://doi.org/10.1614/WS-D-12-00111.1}{10.1614/WS-D-12-00111.1}.
+
+\leavevmode\hypertarget{ref-spaunhorst2018}{}%
+Spaunhorst, D. J., Devkota, P., Johnson, W. G., Smeda, R. J., Meyer, C.
+J., and Norsworthy, J. K. (2018). Phenology of {Five Palmer} amaranth
+({Amaranthus} palmeri) {Populations Grown} in {Northern Indiana} and
+{Arkansas}. \emph{Weed Science} 66, 457--469.
+doi:\href{https://doi.org/10.1017/wsc.2018.12}{10.1017/wsc.2018.12}.
+
+\leavevmode\hypertarget{ref-wang1992}{}%
+Wang, J. L., Klessig, D. F., and Berry, J. O. (1992). Regulation of {C4
+Gene Expression} in {Developing Amaranth Leaves}. \emph{The Plant Cell}
+4, 173--184.
+doi:\href{https://doi.org/10.1105/tpc.4.2.173}{10.1105/tpc.4.2.173}.
 

### 2021-08-04 Manuscript/Manuscript.Rmd
jh

--- 2021-07-30 Manuscript/Manuscript.Rmd
@@
 Palmer amaranth (*Amaranthus palmeri* S.
-Watson) is currently considered one of the most economically damaged weed species to cropping systems in the United States [@ward2013].
-The species has showed a remarkable capacity to evolve resistance to herbicides.
-Palmer amaranth has evolved resistance to eight herbicide sites of action [@heap2021], increasing the weed management complexity [@lindsay2017].
-Uncontrolled Palmer amaranth in competition for water, light and nutrients can drastically reduce crop yields [@berger2015].
-Palmer amaranth is documented with potential to reduce 91%, 68%, and 54% of corn [@massinga2001], soybean [@klingaman1994], and cotton [@morgan2001] yields, respectively.
-Thus, Palmer amaranth poses a risk to sustainable agriculture.
+Watson) is currently considered one of the most economically damaged weed species to cropping systems in the United States.
+Unmanaged Palmer amaranth in competition for water, light and nutrients can drastically impact on crop yields [@berger2015].
+For example, Palmer amaranth is documented with potential to reduce 91%, 68%, and 54% of corn [@massinga2001], soybean [@klingaman1994], and cotton [@morgan2001] yields, respectively.
+Moreover, Palmer amaranth has showed a remarkable capacity to evolve resistance to herbicides.
+To date, Palmer amaranth has evolved resistance to eight herbicide sites of action [@heap2021], increasing the weed management complexity [@lindsay2017].
+Thus, Palmer amaranth poses an economical risk to sustainable agriculture.
 
@@
 Palmer amaranth weediness is likely a result of human-assisted selection in combination with species biology.
-Farm mechanization, conservation agriculture (e.g., no-till), and reliance on herbicides for weed management are the main human-mediated selection of Palmer amaranth into cropping systems.
-On the other hand, Palmer amaranth is a prolific seed producer with a C4 photosynthetic apparatus [@ward2013].
+Farm mechanization, conservation agriculture (e.g., no-till), and reliance on herbicides for weed management are the main human-mediated selection of Palmer amaranth into cropping systems [@ward2013].
+On the other hand, Palmer amaranth is a prolific seed producer with a C4 photosynthetic apparatus [@wang1992].
 With a dioecy nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging adaptive traits among plants [@oliveira2018].
 Also, Palmer amaranth small seed size (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure [@hartzler2016], animals [@farmer2017], and plant propagules [@yu2021].
-Therefore, Palmer amaranth dispersal capacity make the species one of the most successful cases of weed adaption to cropping systems.
+The dispersal capacity of Palmer amaranth make the species one of the most successful cases of weed adaption to cropping systems.
 
-Light and temperature are likely the main environment requirements for Palmer amaranth successful adaptation.
+Light and temperature are likely the main environment requirements for Palmer amaranth successful grow and development.
 Palmer amaranth is reported with an extended germination period [@jha2010].
 Germination of Palmer amaranth is triggered by 18 C soil temperature [@keeley1987], and optimal germination and biomass production occur at 35/30 C day and night temperatures [@guo2003].
+Also, Palmer amaranth germination are human-mediated by tillage timings and preemergence-applied herbicides [@chahal2021].
 Water has not shown to limit Palmer amaranth fitness.
 Under continuous water stress, Palmer amaranth survived and produced at least 14000 seeds plant-1 [@chahal2018].
-Also, seeds from Palmer amaranth growing under water stress conditions were heavier, less dormant, and prompt for germination [@matzrafi2021].
+Also, seeds from Palmer amaranth growing with limited water conditions were heavier, less dormant, and prompt for germination [@matzrafi2021].
 The continuous global temperature warming can impact agriculture and promote niches for Palmer amaranth invasion/adaptation into new environments.
@@
 
-Palmer amaranth is already documented in agronomic crops of South America [@larran2017; @kupper2017] and Southern Europe [@milani2021].
-In the US, Palmer amaranth is established at crop [@garetson2019] and non-crop land [@bagavathiannan2016] in the warm southern United States but its range is expanding to cool temperatures northward.
-For example, herbicide resistant Palmer amaranth is widespread in Nebraska  [@oliveira2021a],  Michigan [@kohrt2017], and Connecticut [@aulakh2021].
-Successful cases of Palmer amaranth invasion and near to eradication is reported in Minnesota [@yu2021].
+Palmer amaranth is already found in agronomic crops of South America [@larran2017; @kupper2017] and Southern Europe [@milani2021].
+In the US, Palmer amaranth is well established at cotton fields [@garetson2019] and non-crop areas [@bagavathiannan2016] in the southern United States but its range is expanding northward.
+For example, herbicide resistant Palmer amaranth is widespread in Nebraska  [@oliveira2021a],  Michigan [@kohrt2017], Connecticut [@aulakh2021], and potential damaged to soybean fields in Illinois is reported [@davis2015].
+Successful cases of Palmer amaranth invasion and near to eradication is documented in Minnesota [@yu2021].
 No Palmer amaranth actively growing was found in Canada; however, Palmer amaranth seeds was detected in sweet potato slips [@page2021].
 Nonetheless, it seems fated the need to manage Palmer amaranth in agronomic crops throughout multiple environments in the near future.
-Therefore, strategies on Palmer amaranth management should encompass the agroecosystem level but not only attempts to eradicate the weed.
+Strategies on Palmer amaranth management should encompass the agroecosystem level but not only attempts to eradicate the weed.
 Most tactics to manage Palmer amaranth are based on technology fixes [@scott2011], which are short-term (e.g., herbicide and/or tillage) rather than long-term weed management.
+Palmer amaranth management should be built on minimizing the species ability to adapt, grow and develop into agroecossystems.
 
-The continuous Palmer amaranth dispersal and potential establishment into northern United States warrant investigations on species morphology growing in such environments.
+In the southeastern US, early growing Palmer amaranth is well known to have higher impact on cotton yields than late established plants [@macrae2013].
+In the northern states, Palmer amaranth impact on the agroecossystem is recent.
+Studies investigating Palmer amaranth in those locations is limited due to the plant classification as noxious weed [@yu2021].
+Nonetheless, the continuous Palmer amaranth dispersal and potential establishment across northern United States is concerning and warrant investigations on species morphology in such environments.
 Understanding Palmer amaranth biology and growing strategies under different agroecossystems can enhance our knowledge on species adaptation.
 It can also aid on designing proactive and ecological tactics to limit the species range expansion, reduce its negative impact, and design resilient and sustainable farming systems [@maclaren2020].
-Therefore, the objective of this study was to investigate the flowering pattern, biomass production, and height of Palmer amaranth growing under in corn, soybean and fallow at two timings across five locations in the mid/upper United States Midwest.
+Therefore, the objective of this study was to investigate the flowering pattern, biomass production, and height of Palmer amaranth growing under corn, soybean and fallow at two timings across five locations in the United States Midwest.
 
@@
 
-Palmer amaranth cumulative flowering from first transplant timing ranged from late June to early September (Figure 2).
-Palmer amaranth growing in corn resulted in a longer flowering pattern compared to fallow and soybean.
+Palmer amaranth growing in corn resulted in a longer flowering pattern compared to fallow and soybean at first cohort (Figure 2A).
+The 10% cumulative Palmer amaranth flowering in soybean, fallow and corn occurred at the end of June.
 Palmer amaranth reached 10% flowering in soybean at doy 180, which was slightly different from fallow (doy 180.9; *P* = 0.01) and corn (doy 181.7; *P* = 0.00).
-Nonetheless, the 10% cumulative Palmer amaranth flowering in soybean, fallow and corn occurred at the end of June.
 The 50% Palmer amaranth cumulative flowering occurred in July.
@@
 
-Palmer amaranth cumulative flowering from second transplanting timing ranged from mid July to mid September (Figure 2).
-All cumulative flowering comparisons among agroecossystems were significant (*P* < 0.00) Palmer amaranth growing in fallow resulted in earlier flowering time compared to soybean and corn.
+Palmer amaranth cumulative flowering at second cohort ranged from mid July to mid September (Figure 2).
+All cumulative flowering comparisons among agroecossystems were significant different (*P* < 0.00) Palmer amaranth growing in fallow resulted in earlier flowering time compared to soybean and corn.
 Palmer amaranth growing in fallow reached 10%, 50%, and 90% flowering time at day 203.8, 214.4, and 232.2, respectively.
@@
 
-Palmer amaranth accumulated more biomass when growing in fallow compared to Palmer amaranth growing in soybean and corn.
-At first transplanting time, Palmer amaranth biomass was 78.3 g plant-1, 26.4 g plant-1 and 14.9 g plant-1 in fallow, soybean and corn, respectively.
-Palmer amaranth produce less biomass at second transplanting time.
-For example, Palmer amaranth growing in fallow resulted in 49.2 g plant-1, 3.4 g plant-1 in soybean and 1.4 g plant when growing in corn.
+Palmer amaranth accumulated more biomass when growing in fallow compared to Palmer amaranth growing in soybean and corn (Figure 3A).
+At first cohort time, Palmer amaranth biomass was 78.3 g plant-1, 26.4 g plant-1 and 14.9 g plant-1 in fallow, soybean and corn, respectively.
+Moreover, Palmer amaranth produce less biomass at cohort time.
+Palmer amaranth growing in fallow resulted in 49.2 g plant-1, 3.4 g plant-1 in soybean and 1.4 g plant when growing in corn.
+
+The Palmer amaranth achieve 92.3 cm tall when growing at first cohort timing in corn (Figure 3B).
 
@@
 # Discussion {-}
+
+Our study showed that Palmer amaranth flowering time, biomass and height varied within agroecossystems and cohort timing.
+In general, Palmer amaranth produced more biomass and taller plants when growing at first cohort rather than second cohort.
+At first cohort, resources (e.g., soil nutrients) and conditions (e.g., light) were more timely available for both species, crop and weed.
+High biomass and taller Palmer amaranth plants is likely a weed strategy to compete for light in between crop rows in absence of canopy.
+In such condition, Palmer amaranth showed an extraordinary plasticity to adapt upon the agroecosystem.
+This is evident when comparing Palmer amaranth shape, and its extended flowering pattern when growing into corn compared to soybean.
+In corn, Palmer amaranth was taller and thinner when growing in corn.
+The Palmer amaranth competition strategy was likely to mimic the crop grow and development (Figure 4).
+These results are concerning as breeding more competitive crop varieties is likely to select more competitive Palmer amaranth biotypes [@bravo2017].
+
+Palmer amaranth grow and development at second cohort was limited due to the crop competitive ability at advanced development stages.
+Palmer amaranth was transplanted when corn canopy was nearly closed, which reduced Palmer amaranth competitiveness.
+As a result, Palmer amaranth height and biomass was lower compared to its first cohort.
+Under crop canopy, Palmer amaranth flowering pattern was near to similar into corn and soybean.
+Palmer amaranth growing without crop competition produced highest amounts of biomass and lowest extended flowering pattern.
+The Palmer amaranth strategy in bareground was to invest biomass in growing plant width and height.
+Nonetheless, Palmer amaranth produced 31% less biomass in second cohort compared to first cohort timing.
+In a bareground study, early emerged Palmer amaranth without competition was 50% taller than late emerged plants [@webster2015].
+These results suggests that crop competition is not the only factor limiting late Palmer amaranth establishment.
+The limited growth of Palmer amaranth at second cohort is likely a reduced plant response to thermal units (e.g, growing degree days).
+It is hypothesize that reduced day length contributed to smaller plants at second cohort as well as shorter flowering period.
+A study in North Carolina and Illinois predicted that less than 10% Palmer amaranth seedlings emergence occurred after June [@piskackova2021].
+In addition, Palmer amaranth negative impact on soybean [@korres2020] and cotton [@webster2015] yields was higher when plants were established near to crop planting.
+
+Seed production was not evaluated due to Palmer amaranth harvest at flowering time.
+Nonetheless, it is well documented a strong positive correlation between Palmer amaranth biomass and seed production [@schwartz2016; @spaunhorst2018].
+I our study, Palmer amaranth growing at first cohort accumulated an overall 50% more biomass when compared to second cohort.
+Therefore, Palmer amaranth plants growing in the second cohort is likely to produce less seeds regardless the agroecossystem.
+Our observation is consistent with the findings that first Palmer amaranth cohort produced 50% more seeds per plant than Palmer amaranth plants established six weeks later in bareground [@webster2015].
+Nonetheless, seed production at second cohort is likely to replenish the soil seedbank.
+Seed production and deposition in the seedbank is a key factor for species perpetuation [@menges1987].
+Palmer amaranth can produce hundred thousands seeds per plant [@schwartz2016; @keeley1987], and stay viable buried in the seedbank for at least 36 months [@sosnoskie2013].
+
+Our study demonstrated the short-term Palmer amaranth plasticity to grow and develop into cropping-systems.
+Palmer amaranth management priority should focus on minimizing Palmer amaranth dispersal into new geographies.
+Early-season management programs would have a large negative effect in Palmer amaranth growth and development.
+Tactics that promote crop advantage against Palmer amaranth, including early crop planting, crop rotation, plant width, preemergence applied herbicide [@sanctis2021], and crop residue (e.g.
+cover crops) would minimize the negative impact of Palmer amaranth in agroecossystems.
 

### 2021-08-04 Manuscript/Manuscript.tex
jh

@@
 considered one of the most economically damaged weed species to cropping
-systems in the United States . The species has showed a remarkable
-capacity to evolve resistance to herbicides. To date, Palmer amaranth
-has evolved resistance to eight herbicide sites of action (Heap, 2021),
-increasing the weed management complexity (Lindsay et al., 2017).
-Unmanaged Palmer amaranth in competition for water, light and nutrients
-can drastically impact on crop yields (Berger et al., 2015). For
-example, Palmer amaranth is documented with potential to reduce 91\%,
-68\%, and 54\% of corn (Massinga et al., 2001), soybean (Klingaman and
-Oliver, 1994), and cotton (Morgan et al., 2001) yields, respectively.
-Thus, Palmer amaranth poses an economical risk to sustainable
+systems in the United States. Unmanaged Palmer amaranth in competition
+for water, light and nutrients can drastically impact on crop yields
+(Berger et al., 2015). For example, Palmer amaranth is documented with
+potential to reduce 91\%, 68\%, and 54\% of corn (Massinga et al.,
+2001), soybean (Klingaman and Oliver, 1994), and cotton (Morgan et al.,
+2001) yields, respectively. Moreover, Palmer amaranth has showed a
+remarkable capacity to evolve resistance to herbicides. To date, Palmer
+amaranth has evolved resistance to eight herbicide sites of action
+(Heap, 2021), increasing the weed management complexity (Lindsay et al.,
+2017). Thus, Palmer amaranth poses an economical risk to sustainable
 agriculture.
@@
 Light and temperature are likely the main environment requirements for
-Palmer amaranth successful adaptation. Palmer amaranth is reported with
-an extended germination period (Jha et al., 2010). Germination of Palmer
-amaranth is triggered by 18 C soil temperature (Keeley et al., 1987),
-and optimal germination and biomass production occur at 35/30 C day and
-night temperatures (Guo and Al-Khatib, 2003). Water has not shown to
-limit Palmer amaranth fitness. Under continuous water stress, Palmer
-amaranth survived and produced at least 14000 seeds plant-1 (Chahal et
-al., 2018). Also, seeds from Palmer amaranth growing under water stress
-conditions were heavier, less dormant, and prompt for germination
-(Matzrafi et al., 2021). The continuous global temperature warming can
-impact agriculture and promote niches for Palmer amaranth
+Palmer amaranth successful grow and development. Palmer amaranth is
+reported with an extended germination period (Jha et al., 2010).
+Germination of Palmer amaranth is triggered by 18 C soil temperature
+(Keeley et al., 1987), and optimal germination and biomass production
+occur at 35/30 C day and night temperatures (Guo and Al-Khatib, 2003).
+Also, Palmer amaranth germination are human-mediated by tillage timings
+and preemergence-applied herbicides (Chahal et al., 2021). Water has not
+shown to limit Palmer amaranth fitness. Under continuous water stress,
+Palmer amaranth survived and produced at least 14000 seeds plant-1
+(Chahal et al., 2018). Also, seeds from Palmer amaranth growing with
+limited water conditions were heavier, less dormant, and prompt for
+germination (Matzrafi et al., 2021). The continuous global temperature
+warming can impact agriculture and promote niches for Palmer amaranth
 invasion/adaptation into new environments. Currently, it is estimated
@@
 
-Palmer amaranth is already documented in agronomic crops of South
-America (Larran et al., 2017; Küpper et al., 2017) and Southern Europe
-(Milani et al., 2021). In the US, Palmer amaranth is well established at
-crop (Garetson et al., 2019) and non-crop land (Bagavathiannan and
-Norsworthy, 2016) in the warm southern United States but its range is
-expanding to cool temperatures northward. For example, herbicide
-resistant Palmer amaranth is widespread in Nebraska (Oliveira et al.,
-2021), Michigan (Kohrt et al., 2017), and Connecticut (Aulakh et al.,
-2021). Successful cases of Palmer amaranth invasion and near to
-eradication is reported in Minnesota (Yu et al., 2021). No Palmer
-amaranth actively growing was found in Canada; however, Palmer amaranth
-seeds was detected in sweet potato slips (Page et al., 2021).
-Nonetheless, it seems fated the need to manage Palmer amaranth in
-agronomic crops throughout multiple environments in the near future.
-Therefore, strategies on Palmer amaranth management should encompass the
-agroecosystem level but not only attempts to eradicate the weed. Most
-tactics to manage Palmer amaranth are based on technology fixes (Scott,
-2011), which are short-term (e.g., herbicide and/or tillage) rather than
-long-term weed management. Strategies to manage Palmer amaranth should
-be built on minimizing the species ability to adapt, grow and develop
-into agroecossystems.
+Palmer amaranth is already found in agronomic crops of South America
+(Larran et al., 2017; Küpper et al., 2017) and Southern Europe (Milani
+et al., 2021). In the US, Palmer amaranth is well established at cotton
+fields (Garetson et al., 2019) and non-crop areas (Bagavathiannan and
+Norsworthy, 2016) in the southern United States but its range is
+expanding northward. For example, herbicide resistant Palmer amaranth is
+widespread in Nebraska (Oliveira et al., 2021), Michigan (Kohrt et al.,
+2017), Connecticut (Aulakh et al., 2021), and potential damaged to
+soybean fields in Illinois is reported (Davis et al., 2015). Successful
+cases of Palmer amaranth invasion and near to eradication is documented
+in Minnesota (Yu et al., 2021). No Palmer amaranth actively growing was
+found in Canada; however, Palmer amaranth seeds was detected in sweet
+potato slips (Page et al., 2021). Nonetheless, it seems fated the need
+to manage Palmer amaranth in agronomic crops throughout multiple
+environments in the near future. Strategies on Palmer amaranth
+management should encompass the agroecosystem level but not only
+attempts to eradicate the weed. Most tactics to manage Palmer amaranth
+are based on technology fixes (Scott, 2011), which are short-term (e.g.,
+herbicide and/or tillage) rather than long-term weed management. Palmer
+amaranth management should be built on minimizing the species ability to
+adapt, grow and develop into agroecossystems.
 
-The continuous Palmer amaranth dispersal and potential establishment
-into northern United States warrant investigations on species morphology
-in such environments. Understanding Palmer amaranth biology and growing
-strategies under different agroecossystems can enhance our knowledge on
-species adaptation. It can also aid on designing proactive and
-ecological tactics to limit the species range expansion, reduce its
-negative impact, and design resilient and sustainable farming systems
-(MacLaren et al., 2020). Therefore, the objective of this study was to
-investigate the flowering pattern, biomass production, and height of
-Palmer amaranth growing under corn, soybean and fallow at two timings
-across five locations in the United States Midwest.
+In the southeastern US, early growing Palmer amaranth is well known to
+have higher impact on cotton yields than late established plants (MacRae
+et al., 2013). In the northern states, Palmer amaranth impact on the
+agroecossystem is recent. Studies investigating Palmer amaranth in those
+locations is limited due to the plant classification as noxious weed (Yu
+et al., 2021). Nonetheless, the continuous Palmer amaranth dispersal and
+potential establishment across northern United States is concerning and
+warrant investigations on species morphology in such environments.
+Understanding Palmer amaranth biology and growing strategies under
+different agroecossystems can enhance our knowledge on species
+adaptation. It can also aid on designing proactive and ecological
+tactics to limit the species range expansion, reduce its negative
+impact, and design resilient and sustainable farming systems (MacLaren
+et al., 2020). Therefore, the objective of this study was to investigate
+the flowering pattern, biomass production, and height of Palmer amaranth
+growing under corn, soybean and fallow at two timings across five
+locations in the United States Midwest.
 
@@
 Palmer amaranth growing in corn resulted in a longer flowering pattern
-compared to fallow and soybean (Figure 2A). Palmer amaranth reached 10\%
-flowering in soybean at doy 180, which was slightly different from
-fallow (doy 180.9; \emph{P} = 0.01) and corn (doy 181.7; \emph{P} =
-0.00). Nonetheless, the 10\% cumulative Palmer amaranth flowering in
-soybean, fallow and corn occurred at the end of June. The 50\% Palmer
+compared to fallow and soybean at first cohort (Figure 2A). The 10\%
+cumulative Palmer amaranth flowering in soybean, fallow and corn
+occurred at the end of June. Palmer amaranth reached 10\% flowering in
+soybean at doy 180, which was slightly different from fallow (doy 180.9;
+\emph{P} = 0.01) and corn (doy 181.7; \emph{P} = 0.00). The 50\% Palmer
 amaranth cumulative flowering occurred in July. For example, Palmer
@@
 
-Palmer amaranth cumulative flowering from second transplanting timing
-ranged from mid July to mid September (Figure 2). All cumulative
-flowering comparisons among agroecossystems were significant (\emph{P}
-\textless{} 0.00) Palmer amaranth growing in fallow resulted in earlier
-flowering time compared to soybean and corn. Palmer amaranth growing in
-fallow reached 10\%, 50\%, and 90\% flowering time at day 203.8, 214.4,
-and 232.2, respectively. Palmer amaranth growing in soybean reached 10\%
+Palmer amaranth cumulative flowering at second cohort ranged from mid
+July to mid September (Figure 2). All cumulative flowering comparisons
+among agroecossystems were significant different (\emph{P} \textless{}
+0.00) Palmer amaranth growing in fallow resulted in earlier flowering
+time compared to soybean and corn. Palmer amaranth growing in fallow
+reached 10\%, 50\%, and 90\% flowering time at day 203.8, 214.4, and
+232.2, respectively. Palmer amaranth growing in soybean reached 10\%
 flowering at doy 210.9, which was 6 days prior to corn (\emph{P}-value =
@@
 Palmer amaranth accumulated more biomass when growing in fallow compared
-to Palmer amaranth growing in soybean and corn (Figure 3). At first
-transplanting time, Palmer amaranth biomass was 78.3 g plant-1, 26.4 g
-plant-1 and 14.9 g plant-1 in fallow, soybean and corn, respectively.
-Palmer amaranth produce less biomass at second transplanting time. For
-example, Palmer amaranth growing in fallow resulted in 49.2 g plant-1,
-3.4 g plant-1 in soybean and 1.4 g plant when growing in corn.
+to Palmer amaranth growing in soybean and corn (Figure 3A). At first
+cohort time, Palmer amaranth biomass was 78.3 g plant-1, 26.4 g plant-1
+and 14.9 g plant-1 in fallow, soybean and corn, respectively. Moreover,
+Palmer amaranth produce less biomass at cohort time. Palmer amaranth
+growing in fallow resulted in 49.2 g plant-1, 3.4 g plant-1 in soybean
+and 1.4 g plant when growing in corn.
 
-For example, Palmer amaranth growing at first cohort accumulated 90\%,
-85\% and 31\% more biomass when compared to second cohort at corn,
-soybean and fallow, respectively
-
-In addition, Palmer amaranth growing in corn and soybean at second
-cohort had a similar end-season flowering pattern.
+The Palmer amaranth achieve 92.3 cm tall when growing at first cohort
+timing in corn (Figure 3B).
 
@@
 
-Our study showed the Palmer amaranth flowering pattern, grow and
-development varied within agroecossystems and timing. In general, Palmer
+Our study showed that Palmer amaranth flowering time, biomass and height
+varied within agroecossystems and cohort timing. In general, Palmer
 amaranth produced more biomass and taller plants when growing at first
@@
 nutrients) and conditions (e.g., light) were more timely available for
-both species, crop and weed. Thus, Palmer amaranth growing strategy was
-to mimic the crop species (Figure 4). In contrast, at second cohort,
-Palmer amaranth grow and development was limited due to the crop
-competitive ability at advanced development stages.
+both species, crop and weed. High biomass and taller Palmer amaranth
+plants is likely a weed strategy to compete for light in between crop
+rows in absence of canopy. In such condition, Palmer amaranth showed an
+extraordinary plasticity to adapt upon the agroecosystem. This is
+evident when comparing Palmer amaranth shape, and its extended flowering
+pattern when growing into corn compared to soybean. In corn, Palmer
+amaranth was taller and thinner when growing in corn. The Palmer
+amaranth competition strategy was likely to mimic the crop grow and
+development (Figure 4). These results are concerning as breeding more
+competitive crop varieties is likely to select more competitive Palmer
+amaranth biotypes (Bravo et al., 2017).
 
-Taller Palmer amaranth plants is a strategy to compete for light in
-between crop rows in absence of canopy. For example, at first cohort,
-corn and Palmer amaranth had abundant light to grow and develop, which
-likely aid its extended the flowering pattern. However, at second
-cohort, Palmer amaranth was transplanted when corn canopy was nearly
-closed, which reduced Palmer amaranth competition. As a result, Palmer
-amaranth height and biomass was nearly 50\% and 90\% lower compared to
-its first cohort, respectively. Similar trend was observed to Palmer
-amaranth growing in soybeans. Palmer amaranth growing without crop
-competition produced highest amount of biomass. The Palmer amaranth
-strategy in fallow was to invest biomass in growing plant width and
-height. Nonetheless, Palmer amaranth produced 31\% less biomass in
-second cohort compared to first cohort timing. Therefore, crop
-competition was not the only factor limiting Palmer amaranth growing in
-the second cohort.
+Palmer amaranth grow and development at second cohort was limited due to
+the crop competitive ability at advanced development stages. Palmer
+amaranth was transplanted when corn canopy was nearly closed, which
+reduced Palmer amaranth competitiveness. As a result, Palmer amaranth
+height and biomass was lower compared to its first cohort. Under crop
+canopy, Palmer amaranth flowering pattern was near to similar into corn
+and soybean. Palmer amaranth growing without crop competition produced
+highest amounts of biomass and lowest extended flowering pattern. The
+Palmer amaranth strategy in bareground was to invest biomass in growing
+plant width and height. Nonetheless, Palmer amaranth produced 31\% less
+biomass in second cohort compared to first cohort timing. In a
+bareground study, early emerged Palmer amaranth without competition was
+50\% taller than late emerged plants (Webster and Grey, 2015). These
+results suggests that crop competition is not the only factor limiting
+late Palmer amaranth establishment. The limited growth of Palmer
+amaranth at second cohort is likely a reduced plant response to thermal
+units (e.g, growing degree days). It is hypothesize that reduced day
+length contributed to smaller plants at second cohort as well as shorter
+flowering period. A study in North Carolina and Illinois predicted that
+less than 10\% Palmer amaranth seedlings emergence occurred after June
+(Piskackova et al., 2021). In addition, Palmer amaranth negative impact
+on soybean (Korres et al., 2020) and cotton (Webster and Grey, 2015)
+yields was higher when plants were established near to crop planting.
 
-It was not evaluate seed production due to Palmer amaranth harvest at
-early flowering. Nonetheless, it is reported a strong positive
+Seed production was not evaluated due to Palmer amaranth harvest at
+flowering time. Nonetheless, it is well documented a strong positive
 correlation between Palmer amaranth biomass and seed production
 (Schwartz et al., 2016; Spaunhorst et al., 2018). I our study, Palmer
-amaranth growing at first cohort accumulated nearly 50\% more overall
+amaranth growing at first cohort accumulated an overall 50\% more
 biomass when compared to second cohort. Therefore, Palmer amaranth
 plants growing in the second cohort is likely to produce less seeds
-regardless the agroecossystem. Seed production and deposition in the
-seedbank is a key factor for species perpetuation (Menges, 1987). Palmer
-amaranth germination is influenced by soil depth and seed viable seeds
-were fount after 36 months buried in the seedbank (Sosnoskie et al.,
-2013). Therefore, tactics that promote crop advantage against Palmer
-amaranth, including early crop planting, preemergence applied herbicide,
-and crop residue (e.g.~cover crops) can minimize the negative impact of
-Palmer amaranth in agroecossystems.
+regardless the agroecossystem. Our observation is consistent with the
+findings that first Palmer amaranth cohort produced 50\% more seeds per
+plant than Palmer amaranth plants established six weeks later in
+bareground (Webster and Grey, 2015). Nonetheless, seed production at
+second cohort is likely to replenish the soil seedbank. Seed production
+and deposition in the seedbank is a key factor for species perpetuation
+(Menges, 1987). Palmer amaranth can produce hundred thousands seeds per
+plant (Schwartz et al., 2016; Keeley et al., 1987), and stay viable
+buried in the seedbank for at least 36 months (Sosnoskie et al., 2013).
 
-Palmer amaranth has shown to be responsive to to thermal units (e.g,
-growing degree days). A study in North Carolina demonstrated that Palmer
-amaranth emergence in response to soil temperature, with seedlings
-emerging before July prior to crop canopy closure (Piskackova et al.,
-2021). In addition, in a bareground study, Palmer emergence emergence
-pattern in Nebraska was influenced by soil management. Early tillage
-resulted in 90\% Palmer amaranth emergence in June
-
-can cause significant soybean yield loss (Korres et al., 2019)
+Our study demonstrated the short-term Palmer amaranth plasticity to grow
+and develop into cropping-systems. Palmer amaranth management priority
+should focus on minimizing Palmer amaranth dispersal into new
+geographies. Early-season management programs would have a large
+negative effect in Palmer amaranth growth and development. Tactics that
+promote crop advantage against Palmer amaranth, including early crop
+planting, crop rotation, plant width, preemergence applied herbicide
+(Sanctis et al., 2021), and crop residue (e.g.~cover crops) would
+minimize the negative impact of Palmer amaranth in agroecossystems.
 
@@
 
+\leavevmode\hypertarget{ref-bravo2017}{}%
+Bravo, W., Leon, R. G., Ferrell, J. A., Mulvaney, M. J., and Wood, C. W.
+(2017). Differentiation of {Life}-{History Traits} among {Palmer
+Amaranth Populations} ({Amaranthus} palmeri) and {Its Relation} to
+{Cropping Systems} and {Glyphosate Sensitivity}. \emph{Weed Science} 65,
+339--349.
+doi:\href{https://doi.org/10.1017/wsc.2017.14}{10.1017/wsc.2017.14}.
+
 \leavevmode\hypertarget{ref-briscoerunquist2019}{}%
@@
 doi:\href{https://doi.org/10.1038/s41598-018-38054-9}{10.1038/s41598-018-38054-9}.
+
+\leavevmode\hypertarget{ref-chahal2021}{}%
+Chahal, P. S., Barnes, E. R., and Jhala, A. J. (2021). Emergence pattern
+of {Palmer} amaranth ({Amaranthus} palmeri) influenced by tillage
+timings and residual herbicides. \emph{Weed Technology} 35, 433--439.
+doi:\href{https://doi.org/10.1017/wet.2020.136}{10.1017/wet.2020.136}.
 
@@
 doi:\href{https://doi.org/10.1186/s13104-021-05622-8}{10.1186/s13104-021-05622-8}.
+
+\leavevmode\hypertarget{ref-davis2015}{}%
+Davis, A. S., Schutte, B. J., Hager, A. G., and Young, B. G. (2015).
+Palmer {Amaranth} ({Amaranthus} palmeri) {Damage Niche} in {Illinois
+Soybean Is Seed Limited}. \emph{Weed Science} 63, 658--668.
+doi:\href{https://doi.org/10.1614/WS-D-14-00177.1}{10.1614/WS-D-14-00177.1}.
 
@@
 
-\leavevmode\hypertarget{ref-korres2019}{}%
-Korres, N. E., Norsworthy, J. K., and Mauromoustakos, A. (2019). Effects
-of {Palmer Amaranth} ({Amaranthus} palmeri) {Establishment Time} and
-{Distance} from the {Crop Row} on {Biological} and {Phenological
-Characteristics} of the {Weed}: {Implications} on {Soybean Yield}.
-\emph{Weed Science} 67, 126--135.
-doi:\href{https://doi.org/10.1017/wsc.2018.84}{10.1017/wsc.2018.84}.
+\leavevmode\hypertarget{ref-korres2020}{}%
+Korres, N. E., Norsworthy, J. K., Mauromoustakos, A., and Williams, M.
+M. (2020). Soybean density and {Palmer} amaranth ({Amaranthus} palmeri)
+establishment time: Effects on weed biology, crop yield, and economic
+returns. \emph{Weed Science} 68, 467--475.
+doi:\href{https://doi.org/10.1017/wsc.2020.41}{10.1017/wsc.2020.41}.
 
@@
 doi:\href{https://doi.org/10.1007/s13593-020-00631-6}{10.1007/s13593-020-00631-6}.
+
+\leavevmode\hypertarget{ref-macrae2013}{}%
+MacRae, A. W., Webster, T. M., Sosnoskie, L. M., Culpepper, A. S., and
+Kichler, J. M. (2013). Cotton {Yield Loss Potential} in {Response} to
+{Length} of {Palmer Amaranth} ({Amaranthus} palmeri) {Interference}. 17,
+6.
 
@@
 
+\leavevmode\hypertarget{ref-sanctis2021}{}%
+Sanctis, J. H. S. de, Barnes, E. R., Knezevic, S. Z., Kumar, V., and
+Jhala, A. J. (2021). Residual herbicides affect critical time of
+{Palmer} amaranth removal in soybean. \emph{Agronomy Journal} 113,
+1920--1933.
+doi:\href{https://doi.org/10.1002/agj2.20615}{10.1002/agj2.20615}.
+
 \leavevmode\hypertarget{ref-sauer1957}{}%
@@
 
+\leavevmode\hypertarget{ref-webster2015}{}%
+Webster, T. M., and Grey, T. L. (2015). Glyphosate-{Resistant Palmer
+Amaranth} ({Amaranthus} palmeri) {Morphology}, {Growth}, and {Seed
+Production} in {Georgia}. \emph{Weed Science} 63, 264--272.
+doi:\href{https://doi.org/10.1614/WS-D-14-00051.1}{10.1614/WS-D-14-00051.1}.
+
 \leavevmode\hypertarget{ref-yu2021}{}%

### 2021-08-05 Manuscript/Manuscript.Rmd
JH

@@
 In the northern states, Palmer amaranth impact on the agroecossystem is recent.
-Studies investigating Palmer amaranth in those locations is limited due to the plant classification as noxious weed [@yu2021].
+Studies investigating Palmer amaranth in those locations is limited due to the plant classification as noxious weed species [@yu2021].
 Nonetheless, the continuous Palmer amaranth dispersal and potential establishment across northern United States is concerning and warrant investigations on species morphology in such environments.
@@
 Analysis of variance was performed with *anova* function from "car" package [@fox2018].
-Marginal means and compact letter display were estimated with *emmeans* and *cld* from packages "emmeans" and multcomp [@hothorn2008].
+Marginal means and compact letter display were estimated with *emmeans* and *cld* from packages "emmeans" and "multcomp" [@hothorn2008].
 
@@
 Palmer amaranth growing in corn resulted in a longer flowering pattern compared to fallow and soybean at first cohort (Figure 2A).
-The 10% cumulative Palmer amaranth flowering in soybean, fallow and corn occurred at the end of June.
+Nonetheless, the 10% cumulative Palmer amaranth flowering in soybean, fallow and corn occurred at the end of June.
 Palmer amaranth reached 10% flowering in soybean at doy 180, which was slightly different from fallow (doy 180.9; *P* = 0.01) and corn (doy 181.7; *P* = 0.00).
 The 50% Palmer amaranth cumulative flowering occurred in July.
-For example, Palmer amaranth reached 50% in fallow at doy 193.4, followed by soybean (doy 194.8), corn (doy 206.6).
+For example, Palmer amaranth reached 50% flowering in fallow at doy 193.4, followed by soybean (doy 194.8), corn (doy 206.6).
 Similar trend was observed at 90% Palmer amaranth cumulative flowering.
@@
 
-Palmer amaranth cumulative flowering at second cohort ranged from mid July to mid September (Figure 2).
-All cumulative flowering comparisons among agroecossystems were significant different (*P* < 0.00) Palmer amaranth growing in fallow resulted in earlier flowering time compared to soybean and corn.
+Palmer amaranth cumulative flowering at second cohort ranged from mid July to mid September (Figure 2B).
+Palmer amaranth growing in fallow resulted in earlier flowering time compared to soybean and corn.
 Palmer amaranth growing in fallow reached 10%, 50%, and 90% flowering time at day 203.8, 214.4, and 232.2, respectively.
@@
 At first cohort, resources (e.g., soil nutrients) and conditions (e.g., light) were more timely available for both species, crop and weed.
-High biomass and taller Palmer amaranth plants is likely a weed strategy to compete for light in between crop rows in absence of canopy.
+High biomass and taller Palmer amaranth plants are likely a weed strategy to compete for light in between crop rows in absence of canopy.
 In such condition, Palmer amaranth showed an extraordinary plasticity to adapt upon the agroecosystem.
 This is evident when comparing Palmer amaranth shape, and its extended flowering pattern when growing into corn compared to soybean.
-In corn, Palmer amaranth was taller and thinner when growing in corn.
+Palmer amaranth was taller and thinner when growing within corn.
 The Palmer amaranth competition strategy was likely to mimic the crop grow and development (Figure 4).
@@
 As a result, Palmer amaranth height and biomass was lower compared to its first cohort.
-Under crop canopy, Palmer amaranth flowering pattern was near to similar into corn and soybean.
-Palmer amaranth growing without crop competition produced highest amounts of biomass and lowest extended flowering pattern.
+Under crop canopy, Palmer amaranth flowering pattern was near to similar in corn and soybean.
+Palmer amaranth growing without crop competition produced highest amounts of biomass and less extended flowering pattern.
 The Palmer amaranth strategy in bareground was to invest biomass in growing plant width and height.
@@
 Palmer amaranth can produce hundred thousands seeds per plant [@schwartz2016; @keeley1987], and stay viable buried in the seedbank for at least 36 months [@sosnoskie2013].
+
+Palmer amaranth flowering pattern was variable
 

### 2021-08-05 Manuscript/Manuscript.tex
JH

@@
 agroecossystem is recent. Studies investigating Palmer amaranth in those
-locations is limited due to the plant classification as noxious weed (Yu
-et al., 2021). Nonetheless, the continuous Palmer amaranth dispersal and
-potential establishment across northern United States is concerning and
-warrant investigations on species morphology in such environments.
-Understanding Palmer amaranth biology and growing strategies under
-different agroecossystems can enhance our knowledge on species
-adaptation. It can also aid on designing proactive and ecological
-tactics to limit the species range expansion, reduce its negative
-impact, and design resilient and sustainable farming systems (MacLaren
-et al., 2020). Therefore, the objective of this study was to investigate
-the flowering pattern, biomass production, and height of Palmer amaranth
-growing under corn, soybean and fallow at two timings across five
-locations in the United States Midwest.
+locations is limited due to the plant classification as noxious weed
+species (Yu et al., 2021). Nonetheless, the continuous Palmer amaranth
+dispersal and potential establishment across northern United States is
+concerning and warrant investigations on species morphology in such
+environments. Understanding Palmer amaranth biology and growing
+strategies under different agroecossystems can enhance our knowledge on
+species adaptation. It can also aid on designing proactive and
+ecological tactics to limit the species range expansion, reduce its
+negative impact, and design resilient and sustainable farming systems
+(MacLaren et al., 2020). Therefore, the objective of this study was to
+investigate the flowering pattern, biomass production, and height of
+Palmer amaranth growing under corn, soybean and fallow at two timings
+across five locations in the United States Midwest.
 
@@
 Palmer amaranth growing in corn resulted in a longer flowering pattern
-compared to fallow and soybean at first cohort (Figure 2A). The 10\%
-cumulative Palmer amaranth flowering in soybean, fallow and corn
-occurred at the end of June. Palmer amaranth reached 10\% flowering in
-soybean at doy 180, which was slightly different from fallow (doy 180.9;
-\emph{P} = 0.01) and corn (doy 181.7; \emph{P} = 0.00). The 50\% Palmer
-amaranth cumulative flowering occurred in July. For example, Palmer
-amaranth reached 50\% in fallow at doy 193.4, followed by soybean (doy
-194.8), corn (doy 206.6). Similar trend was observed at 90\% Palmer
-amaranth cumulative flowering. Palmer amaranth growing in corn reached
-90\% flowering at doy 252.6 (early September), which was 37.8 and 32.2
-days after Palmer amaranth 90\% flowering in fallow and soybean,
-respectively.
+compared to fallow and soybean at first cohort (Figure 2A). Nonetheless,
+the 10\% cumulative Palmer amaranth flowering in soybean, fallow and
+corn occurred at the end of June. Palmer amaranth reached 10\% flowering
+in soybean at doy 180, which was slightly different from fallow (doy
+180.9; \emph{P} = 0.01) and corn (doy 181.7; \emph{P} = 0.00). The 50\%
+Palmer amaranth cumulative flowering occurred in July. For example,
+Palmer amaranth reached 50\% flowering in fallow at doy 193.4, followed
+by soybean (doy 194.8), corn (doy 206.6). Similar trend was observed at
+90\% Palmer amaranth cumulative flowering. Palmer amaranth growing in
+corn reached 90\% flowering at doy 252.6 (early September), which was
+37.8 and 32.2 days after Palmer amaranth 90\% flowering in fallow and
+soybean, respectively.
 
 Palmer amaranth cumulative flowering at second cohort ranged from mid
-July to mid September (Figure 2). All cumulative flowering comparisons
-among agroecossystems were significant different (\emph{P} \textless{}
-0.00) Palmer amaranth growing in fallow resulted in earlier flowering
-time compared to soybean and corn. Palmer amaranth growing in fallow
-reached 10\%, 50\%, and 90\% flowering time at day 203.8, 214.4, and
-232.2, respectively. Palmer amaranth growing in soybean reached 10\%
-flowering at doy 210.9, which was 6 days prior to corn (\emph{P}-value =
-0.00). Similar trend was observed at 50\% flowering, whereas Palmer
-amaranth reached 50\% flowering in corn (doy 233.0) 4 days after
-soybeans (doy 228.9; \emph{P} = 0.00). The 90\% Palmer amaranth
-cumulative flowering occurred at same day in corn (260.9) and soybean
-(260.5; \emph{P} = 0.66).
+July to mid September (Figure 2B). Palmer amaranth growing in fallow
+resulted in earlier flowering time compared to soybean and corn. Palmer
+amaranth growing in fallow reached 10\%, 50\%, and 90\% flowering time
+at day 203.8, 214.4, and 232.2, respectively. Palmer amaranth growing in
+soybean reached 10\% flowering at doy 210.9, which was 6 days prior to
+corn (\emph{P}-value = 0.00). Similar trend was observed at 50\%
+flowering, whereas Palmer amaranth reached 50\% flowering in corn (doy
+233.0) 4 days after soybeans (doy 228.9; \emph{P} = 0.00). The 90\%
+Palmer amaranth cumulative flowering occurred at same day in corn
+(260.9) and soybean (260.5; \emph{P} = 0.66).
 
@@
 both species, crop and weed. High biomass and taller Palmer amaranth
-plants is likely a weed strategy to compete for light in between crop
+plants are likely a weed strategy to compete for light in between crop
 rows in absence of canopy. In such condition, Palmer amaranth showed an
@@
 evident when comparing Palmer amaranth shape, and its extended flowering
-pattern when growing into corn compared to soybean. In corn, Palmer
-amaranth was taller and thinner when growing in corn. The Palmer
-amaranth competition strategy was likely to mimic the crop grow and
-development (Figure 4). These results are concerning as breeding more
-competitive crop varieties is likely to select more competitive Palmer
-amaranth biotypes (Bravo et al., 2017).
+pattern when growing into corn compared to soybean. Palmer amaranth was
+taller and thinner when growing within corn. The Palmer amaranth
+competition strategy was likely to mimic the crop grow and development
+(Figure 4). These results are concerning as breeding more competitive
+crop varieties is likely to select more competitive Palmer amaranth
+biotypes (Bravo et al., 2017).
 
@@
 height and biomass was lower compared to its first cohort. Under crop
-canopy, Palmer amaranth flowering pattern was near to similar into corn
+canopy, Palmer amaranth flowering pattern was near to similar in corn
 and soybean. Palmer amaranth growing without crop competition produced
-highest amounts of biomass and lowest extended flowering pattern. The
+highest amounts of biomass and less extended flowering pattern. The
 Palmer amaranth strategy in bareground was to invest biomass in growing

### 2021-08-09 Manuscript/Manuscript.Rmd
as

@@
 
+```{r fig5, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../data analysis/figures/Figure 4.png")
+```
+
 # Discussion {-}
@@
 ```
+
+```{r, Figure-5, ref.label = "fig5", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "120mm", out.height = "180mm"}
+# You can also refer to code chunks from above to place figures at the bottom.
+```

### 2021-08-09 Manuscript/Manuscript.tex
as

@@
 compact letter display were estimated with \emph{emmeans} and \emph{cld}
-from packages ``emmeans'' and multcomp (Hothorn et al., 2008).
+from packages ``emmeans'' and ``multcomp'' (Hothorn et al., 2008).
 
@@
 
+Palmer amaranth flowering pattern was variable
+
 Our study demonstrated the short-term Palmer amaranth plasticity to grow
@@
 \caption{Harvest Palmer amaranth plants at 40 days after first transplant timing. From left to right, Palmer amaranth growing in fallow, soybean and corn in Arlington, Wisconsin}\label{fig:Figure-4}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=120mm,height=180mm]{../data analysis/figures/Figure 4}
+
+}
+
+\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)}\label{fig:Figure-5}
 \end{figure}

### 2021-08-10 data analysis/new_analyses.Rmd
as

@@
 ```
+
+# GAM try
+
+```{r}
+data4 %>% 
+  filter(!is.na(doyh)) %>% 
+  summarise(min(doyh), max(doyh))
+```
+
+```{r}
+library(mgcv)
+
+data2 %>% 
+    mutate(gender_A = case_when(
+    gender == "m" ~ 0,
+    gender == "f" ~ 1,
+    TRUE ~ NA_real_
+  )) %>% 
+  filter(!is.na(gender)) %>% 
+  mutate_if(is_character, as.factor) %>% 
+  filter(month == "June") -> june
+```
+
+```{r}
+gam_doy1 <- gam(gender_A ~ 
+                 s(doyh, by = crop) + crop + 
+                 s(location, by = year, bs = "re"), 
+               family = "binomial", method = "REML", 
+               data = june)
+summary(gam_doy1)
+plogis(coef(gam_doy1)[1])
+
+plot(gam_doy1, residuals = FALSE, pch = 1, ylim = c(0,1), cex = 1,
+     trans = plogis,  
+     seWithMean = TRUE,
+     shade = TRUE, 
+     shift = coef(gam_doy)[1],
+     shade.col = "lightgreen")
+
+mgcv::predict.gam(gam_doy1, type = "response", se.fit = TRUE)
+```
+
+```{r}
+visreg(gam_doy2, "doyh", "crop", 
+       gg= TRUE, overlay = TRUE,
+       scale="response",
+       jitter=TRUE, 
+#       rug=2
+#       fill.par = list(fill = c("#30123bFF"), alpha = 0.1),
+#       line=list(color = c("#30123bFF"), alpha = 0.1)
+       ) +
+  labs(x = "Day of year", y = " Probability (%)") +
+  facet_grid(scales = "free") +
+  scale_y_continuous(limits = c(0, 1)) +
+  scale_color_viridis_d(option = "C") +
+  scale_fill_viridis_d(option = "C", alpha = 0.05) +
+  annotate("text", x = 205, y = 0.90, label = "Second cohort",
+            size = 5, color = 1) +
+  theme_test() 
+```
+
+```{r}
+gam1 <- gam(gender_A ~ 
+                  s(doyh, by = crop) + 
+                  s(height, by = crop) + 
+                  s(weight, by = crop) + 
+                  month +
+                  s(location, by = year, bs = "re"), 
+               family = "binomial", method = "REML", 
+               data = data4)
+
+gam2 <- gam(gender_A ~ 
+                  s(doyh) + 
+                  s(height, by = crop) + 
+                  s(weight, by = crop) + 
+#                  month +
+                  s(location, by = year, bs = "re"), 
+               family = "binomial", method = "REML", 
+               data = data4)
+plot(gam2, residuals = FALSE, pch = 1, ylim = c(0,1), cex = 1,
+     trans = plogis,  
+     seWithMean = TRUE,
+     shade = TRUE, 
+     shift = coef(gam_doy)[1],
+     shade.col = "lightgreen")
+anova(gam1, gam2, test = "F")
+summary(gam2)
+```
+
+```{r}
+gam_doy2 <- gam(gender_A ~ 
+                 s(height, by = crop) + 
+                  s(height, by = crop) + 
+                  s(weight, by = crop) + 
+                 s(location, by = year, bs = "re"), 
+               family = "binomial", method = "REML", 
+               data = data4 %>% filter(month == "July"))
+summary(gam_doy2)
+plogis(coef(gam_doy2)[2])
+
+plot(gam_doy2, residuals = FALSE, pch = 1, ylim = c(0,1), cex = 1,
+     trans = plogis,  
+     seWithMean = TRUE,
+     shade = TRUE, 
+     shift = coef(gam_doy)[1],
+     shade.col = "lightgreen")
+
+mgcv::predict.gam(gam_doy2, type = "response", se.fit = TRUE)
+```
+
+```{r}
+visreg(gam_doy2, "doyh", "crop", 
+       gg= TRUE, overlay = TRUE,
+       scale="response",
+       jitter=TRUE, 
+#       rug=2
+#       fill.par = list(fill = c("#30123bFF"), alpha = 0.1),
+#       line=list(color = c("#30123bFF"), alpha = 0.1)
+       ) +
+  labs(x = "Day of year", y = " Probability (%)") +
+  facet_grid(scales = "free") +
+  scale_y_continuous(limits = c(0, 1)) +
+  scale_color_viridis_d(option = "C") +
+  scale_fill_viridis_d(option = "C", alpha = 0.05) +
+  annotate("text", x = 205, y = 0.90, label = "Second cohort",
+            size = 5, color = 1) +
+  theme_test() 
+```
+
+```{r}
+data3 %>% 
+  ggplot(aes(x = doyh, y = gender_A)) +
+  geom_point(alpha = 0.1, aes(color = crop)) +
+  geom_emoji(data = cohortf1, aes(x = x, y = y), emoji = "2640") +
+  geom_emoji(data = cohortm1, aes(x = x, y = y), emoji = "2642") +
+  geom_smooth(aes(color = crop), 
+              method = "glm", 
+              se = TRUE, alpha = 0.05,
+              method.args=list(family="binomial"),
+              stat = "smooth") +
+  theme_test(base_family = "roboto") +
+  facet_grid(~month) +
+#  scale_color_manual(values = c(2,1)) +
+#  scale_fill_manual(values = c(2,1)) +
+  scale_color_viridis_d(option = "C") +
+  scale_fill_viridis_d(option = "C", alpha = 0.05) +
+  scale_x_continuous(limits = c(170, 260), 
+                     breaks = seq(170, 260, 15)) +
+  scale_y_continuous( breaks = seq(0, 1, by = 0.2)) +
+  labs(x = "Day of year", y = NULL) +
+  geom_text(data = first, mapping = aes(x = x, y = y), label = first$label,
+            size = 10, color = 1) +
+  geom_text(data = second, mapping = aes(x = x, y = y), label = second$label,
+            size = 10, color = 2) +
+  theme(legend.position = "none",
+        axis.title = element_markdown(size = 35),
+        axis.text = element_markdown(size = 25),
+        strip.text = element_markdown(face = "bold", size = 30)) -> pl_doyh
+#  xlim(175, 270)
+
+ggsave("figures/fig.png", width = 9, height = 4)
+```

### 2021-08-10 Manuscript/Manuscript.Rmd
as

@@
 Marginal means and compact letter display were estimated with *emmeans* and *cld* from packages "emmeans" and "multcomp" [@hothorn2008].
+
+Palmer amaranth gender was fitted to a binary logistic regression [@bangdiwala2018].
+Binary logistic regression is used for predicting binary classes, such gender of Palmer amaranth plants.
+The whole dataset was splitted into 80% train and 20% test data.
+The 80% train is used for the model training and the rest 20% is used for checking how the model generalized on unseen dataset.
+With 80% dataset, a binary response variable, male (0) and female (1), was fitted to a generalized linear model (*glm* function) including day of year harvest, height, weight, crop and month as independent variables.
+The model family was binomial with a logit function.
+The model fit was assessed through pseudo R squared values (McFadden, Cox and Snell, Cragg and Uhler) and likelihood ratio using *nagelkerke* function ("rcompanion" package).
+The marginal effects computation was performed with Average Marginal Effects at every observed value of x and average across the results (AMEs) [@leeper2017] using *margins* function from "margins" package.
+The rest 20% dataset was predicted using *predict* function with a cutoff estimation for male or female using *performance* function.
 

### 2021-08-11 Manuscript/Manuscript.Rmd
ko

@@
 
-Palmer amaranth height and biomass were performed with a linear mixed model using *lmer* function from "lme4" package [@bates2015].
-Plant height and biomass were transformed to meet model assumption of normality.
-In the model, agroecosystem (crop, soybean, fallow) was the fixed effect and year nested with location the random effects.
-Analysis of variance was performed with *anova* function from "car" package [@fox2018].
-Marginal means and compact letter display were estimated with *emmeans* and *cld* from packages "emmeans" and "multcomp" [@hothorn2008].
-
 Palmer amaranth gender was fitted to a binary logistic regression [@bangdiwala2018].
 Binary logistic regression is used for predicting binary classes, such gender of Palmer amaranth plants.
+Prior to the analysis, all missing values were removed from the dataset.
+Also, data from Grant was not used in this analysis due to the uniform plant harvesting at that location.
 The whole dataset was splitted into 80% train and 20% test data.
@@
 The rest 20% dataset was predicted using *predict* function with a cutoff estimation for male or female using *performance* function.
+The model quality prediction from the classification algorithm was measure with precision (*precision* function), recall (*recall* function) and F1 score (*f_meas* function) using "yardstick" package.
+The precision determines the accuracy of positive predictions (female plants), recall determines the fraction of positives that were correctly identified, and F1 score is a weighted harmonic mean of precision and recall with the best score of 1 and the worst score of 0.
+F1 score conveys the balance between the precision and the recall.
+
+Palmer amaranth height and biomass were performed with a linear mixed model using *lmer* function from "lme4" package [@bates2015].
+Plant height and biomass were transformed to meet model assumption of normality.
+In the model, agroecosystem (crop, soybean, fallow) was the fixed effect and year nested with location the random effects.
+Analysis of variance was performed with *anova* function from "car" package [@fox2018].
+Marginal means and compact letter display were estimated with *emmeans* and *cld* from packages "emmeans" and "multcomp" [@hothorn2008].
 
@@
 
-## Cumulative flowering {-}
+## Palmer amaranth cumulative flowering {-}
 
@@
 
-## Height and biomass {-}
+## Palmer amaranth gender {-}
+
+The model fit was 0.11, 0.14, 0.18 with using pseudo R squared test from McFadden, Cox and Snell, and Cragg and Uhler, respectively.
+The likelihood ratio test showed a p-value of < 0.00.
+The marginal effects showed that increasing a weight unit by 1 g increases the probability of having a female plant by 0.08%.
+Similar trend is observed to height as well as doy, whereas the probability of being female increase by 0.16% and 0.4% when a unit of height (cm) and day increases.
+
+The model evaluation accuracy in the unseen 20% dataset was 0.61 with a cutoff value for female and male plants of 0.49.
+The model classification showed a precision of 0.59, recall of 0.76, and a F1 score of 0.66.
+In addition, the area under the curve was 0.61.
+
+## Palmer amaranth height and biomass {-}
 
 Palmer amaranth accumulated more biomass when growing in fallow compared to Palmer amaranth growing in soybean and corn (Figure 3A).
-At first cohort time, Palmer amaranth biomass was 78.3 g plant-1, 26.4 g plant-1 and 14.9 g plant-1 in fallow, soybean and corn, respectively.
-Moreover, Palmer amaranth produce less biomass at cohort time.
-Palmer amaranth growing in fallow resulted in 49.2 g plant-1, 3.4 g plant-1 in soybean and 1.4 g plant when growing in corn.
+At first cohort time, Palmer amaranth biomass was 75.5 g plant-1, 28.3 g plant-1 and 16.3 g plant-1 in fallow, soybean and corn, respectively.
+At second cohort timing, Palmer amaranth produced 62.6 g plant in fallow, followed by 6.3 g plant in soybean, and 1.4 g plant in corn.
 
-The Palmer amaranth achieve 92.3 cm tall when growing at first cohort timing in corn (Figure 3B).
+Palmer amaranth height was more uniform across cohort timings, except when growing in corn (Figure 3B).
+Palmer amaranth achieve achieve 69.2 cm tall when growing at first cohort timing in bareground, which was not different to 70.7 cm tall at second cohort timing (P = 0.74).
+In addition, no difference in Palmer amaranth height (69.3 cm) was found in soybean at first cohort compared to fallow (P > 0.75).
+At second cohort in soybean, Palmer amaranth was near to 10 cm lower compared to its first cohort time (P = 0.04).
+The tallest and smallest Palmer amaranth plants were found in corn.
+Palmer amaranth reached 85.2 cm tall at first cohort and 38.2 cm at second cohort timings.
 
@@
 The Palmer amaranth strategy in bareground was to invest biomass in growing plant width and height.
-Nonetheless, Palmer amaranth produced 31% less biomass in second cohort compared to first cohort timing.
+Nonetheless, Palmer amaranth produced 21% less biomass in second cohort compared to first cohort timing.
 In a bareground study, early emerged Palmer amaranth without competition was 50% taller than late emerged plants [@webster2015].
@@
 Nonetheless, it is well documented a strong positive correlation between Palmer amaranth biomass and seed production [@schwartz2016; @spaunhorst2018].
-I our study, Palmer amaranth growing at first cohort accumulated an overall 50% more biomass when compared to second cohort.
+I our study, Palmer amaranth growing at first cohort accumulated an overall 36% more biomass when compared to second cohort.
 Therefore, Palmer amaranth plants growing in the second cohort is likely to produce less seeds regardless the agroecossystem.
@@
 
-Palmer amaranth flowering pattern was variable
+Palmer amaranth flowering pattern was slightly influenced by cropping systems and cohort timings.
+Palmer amaranth growing in fallow and corn resulted in the overall shortest and longest flowering pattern.
+When growing in soybean, Palmer amaranth flowering pattern was similar to bareground and corn agroecossystems at first and second cohort timings, respectively.
+It seems that Palmer amaranth cohorts tend to have an extended flowering pattern when in high competition.
+The agroecossystems influenced Palmer amaranth gender.
+When growing in corn, the probability of having Palmer amaranth female plants reduced 18.1%.
 
@@
 Early-season management programs would have a large negative effect in Palmer amaranth growth and development.
-Tactics that promote crop advantage against Palmer amaranth, including early crop planting, crop rotation, plant width, preemergence applied herbicide [@sanctis2021], and crop residue (e.g.
+Tactics that promote early-season crop advantage against Palmer amaranth, including early crop planting, crop rotation, plant width, preemergence applied herbicide [@sanctis2021], and crop residue (e.g.
 cover crops) would minimize the negative impact of Palmer amaranth in agroecossystems.

### 2021-08-11 Manuscript/Manuscript.tex
ko

@@
 
+Palmer amaranth gender was fitted to a binary logistic regression
+(Bangdiwala, 2018). Binary logistic regression is used for predicting
+binary classes, such gender of Palmer amaranth plants. Prior to the
+analysis, all missing values were removed from the dataset. Also, data
+from Grant was not used in this analysis due to the uniform plant
+harvesting at that location. The whole dataset was splitted into 80\%
+train and 20\% test data. The 80\% train is used for the model training
+and the rest 20\% is used for checking how the model generalized on
+unseen dataset. With 80\% dataset, a binary response variable, male (0)
+and female (1), was fitted to a generalized linear model (\emph{glm}
+function) including day of year harvest, height, weight, crop and month
+as independent variables. The model family was binomial with a logit
+function. The model fit was assessed through pseudo R squared values
+(McFadden, Cox and Snell, Cragg and Uhler) and likelihood ratio using
+\emph{nagelkerke} function (``rcompanion'' package). The marginal
+effects computation was performed with Average Marginal Effects at every
+observed value of x and average across the results (AMEs) (Leeper, 2017)
+using \emph{margins} function from ``margins'' package. The rest 20\%
+dataset was predicted using \emph{predict} function with a cutoff
+estimation for male or female using \emph{performance} function. The
+model quality prediction from the classification algorithm was measure
+with precision (\emph{precision} function), recall (\emph{recall}
+function) and F1 score (\emph{f\_meas} function) using ``yardstick''
+package. The precision determines the accuracy of positive predictions
+(female plants), recall determines the fraction of positives that were
+correctly identified, and F1 score is a weighted harmonic mean of
+precision and recall with the best score of 1 and the worst score of 0.
+F1 score conveys the balance between the precision and the recall.
+
 Palmer amaranth height and biomass were performed with a linear mixed
@@
 
-\hypertarget{cumulative-flowering}{%
-\subsection*{Cumulative flowering}\label{cumulative-flowering}}
-\addcontentsline{toc}{subsection}{Cumulative flowering}
+\hypertarget{palmer-amaranth-cumulative-flowering}{%
+\subsection*{Palmer amaranth cumulative
+flowering}\label{palmer-amaranth-cumulative-flowering}}
+\addcontentsline{toc}{subsection}{Palmer amaranth cumulative flowering}
 
@@
 
-\hypertarget{height-and-biomass}{%
-\subsection*{Height and biomass}\label{height-and-biomass}}
-\addcontentsline{toc}{subsection}{Height and biomass}
+\hypertarget{palmer-amaranth-gender}{%
+\subsection*{Palmer amaranth gender}\label{palmer-amaranth-gender}}
+\addcontentsline{toc}{subsection}{Palmer amaranth gender}
+
+The model fit was 0.11, 0.14, 0.18 with using pseudo R squared test from
+McFadden, Cox and Snell, and Cragg and Uhler, respectively. The
+likelihood ratio test showed a p-value of \textless{} 0.00. The marginal
+effects showed that increasing a weight unit by 1 g increases the
+probability of having a female plant by 0.08\%. Similar trend is
+observed to height as well as doy, whereas the probability of being
+female increase by 0.16\% and 0.4\% when a unit of height (cm) and day
+increases.
+
+The model evaluation accuracy in the unseen 20\% dataset was 0.61 with a
+cutoff value for female and male plants of 0.49. The model
+classification showed a precision of 0.59, recall of 0.76, and a F1
+score of 0.66. In addition, the area under the curve was 0.61.
+
+\hypertarget{palmer-amaranth-height-and-biomass}{%
+\subsection*{Palmer amaranth height and
+biomass}\label{palmer-amaranth-height-and-biomass}}
+\addcontentsline{toc}{subsection}{Palmer amaranth height and biomass}
 
@@
 to Palmer amaranth growing in soybean and corn (Figure 3A). At first
-cohort time, Palmer amaranth biomass was 78.3 g plant-1, 26.4 g plant-1
-and 14.9 g plant-1 in fallow, soybean and corn, respectively. Moreover,
-Palmer amaranth produce less biomass at cohort time. Palmer amaranth
-growing in fallow resulted in 49.2 g plant-1, 3.4 g plant-1 in soybean
-and 1.4 g plant when growing in corn.
+cohort time, Palmer amaranth biomass was 75.5 g plant-1, 28.3 g plant-1
+and 16.3 g plant-1 in fallow, soybean and corn, respectively. At second
+cohort timing, Palmer amaranth produced 62.6 g plant in fallow, followed
+by 6.3 g plant in soybean, and 1.4 g plant in corn.
 
-The Palmer amaranth achieve 92.3 cm tall when growing at first cohort
-timing in corn (Figure 3B).
+Palmer amaranth height was more uniform across cohort timings, except
+when growing in corn (Figure 3B). Palmer amaranth achieve achieve 69.2
+cm tall when growing at first cohort timing in bareground, which was not
+different to 70.7 cm tall at second cohort timing (P = 0.74). In
+addition, no difference in Palmer amaranth height (69.3 cm) was found in
+soybean at first cohort compared to fallow (P \textgreater{} 0.75). At
+second cohort in soybean, Palmer amaranth was near to 10 cm lower
+compared to its first cohort time (P = 0.04). The tallest and smallest
+Palmer amaranth plants were found in corn. Palmer amaranth reached 85.2
+cm tall at first cohort and 38.2 cm at second cohort timings.
 
@@
 
+\leavevmode\hypertarget{ref-bangdiwala2018}{}%
+Bangdiwala, S. I. (2018). Regression: Binary logistic.
+\emph{International Journal of Injury Control and Safety Promotion} 25,
+336--338.
+doi:\href{https://doi.org/10.1080/17457300.2018.1486503}{10.1080/17457300.2018.1486503}.
+
 \leavevmode\hypertarget{ref-bates2015}{}%
@@
 doi:\href{https://doi.org/10.1002/ps.4662}{10.1002/ps.4662}.
+
+\leavevmode\hypertarget{ref-leeper2017}{}%
+Leeper, T. J. (2017). Interpreting {Regression Results} using {Average
+Marginal Eﬀects} with {R}'s margins. 31.
 

### 2021-08-11 data analysis/Analysis.Rmd
ko

--- 2020-11-27 Analysis.Rmd
@@
 ```
+
+```{r}
+library(popbio)
+?logi.hist.plot()
+```
+
+```{r}
+data2 %>% 
+  mutate(gender_A = case_when(
+    gender == "m" ~ TRUE,
+    gender == "f" ~ FALSE,
+    TRUE ~ NA
+  )) %>% 
+  mutate_if(is_character, as.factor) %>% 
+  filter(!is.na(gender_A)) %>% 
+  mutate(doyh = as.numeric(doyh)) -> data3
+```
+
+```{r}
+logi.hist.plot(data3$weight, data3$gender_A, 
+               logi.mod = 1, boxp=FALSE, counts=FALSE, int= 10,
+               type = "hist", rug = FALSE, col="gray")
+
+ggsave("fig.png")
+```
+
+```{r}
+aq.trans$survived <- aq.trans$fate!="dead"
+a <- subset(aq.trans, leaf<50 & stage!="recruit", c(leaf,survived))
+logi.hist.plot(a$leaf,  a$survived,
+  type="hist", boxp=FALSE, counts=TRUE, int=10,
+  ylabel="Survival probability", ylabel2="Number of plants",
+  xlab="Number of leaves")
+b <- glm(survived ~ leaf, binomial, data=a)
+summary(b)
+```
+
+```{r}
+data2 %>% 
+  mutate(gender_A = case_when(
+    gender == "m" ~ 0,
+    gender == "f" ~ 1,
+    TRUE ~ NA_real_
+  )) %>% 
+  mutate_if(is_character, as.factor) %>% 
+  filter(!is.na(gender_A)) %>% 
+  filter(!is.na(doyh)) %>% 
+  filter(!is.na(weight)) %>% 
+  filter(!is.na(height)) %>% 
+  mutate(doyh = as.numeric(doyh)) -> data3
+```
+
+```{r}
+seq(min(mtcars$hp), max(mtcars$hp),len=100)
+```
+
+```{r}
+fit = glm(gender_A ~ doyh, data = data3, family=binomial)
+newdat <- data.frame(doyh=seq(min(data3$doyh), max(data3$doyh), len=length(data3$doyh)))
+newdat$vs = predict(fit, newdata=newdat, type="response")
+plot(gender_A ~ doyh, data=data3, col="red4")
+lines(gender_A ~ doyh, newdat, col="green4", lwd=2)
+```
+
+```{r}
+label1 <- tibble(ed = "10% flowering", 
+                   x = 2.35, y = 156.5, label = "June 01")
+rect1 <- tibble(ed = "50% flowering", 
+               x = 2.85, y = 195, xend = 2.7, yend = 199)
+
+label2 <- tibble(ed = "10% flowering", 
+                   x = 2.5, y = 186.5, label = "July 01")
+rect2 <- tibble(ed = "50% flowering",  
+               x = 3.05, y = 229, xend = 2.9, yend = 235)
+
+month <- tibble(ed = "50% flowering",
+                x = c(0.5, 0.5, 0.5, 0.5),
+                y = c(168, 197, 228.5, 259), 
+                label = c("June", "July", "August", "September"))
+
+label3 <- tibble(ed = "50% flowering", 
+                   x = 2.9, y = 240, label = "Second \ncohort")
+
+label4 <- tibble(ed = "50% flowering", 
+                   x = 2.7, y = 205, label = "First \ncohort")
+```
+
+```{r eval = FALSE, include=FALSE}
+result3 %>% 
+  mutate(ed = case_when(
+    ed == "10" ~ "10% flowering",
+    ed == "50" ~ "50% flowering",
+    ed == "90" ~ "90% flowering",
+    TRUE ~ "ed")) %>% 
+  mutate(crop = fct_recode(crop,
+          "<b style='color:#30123bFF;'>corn</b>" = "corn",
+          "<b style='color:#fb832d;'>bareground</b>" = "bareground",
+          "<b style='color:#28bbecff;'>soybean</b>" = "soybean")) %>% 
+  ggplot(aes(x = crop, y = estimate)) +
+  geom_segment(aes(x = 0.8, xend = 3.2, y = 153, yend = 153), 
+             color = 1,
+             alpha = 0.5) +
+  geom_segment(aes(x = 0.8, xend = 3.2, y = 183, yend = 183), 
+             color = 2,
+             alpha = 0.5) +
+  scale_x_discrete(expand = c(0.1, 0.1)) +
+  geom_rect(mapping=aes(xmin= 0.8, xmax= 3.2,
+                        ymin= 153, ymax= 183), 
+            fill = "#FFFFF0", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.8, xmax= 3.2,
+                        ymin= 183, ymax= 214), 
+            fill = "beige", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.8, xmax= 3.2,
+                        ymin= 214, ymax= 244), 
+            fill = "#F0FFF0", alpha=0.05) +
+  geom_rect(mapping=aes(xmin= 0.8, xmax= 3.2,
+                        ymin= 244, ymax= 274), 
+            fill = "#F8F8FF", alpha=0.5) +
+  geom_text(data = month, aes(x = x, y = y), label = month$label,
+            size = 2, color = c("#8B8000", "#80461b", 
+                                "darkgreen", "blue")) +
+  scale_color_manual(values = c(1, 2)) +
+  labs(x = NULL, y = "Day of year",
+       title = NULL) +
+  geom_curve(data = rect1, aes(x = x, y = y, 
+                               xend = xend, yend = yend),
+             arrow = arrow(length = unit(0.07, "inch")), 
+             color = 1,
+             size = 0.4, curvature = 0.3) + 
+  geom_curve(data = rect2, aes(x = x, y = y, 
+                               xend = xend, yend = yend),
+             color = 2,
+             arrow = arrow(length = unit(0.07, "inch")), 
+             size = 0.4, curvature = 0.3) + 
+  geom_text(data = label1, 
+            aes(x = x, y = y), 
+            size = 2, hjust = 0, angle = 270,
+            label = label1$label, color = 1) +
+  geom_text(data = label2, 
+            aes(x = x, y = y), 
+            size = 2, hjust = 0, angle = 270,
+            label = label2$label, color = 2) +
+  geom_text(data = label3, 
+            aes(x = x, y = y), 
+            size = 2, hjust = 0, 
+            label = label3$label, color = 2) +
+  geom_text(data = label4, 
+            aes(x = x, y = y), 
+            size = 2, hjust = 0, 
+            label = label4$label, color = 1) +
+  facet_grid(~ ed, scales = "free_y") +
+  geom_point(aes(color = month), 
+             size = 1,
+             position = position_dodge2(width = 0.4)) +
+  geom_linerange(aes(ymin = lower, ymax = upper, color = month), 
+                 size = 1.3, position = position_dodge2(width = 0.4)) +
+  coord_flip() +
+  theme_test(base_family = "roboto") +
+  theme(legend.position = "none",
+        strip.text = element_markdown(face = "bold"),
+        plot.title = element_markdown(),
+        plot.title.position = "plot",
+        axis.text.x = element_markdown(angle = 30),
+        axis.text.y = element_markdown()) -> figure
+
+#ggsave("figures/Combined1.png", width = 8, height = 4)
+```
+
+```{r eval = FALSE, include=FALSE}
+(wrap_elements(full = curv1 | curv2)) / wrap_elements(full = figure) +
+   plot_layout(heights = c(1.5, 1.5)) + plot_annotation(tag_levels = 'A')
+
+#ggsave("figures/Combined2.png", height = 6)
+```

### 2021-08-11 data analysis/gender_analysis.Rmd
ko

---
title: "gender"
author: "Maxwel Coura Oliveira"
date: "8/11/2021"
output: html_document
---

```{r}
library(tidyverse)
library(lubridate)
library(tidymodels)
library(yardstick)
library(broom)       # for making model summary tidy
library(visreg)      # for potting logodds and probability 
library(margins)     # to calculate Average Marginal Effects
library(rcompanion)  # to calculate pseudo R2
library(ROCR) 
#https://towardsdatascience.com/modelling-binary-logistic-regression-using-tidymodels-library-in-r-part-1-c1bdce0ac055
#https://towardsdatascience.com/modelling-binary-logistic-regression-using-r-research-oriented-modelling-and-interpretation-f67b3a954101
```

```{r}
read_csv("../data/data2.csv") %>% 
#  filter(location != "grant") %>% 
  mutate(location = str_replace(location, "havelock", "lincoln")) %>% 
  mutate(harvest = case_when(
    location == "clay center" & doyh == 169 ~ "8/18/2018",
    TRUE ~ harvest
  )) -> data1
```

```{r}
data1 %>% 
  rename(month = planting,
         planting = planting_1) %>% 
  mutate(planting = mdy(planting),
         harvest = mdy(harvest),
         doyp = yday(planting),
         doyh = yday(harvest)) %>% 
  mutate_if(is_character, as_factor) %>% 
  filter(!is.na(gender)) %>% 
#  filter(!is.na(weight)) %>% 
  mutate(fem = as_factor(if_else(gender == "f", 1, 0))) %>% 
  mutate(crop = fct_recode(crop, 
                           "bareground" = "fallow")) %>% 
  filter(location != "grant") %>% 
  dplyr::select(fem, gender, location, year, month, crop, height, weight, doyh) -> data2
  na.omit(data2) -> data2
```

```{r}
# Total number of rows in the data frame
n <- nrow(data2)
# Number of rows for the training set (80% of the dataset)
n_train <- round(0.80 * n)
# Create a vector of indices which is an 80% random sample
set.seed(123)
train_indices <- sample(1:n, n_train)
# Subset the data frame to training indices only
train <- data2[train_indices, ]
# Exclude the training indices to create the test set
test <- data2[-train_indices, ]
```

```{r}
paste("train sample size: ", dim(train)[1])
paste("test sample size: ", dim(test)[1])
```

```{r}
library(recipes)
recipe(fem ~ doyh + crop + month + height + weight, data = train) %>% 
  recipes::step_dummy(all_nominal(), -all_outcomes()) -> recip1

#preprocessing
prepped_rec <- prep(recip1, verbose = TRUE, retain = TRUE )

# new data set with juice
juice <- juice(prepped_rec)

# bake to see the preprocessed training data
preproc_train <- recipes::bake(prepped_rec, new_data = NULL)
glimpse(preproc_train)
# bake to see the preprocessed testing data
preproc_test <- recipes::bake(prepped_rec, new_data = test)
glimpse(preproc_test)
```

```{r}
#Fitting a binary logistic regression
model_logi <- glm(fem ~ ., data = preproc_train, family = "binomial")
#Model summary
summary(model_logi)
```

```{r}
# Pseudo R_squared values and Likelyhood ratio test
rcompanion::nagelkerke(model_logi)
```

```{r}
# Probabilities of female with doy
visreg(model_logi, "doyh", scale="response", rug=2, 
       xlab="Day of year",
       ylab="P(female)",
       ylim = c(0,1)) +
  theme_test() -> doyf
```

```{r}
# Probabilities of female with weight
visreg(model_logi, "weight", scale="response", rug=2, 
       xlab=expression(paste("Weight (g plant "^"-1",")")),
       ylab="P(female)",
       ylim = c(0,1)) +
  theme_test() -> weight_f
```

```{r}
# Probabilities of female with weight
visreg(model_logi, "height", scale="response", rug=2, 
       xlab=expression(paste("Height (g plant "^"-1",")")),
       ylab="P(female)",
       ylim = c(0,1)) +
  theme_test() -> height_f
```

```{r}
tidy(model_logi, exponentiate = TRUE, conf.level = 0.95) #odds ratio
```

```{r}
# Calculate average marginal effect
effects_logit_dia = margins(model_logi)
# Summary of marginal effect
summary(effects_logit_dia)
#For example, the AME value of height is 0.0016 which can be interpreted as a unit increase in height value increases the probability of having female by 0.16%.
```

```{r}
plot(effects_logit_dia) #p
```

```{r}
# predict the preproc_test dataset
pred <- predict(model_logi, preproc_test, type="response") 
predicted <- round(pred) # round of the value; >0.5 will convert to 
                         # 1 else 0
# Creating a contigency table
tab <- table(Predicted = predicted, Reference = preproc_test$fem)
tab
```

```{r}
library(yardstick)
# Creating a dataframe of observed and predicted data
act_pred <- data.frame(observed = preproc_test$fem, predicted = 
                      factor(predicted))
# Calculating Accuracy
accuracy_est <- accuracy(act_pred, observed, predicted)
print(accuracy_est)
```

```{r}
pred.rocr <- prediction(pred, preproc_test$fem)
eval <- performance(pred.rocr, "acc")
plot(eval)
```

```{r}
# Identifying the best cutoff value that maximizes accuracy
max <- which.max(slot(eval, "y.values")[[1]])
acc <- slot(eval, "y.values")[[1]][max] #y.values are accuracy 
                                        #measures
cut <- slot(eval, "x.values")[[1]][max] #x.values are cutoff 
                                        #measures
print(c(Accuracy = acc, Cutoff = cut))
```

```{r}
library(yardstick)
# Creating a actual/observed vs predicted dataframe
act_pred <- data.frame(observed = preproc_test$fem, predicted =  
                       factor(predicted))
# Calculating precision, recall and F1_score
prec <- precision(act_pred, observed, predicted)
rec <- recall(act_pred, observed, predicted)
F1_score <- f_meas(act_pred, observed, predicted) #called f_measure
print(prec)
print(rec)
print(F1_score)
```

```{r}
perf_rocr <- performance(pred.rocr, measure = "auc",
                         x.measure = "cutoff")
perf_rocr@y.values[[1]] <- round(perf_rocr@y.values[[1]], digits = 
                                 4)
perf.tpr.fpr.rocr <- performance(pred.rocr, "tpr", "fpr")
plot(perf.tpr.fpr.rocr, colorize=T, 
     main = paste("AUC:", (perf_rocr@y.values)))
abline(a = 0, b = 1)
```

### 2021-08-11 data analysis/weight.Rmd
ko

---
title: "weight"
author: "Maxwel Coura Oliveira"
date: "8/11/2021"
output: html_document
---

###------------
```{r}
library(tidyverse)
library(lubridate)
library(lme4)
library(lmerTest)
library(emmeans)
library(tidytext)
library(ggthemes)
library(ggtext)
library(patchwork)
library(extrafont)
library(showtext)
showtext_auto()
font_add_google("Roboto", "roboto") 
```

```{r}
data <- read_csv("../data/data2.csv")
```

###--------------------------

```{r}
data %>% 
  mutate(location = str_replace(location, "havelock", "lincoln")) %>% 
  mutate(harvest = case_when(
    location == "clay center" & doyh == 169 ~ "8/18/2018",
    TRUE ~ harvest
  )) -> data1
```

```{r}
data1 %>% 
  rename(month = planting) %>% 
  mutate_at(c("year", "location"), as_factor) %>% 
  mutate(crop = fct_recode(crop, 
                           "bareground" = "fallow")) -> data2
```

```{r}
model <- lmer(log(weight) ~ crop * month + 
                                 (1|location/year), data = data2)
```

```{r}
anova(model)
```

```{r}
emmeans(model, ~ crop * month, type = "response") -> emmeans1
```

```{r}
multcomp::cld(emmeans1, 
              alpha=0.05, 
              Letters=letters, 
              adjust="none", reversed = TRUE) %>% 
  as_tibble() -> weight
```

```{r}
weight %>% 
  mutate(crop = fct_recode(crop,
          "<b style='color:#30123bFF;'>corn</b>" = "corn",
          "<b style='color:#fb832d;'>bareground</b>" = "bareground",
          "<b style='color:#28bbecff;'>soybean</b>" = "soybean")) %>%  
  ggplot(aes(y = crop, 
             x = response,
             shape = month,
             color = month)) +
  scale_y_discrete(expand = c(0.1, 0.1)) +
  geom_jitter(data = data2 %>% 
                rename(response = weight) %>% 
                mutate(crop = fct_recode(crop,
          "<b style='color:#30123bFF;'>corn</b>" = "corn",
          "<b style='color:#fb832d;'>bareground</b>" = "bareground",
          "<b style='color:#28bbecff;'>soybean</b>" = "soybean")),
#                filter(month == "First transplanting"), 
              alpha = 0.01, aes(color = month, shape = month), size = 1,
              position = position_dodge2(width = 0.4, reverse = TRUE)) +
  geom_linerange(aes(xmin = lower.CL, xmax = upper.CL),
                 position = position_dodge2(width = 0.4, reverse = TRUE)) +
  scale_color_manual(values = c(2,1)) +
  geom_text(aes(label = round(response,1), x = upper.CL + 130, color = month),
            size = 8, fontface = "bold",
            position = position_dodge2(width = 0.4, reverse = TRUE)) +
  geom_point(size = 2,
             position = position_dodge2(width = 0.4, reverse = TRUE)) +
  scale_x_continuous(limits = c(0,1201), breaks = seq(0, 1200, 300)) +
  scale_shape_manual(values = c(1,16)) +
  labs(x = expression(paste("Weight (g plant "^"-1",")")), 
       y = NULL, color = NULL) +
  annotate("text", x = 300, y = 1.7, 
           label = "First cohort",
           size = 7, color = 1,
           fontface = "bold") + 
  annotate("text", x = 300, y = 2.3, 
           label = "Second cohort",
           size = 7, color = 2,
           fontface = "bold") + 
  theme_test(base_family = "roboto") +
  theme(legend.position = "none",
        axis.text.y = element_markdown(size = 25),
        axis.title = element_markdown(size = 30),
        axis.text.x = element_markdown(size = 25),
        legend.text = element_markdown(size = 25)) -> figA

#ggsave("figures/biomass.png", width = 4, height = 4)
```

```{r}
model2 = lmer(log(height) ~ crop * month + 
                                 (1|location/year), data = data2)
anova(model2)
```

```{r}
emmeans(model2, ~ crop * month, type = "response", cont="pairwise", adjust="none") -> emmeans2
```

```{r}
multcomp::cld(emmeans2$emmeans, 
              alpha=0.05, 
              Letters=letters, 
              adjust="none", reversed = TRUE) %>% 
  as_tibble() -> height
```

```{r warning = FALSE}
height %>% 
  mutate(crop = fct_recode(crop,
          "<b style='color:#fb832d;'>bareground</b>" = "bareground",
          "<b style='color:#30123bFF;'>corn</b>" = "corn",
          "<b style='color:#28bbecff;'>soybean</b>" = "soybean")) %>%
  ggplot(aes(y = crop,
             x = response,
             shape = month,
             color = month)) +
  scale_y_discrete(expand = c(0.1, 0.1)) +
  geom_jitter(data = data2 %>% 
                rename(response = height) %>% 
                mutate(crop = fct_recode(crop,
          "<b style='color:#fb832d;'>bareground</b>" = "bareground",
          "<b style='color:#30123bFF;'>corn</b>" = "corn",
          "<b style='color:#28bbecff;'>soybean</b>" = "soybean")),
#                filter(month == "First transplanting"), 
              alpha = 0.02, aes(color = month, shape = month), size = 1,
              position = position_dodge2(width = 0.4, reverse = TRUE)) +
  geom_linerange(aes(xmin = lower.CL, xmax = upper.CL),
                 position = position_dodge2(width = 0.4, reverse = TRUE)) +
  scale_color_manual(values = c(2,1)) +
  geom_text(aes(label = round(response,1), x = upper.CL + 20, color = month),
            size = 8, fontface = "bold",
            position = position_dodge2(width = 0.4, reverse = TRUE)) +
  geom_point(size = 2,
             position = position_dodge2(width = 0.4, reverse = TRUE)) +
  scale_shape_manual(values = c(1,16)) +
#  scale_x_continuous(limits = c(0,300), breaks = c(0,25,50,75,100)) +
  labs(x = expression(paste("Height (cm plant "^"-1",")")), 
       y = NULL, color = NULL) +
  theme_test(base_family = "roboto") +
  theme(legend.position = "none",
        axis.text.y = element_markdown(size = 25),
        axis.title = element_markdown(size = 30),
        axis.text.x = element_markdown(size = 25),
        legend.text = element_markdown(size = 25)) -> figB

#ggsave("figures/height.png", width = 4, height = 4)
```

```{r}
(figA | figB) +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 30, hjust = 0, vjust = 0))

ggsave("figures/Figure 3.png", width = 6, height = 3)
```

```{r eval=FALSE}
wrap_elements(full = curve1) / (figure) / (figA | figB) +
   plot_layout(heights = c(1.5, 1.25, 1.25))

#ggsave("figures/combine 3.png", height = 9)
```

```{r}
library(emoGG)
```

```{r}
data2 %>% 
  mutate(gender_A = case_when(
    gender == "m" ~ 0,
    gender == "f" ~ 1,
    TRUE ~ NA_real_
  )) %>% 
  mutate_if(is_character, as.factor) %>% 
  filter(!is.na(gender_A)) %>% 
  filter(!is.na(doyh)) %>% 
  filter(!is.na(weight)) %>% 
  filter(!is.na(height)) %>% 
  mutate(doyh = as.numeric(doyh)) %>% 
  mutate(crop = fct_recode(crop, 
                           "bareground" = "fallow")) -> data3
```

```{r}
first <- tibble(crop = "bareground",
                x = 200, y = 0.73,
                label = "First cohort")
second <- tibble(crop = "bareground",
                 x = 235, y = 0.4,
                label = "Second cohort")
cohortf1 <- tibble(crop = "corn",
                 x = 215, y = 0.94)
cohortm1 <- tibble(crop = "corn",
                 x = 215, y = 0.06)
```

```{r}
data3 %>% 
  ggplot(aes(x = doyh, y = gender_A)) +
  geom_point(alpha = 0.1, aes(color = month)) +
  geom_emoji(data = cohortf1, aes(x = x, y = y), emoji = "2640") +
  geom_emoji(data = cohortm1, aes(x = x, y = y), emoji = "2642") +
  geom_smooth(aes(color = month), 
              method = "gam", 
              se = TRUE, 
              method.args=list(family="binomial", 
                               type = "response", 
                               trans = "plogis"),
              stat = "smooth") +
  theme_test(base_family = "roboto") +
  facet_grid(~crop) +
  scale_color_manual(values = c(2,1)) +
  scale_fill_manual(values = c(2,1)) +
  scale_x_continuous(limits = c(170, 260), 
                     breaks = seq(170, 260, 15)) +
  scale_y_continuous( breaks = seq(0, 1, by = 0.2)) +
  labs(x = "Day of year", y = NULL) +
  geom_text(data = first, mapping = aes(x = x, y = y), label = first$label,
            size = 10, color = 1) +
  geom_text(data = second, mapping = aes(x = x, y = y), label = second$label,
            size = 10, color = 2) +
  theme(legend.position = "none",
        axis.title = element_markdown(size = 35),
        axis.text = element_markdown(size = 25),
        strip.text = element_markdown(face = "bold", size = 30)) -> pl_doyh
#  xlim(175, 270)

ggsave("figures/fig.png", width = 9, height = 4)
```

```{r}
cohortf2 <- tibble(crop = "corn",
                 x = 92, y = 0.94)
cohortm2 <- tibble(crop = "corn",
                 x = 92, y = 0.06)
```

```{r}
data3 %>% 
  mutate(crop = fct_relevel(crop,
                            levels = c("bareground", "corn", "soybean"))) %>% 
  ggplot(aes(x = weight, y = gender_A)) +
  geom_point(alpha = 0.1, aes(color = month)) +
  geom_emoji(data = cohortf2, aes(x = x, y = y), emoji = "2640") +
  geom_emoji(data = cohortm2, aes(x = x, y = y), emoji = "2642") +
  geom_smooth(aes(color = month), 
              method = "gam", 
              se = FALSE, 
              stat = "smooth") +
  theme_test(base_family = "roboto") +
  facet_grid(~crop, scales = "free") +
  scale_color_manual(values = c(2,1)) +
  scale_fill_manual(values = c(2,1)) +
  scale_y_continuous( breaks = seq(0, 1, by = 0.2)) +
  labs(x = expression(paste("Weight (g plant "^"-1",")")), y = NULL) +
  theme(legend.position = "none",
        axis.title = element_markdown(size = 35),
        axis.text = element_markdown(size = 25),
        strip.text = element_markdown(face = "bold", size = 30)) -> pl_weight
#  xlim(175, 270)

ggsave("fig.png", width = 9, height = 4)
```
```{r}
cohortf2 <- tibble(crop = "corn",
                 x = 110, y = 0.94)
cohortm2 <- tibble(crop = "corn",
                 x = 110, y = 0.06)
```

```{r}
data3 %>% 
  mutate(crop = fct_relevel(crop,
                            levels = c("bareground", "corn", "soybean"))) %>% 
  ggplot(aes(x = height, y = gender_A)) +
  geom_point(alpha = 0.1, aes(color = month)) +
  geom_emoji(data = cohortf2, aes(x = x, y = y), emoji = "2640") +
  geom_emoji(data = cohortm2, aes(x = x, y = y), emoji = "2642") +
  geom_smooth(aes(color = month), 
              method = "gam", 
              se = FALSE, 
              stat = "smooth") +
  theme_test(base_family = "roboto") +
  facet_grid(~crop) +
  scale_color_manual(values = c(2,1)) +
  scale_fill_manual(values = c(2,1)) +
  scale_y_continuous( breaks = seq(0, 1, by = 0.2)) +
  labs(x = expression(paste("Height (cm plant "^"-1",")")), y = NULL) +
    theme(legend.position = "none",
        axis.title = element_markdown(size = 35),
        axis.text = element_markdown(size = 25),
        strip.text = element_markdown(face = "bold", size = 30)) -> pl_height
#  xlim(175, 270)

ggsave("fig.png", width = 9, height = 4)
```

```{r}
pl_doyh / pl_weight / pl_height

ggsave("figures/Figure 4.png", width = 6, height = 8)
```

### 2021-08-12 Manuscript/Manuscript.Rmd
oi

@@
 The likelihood ratio test showed a p-value of < 0.00.
-The marginal effects showed that increasing a weight unit by 1 g increases the probability of having a female plant by 0.08%.
-Similar trend is observed to height as well as doy, whereas the probability of being female increase by 0.16% and 0.4% when a unit of height (cm) and day increases.
+The average marginal effects showed that increasing a weight unit by 1 g increases the probability of having a female plant by 0.08% (P = 0.01).
+Similar trend is observed to height as well as doy, whereas the probability of being female increase by 0.2% (P = 0.01) and 0.4% (P = 0.00) when a unit of height (cm) and day increases, respectively.
 
-The model evaluation accuracy in the unseen 20% dataset was 0.61 with a cutoff value for female and male plants of 0.49.
+The model accuracy evaluation accuracy in the unseen 20% dataset was 0.61 with a cutoff value for female and male plants of 0.49.
 The model classification showed a precision of 0.59, recall of 0.76, and a F1 score of 0.66.
@@
 
-Our study showed that Palmer amaranth flowering time, biomass and height varied within agroecossystems and cohort timing.
+Our study showed that Palmer amaranth biomass, height, flowering pattern and gender varied within agroecossystems and cohort timings.
 In general, Palmer amaranth produced more biomass and taller plants when growing at first cohort rather than second cohort.
@@
 In such condition, Palmer amaranth showed an extraordinary plasticity to adapt upon the agroecosystem.
-This is evident when comparing Palmer amaranth shape, and its extended flowering pattern when growing into corn compared to soybean.
+This is evident when comparing Palmer amaranth canopy shape, and its extended flowering pattern when growing into corn compared to soybean.
 Palmer amaranth was taller and thinner when growing within corn.
-The Palmer amaranth competition strategy was likely to mimic the crop grow and development (Figure 4).
-These results are concerning as breeding more competitive crop varieties is likely to select more competitive Palmer amaranth biotypes [@bravo2017].
+The Palmer amaranth competition strategy was to mimic the crop grow and development (Figure 4).
+These results suggests that Palmer amaranth can fast evolve life-history traits to adapt  into agroecossystems and cultural practices, including nitrogen fertilization [@bravo2018].
+Our result highlight the threat of Palmer amaranth to agroecossystems as breeding more competitive crop varieties is likely to select more competitive biotypes [@bravo2017].
 
@@
 
-Seed production was not evaluated due to Palmer amaranth harvest at flowering time.
+Seed production was not evaluated due to plant harvest at initiation of flowering.
 Nonetheless, it is well documented a strong positive correlation between Palmer amaranth biomass and seed production [@schwartz2016; @spaunhorst2018].
@@
 Our observation is consistent with the findings that first Palmer amaranth cohort produced 50% more seeds per plant than Palmer amaranth plants established six weeks later in bareground [@webster2015].
-Nonetheless, seed production at second cohort is likely to replenish the soil seedbank.
-Seed production and deposition in the seedbank is a key factor for species perpetuation [@menges1987].
+Still, seed production at second cohort is likely to replenish the soil seedbank.
+Seed production and deposition in the seedbank is also a key factor for species perpetuation [@menges1987].
 Palmer amaranth can produce hundred thousands seeds per plant [@schwartz2016; @keeley1987], and stay viable buried in the seedbank for at least 36 months [@sosnoskie2013].
+Therefore, avoid Palmer amaranth seed production is also a key strategy to minimize weed impact into agroecossystems.
 
-Palmer amaranth flowering pattern was slightly influenced by cropping systems and cohort timings.
-Palmer amaranth growing in fallow and corn resulted in the overall shortest and longest flowering pattern.
-When growing in soybean, Palmer amaranth flowering pattern was similar to bareground and corn agroecossystems at first and second cohort timings, respectively.
-It seems that Palmer amaranth cohorts tend to have an extended flowering pattern when in high competition.
-The agroecossystems influenced Palmer amaranth gender.
-When growing in corn, the probability of having Palmer amaranth female plants reduced 18.1%.
+An ecological approach to reduce seed production in Palmer amaranth is understanding its flowering pattern.
+Our study suggests that Palmer amaranth flowering pattern was slightly influenced by cropping systems and cohort timings.
+Palmer amaranth growing in fallow and corn resulted in the overall shortest and longest flowering pattern, respectively.
+When growing in soybean, Palmer amaranth flowering pattern was similar to bareground at first cohort but similar to corn at second cohort timing.
+Plant flowering initiation is complex and depends on biological and ecological factors [@lang1965].
+We hypothesize that when growing in high competition (e.g., second cohort), Palmer amaranth plants tend to initiate flowering early, as well as having an extended flowering pattern.
+Early flower initiation is plant strategy when growing under stress.
+For example, when growing under water stress, early flowering in Palmer amaranth resulted in a mismatch between female and male plants by seven days [@mesgaran2021].
+A mismatch in Palmer amaranth male and female flowering period can minimize plant outcross, and thus reduce plant seed production.
+Sex dimorphism manipulation is considered a potential ecological pest control [@mcfarlane2018; @schliekelman2005].
+
+The mechanisms of sex-determination in plant species is intriguing and arouse the curiosity of many scientists, including Darwin [@darwin1888].
+In our study, the gender model performance was decent considering the biology of plant flowering.
+A 1:1 male and female sex ratio is a general evolutionary stable strategy for plant species perpetuation [@fisher1930].
+However, a slight deviation from 1:1 sex ratio might occur in some dioecious species.
+For example, the dioecious *Halophila stipulacea* is a female-biased plant in its native habitat, but the naturalized *H.
+stipulacea* have a 1:1 ratio [@nguyen2018].
+Naturalized of *H.
+stipulacea* reduced female-male ratio to expand into its non-native habitat.
+Also, biotic and/or abiotic stress can influence plant sex determination.
+Palmer amaranth  male-to-female ratio was greater under high plant densities [@korres2017a] and after herbicide application [@rumpa2019].
+Our model estimated that late flowering, heavier and taller Palmer amaranth plants deviated from 1:1 ratio in favor to female plants.
+It was reported that female Palmer amaranth plants invested more in height, stem and biomass while male invested more in leaf area and leaf dry weight under nutrient deficiency [@korres2017].
+Our model also estimated more female plants in soybean and bareground compared to corn, which might be a result of plant biomass in each agroecossystem.
+Our results showed the influence of life-history and ecological traits on sexual dimorphism in Palmer amaranth.
+Sexual dimorphism is documented in other dioecious species [@barrett2013].
+For example, stronger female plant competition and greater male tolerance to herbivory was reported in *Spinacia oleracea* [@perez-llorca2019].
+Research on candidate genes for sex determination in *Amaranthus* species are currently underway but it is far to complete [@montgomery2021; @montgomery2019].
+Further studies are also needed to understand the ecological basis of Palmer amaranth flowering.
 
@@
 Palmer amaranth management priority should focus on minimizing Palmer amaranth dispersal into new geographies.
-Early-season management programs would have a large negative effect in Palmer amaranth growth and development.
-Tactics that promote early-season crop advantage against Palmer amaranth, including early crop planting, crop rotation, plant width, preemergence applied herbicide [@sanctis2021], and crop residue (e.g.
+Flower manipulation is a promising ecological Palmer amaranth but it is not well understood.
+Nonetheless, early-season management programs would have a large negative effect in Palmer amaranth growth and development.
+Tactics that promote early-season crop advantage against Palmer amaranth, including early crop planting, crop rotation [@oliveira2021a], plant width, preemergence applied herbicide [@sanctis2021], and crop residue (e.g.
 cover crops) would minimize the negative impact of Palmer amaranth in agroecossystems.

### 2021-08-12 Manuscript/Manuscript.tex
oi

@@
 McFadden, Cox and Snell, and Cragg and Uhler, respectively. The
-likelihood ratio test showed a p-value of \textless{} 0.00. The marginal
-effects showed that increasing a weight unit by 1 g increases the
-probability of having a female plant by 0.08\%. Similar trend is
-observed to height as well as doy, whereas the probability of being
-female increase by 0.16\% and 0.4\% when a unit of height (cm) and day
-increases.
-
-The model evaluation accuracy in the unseen 20\% dataset was 0.61 with a
-cutoff value for female and male plants of 0.49. The model
+likelihood ratio test showed a p-value of \textless{} 0.00. The average
+marginal effects showed that increasing a weight unit by 1 g increases
+the probability of having a female plant by 0.08\% (P = 0.01). Similar
+trend is observed to height as well as doy, whereas the probability of
+being female increase by 0.2\% (P = 0.01) and 0.4\% (P = 0.00) when a
+unit of height (cm) and day increases, respectively.
+
+The model accuracy evaluation accuracy in the unseen 20\% dataset was
+0.61 with a cutoff value for female and male plants of 0.49. The model
 classification showed a precision of 0.59, recall of 0.76, and a F1
@@
 
-Our study showed that Palmer amaranth flowering time, biomass and height
-varied within agroecossystems and cohort timing. In general, Palmer
-amaranth produced more biomass and taller plants when growing at first
-cohort rather than second cohort. At first cohort, resources (e.g., soil
-nutrients) and conditions (e.g., light) were more timely available for
-both species, crop and weed. High biomass and taller Palmer amaranth
-plants are likely a weed strategy to compete for light in between crop
-rows in absence of canopy. In such condition, Palmer amaranth showed an
-extraordinary plasticity to adapt upon the agroecosystem. This is
-evident when comparing Palmer amaranth shape, and its extended flowering
-pattern when growing into corn compared to soybean. Palmer amaranth was
-taller and thinner when growing within corn. The Palmer amaranth
-competition strategy was likely to mimic the crop grow and development
-(Figure 4). These results are concerning as breeding more competitive
-crop varieties is likely to select more competitive Palmer amaranth
-biotypes (Bravo et al., 2017).
+Our study showed that Palmer amaranth biomass, height, flowering pattern
+and gender varied within agroecossystems and cohort timings. In general,
+Palmer amaranth produced more biomass and taller plants when growing at
+first cohort rather than second cohort. At first cohort, resources
+(e.g., soil nutrients) and conditions (e.g., light) were more timely
+available for both species, crop and weed. High biomass and taller
+Palmer amaranth plants are likely a weed strategy to compete for light
+in between crop rows in absence of canopy. In such condition, Palmer
+amaranth showed an extraordinary plasticity to adapt upon the
+agroecosystem. This is evident when comparing Palmer amaranth canopy
+shape, and its extended flowering pattern when growing into corn
+compared to soybean. Palmer amaranth was taller and thinner when growing
+within corn. The Palmer amaranth competition strategy was to mimic the
+crop grow and development (Figure 4). These results suggests that Palmer
+amaranth can fast evolve life-history traits to adapt into
+agroecossystems and cultural practices, including nitrogen fertilization
+(Bravo et al., 2018). Our result highlight the threat of Palmer amaranth
+to agroecossystems as breeding more competitive crop varieties is likely
+to select more competitive biotypes (Bravo et al., 2017).
 
@@
 Palmer amaranth strategy in bareground was to invest biomass in growing
-plant width and height. Nonetheless, Palmer amaranth produced 31\% less
+plant width and height. Nonetheless, Palmer amaranth produced 21\% less
 biomass in second cohort compared to first cohort timing. In a
@@
 
-Seed production was not evaluated due to Palmer amaranth harvest at
-flowering time. Nonetheless, it is well documented a strong positive
+Seed production was not evaluated due to plant harvest at initiation of
+flowering. Nonetheless, it is well documented a strong positive
 correlation between Palmer amaranth biomass and seed production
 (Schwartz et al., 2016; Spaunhorst et al., 2018). I our study, Palmer
-amaranth growing at first cohort accumulated an overall 50\% more
+amaranth growing at first cohort accumulated an overall 36\% more
 biomass when compared to second cohort. Therefore, Palmer amaranth
@@
 plant than Palmer amaranth plants established six weeks later in
-bareground (Webster and Grey, 2015). Nonetheless, seed production at
-second cohort is likely to replenish the soil seedbank. Seed production
-and deposition in the seedbank is a key factor for species perpetuation
+bareground (Webster and Grey, 2015). Still, seed production at second
+cohort is likely to replenish the soil seedbank. Seed production and
+deposition in the seedbank is also a key factor for species perpetuation
 (Menges, 1987). Palmer amaranth can produce hundred thousands seeds per
@@
 
-Palmer amaranth flowering pattern was variable
+Palmer amaranth flowering pattern was slightly influenced by cropping
+systems and cohort timings. Palmer amaranth growing in fallow and corn
+resulted in the overall shortest and longest flowering pattern,
+respectively. When growing in soybean, Palmer amaranth flowering pattern
+was similar to bareground at first cohort but similar to corn at second
+cohort timing. Plant flowering initiation is complex and depends on
+biological and ecological factors (Lang, 1965). We hypothesize that when
+growing in high competition (e.g., second cohort), Palmer amaranth
+plants tend to initiate flowering early, as well as having an extended
+flowering pattern. Early flower initiation is plant strategy when
+growing under stress. For example, when growing under water stress,
+early flowering in Palmer amaranth resulted in a mismatch between female
+and male plants by seven days (Mesgaran et al., 2021). A mismatch in
+Palmer amaranth male and female flowering period can minimize plant
+outcross, and thus reduce plant seed production.
+
+The gender model performance was decent considering the biology of plant
+flowering. A 1:1 male and female sex ratio is a general evolutionary
+stable strategy for plant species perpetuation (Fisher, 1930). However,
+a slight deviation from 1:1 sex ratio might occur in some dioecious
+species. For example, the dioecious \emph{Halophila stipulacea} is a
+female-biased plant in its native habitat, but the naturalized \emph{H.
+stipulacea} have a 1:1 ratio (Nguyen et al., 2018). Naturalized of
+\emph{H. stipulacea} reduced female-male ratio to expand into its
+non-native habitat. Also, biotic and/or abiotic stress can influence
+plant sex determination. Palmer amaranth male-to-female ratio was
+greater under high plant densities (Korres and Norsworthy, 2017) and
+after herbicide application (Rumpa et al., 2019). Our model estimated
+that late flowering, heavier and taller Palmer amaranth plants deviated
+from 1:1 ratio in favor to female plants. It was reported that female
+Palmer amaranth plants invested more in height, stem and biomass while
+male invested more in leaf area and leaf dry weight under nutrient
+deficiency (Korres et al., 2017). Our model also estimated more female
+plants in soybean and bareground compared to corn, which might be a
+result of plant biomass in each agroecossystem. Our results showed the
+influence of life-history and ecological traits on sexual dimorphism in
+Palmer amaranth. Sexual dimorphism is documented in other dioecious
+species (Barrett and Hough, 2013). For example, stronger female plant
+competition and greater male tolerance to herbivory was reported in
+\emph{Spinacia oleracea} (Pérez-Llorca and Sánchez Vilas, 2019). The
+mechanisms of sex-determination in plant species is intriguing and
+arouse the curiosity of many scientists, including Darwin (Darwin,
+1888). Research on candidate genes for sex determination in
+\emph{Amaranthus} species are currently underway (Montgomery et al.,
+2021, 2019). Sex dimorphism manipulation is considered a potential
+ecological pest control (McFarlane et al., 2018; Schliekelman et al.,
+2005).
 
@@
 negative effect in Palmer amaranth growth and development. Tactics that
-promote crop advantage against Palmer amaranth, including early crop
-planting, crop rotation, plant width, preemergence applied herbicide
-(Sanctis et al., 2021), and crop residue (e.g.~cover crops) would
-minimize the negative impact of Palmer amaranth in agroecossystems.
+promote early-season crop advantage against Palmer amaranth, including
+early crop planting, crop rotation, plant width, preemergence applied
+herbicide (Sanctis et al., 2021), and crop residue (e.g.~cover crops)
+would minimize the negative impact of Palmer amaranth in
+agroecossystems.
 
@@
 
+\leavevmode\hypertarget{ref-barrett2013}{}%
+Barrett, S. C. H., and Hough, J. (2013). Sexual dimorphism in flowering
+plants. \emph{Journal of Experimental Botany} 64, 67--82.
+doi:\href{https://doi.org/10.1093/jxb/ers308}{10.1093/jxb/ers308}.
+
 \leavevmode\hypertarget{ref-bates2015}{}%
@@
 
+\leavevmode\hypertarget{ref-bravo2018}{}%
+Bravo, W., Leon, R. G., Ferrell, J. A., Mulvaney, M. J., and Wood, C. W.
+(2018). Evolutionary {Adaptations} of {Palmer Amaranth} ({Amaranthus}
+palmeri) to {Nitrogen Fertilization} and {Crop Rotation History Affect
+Morphology} and {Nutrient}-{Use Efficiency}. \emph{Weed Science} 66,
+180--189.
+doi:\href{https://doi.org/10.1017/wsc.2017.73}{10.1017/wsc.2017.73}.
+
 \leavevmode\hypertarget{ref-briscoerunquist2019}{}%
@@
 
+\leavevmode\hypertarget{ref-darwin1888}{}%
+Darwin, C. (1888). \emph{The {Different Forms} of {Flowers} on {Plants}
+of the {Same Species}}. {J. Murray} Available at:
+\url{http://books.google.com?id=7uMEAAAAYAAJ}.
+
 \leavevmode\hypertarget{ref-davis2015}{}%
@@
 2592--2603. doi:\href{https://doi.org/10.1002/ps.4710}{10.1002/ps.4710}.
+
+\leavevmode\hypertarget{ref-fisher1930}{}%
+Fisher, R. A. (1930). The genetical theory of natural selection.
+\emph{Eugen Rev} 22, 127--130. Available at:
+\url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2984947/} {[}Accessed
+August 12, 2021{]}.
 
@@
 doi:\href{https://doi.org/10.1017/wsc.2017.2}{10.1017/wsc.2017.2}.
+
+\leavevmode\hypertarget{ref-korres2017a}{}%
+Korres, N. E., and Norsworthy, J. K. (2017). Palmer {Amaranth}
+({Amaranthus} palmeri) {Demographic} and {Biological Characteristics} in
+{Wide}-{Row Soybean}. \emph{Weed Science} 65, 491--503.
+doi:\href{https://doi.org/10.1017/wsc.2017.12}{10.1017/wsc.2017.12}.
+
+\leavevmode\hypertarget{ref-korres2017}{}%
+Korres, N. E., Norsworthy, J. K., FitzSimons, T., Roberts, T. L., and
+Oosterhuis, D. M. (2017). Differential {Response} of {Palmer Amaranth}
+({Amaranthus} palmeri) {Gender} to {Abiotic Stress}. \emph{Weed Science}
+65, 213--227.
+doi:\href{https://doi.org/10.1017/wsc.2016.34}{10.1017/wsc.2016.34}.
 
@@
 
+\leavevmode\hypertarget{ref-lang1965}{}%
+Lang, A. (1965). {``Physiology of flower initiation,''} in
+\emph{Differenzierung und {Entwicklung} / {Differentiation} and
+{Development}} Handbuch der {Pflanzenphysiologie} / {Encyclopedia} of
+{Plant Physiology}., ed. A. Lang ({Berlin, Heidelberg}: {Springer}),
+1380--1536.
+doi:\href{https://doi.org/10.1007/978-3-642-50088-6_39}{10.1007/978-3-642-50088-6\_39}.
+
 \leavevmode\hypertarget{ref-larran2017}{}%
@@
 
+\leavevmode\hypertarget{ref-mcfarlane2018}{}%
+McFarlane, G. R., Whitelaw, C. B. A., and Lillico, S. G. (2018).
+{CRISPR}-{Based Gene Drives} for {Pest Control}. \emph{Trends in
+Biotechnology} 36, 130--133.
+doi:\href{https://doi.org/10.1016/j.tibtech.2017.10.001}{10.1016/j.tibtech.2017.10.001}.
+
 \leavevmode\hypertarget{ref-menges1987}{}%
@@
 \url{http://www.jstor.org/stable/4044593}.
+
+\leavevmode\hypertarget{ref-mesgaran2021}{}%
+Mesgaran, M. B., Matzrafi, M., and Ohadi, S. (2021). Sex dimorphism in
+dioecious {Palmer} amaranth ({Amaranthus} palmeri) in response to water
+stress. \emph{Planta} 254, 17.
+doi:\href{https://doi.org/10.1007/s00425-021-03664-7}{10.1007/s00425-021-03664-7}.
 
@@
 
+\leavevmode\hypertarget{ref-montgomery2021}{}%
+Montgomery, J. S., Giacomini, D. A., Weigel, D., and Tranel, P. J.
+(2021). Male-specific {Y}-chromosomal regions in waterhemp ({Amaranthus}
+tuberculatus) and {Palmer} amaranth ({Amaranthus} palmeri). \emph{New
+Phytologist} 229, 3522--3533.
+doi:\href{https://doi.org/10.1111/nph.17108}{10.1111/nph.17108}.
+
+\leavevmode\hypertarget{ref-montgomery2019}{}%
+Montgomery, J. S., Sadeque, A., Giacomini, D. A., Brown, P. J., and
+Tranel, P. J. (2019). Sex-specific markers for waterhemp ({Amaranthus}
+tuberculatus) and {Palmer} amaranth ({Amaranthus} palmeri). \emph{Weed
+Science} 67, 412--418.
+doi:\href{https://doi.org/10.1017/wsc.2019.27}{10.1017/wsc.2019.27}.
+
 \leavevmode\hypertarget{ref-morgan2001}{}%
@@
 doi:\href{https://doi.org/10.1614/0890-037X(2001)015\%5B0408:CIOPAA\%5D2.0.CO;2}{10.1614/0890-037X(2001)015{[}0408:CIOPAA{]}2.0.CO;2}.
+
+\leavevmode\hypertarget{ref-nguyen2018}{}%
+Nguyen, H. M., Kleitou, P., Kletou, D., Sapir, Y., and Winters, G.
+(2018). Differences in flowering sex ratios between native and invasive
+populations of the seagrass {Halophila} stipulacea. \emph{Botanica
+Marina} 61, 337--342.
+doi:\href{https://doi.org/10.1515/bot-2018-0015}{10.1515/bot-2018-0015}.
 
@@
 
+\leavevmode\hypertarget{ref-perez-llorca2019}{}%
+Pérez-Llorca, M., and Sánchez Vilas, J. (2019). Sexual dimorphism in
+response to herbivory and competition in the dioecious herb {Spinacia}
+oleracea. \emph{Plant Ecol} 220, 57--68.
+doi:\href{https://doi.org/10.1007/s11258-018-0902-7}{10.1007/s11258-018-0902-7}.
+
 \leavevmode\hypertarget{ref-piskackova2021}{}%
@@
 
+\leavevmode\hypertarget{ref-rumpa2019}{}%
+Rumpa, M. M., Krausz, R. F., Gibson, D. J., and Gage, K. L. (2019).
+Effect of {PPO}-{Inhibiting Herbicides} on the {Growth} and {Sex Ratio}
+of a {Dioecious Weed Species Amaranthus} palmeri ({Palmer Amaranth}).
+\emph{Agronomy} 9, 275.
+doi:\href{https://doi.org/10.3390/agronomy9060275}{10.3390/agronomy9060275}.
+
 \leavevmode\hypertarget{ref-sanctis2021}{}%
@@
 \url{http://www.jstor.org/stable/41423815}.
+
+\leavevmode\hypertarget{ref-schliekelman2005}{}%
+Schliekelman, P., Ellner, S., and Gould, F. (2005). Pest {Control} by
+{Genetic Manipulation} of {Sex Ratio}. \emph{Journal of Economic
+Entomology} 98, 18--34.
+doi:\href{https://doi.org/10.1093/jee/98.1.18}{10.1093/jee/98.1.18}.
 

### 2021-08-12 Manuscript/Manuscript.Rmd
jk

@@
 Germination of Palmer amaranth is triggered by 18 C soil temperature [@keeley1987], and optimal germination and biomass production occur at 35/30 C day and night temperatures [@guo2003].
-Also, Palmer amaranth germination are human-mediated by tillage timings and preemergence-applied herbicides [@chahal2021].
+Palmer amaranth germination is also human-mediated by tillage timings and preemergence-applied herbicides [@chahal2021].
 Water has not shown to limit Palmer amaranth fitness.
 Under continuous water stress, Palmer amaranth survived and produced at least 14000 seeds plant-1 [@chahal2018].
-Also, seeds from Palmer amaranth growing with limited water conditions were heavier, less dormant, and prompt for germination [@matzrafi2021].
+Seeds from Palmer amaranth growing with limited water conditions were heavier, less dormant, and prompt for germination [@matzrafi2021].
+Also, growing conditions and management can impact sex dimorphism and flowering pattern in Palmer amaranth [@korres2017; @rumpa2019].
 The continuous global temperature warming can impact agriculture and promote niches for Palmer amaranth invasion/adaptation into new environments.
@@
 For example, herbicide resistant Palmer amaranth is widespread in Nebraska  [@oliveira2021a],  Michigan [@kohrt2017], Connecticut [@aulakh2021], and potential damaged to soybean fields in Illinois is reported [@davis2015].
+In Iowa, a study showed that Palmer amaranth is not well adapted compared to waterhemp (*Amaranthus tuberculatus*) [@baker2021].
 Successful cases of Palmer amaranth invasion and near to eradication is documented in Minnesota [@yu2021].
@@
 
-The study was performed with a *A.
-palmeri* accession (Per1) from Perkins County, Nebraska.
+The study was performed with a Palmer amaranth accession (Per1) from Perkins County, Nebraska.
 Per1 accession collection is documented with no reported herbicide resistance [@oliveira2021a].
@@
 
+## Palmer amaranth height and biomass {-}
+
+Palmer amaranth accumulated more biomass when growing in fallow compared to Palmer amaranth growing in soybean and corn (Figure 3A).
+At first cohort time, Palmer amaranth biomass was 75.5 g plant-1, 28.3 g plant-1 and 16.3 g plant-1 in fallow, soybean and corn, respectively.
+At second cohort timing, Palmer amaranth produced 62.6 g plant in fallow, followed by 6.3 g plant in soybean, and 1.4 g plant in corn.
+
+Palmer amaranth height was more uniform across cohort timings, except when growing in corn (Figure 3B).
+Palmer amaranth achieve achieve 69.2 cm tall when growing at first cohort timing in bareground, which was not different to 70.7 cm tall at second cohort timing (P = 0.74).
+In addition, no difference in Palmer amaranth height (69.3 cm) was found in soybean at first cohort compared to fallow (P > 0.75).
+At second cohort in soybean, Palmer amaranth was near to 10 cm lower compared to its first cohort time (P = 0.04).
+The tallest and smallest Palmer amaranth plants were found in corn.
+Palmer amaranth reached 85.2 cm tall at first cohort and 38.2 cm at second cohort timings.
+
 ## Palmer amaranth cumulative flowering {-}
@@
 
-## Palmer amaranth height and biomass {-}
-
-Palmer amaranth accumulated more biomass when growing in fallow compared to Palmer amaranth growing in soybean and corn (Figure 3A).
-At first cohort time, Palmer amaranth biomass was 75.5 g plant-1, 28.3 g plant-1 and 16.3 g plant-1 in fallow, soybean and corn, respectively.
-At second cohort timing, Palmer amaranth produced 62.6 g plant in fallow, followed by 6.3 g plant in soybean, and 1.4 g plant in corn.
-
-Palmer amaranth height was more uniform across cohort timings, except when growing in corn (Figure 3B).
-Palmer amaranth achieve achieve 69.2 cm tall when growing at first cohort timing in bareground, which was not different to 70.7 cm tall at second cohort timing (P = 0.74).
-In addition, no difference in Palmer amaranth height (69.3 cm) was found in soybean at first cohort compared to fallow (P > 0.75).
-At second cohort in soybean, Palmer amaranth was near to 10 cm lower compared to its first cohort time (P = 0.04).
-The tallest and smallest Palmer amaranth plants were found in corn.
-Palmer amaranth reached 85.2 cm tall at first cohort and 38.2 cm at second cohort timings.
-
 ```{r eval = FALSE, include = FALSE}
@@
 This is evident when comparing Palmer amaranth canopy shape, and its extended flowering pattern when growing into corn compared to soybean.
-Palmer amaranth was taller and thinner when growing within corn.
 The Palmer amaranth competition strategy was to mimic the crop grow and development (Figure 4).
 These results suggests that Palmer amaranth can fast evolve life-history traits to adapt  into agroecossystems and cultural practices, including nitrogen fertilization [@bravo2018].
-Our result highlight the threat of Palmer amaranth to agroecossystems as breeding more competitive crop varieties is likely to select more competitive biotypes [@bravo2017].
+Our results highlight the Palmer amaranth threat to agroecossystems as breeding more competitive crop varieties is likely to select more competitive biotypes [@bravo2017].
 
@@
 As a result, Palmer amaranth height and biomass was lower compared to its first cohort.
-Under crop canopy, Palmer amaranth flowering pattern was near to similar in corn and soybean.
+Under crop canopy (e.g., second cohort), Palmer amaranth flowering pattern was near to similar in corn and soybean.
 Palmer amaranth growing without crop competition produced highest amounts of biomass and less extended flowering pattern.
@@
 Palmer amaranth can produce hundred thousands seeds per plant [@schwartz2016; @keeley1987], and stay viable buried in the seedbank for at least 36 months [@sosnoskie2013].
-Therefore, avoid Palmer amaranth seed production is also a key strategy to minimize weed impact into agroecossystems.
+Therefore, preventing Palmer amaranth seed production or/and seed migration to its non-native habitat is an essential strategy to minimize weed impact into agroecossystem [@davis2015].
 
@@
 When growing in soybean, Palmer amaranth flowering pattern was similar to bareground at first cohort but similar to corn at second cohort timing.
-Plant flowering initiation is complex and depends on biological and ecological factors [@lang1965].
+Plant flowering initiation is complex and it depends on biological and ecological factors [@lang1965].
 We hypothesize that when growing in high competition (e.g., second cohort), Palmer amaranth plants tend to initiate flowering early, as well as having an extended flowering pattern.
-Early flower initiation is plant strategy when growing under stress.
+Early flower initiation is plant strategy when growing in stress conditions.
 For example, when growing under water stress, early flowering in Palmer amaranth resulted in a mismatch between female and male plants by seven days [@mesgaran2021].
@@
 Naturalized of *H.
-stipulacea* reduced female-male ratio to expand into its non-native habitat.
+stipulacea* reduced female-male ratio to expand into its non-native habitat [@nguyen2018].
 Also, biotic and/or abiotic stress can influence plant sex determination.
@@
 Research on candidate genes for sex determination in *Amaranthus* species are currently underway but it is far to complete [@montgomery2021; @montgomery2019].
-Further studies are also needed to understand the ecological basis of Palmer amaranth flowering.
+Further studies are also needed to understand the ecological basis of Palmer amaranth flowering, including the plant behavior under climate change.
 

### 2021-08-12 Manuscript/Manuscript.tex
jk

@@
 occur at 35/30 C day and night temperatures (Guo and Al-Khatib, 2003).
-Also, Palmer amaranth germination are human-mediated by tillage timings
+Palmer amaranth germination is also human-mediated by tillage timings
 and preemergence-applied herbicides (Chahal et al., 2021). Water has not
@@
 Palmer amaranth survived and produced at least 14000 seeds plant-1
-(Chahal et al., 2018). Also, seeds from Palmer amaranth growing with
-limited water conditions were heavier, less dormant, and prompt for
-germination (Matzrafi et al., 2021). The continuous global temperature
+(Chahal et al., 2018). Seeds from Palmer amaranth growing with limited
+water conditions were heavier, less dormant, and prompt for germination
+(Matzrafi et al., 2021). Also, growing conditions and management can
+impact sex dimorphism and flowering pattern in Palmer amaranth (Korres
+et al., 2017; Rumpa et al., 2019). The continuous global temperature
 warming can impact agriculture and promote niches for Palmer amaranth
@@
 2017), Connecticut (Aulakh et al., 2021), and potential damaged to
-soybean fields in Illinois is reported (Davis et al., 2015). Successful
-cases of Palmer amaranth invasion and near to eradication is documented
-in Minnesota (Yu et al., 2021). No Palmer amaranth actively growing was
+soybean fields in Illinois is reported (Davis et al., 2015). In Iowa, a
+study showed that Palmer amaranth is not well adapted compared to
+waterhemp (\emph{Amaranthus palmeri}) (Baker, 2021). Successful cases of
+Palmer amaranth invasion and near to eradication is documented in
+Minnesota (Yu et al., 2021). No Palmer amaranth actively growing was
 found in Canada; however, Palmer amaranth seeds was detected in sweet
@@
 buried in the seedbank for at least 36 months (Sosnoskie et al., 2013).
-
-Palmer amaranth flowering pattern was slightly influenced by cropping
-systems and cohort timings. Palmer amaranth growing in fallow and corn
-resulted in the overall shortest and longest flowering pattern,
-respectively. When growing in soybean, Palmer amaranth flowering pattern
-was similar to bareground at first cohort but similar to corn at second
-cohort timing. Plant flowering initiation is complex and depends on
-biological and ecological factors (Lang, 1965). We hypothesize that when
-growing in high competition (e.g., second cohort), Palmer amaranth
-plants tend to initiate flowering early, as well as having an extended
-flowering pattern. Early flower initiation is plant strategy when
-growing under stress. For example, when growing under water stress,
-early flowering in Palmer amaranth resulted in a mismatch between female
-and male plants by seven days (Mesgaran et al., 2021). A mismatch in
-Palmer amaranth male and female flowering period can minimize plant
-outcross, and thus reduce plant seed production.
-
-The gender model performance was decent considering the biology of plant
-flowering. A 1:1 male and female sex ratio is a general evolutionary
-stable strategy for plant species perpetuation (Fisher, 1930). However,
-a slight deviation from 1:1 sex ratio might occur in some dioecious
-species. For example, the dioecious \emph{Halophila stipulacea} is a
-female-biased plant in its native habitat, but the naturalized \emph{H.
-stipulacea} have a 1:1 ratio (Nguyen et al., 2018). Naturalized of
-\emph{H. stipulacea} reduced female-male ratio to expand into its
-non-native habitat. Also, biotic and/or abiotic stress can influence
-plant sex determination. Palmer amaranth male-to-female ratio was
-greater under high plant densities (Korres and Norsworthy, 2017) and
-after herbicide application (Rumpa et al., 2019). Our model estimated
-that late flowering, heavier and taller Palmer amaranth plants deviated
-from 1:1 ratio in favor to female plants. It was reported that female
-Palmer amaranth plants invested more in height, stem and biomass while
-male invested more in leaf area and leaf dry weight under nutrient
-deficiency (Korres et al., 2017). Our model also estimated more female
-plants in soybean and bareground compared to corn, which might be a
-result of plant biomass in each agroecossystem. Our results showed the
-influence of life-history and ecological traits on sexual dimorphism in
-Palmer amaranth. Sexual dimorphism is documented in other dioecious
-species (Barrett and Hough, 2013). For example, stronger female plant
-competition and greater male tolerance to herbivory was reported in
-\emph{Spinacia oleracea} (Pérez-Llorca and Sánchez Vilas, 2019). The
-mechanisms of sex-determination in plant species is intriguing and
+Therefore, preventing Palmer amaranth seed production or/and seed
+migration to its non-native habitat is a key strategy to minimize weed
+impact into agroecossystem (Davis et al., 2015).
+
+An ecological approach to reduce seed production in Palmer amaranth is
+understanding its flowering pattern. Our study suggests that Palmer
+amaranth flowering pattern was slightly influenced by cropping systems
+and cohort timings. Palmer amaranth growing in fallow and corn resulted
+in the overall shortest and longest flowering pattern, respectively.
+When growing in soybean, Palmer amaranth flowering pattern was similar
+to bareground at first cohort but similar to corn at second cohort
+timing. Plant flowering initiation is complex and depends on biological
+and ecological factors (Lang, 1965). We hypothesize that when growing in
+high competition (e.g., second cohort), Palmer amaranth plants tend to
+initiate flowering early, as well as having an extended flowering
+pattern. Early flower initiation is plant strategy when growing under
+stress. For example, when growing under water stress, early flowering in
+Palmer amaranth resulted in a mismatch between female and male plants by
+seven days (Mesgaran et al., 2021). A mismatch in Palmer amaranth male
+and female flowering period can minimize plant outcross, and thus reduce
+plant seed production. Sex dimorphism manipulation is considered a
+potential ecological pest control (McFarlane et al., 2018; Schliekelman
+et al., 2005).
+
+The mechanisms of sex-determination in plant species is intriguing and
 arouse the curiosity of many scientists, including Darwin (Darwin,
-1888). Research on candidate genes for sex determination in
-\emph{Amaranthus} species are currently underway (Montgomery et al.,
-2021, 2019). Sex dimorphism manipulation is considered a potential
-ecological pest control (McFarlane et al., 2018; Schliekelman et al.,
-2005).
+1888). In our study, the gender model performance was decent considering
+the biology of plant flowering. A 1:1 male and female sex ratio is a
+general evolutionary stable strategy for plant species perpetuation
+(Fisher, 1930). However, a slight deviation from 1:1 sex ratio might
+occur in some dioecious species. For example, the dioecious
+\emph{Halophila stipulacea} is a female-biased plant in its native
+habitat, but the naturalized \emph{H. stipulacea} have a 1:1 ratio
+(Nguyen et al., 2018). Naturalized of \emph{H. stipulacea} reduced
+female-male ratio to expand into its non-native habitat. Also, biotic
+and/or abiotic stress can influence plant sex determination. Palmer
+amaranth male-to-female ratio was greater under high plant densities
+(Korres and Norsworthy, 2017) and after herbicide application (Rumpa et
+al., 2019). Our model estimated that late flowering, heavier and taller
+Palmer amaranth plants deviated from 1:1 ratio in favor to female
+plants. It was reported that female Palmer amaranth plants invested more
+in height, stem and biomass while male invested more in leaf area and
+leaf dry weight under nutrient deficiency (Korres et al., 2017). Our
+model also estimated more female plants in soybean and bareground
+compared to corn, which might be a result of plant biomass in each
+agroecossystem. Our results showed the influence of life-history and
+ecological traits on sexual dimorphism in Palmer amaranth. Sexual
+dimorphism is documented in other dioecious species (Barrett and Hough,
+2013). For example, stronger female plant competition and greater male
+tolerance to herbivory was reported in \emph{Spinacia oleracea}
+(Pérez-Llorca and Sánchez Vilas, 2019). Research on candidate genes for
+sex determination in \emph{Amaranthus} species are currently underway
+but it is far to complete (Montgomery et al., 2021, 2019). Further
+studies are also needed to understand the ecological basis of Palmer
+amaranth flowering, including the plant behavior under climate change.
 
@@
 should focus on minimizing Palmer amaranth dispersal into new
-geographies. Early-season management programs would have a large
-negative effect in Palmer amaranth growth and development. Tactics that
-promote early-season crop advantage against Palmer amaranth, including
-early crop planting, crop rotation, plant width, preemergence applied
+geographies. Flower manipulation is a promising ecological Palmer
+amaranth but it is not well understood. Nonetheless, early-season
+management programs would have a large negative effect in Palmer
+amaranth growth and development. Tactics that promote early-season crop
+advantage against Palmer amaranth, including early crop planting, crop
+rotation (Oliveira et al., 2021), plant width, preemergence applied
 herbicide (Sanctis et al., 2021), and crop residue (e.g.~cover crops)
@@
 
+\leavevmode\hypertarget{ref-baker2021}{}%
+Baker, R. (2021). Comparative analysis of {Palmer} amaranth
+({Amaranthus} palmeri) and waterhemp ({A}. Tuberculatus) in {Iowa}.
+doi:\href{https://doi.org/10.31274/etd-20210609-11}{10.31274/etd-20210609-11}.
+
 \leavevmode\hypertarget{ref-bangdiwala2018}{}%

### 2021-08-12 Manuscript/Manuscript.Rmd
jhkj

@@
 
-```{r, Figure-5, ref.label = "fig5", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "120mm", out.height = "180mm"}
+```{r, Figure-5, ref.label = "fig5", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "170mm", out.height = "70mm"}
 # You can also refer to code chunks from above to place figures at the bottom.

### 2021-08-12 Manuscript/Manuscript.tex
jhkj

@@
 study showed that Palmer amaranth is not well adapted compared to
-waterhemp (\emph{Amaranthus palmeri}) (Baker, 2021). Successful cases of
-Palmer amaranth invasion and near to eradication is documented in
-Minnesota (Yu et al., 2021). No Palmer amaranth actively growing was
+waterhemp (\emph{Amaranthus tuberculatus}) (Baker, 2021). Successful
+cases of Palmer amaranth invasion and near to eradication is documented
+in Minnesota (Yu et al., 2021). No Palmer amaranth actively growing was
 found in Canada; however, Palmer amaranth seeds was detected in sweet
@@
 
-The study was performed with a \emph{A. palmeri} accession (Per1) from
+The study was performed with a Palmer amaranth accession (Per1) from
 Perkins County, Nebraska. Per1 accession collection is documented with
@@
 
+\hypertarget{palmer-amaranth-height-and-biomass}{%
+\subsection*{Palmer amaranth height and
+biomass}\label{palmer-amaranth-height-and-biomass}}
+\addcontentsline{toc}{subsection}{Palmer amaranth height and biomass}
+
+Palmer amaranth accumulated more biomass when growing in fallow compared
+to Palmer amaranth growing in soybean and corn (Figure 3A). At first
+cohort time, Palmer amaranth biomass was 75.5 g plant-1, 28.3 g plant-1
+and 16.3 g plant-1 in fallow, soybean and corn, respectively. At second
+cohort timing, Palmer amaranth produced 62.6 g plant in fallow, followed
+by 6.3 g plant in soybean, and 1.4 g plant in corn.
+
+Palmer amaranth height was more uniform across cohort timings, except
+when growing in corn (Figure 3B). Palmer amaranth achieve achieve 69.2
+cm tall when growing at first cohort timing in bareground, which was not
+different to 70.7 cm tall at second cohort timing (P = 0.74). In
+addition, no difference in Palmer amaranth height (69.3 cm) was found in
+soybean at first cohort compared to fallow (P \textgreater{} 0.75). At
+second cohort in soybean, Palmer amaranth was near to 10 cm lower
+compared to its first cohort time (P = 0.04). The tallest and smallest
+Palmer amaranth plants were found in corn. Palmer amaranth reached 85.2
+cm tall at first cohort and 38.2 cm at second cohort timings.
+
 \hypertarget{palmer-amaranth-cumulative-flowering}{%
@@
 
-\hypertarget{palmer-amaranth-height-and-biomass}{%
-\subsection*{Palmer amaranth height and
-biomass}\label{palmer-amaranth-height-and-biomass}}
-\addcontentsline{toc}{subsection}{Palmer amaranth height and biomass}
-
-Palmer amaranth accumulated more biomass when growing in fallow compared
-to Palmer amaranth growing in soybean and corn (Figure 3A). At first
-cohort time, Palmer amaranth biomass was 75.5 g plant-1, 28.3 g plant-1
-and 16.3 g plant-1 in fallow, soybean and corn, respectively. At second
-cohort timing, Palmer amaranth produced 62.6 g plant in fallow, followed
-by 6.3 g plant in soybean, and 1.4 g plant in corn.
-
-Palmer amaranth height was more uniform across cohort timings, except
-when growing in corn (Figure 3B). Palmer amaranth achieve achieve 69.2
-cm tall when growing at first cohort timing in bareground, which was not
-different to 70.7 cm tall at second cohort timing (P = 0.74). In
-addition, no difference in Palmer amaranth height (69.3 cm) was found in
-soybean at first cohort compared to fallow (P \textgreater{} 0.75). At
-second cohort in soybean, Palmer amaranth was near to 10 cm lower
-compared to its first cohort time (P = 0.04). The tallest and smallest
-Palmer amaranth plants were found in corn. Palmer amaranth reached 85.2
-cm tall at first cohort and 38.2 cm at second cohort timings.
-
 \hypertarget{discussion}{%
@@
 shape, and its extended flowering pattern when growing into corn
-compared to soybean. Palmer amaranth was taller and thinner when growing
-within corn. The Palmer amaranth competition strategy was to mimic the
-crop grow and development (Figure 4). These results suggests that Palmer
-amaranth can fast evolve life-history traits to adapt into
+compared to soybean. The Palmer amaranth competition strategy was to
+mimic the crop grow and development (Figure 4). These results suggests
+that Palmer amaranth can fast evolve life-history traits to adapt into
 agroecossystems and cultural practices, including nitrogen fertilization
-(Bravo et al., 2018). Our result highlight the threat of Palmer amaranth
+(Bravo et al., 2018). Our results highlight the Palmer amaranth threat
 to agroecossystems as breeding more competitive crop varieties is likely
@@
 height and biomass was lower compared to its first cohort. Under crop
-canopy, Palmer amaranth flowering pattern was near to similar in corn
-and soybean. Palmer amaranth growing without crop competition produced
-highest amounts of biomass and less extended flowering pattern. The
-Palmer amaranth strategy in bareground was to invest biomass in growing
-plant width and height. Nonetheless, Palmer amaranth produced 21\% less
-biomass in second cohort compared to first cohort timing. In a
-bareground study, early emerged Palmer amaranth without competition was
-50\% taller than late emerged plants (Webster and Grey, 2015). These
-results suggests that crop competition is not the only factor limiting
-late Palmer amaranth establishment. The limited growth of Palmer
-amaranth at second cohort is likely a reduced plant response to thermal
-units (e.g, growing degree days). It is hypothesize that reduced day
-length contributed to smaller plants at second cohort as well as shorter
-flowering period. A study in North Carolina and Illinois predicted that
-less than 10\% Palmer amaranth seedlings emergence occurred after June
-(Piskackova et al., 2021). In addition, Palmer amaranth negative impact
-on soybean (Korres et al., 2020) and cotton (Webster and Grey, 2015)
-yields was higher when plants were established near to crop planting.
+canopy (e.g., second cohort), Palmer amaranth flowering pattern was near
+to similar in corn and soybean. Palmer amaranth growing without crop
+competition produced highest amounts of biomass and less extended
+flowering pattern. The Palmer amaranth strategy in bareground was to
+invest biomass in growing plant width and height. Nonetheless, Palmer
+amaranth produced 21\% less biomass in second cohort compared to first
+cohort timing. In a bareground study, early emerged Palmer amaranth
+without competition was 50\% taller than late emerged plants (Webster
+and Grey, 2015). These results suggests that crop competition is not the
+only factor limiting late Palmer amaranth establishment. The limited
+growth of Palmer amaranth at second cohort is likely a reduced plant
+response to thermal units (e.g, growing degree days). It is hypothesize
+that reduced day length contributed to smaller plants at second cohort
+as well as shorter flowering period. A study in North Carolina and
+Illinois predicted that less than 10\% Palmer amaranth seedlings
+emergence occurred after June (Piskackova et al., 2021). In addition,
+Palmer amaranth negative impact on soybean (Korres et al., 2020) and
+cotton (Webster and Grey, 2015) yields was higher when plants were
+established near to crop planting.
 
@@
 Therefore, preventing Palmer amaranth seed production or/and seed
-migration to its non-native habitat is a key strategy to minimize weed
-impact into agroecossystem (Davis et al., 2015).
+migration to its non-native habitat is an essential strategy to minimize
+weed impact into agroecossystem (Davis et al., 2015).
 
@@
 to bareground at first cohort but similar to corn at second cohort
-timing. Plant flowering initiation is complex and depends on biological
-and ecological factors (Lang, 1965). We hypothesize that when growing in
-high competition (e.g., second cohort), Palmer amaranth plants tend to
-initiate flowering early, as well as having an extended flowering
-pattern. Early flower initiation is plant strategy when growing under
-stress. For example, when growing under water stress, early flowering in
-Palmer amaranth resulted in a mismatch between female and male plants by
-seven days (Mesgaran et al., 2021). A mismatch in Palmer amaranth male
-and female flowering period can minimize plant outcross, and thus reduce
-plant seed production. Sex dimorphism manipulation is considered a
-potential ecological pest control (McFarlane et al., 2018; Schliekelman
-et al., 2005).
+timing. Plant flowering initiation is complex and it depends on
+biological and ecological factors (Lang, 1965). We hypothesize that when
+growing in high competition (e.g., second cohort), Palmer amaranth
+plants tend to initiate flowering early, as well as having an extended
+flowering pattern. Early flower initiation is plant strategy when
+growing in stress conditions. For example, when growing under water
+stress, early flowering in Palmer amaranth resulted in a mismatch
+between female and male plants by seven days (Mesgaran et al., 2021). A
+mismatch in Palmer amaranth male and female flowering period can
+minimize plant outcross, and thus reduce plant seed production. Sex
+dimorphism manipulation is considered a potential ecological pest
+control (McFarlane et al., 2018; Schliekelman et al., 2005).
 
@@
 (Nguyen et al., 2018). Naturalized of \emph{H. stipulacea} reduced
-female-male ratio to expand into its non-native habitat. Also, biotic
-and/or abiotic stress can influence plant sex determination. Palmer
-amaranth male-to-female ratio was greater under high plant densities
-(Korres and Norsworthy, 2017) and after herbicide application (Rumpa et
-al., 2019). Our model estimated that late flowering, heavier and taller
-Palmer amaranth plants deviated from 1:1 ratio in favor to female
-plants. It was reported that female Palmer amaranth plants invested more
-in height, stem and biomass while male invested more in leaf area and
-leaf dry weight under nutrient deficiency (Korres et al., 2017). Our
-model also estimated more female plants in soybean and bareground
-compared to corn, which might be a result of plant biomass in each
-agroecossystem. Our results showed the influence of life-history and
-ecological traits on sexual dimorphism in Palmer amaranth. Sexual
-dimorphism is documented in other dioecious species (Barrett and Hough,
-2013). For example, stronger female plant competition and greater male
-tolerance to herbivory was reported in \emph{Spinacia oleracea}
-(Pérez-Llorca and Sánchez Vilas, 2019). Research on candidate genes for
-sex determination in \emph{Amaranthus} species are currently underway
-but it is far to complete (Montgomery et al., 2021, 2019). Further
-studies are also needed to understand the ecological basis of Palmer
-amaranth flowering, including the plant behavior under climate change.
+female-male ratio to expand into its non-native habitat (Nguyen et al.,
+2018). Also, biotic and/or abiotic stress can influence plant sex
+determination. Palmer amaranth male-to-female ratio was greater under
+high plant densities (Korres and Norsworthy, 2017) and after herbicide
+application (Rumpa et al., 2019). Our model estimated that late
+flowering, heavier and taller Palmer amaranth plants deviated from 1:1
+ratio in favor to female plants. It was reported that female Palmer
+amaranth plants invested more in height, stem and biomass while male
+invested more in leaf area and leaf dry weight under nutrient deficiency
+(Korres et al., 2017). Our model also estimated more female plants in
+soybean and bareground compared to corn, which might be a result of
+plant biomass in each agroecossystem. Our results showed the influence
+of life-history and ecological traits on sexual dimorphism in Palmer
+amaranth. Sexual dimorphism is documented in other dioecious species
+(Barrett and Hough, 2013). For example, stronger female plant
+competition and greater male tolerance to herbivory was reported in
+\emph{Spinacia oleracea} (Pérez-Llorca and Sánchez Vilas, 2019).
+Research on candidate genes for sex determination in \emph{Amaranthus}
+species are currently underway but it is far to complete (Montgomery et
+al., 2021, 2019). Further studies are also needed to understand the
+ecological basis of Palmer amaranth flowering, including the plant
+behavior under climate change.
 
@@
 
-{\centering \includegraphics[width=120mm,height=180mm]{../data analysis/figures/Figure 4}
+{\centering \includegraphics[width=170mm,height=70mm]{../data analysis/figures/Figure 4}
 

### 2021-08-12 data analysis/gender_analysis.Rmd
oi
jhkj

@@
 ```{r}
+library(emoGG)
 library(tidyverse)
@@
 library(recipes)
-recipe(fem ~ doyh + crop + month + height + weight, data = train) %>% 
+recipe(fem ~ doyh + crop + height + weight, data = train) %>% 
   recipes::step_dummy(all_nominal(), -all_outcomes()) -> recip1
@@
 # Probabilities of female with doy
-visreg(model_logi, "doyh", scale="response", rug=2, 
+visreg(model_logi, "doyh", scale="response", rug=2, gg=TRUE,
        xlab="Day of year",
        ylab="P(female)",
-       ylim = c(0,1)) +
-  theme_test() -> doyf
+       line = list(col="1"),
+       fill=list(fill= "green", alpha = 0.2),
+       points=list(size=1, col = 1, pch=1),
+       ylim = c(0,1)) -> doy_f
+```
+```{r}
+doy_f +
+  scale_y_continuous(limits = c(0, 1)) +
+  scale_x_continuous(limits = c(170, 270), breaks = seq(170, 270, 20)) +
+  theme_test() +
+  geom_emoji(aes(x = 220, y = 0.95), emoji = "2640") +
+  geom_emoji(aes(x = 220, y = 0.05), emoji = "2642") -> doyf
+
+ggsave("figures/fig2.png")
 ```
@@
 # Probabilities of female with weight
-visreg(model_logi, "weight", scale="response", rug=2, 
+visreg(model_logi, "weight", scale="response", rug=2, gg = TRUE,
        xlab=expression(paste("Weight (g plant "^"-1",")")),
        ylab="P(female)",
-       ylim = c(0,1)) +
-  theme_test() -> weight_f
+       line = list(col="1"),
+       fill=list(fill= "green", alpha = 0.2),
+       points=list(size=1, col = 1, pch=1),
+       ylim = c(0,1)) -> weight_f
+```
+```{r}
+weight_f +
+  scale_y_continuous(limits = c(0, 1)) +
+  theme_test() +
+  scale_x_continuous(limits = c(0, 1100), breaks = seq(0, 1100, 150)) +
+  geom_emoji(aes(x = 105, y = 0.95), emoji = "2640") +
+  geom_emoji(aes(x = 105, y = 0.05), emoji = "2642") -> weightf
+
+ggsave("figures/fig1.png")
 ```
@@
 # Probabilities of female with weight
-visreg(model_logi, "height", scale="response", rug=2, 
-       xlab=expression(paste("Height (g plant "^"-1",")")),
+visreg(model_logi, "height", scale="response", rug=2, gg = TRUE,
+       xlab=expression(paste("Height (cm plant "^"-1",")")),
        ylab="P(female)",
-       ylim = c(0,1)) +
-  theme_test() -> height_f
+       line = list(col="1"),
+       fill=list(fill= "green", alpha = 0.2),
+       points=list(size=1, col = 1, pch=1),
+       ylim = c(0,1)) -> height_f
+height_f
+```
+
+```{r}
+height_f +
+  scale_y_continuous(limits = c(0, 1)) +
+  theme_test() +
+  geom_emoji(aes(x = 105, y = 0.95), emoji = "2640") +
+  geom_emoji(aes(x = 105, y = 0.05), emoji = "2642") -> heightf
+
+ggsave("figures/fig.png")
+```
+
+```{r}
+library(patchwork)
+
+doyf + weightf + heightf +
+  plot_annotation(tag_levels = 'A') & 
+  theme(plot.tag = element_text(size = 10, hjust = 0, vjust = 0))
+
+ggsave("figures/Figure 4.png", width = 9, height = 3)
 ```
@@
 # Summary of marginal effect
-summary(effects_logit_dia)
+summary(effects_logit_dia) %>% 
 #For example, the AME value of height is 0.0016 which can be interpreted as a unit increase in height value increases the probability of having female by 0.16%.
+tidy(effects_logit_dia) 
+```
+```{r}
+library(reactablefmtr)
+options(scipen = 999)
+summary(effects_logit_dia) %>%
+#  as.data.frame() %>% 
+  mutate_if(is.double, ~round(., 3)) %>% 
+  rename(Term = factor,
+         `P-value` = p,
+         Lower = lower,
+         Upper = upper) %>% 
+  dplyr::select(Term, AME, SE, Lower, Upper, `P-value`) %>% 
+  reactable()
 ```

### 2021-08-13 Manuscript/Manuscript.Rmd
as

@@
 Seeds from Palmer amaranth growing with limited water conditions were heavier, less dormant, and prompt for germination [@matzrafi2021].
-Also, growing conditions and management can impact sex dimorphism and flowering pattern in Palmer amaranth [@korres2017; @rumpa2019].
+Growing conditions and management practices also influence Palmer amaranth sex dimorphism and flowering pattern [@korres2017; @rumpa2019].
 The continuous global temperature warming can impact agriculture and promote niches for Palmer amaranth invasion/adaptation into new environments.
@@
 For example, herbicide resistant Palmer amaranth is widespread in Nebraska  [@oliveira2021a],  Michigan [@kohrt2017], Connecticut [@aulakh2021], and potential damaged to soybean fields in Illinois is reported [@davis2015].
-In Iowa, a study showed that Palmer amaranth is not well adapted compared to waterhemp (*Amaranthus tuberculatus*) [@baker2021].
+In Iowa, a study showed that Palmer amaranth is still not well adapted compared to waterhemp (*Amaranthus tuberculatus*) [@baker2021].
 Successful cases of Palmer amaranth invasion and near to eradication is documented in Minnesota [@yu2021].
@@
 It can also aid on designing proactive and ecological tactics to limit the species range expansion, reduce its negative impact, and design resilient and sustainable farming systems [@maclaren2020].
-Therefore, the objective of this study was to investigate the flowering pattern, biomass production, and height of Palmer amaranth growing under corn, soybean and fallow at two timings across five locations in the United States Midwest.
+Therefore, the objective of this study was to investigate the flowering pattern, gender, biomass production, and height of Palmer amaranth growing under corn, soybean and fallow at two timings across five locations in the United States Midwest.
 
@@
 Palmer amaranth height was more uniform across cohort timings, except when growing in corn (Figure 3B).
-Palmer amaranth achieve achieve 69.2 cm tall when growing at first cohort timing in bareground, which was not different to 70.7 cm tall at second cohort timing (P = 0.74).
+Palmer amaranth achieved 69.2 cm tall when growing at first cohort timing in bareground, which was not different to 70.7 cm tall at second cohort timing (P = 0.74).
 In addition, no difference in Palmer amaranth height (69.3 cm) was found in soybean at first cohort compared to fallow (P > 0.75).
@@
 These results suggests that Palmer amaranth can fast evolve life-history traits to adapt  into agroecossystems and cultural practices, including nitrogen fertilization [@bravo2018].
-Our results highlight the Palmer amaranth threat to agroecossystems as breeding more competitive crop varieties is likely to select more competitive biotypes [@bravo2017].
+Our results highlight the Palmer amaranth as a threat to agroecossystems as breeding more competitive crop varieties is likely to select more competitive biotypes [@bravo2017].
 
@@
 It was reported that female Palmer amaranth plants invested more in height, stem and biomass while male invested more in leaf area and leaf dry weight under nutrient deficiency [@korres2017].
-Our model also estimated more female plants in soybean and bareground compared to corn, which might be a result of plant biomass in each agroecossystem.
+Our model also estimated more female plants in soybean and bareground compared to corn, which might linked to plant competition strategy in each agroecossystem.
 Our results showed the influence of life-history and ecological traits on sexual dimorphism in Palmer amaranth.

### 2021-08-13 Manuscript/Manuscript.Rmd
as

@@
 ---
+
+```{r include=FALSE}
+library(kableExtra)
+library(tidyverse)
+```
 
@@
 It can also aid on designing proactive and ecological tactics to limit the species range expansion, reduce its negative impact, and design resilient and sustainable farming systems [@maclaren2020].
-Therefore, the objective of this study was to investigate the flowering pattern, gender, biomass production, and height of Palmer amaranth growing under corn, soybean and fallow at two timings across five locations in the United States Midwest.
+Therefore, the objective of this study was to investigate the flowering pattern, gender, biomass production, and height of Palmer amaranth growing under corn, soybean and bareground at two timings across five locations in the United States Midwest.
 
@@
 
+\vspace{0.4cm}
+
 ## Field study {-}
 
-The experiment was conducted in 2018 and 2019 under field conditions at five locations: Arlington (Washington County, Wisconsin), Clay Center (Clay County, Nebraska), Grant (Perkins County, Nebraska), Lincoln (Lancaster County, Nebraska), and Macomb (McDonough County, Illinois).
+The experiment was conducted in 2018 and 2019 under field conditions at five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE ('N, 'W), Grant (('N, 'W)), Lincoln (('N, 'W)), and Macomb (('N, 'W)).
 
-A glyphosate-resistant soybean cultivar (DSR-1950 R2Y at 296,400 seeds ha 1), and a corn hybrid were planted at
-
+Fields were conventionally tilled and disked prior to planting.
+Corn hybrid and soybean varieties were planted in 76-cm row spacing (Table 1).
 Monthly mean air temperature and sum precipitation were obtained using Daymet weather data from June through September across the five locations in 2018 and 2019 [@correndo2021] (Figure 1)
 
-```{r fig1, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/weather/Figure 1.png")
+```{r echo=FALSE, warning=FALSE, message=FALSE}
+options(knitr.kable.NA = '')
+read_csv("../data/table.csv") %>% 
+  janitor::row_to_names(row_number = 1) %>% 
+  kable(booktabs = TRUE, caption = "Field study attributes ",
+        col.names = c("Attributes",
+                           "",
+                           "Arlington, WI",
+                           "Clay Center, NE",
+                           "Grant, NE",
+                      "Lincoln, NE", "Macomb, IL")) %>% 
+#  pack_rows(
+ #   index = c("Bareground" = 2, "Corn" = 4, "Soybean" = 8)) %>% 
+  kable_styling(font_size = 12, 
+                full_width = FALSE, 
+                position = "center",
+                latex_options = c("scale_down", "striped", "hold_position")) %>% 
+  collapse_rows(columns = 1, latex_hline = "none", valign = "middle") 
 ```
@@
 The field experimental unit were six adjacent 9.1 m wide (12 rows at 72.2 cm row spacing) by 10.7 m long.
-Each experimental unit was planted with corn or soybean (DSR-1950 R2Y at 296,400 seeds ha 1), or under fallow condition.
+Each experimental unit was planted with corn or soybean, or under bareground condition.
 Palmer amaranth seedlings (potting mix + two seedlings) were and gently transferring to the ground (6 cm deep and 8 cm wide).
@@
 
+```{r table1, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../data analysis/weather/Figure 1.png")
+```
+
 After transplanting, Palmer amaranth flowering was monitored until the end of the study.
@@
 Then, aboveground plant biomass was harvest near soil surface and oven dried at 65 C until reaching constant weight before the weight of biomass (g plant 1) was recorded.
+
+\vspace{0.4cm}
 
@@
 Plant height and biomass were transformed to meet model assumption of normality.
-In the model, agroecosystem (crop, soybean, fallow) was the fixed effect and year nested with location the random effects.
+In the model, agroecosystem (crop, soybean, bareground) was the fixed effect and year nested with location the random effects.
 Analysis of variance was performed with *anova* function from "car" package [@fox2018].
@@
 
-Palmer amaranth accumulated more biomass when growing in fallow compared to Palmer amaranth growing in soybean and corn (Figure 3A).
-At first cohort time, Palmer amaranth biomass was 75.5 g plant-1, 28.3 g plant-1 and 16.3 g plant-1 in fallow, soybean and corn, respectively.
-At second cohort timing, Palmer amaranth produced 62.6 g plant in fallow, followed by 6.3 g plant in soybean, and 1.4 g plant in corn.
+Palmer amaranth accumulated more biomass when growing in bareground compared to Palmer amaranth growing in soybean and corn (figure 2A).
+At first cohort time, Palmer amaranth biomass was 75.5 g plant-1, 28.3 g plant-1 and 16.3 g plant-1 in bareground, soybean and corn, respectively.
+At second cohort timing, Palmer amaranth produced 62.6 g plant in bareground, followed by 6.3 g plant in soybean, and 1.4 g plant in corn.
 
-Palmer amaranth height was more uniform across cohort timings, except when growing in corn (Figure 3B).
+Palmer amaranth height was more uniform across cohort timings, except when growing in corn (figure 2B).
 Palmer amaranth achieved 69.2 cm tall when growing at first cohort timing in bareground, which was not different to 70.7 cm tall at second cohort timing (P = 0.74).
-In addition, no difference in Palmer amaranth height (69.3 cm) was found in soybean at first cohort compared to fallow (P > 0.75).
+In addition, no difference in Palmer amaranth height (69.3 cm) was found in soybean at first cohort compared to bareground (P > 0.75).
 At second cohort in soybean, Palmer amaranth was near to 10 cm lower compared to its first cohort time (P = 0.04).
@@
 
+\vspace{0.4cm}
+
 ## Palmer amaranth cumulative flowering {-}
 
-Palmer amaranth growing in corn resulted in a longer flowering pattern compared to fallow and soybean at first cohort (Figure 2A).
-Nonetheless, the 10% cumulative Palmer amaranth flowering in soybean, fallow and corn occurred at the end of June.
-Palmer amaranth reached 10% flowering in soybean at doy 180, which was slightly different from fallow (doy 180.9; *P* = 0.01) and corn (doy 181.7; *P* = 0.00).
+Palmer amaranth growing in corn resulted in a longer flowering pattern compared to bareground and soybean at first cohort (figure3A).
+Nonetheless, the 10% cumulative Palmer amaranth flowering in soybean, bareground and corn occurred at the end of June.
+Palmer amaranth reached 10% flowering in soybean at doy 180, which was slightly different from bareground (doy 180.9; *P* = 0.01) and corn (doy 181.7; *P* = 0.00).
 The 50% Palmer amaranth cumulative flowering occurred in July.
-For example, Palmer amaranth reached 50% flowering in fallow at doy 193.4, followed by soybean (doy 194.8), corn (doy 206.6).
+For example, Palmer amaranth reached 50% flowering in bareground at doy 193.4, followed by soybean (doy 194.8), corn (doy 206.6).
 Similar trend was observed at 90% Palmer amaranth cumulative flowering.
-Palmer amaranth growing in corn reached 90% flowering at doy 252.6 (early September), which was 37.8 and 32.2 days after Palmer amaranth 90% flowering in fallow and soybean, respectively.
+Palmer amaranth growing in corn reached 90% flowering at doy 252.6 (early September), which was 37.8 and 32.2 days after Palmer amaranth 90% flowering in bareground and soybean, respectively.
 
-```{r fig2, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/Figure 2.png")
-```
-
-Palmer amaranth cumulative flowering at second cohort ranged from mid July to mid September (Figure 2B).
-Palmer amaranth growing in fallow resulted in earlier flowering time compared to soybean and corn.
-Palmer amaranth growing in fallow reached 10%, 50%, and 90% flowering time at day 203.8, 214.4, and 232.2, respectively.
+Palmer amaranth cumulative flowering at second cohort ranged from mid July to mid September (figure3B).
+Palmer amaranth growing in bareground resulted in earlier flowering time compared to soybean and corn.
+Palmer amaranth growing in bareground reached 10%, 50%, and 90% flowering time at day 203.8, 214.4, and 232.2, respectively.
 Palmer amaranth growing in soybean reached 10% flowering at doy 210.9, which was 6 days prior to corn (*P*-value = 0.00).
@@
 
-```{r eval = FALSE, include = FALSE}
-options(scipen = 7)
-EDcomp(model4, c(90, 90))
-
-result3 %>% 
-  filter(ed == "90" & month == "Second transplanting")
-```
+\vspace{0.4cm}
 
@@
 
-```{r eval = FALSE, include = FALSE}
-multcomp::cld(analysis$emmeans[[2]], 
-              alpha=0.05, Letters=letters, 
-              adjust="none", reversed = TRUE)
+```{r fig2, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../data analysis/figures/Figure 3.png")
 ```
@@
 ```{r fig3, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/Figure 3.png")
+knitr::include_graphics("../data analysis/figures/image.jpeg")
 ```
@@
 ```{r fig4, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/image.jpeg")
+knitr::include_graphics("../data analysis/figures/Figure 2.png")
 ```
@@
 Our study suggests that Palmer amaranth flowering pattern was slightly influenced by cropping systems and cohort timings.
-Palmer amaranth growing in fallow and corn resulted in the overall shortest and longest flowering pattern, respectively.
+Palmer amaranth growing in bareground and corn resulted in the overall shortest and longest flowering pattern, respectively.
 When growing in soybean, Palmer amaranth flowering pattern was similar to bareground at first cohort but similar to corn at second cohort timing.
@@
 Our study demonstrated the short-term Palmer amaranth plasticity to grow and develop into cropping-systems.
-Palmer amaranth management priority should focus on minimizing Palmer amaranth dispersal into new geographies.
-Flower manipulation is a promising ecological Palmer amaranth but it is not well understood.
-Nonetheless, early-season management programs would have a large negative effect in Palmer amaranth growth and development.
+Is likely that Palmer amaranth range will continue to expanding to new geographies.
+Therefore, preventive management is a priority to minimizing Palmer amaranth dispersal.
+Reactive management should focus on early-season management programs, which would have a large negative effect on Palmer amaranth growth and development.
 Tactics that promote early-season crop advantage against Palmer amaranth, including early crop planting, crop rotation [@oliveira2021a], plant width, preemergence applied herbicide [@sanctis2021], and crop residue (e.g.
 cover crops) would minimize the negative impact of Palmer amaranth in agroecossystems.
+The aggressiveness and differential Palmer amaranth adaptation to agroecosystem is striking and require national efforts to minimize the species impact on economy and sustainability.
 
 # Disclosure/Conflict-of-Interest Statement {-}  
-
+Flower manipulation is a promising ecological Palmer amaranth but it is not well understood.
 <!--  
@@
 
-```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "130mm", out.height = "130mm"}
+```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL', out.width = "150mm", out.height = "90mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 
-```{r, Figure-3, ref.label = "fig3", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Palmer amaranth biomass (A) and height (B) growing in corn, fallow, and soybean across Arlington, Clay Center, Grant, Lincoln, and Macomb', out.width = "130mm", out.height = "70mm"}
+```{r, Figure-3, ref.label = "fig3", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Harvest Palmer amaranth plants at 40 days after first transplant timing.
+From left to right, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin', out.width = "90mm", out.height = "90mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 
-```{r, Figure-4, ref.label = "fig4", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Harvest Palmer amaranth plants at 40 days after first transplant timing.
-From left to right, Palmer amaranth growing in fallow, soybean and corn in Arlington, Wisconsin', out.width = "70mm", out.height = "70mm"}
+```{r, Figure-4, ref.label = "fig4", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "150mm", out.height = "150mm"}
 # You can also refer to code chunks from above to place figures at the bottom.

### 2021-08-13 Manuscript/Manuscript.tex
as

@@
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 
@@
 water conditions were heavier, less dormant, and prompt for germination
-(Matzrafi et al., 2021). Also, growing conditions and management can
-impact sex dimorphism and flowering pattern in Palmer amaranth (Korres
-et al., 2017; Rumpa et al., 2019). The continuous global temperature
-warming can impact agriculture and promote niches for Palmer amaranth
-invasion/adaptation into new environments. Currently, it is estimated
-that the greatest climatic risk of Palmer amaranth establishment are
-agronomic crops in Australia and Sub-Sahara Africa (Kistner and
-Hatfield, 2018). Temperature is a key factor limiting Palmer amaranth
-expansion to cooler geographies (Briscoe Runquist et al., 2019);
-however, under future climate change Palmer amaranth is likely to expand
-northward into Canada and Northern Europe (Kistner and Hatfield, 2018;
-Briscoe Runquist et al., 2019).
+(Matzrafi et al., 2021). Growing conditions and management practices
+also influence Palmer amaranth sex dimorphism and flowering pattern
+(Korres et al., 2017; Rumpa et al., 2019). The continuous global
+temperature warming can impact agriculture and promote niches for Palmer
+amaranth invasion/adaptation into new environments. Currently, it is
+estimated that the greatest climatic risk of Palmer amaranth
+establishment are agronomic crops in Australia and Sub-Sahara Africa
+(Kistner and Hatfield, 2018). Temperature is a key factor limiting
+Palmer amaranth expansion to cooler geographies (Briscoe Runquist et
+al., 2019); however, under future climate change Palmer amaranth is
+likely to expand northward into Canada and Northern Europe (Kistner and
+Hatfield, 2018; Briscoe Runquist et al., 2019).
 
@@
 soybean fields in Illinois is reported (Davis et al., 2015). In Iowa, a
-study showed that Palmer amaranth is not well adapted compared to
+study showed that Palmer amaranth is still not well adapted compared to
 waterhemp (\emph{Amaranthus tuberculatus}) (Baker, 2021). Successful
@@
 (MacLaren et al., 2020). Therefore, the objective of this study was to
-investigate the flowering pattern, biomass production, and height of
-Palmer amaranth growing under corn, soybean and fallow at two timings
-across five locations in the United States Midwest.
+investigate the flowering pattern, gender, biomass production, and
+height of Palmer amaranth growing under corn, soybean and bareground at
+two timings across five locations in the United States Midwest.
 
@@
 
+\vspace{0.4cm}
+
 \hypertarget{field-study}{%
@@
 The experiment was conducted in 2018 and 2019 under field conditions at
-five locations: Arlington (Washington County, Wisconsin), Clay Center
-(Clay County, Nebraska), Grant (Perkins County, Nebraska), Lincoln
-(Lancaster County, Nebraska), and Macomb (McDonough County, Illinois).
-
-A glyphosate-resistant soybean cultivar (DSR-1950 R2Y at 296,400 seeds
-ha 1), and a corn hybrid were planted at
-
-Monthly mean air temperature and sum precipitation were obtained using
-Daymet weather data from June through September across the five
+five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE ('N,
+'W), Grant (('N, 'W)), Lincoln (('N, 'W)), and Macomb (('N, 'W)).
+
+Fields were conventionally tilled and disked prior to planting. Corn
+hybrid and soybean varieties were planted in 76-cm row spacing (Table
+1). Monthly mean air temperature and sum precipitation were obtained
+using Daymet weather data from June through September across the five
 locations in 2018 and 2019 (Correndo et al., 2021) (Figure 1)
+
+\begin{table}[!h]
+
+\caption{\label{tab:unnamed-chunk-2}Field study attributes }
+\centering
+\resizebox{\linewidth}{!}{
+\fontsize{12}{14}\selectfont
+\begin{tabular}[t]{lllllll}
+\toprule
+Attributes &  & Arlington, WI & Clay Center, NE & Grant, NE & Lincoln, NE & Macomb, IL\\
+\midrule
+\cellcolor{gray!6}{Bareground} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{}\\
+
+ & Weed control & glyphosate & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
+
+\cellcolor{gray!6}{Corn} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{}\\
+
+ & Hybrid &  & DKC60-67 &  &  & \\
+
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Seeding rate} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{86487} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{}\\
+
+ & Weed control & glyphosate / S-metolachor & S-metolachlor + trazine + mesotrione, + bicyclopyrone &  &  & \\
+
+\cellcolor{gray!6}{Soybean} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{}\\
+
+ & Variety & DSR-1950 & AG21X8 &  &  & \\
+
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Seeding rate} & \cellcolor{gray!6}{296400} & \cellcolor{gray!6}{321237} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{}\\
+
+ & Weed control & glyphosate / S-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
+
+\cellcolor{gray!6}{Planting day} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{May 10 to 20} & \cellcolor{gray!6}{May 10 to 14} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{}\\
+
+Soil &  &  &  &  &  & \\
+
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Type} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{Crete Silt Loam} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{}\\
+
+ & Ratio (sand-clay-silt) &  & 58-25-6.5 &  &  & \\
+
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{pH} & \cellcolor{gray!6}{6.6} & \cellcolor{gray!6}{6.5} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{}\\
+
+ & Organic matter (\%) &  & 3 &  &  & \\
+\bottomrule
+\end{tabular}}
+\end{table}
 
@@
 72.2 cm row spacing) by 10.7 m long. Each experimental unit was planted
-with corn or soybean (DSR-1950 R2Y at 296,400 seeds ha 1), or under
-fallow condition. Palmer amaranth seedlings (potting mix + two
-seedlings) were and gently transferring to the ground (6 cm deep and 8
-cm wide). Twenty-four plants were equidistantly placed (0.76 m apart)
-between rows within each agroecossystems. After a week, one was
-eliminated and one was kept. There were two transplant timing: first
-(June 1) and second (July 1). There were 24 Palmer amaranth plants in
-each experimental unit, with a total of 144 plants for each location.
-The study was repeated twice.
+with corn or soybean, or under bareground condition. Palmer amaranth
+seedlings (potting mix + two seedlings) were and gently transferring to
+the ground (6 cm deep and 8 cm wide). Twenty-four plants were
+equidistantly placed (0.76 m apart) between rows within each
+agroecossystems. After a week, one was eliminated and one was kept.
+There were two transplant timing: first (June 1) and second (July 1).
+There were 24 Palmer amaranth plants in each experimental unit, with a
+total of 144 plants for each location. The study was repeated twice.
 
@@
 recorded.
+
+\vspace{0.4cm}
 
@@
 assumption of normality. In the model, agroecosystem (crop, soybean,
-fallow) was the fixed effect and year nested with location the random
-effects. Analysis of variance was performed with \emph{anova} function
-from ``car'' package (Fox and Weisberg, 2018). Marginal means and
-compact letter display were estimated with \emph{emmeans} and \emph{cld}
-from packages ``emmeans'' and ``multcomp'' (Hothorn et al., 2008).
+bareground) was the fixed effect and year nested with location the
+random effects. Analysis of variance was performed with \emph{anova}
+function from ``car'' package (Fox and Weisberg, 2018). Marginal means
+and compact letter display were estimated with \emph{emmeans} and
+\emph{cld} from packages ``emmeans'' and ``multcomp'' (Hothorn et al.,
+2008).
 
@@
 
-Palmer amaranth accumulated more biomass when growing in fallow compared
-to Palmer amaranth growing in soybean and corn (Figure 3A). At first
-cohort time, Palmer amaranth biomass was 75.5 g plant-1, 28.3 g plant-1
-and 16.3 g plant-1 in fallow, soybean and corn, respectively. At second
-cohort timing, Palmer amaranth produced 62.6 g plant in fallow, followed
-by 6.3 g plant in soybean, and 1.4 g plant in corn.
+Palmer amaranth accumulated more biomass when growing in bareground
+compared to Palmer amaranth growing in soybean and corn (figure 2A). At
+first cohort time, Palmer amaranth biomass was 75.5 g plant-1, 28.3 g
+plant-1 and 16.3 g plant-1 in bareground, soybean and corn,
+respectively. At second cohort timing, Palmer amaranth produced 62.6 g
+plant in bareground, followed by 6.3 g plant in soybean, and 1.4 g plant
+in corn.
 
 Palmer amaranth height was more uniform across cohort timings, except
-when growing in corn (Figure 3B). Palmer amaranth achieve achieve 69.2
-cm tall when growing at first cohort timing in bareground, which was not
+when growing in corn (figure 2B). Palmer amaranth achieved 69.2 cm tall
+when growing at first cohort timing in bareground, which was not
 different to 70.7 cm tall at second cohort timing (P = 0.74). In
 addition, no difference in Palmer amaranth height (69.3 cm) was found in
-soybean at first cohort compared to fallow (P \textgreater{} 0.75). At
-second cohort in soybean, Palmer amaranth was near to 10 cm lower
+soybean at first cohort compared to bareground (P \textgreater{} 0.75).
+At second cohort in soybean, Palmer amaranth was near to 10 cm lower
 compared to its first cohort time (P = 0.04). The tallest and smallest
@@
 cm tall at first cohort and 38.2 cm at second cohort timings.
+
+\vspace{0.4cm}
 
@@
 Palmer amaranth growing in corn resulted in a longer flowering pattern
-compared to fallow and soybean at first cohort (Figure 2A). Nonetheless,
-the 10\% cumulative Palmer amaranth flowering in soybean, fallow and
-corn occurred at the end of June. Palmer amaranth reached 10\% flowering
-in soybean at doy 180, which was slightly different from fallow (doy
-180.9; \emph{P} = 0.01) and corn (doy 181.7; \emph{P} = 0.00). The 50\%
-Palmer amaranth cumulative flowering occurred in July. For example,
-Palmer amaranth reached 50\% flowering in fallow at doy 193.4, followed
-by soybean (doy 194.8), corn (doy 206.6). Similar trend was observed at
-90\% Palmer amaranth cumulative flowering. Palmer amaranth growing in
-corn reached 90\% flowering at doy 252.6 (early September), which was
-37.8 and 32.2 days after Palmer amaranth 90\% flowering in fallow and
-soybean, respectively.
+compared to bareground and soybean at first cohort (figure3A).
+Nonetheless, the 10\% cumulative Palmer amaranth flowering in soybean,
+bareground and corn occurred at the end of June. Palmer amaranth reached
+10\% flowering in soybean at doy 180, which was slightly different from
+bareground (doy 180.9; \emph{P} = 0.01) and corn (doy 181.7; \emph{P} =
+0.00). The 50\% Palmer amaranth cumulative flowering occurred in July.
+For example, Palmer amaranth reached 50\% flowering in bareground at doy
+193.4, followed by soybean (doy 194.8), corn (doy 206.6). Similar trend
+was observed at 90\% Palmer amaranth cumulative flowering. Palmer
+amaranth growing in corn reached 90\% flowering at doy 252.6 (early
+September), which was 37.8 and 32.2 days after Palmer amaranth 90\%
+flowering in bareground and soybean, respectively.
 
 Palmer amaranth cumulative flowering at second cohort ranged from mid
-July to mid September (Figure 2B). Palmer amaranth growing in fallow
+July to mid September (figure3B). Palmer amaranth growing in bareground
 resulted in earlier flowering time compared to soybean and corn. Palmer
-amaranth growing in fallow reached 10\%, 50\%, and 90\% flowering time
-at day 203.8, 214.4, and 232.2, respectively. Palmer amaranth growing in
-soybean reached 10\% flowering at doy 210.9, which was 6 days prior to
-corn (\emph{P}-value = 0.00). Similar trend was observed at 50\%
-flowering, whereas Palmer amaranth reached 50\% flowering in corn (doy
-233.0) 4 days after soybeans (doy 228.9; \emph{P} = 0.00). The 90\%
+amaranth growing in bareground reached 10\%, 50\%, and 90\% flowering
+time at day 203.8, 214.4, and 232.2, respectively. Palmer amaranth
+growing in soybean reached 10\% flowering at doy 210.9, which was 6 days
+prior to corn (\emph{P}-value = 0.00). Similar trend was observed at
+50\% flowering, whereas Palmer amaranth reached 50\% flowering in corn
+(doy 233.0) 4 days after soybeans (doy 228.9; \emph{P} = 0.00). The 90\%
 Palmer amaranth cumulative flowering occurred at same day in corn
 (260.9) and soybean (260.5; \emph{P} = 0.66).
+
+\vspace{0.4cm}
 
@@
 agroecossystems and cultural practices, including nitrogen fertilization
-(Bravo et al., 2018). Our results highlight the Palmer amaranth threat
-to agroecossystems as breeding more competitive crop varieties is likely
-to select more competitive biotypes (Bravo et al., 2017).
+(Bravo et al., 2018). Our results highlight the Palmer amaranth as a
+threat to agroecossystems as breeding more competitive crop varieties is
+likely to select more competitive biotypes (Bravo et al., 2017).
 
@@
 amaranth flowering pattern was slightly influenced by cropping systems
-and cohort timings. Palmer amaranth growing in fallow and corn resulted
-in the overall shortest and longest flowering pattern, respectively.
-When growing in soybean, Palmer amaranth flowering pattern was similar
-to bareground at first cohort but similar to corn at second cohort
-timing. Plant flowering initiation is complex and it depends on
+and cohort timings. Palmer amaranth growing in bareground and corn
+resulted in the overall shortest and longest flowering pattern,
+respectively. When growing in soybean, Palmer amaranth flowering pattern
+was similar to bareground at first cohort but similar to corn at second
+cohort timing. Plant flowering initiation is complex and it depends on
 biological and ecological factors (Lang, 1965). We hypothesize that when
@@
 (Korres et al., 2017). Our model also estimated more female plants in
-soybean and bareground compared to corn, which might be a result of
-plant biomass in each agroecossystem. Our results showed the influence
-of life-history and ecological traits on sexual dimorphism in Palmer
-amaranth. Sexual dimorphism is documented in other dioecious species
-(Barrett and Hough, 2013). For example, stronger female plant
+soybean and bareground compared to corn, which might linked to plant
+competition strategy in each agroecossystem. Our results showed the
+influence of life-history and ecological traits on sexual dimorphism in
+Palmer amaranth. Sexual dimorphism is documented in other dioecious
+species (Barrett and Hough, 2013). For example, stronger female plant
 competition and greater male tolerance to herbivory was reported in
@@
 Our study demonstrated the short-term Palmer amaranth plasticity to grow
-and develop into cropping-systems. Palmer amaranth management priority
-should focus on minimizing Palmer amaranth dispersal into new
-geographies. Flower manipulation is a promising ecological Palmer
-amaranth but it is not well understood. Nonetheless, early-season
-management programs would have a large negative effect in Palmer
-amaranth growth and development. Tactics that promote early-season crop
-advantage against Palmer amaranth, including early crop planting, crop
-rotation (Oliveira et al., 2021), plant width, preemergence applied
-herbicide (Sanctis et al., 2021), and crop residue (e.g.~cover crops)
-would minimize the negative impact of Palmer amaranth in
-agroecossystems.
+and develop into cropping-systems. Is likely that Palmer amaranth range
+will continue to expanding to new geographies. Therefore, preventive
+management is a priority to minimizing Palmer amaranth dispersal.
+Reactive management should focus on early-season management programs,
+which would have a large negative effect on Palmer amaranth growth and
+development. Tactics that promote early-season crop advantage against
+Palmer amaranth, including early crop planting, crop rotation (Oliveira
+et al., 2021), plant width, preemergence applied herbicide (Sanctis et
+al., 2021), and crop residue (e.g.~cover crops) would minimize the
+negative impact of Palmer amaranth in agroecossystems. The
+aggressiveness and differential Palmer amaranth adaptation to
+agroecosystem is striking and require national efforts to minimize the
+species impact on economy and sustainability.
 
@@
 Statement}
+
+Flower manipulation is a promising ecological Palmer amaranth but it is
+not well understood.
 
@@
 
-{\centering \includegraphics[width=160mm,height=100mm]{../data analysis/weather/Figure 1}
+{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 3}
 
@@
 
-\caption{Mean average temperature (C) and montly sum precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-1}
+\caption{Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-2}
 \end{figure}
@@
 
-{\centering \includegraphics[width=130mm,height=130mm]{../data analysis/figures/Figure 2}
+{\centering \includegraphics[width=90mm,height=90mm]{../data analysis/figures/image}
 
@@
 
-\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)}\label{fig:Figure-2}
+\caption{Harvest Palmer amaranth plants at 40 days after first transplant timing.
+From left to right, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin}\label{fig:Figure-3}
 \end{figure}
@@
 
-{\centering \includegraphics[width=130mm,height=70mm]{../data analysis/figures/Figure 3}
+{\centering \includegraphics[width=150mm,height=150mm]{../data analysis/figures/Figure 2}
 
@@
 
-\caption{Palmer amaranth biomass (A) and height (B) growing in corn, fallow, and soybean across Arlington, Clay Center, Grant, Lincoln, and Macomb}\label{fig:Figure-3}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=70mm,height=70mm]{../data analysis/figures/image}
-
-}
-
-\caption{Harvest Palmer amaranth plants at 40 days after first transplant timing. From left to right, Palmer amaranth growing in fallow, soybean and corn in Arlington, Wisconsin}\label{fig:Figure-4}
+\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)}\label{fig:Figure-4}
 \end{figure}

### 2021-08-13 data analysis/Analysis.Rmd
as

@@
 ```
+
+```{r eval=FALSE}
+adaptation_weather %>% 
+  filter(month %in% c(6, 7, 8, 9)) %>% 
+  group_by(month_lab, site, year) %>% 
+  summarise(pp = sum(pp)) %>% 
+  mutate(year = as.character(year),
+         highlight = ifelse(year %in% c("2018", "2019"), TRUE, FALSE),
+         variable = if_else(highlight == TRUE, year, "NA")) %>% 
+  ggplot(aes(x = month_lab, y = pp)) +
+  geom_jitter(aes(color = variable), alpha = 0.2) +
+  scale_color_manual(values = c("red", "blue", "gray90")) +
+  geom_line(adaptation_weather %>% 
+               filter(month %in% c(6, 7, 8, 9) & year %in% c(2018)) %>% 
+               group_by(site, month_lab) %>% 
+               summarise(pp = sum(pp)),
+             mapping = aes(x = month_lab, y = pp, group = 1), 
+             color = "red") +
+  geom_line(adaptation_weather %>% 
+               filter(month %in% c(6, 7, 8, 9) & year %in% c(2019)) %>% 
+               group_by(site, month_lab) %>% 
+               summarise(pp = sum(pp)),
+             mapping = aes(x = month_lab, y = pp, group = 1), 
+             color = "blue") +
+  geom_point(adaptation_weather %>% 
+               filter(month %in% c(6, 7, 8, 9) & year %in% c(2018)) %>% 
+               group_by(site, month_lab) %>% 
+               summarise(pp = sum(pp)),
+             mapping = aes(x = month_lab, y = pp), 
+             fill = "red", shape = 21, size = 3) +
+  geom_point(adaptation_weather %>% 
+               filter(month %in% c(6, 7, 8, 9) & year %in% c(2019)) %>% 
+               group_by(site, month_lab) %>% 
+               summarise(pp = sum(pp)),
+             mapping = aes(x = month_lab, y = pp), 
+             fill = "blue", shape = 21, size = 3) +
+  geom_richtext(data = text3, aes(x = 2.6, y = 330), 
+            label = "Monthly sum precipitation <br>in <b style='color:red;'>2018</b> and <b style='color:blue;'>2019</b>", size = 2, 
+            label.color = NA, fill = NA) +
+  geom_richtext(data = text4, aes(x = 3.6, y = 320), 
+            label = "30 years <br>monthly sum <br>precipitation", size = 2, 
+             label.color = NA, fill = NA, color = "gray90") +
+  geom_curve(data = text4, 
+             x = 2.6, y = 310, 
+             xend = 2, yend = 320,
+             arrow = arrow(length = unit(0.07, "inch")), 
+             size = 0.4, curvature = -0.3, color = "gray90") + 
+  facet_grid(~ site) +
+  labs(x = NULL, y = "Precipitation (mm)") +
+  theme_test() +
+  theme(legend.position = "none",
+        strip.text = element_markdown(size = 25),
+        axis.text.y = element_markdown(size = 25),
+        axis.title = element_markdown(size = 25),
+        axis.text.x = element_markdown(size = 25),
+        legend.text = element_markdown(size = 25)) -> ppp
+
+ggsave("ppp.png", height = 3, width = 6)
+```

### 2021-08-13 data analysis/weather/weather.Rmd
as

@@
          highlight = ifelse(year %in% c("2018", "2019"), TRUE, FALSE),
-         variable = if_else(highlight == TRUE, year, "NA")) 
+         variable = if_else(highlight == TRUE, year, "NA")) -> adaptation_weather1
 ```
@@
 ```{r}
-text1 <- tibble(site = "Arlington")
-text2 <- tibble(site = "Clay Center")
+adaptation_weather1 %>% 
+  filter(year %in% c(2018, 2019)) %>% 
+  group_by(site, year, month) %>% 
+  summarise(prec = sum(pp), tmax = mean(tmax), tmin = mean(tmin), tmean = mean(tmean)) %>% 
+  mutate_if(is.double, ~round(., 1))
+```
+
+```{r}
+adaptation_weather %>% 
+  mutate(site = fct_recode(site, 
+                           "Arlington, WI" = "Arlington",
+                           "Clay Center, NE" = "Clay Center",
+                           "Grant, NE" = "Grant",
+                           "Lincoln, NE" = "Lincoln",
+                           "Macomb, IL" = "Macomb")) -> adaptation_weather
+```
+
+```{r}
+text1 <- tibble(site = "Arlington, WI")
+text2 <- tibble(site = "Clay Center, NE")
 ```
@@
              fill = "blue", shape = 22, size = 3) +
-  geom_richtext(data = text1, aes(x = 2, y = 7), 
-            label = "Mean air temperature <br>in <b style='color:red;'>2018</b> and <b style='color:blue;'>2019</b>", size = 2, 
+  geom_richtext(data = text1, aes(x = 2, y = 7), lineheight = 0.4, 
+            label = "Mean air temperature <br>in <b style='color:red;'>2018</b> and <b style='color:blue;'>2019</b>", size = 6, 
             label.color = NA, fill = NA) +
-  geom_richtext(data = text2, aes(x = 2.2, y = 10), 
-            label = "30 years <br>daily mean <br>air temperature", size = 2, 
+  geom_richtext(data = text2, aes(x = 2.2, y = 10), lineheight = 0.4, 
+            label = "30 years <br>daily mean <br>air temperature", size = 6, 
              label.color = NA, fill = NA, color = "gray90") +
@@
              size = 0.4, curvature = -0.3, color = "gray90") + 
+  scale_y_continuous(limits = c(0, 32)) +
   facet_grid(~ site) +
@@
   theme_test() +
-  theme(legend.position = "none") -> temp
+  theme(legend.position = "none",
+        strip.text = element_markdown(size = 28),
+        axis.text.y = element_markdown(size = 25),
+        axis.title = element_markdown(size = 27),
+        axis.text.x = element_markdown(size = 25),
+        legend.text = element_markdown(size = 25))  -> temp
 
@@
 ```{r}
-text3 <- tibble(site = "Grant")
-text4 <- tibble(site = "Lincoln")
+text3 <- tibble(site = "Grant, NE")
+text4 <- tibble(site = "Lincoln, NE")
 ```
@@
          highlight = ifelse(year %in% c("2018", "2019"), TRUE, FALSE),
-         variable = if_else(highlight == TRUE, year, "NA")) %>% 
+         variable = if_else(highlight == TRUE, year, "NA")) -> adaptation_pp
+```
+
+```{r message = FALSE, warning = FALSE}
+adaptation_pp %>% 
 ggplot(aes(x = month_lab, y = pp)) +
-  geom_jitter(aes(color = variable), alpha = 0.2) +
-  scale_color_manual(values = c("red", "blue", "gray90")) +
-  geom_line(adaptation_weather %>% 
-               filter(month %in% c(6, 7, 8, 9) & year %in% c(2018)) %>% 
-               group_by(site, month_lab) %>% 
-               summarise(pp = sum(pp)),
-             mapping = aes(x = month_lab, y = pp, group = 1), 
-             color = "red") +
-  geom_line(adaptation_weather %>% 
-               filter(month %in% c(6, 7, 8, 9) & year %in% c(2019)) %>% 
-               group_by(site, month_lab) %>% 
-               summarise(pp = sum(pp)),
-             mapping = aes(x = month_lab, y = pp, group = 1), 
-             color = "blue") +
-  geom_point(adaptation_weather %>% 
-               filter(month %in% c(6, 7, 8, 9) & year %in% c(2018)) %>% 
-               group_by(site, month_lab) %>% 
-               summarise(pp = sum(pp)),
-             mapping = aes(x = month_lab, y = pp), 
-             fill = "red", shape = 21, size = 3) +
-  geom_point(adaptation_weather %>% 
-               filter(month %in% c(6, 7, 8, 9) & year %in% c(2019)) %>% 
-               group_by(site, month_lab) %>% 
-               summarise(pp = sum(pp)),
-             mapping = aes(x = month_lab, y = pp), 
-             fill = "blue", shape = 21, size = 3) +
-  geom_richtext(data = text3, aes(x = 2.6, y = 330), 
-            label = "Monthly sum precipitation <br>in <b style='color:red;'>2018</b> and <b style='color:blue;'>2019</b>", size = 2, 
+  geom_bar(data = adaptation_pp %>% filter(year == c(2018, 2019)), 
+           mapping = aes(fill = year),
+           stat = "identity", position = "dodge", alpha = 1) +
+  geom_point(data = adaptation_pp %>% filter(year != c(2018, 2019)), 
+              position = position_jitter(seed = 1),
+              alpha = 0.2, color = "gray50", 
+              show.legend = FALSE) +
+  scale_fill_manual(values = c("red", "blue")) +
+  facet_grid(~ site) +
+  geom_richtext(data = text3, aes(x = 2.6, y = 330), lineheight = 0.4, 
+            label = "Monthly sum precipitation <br>in <b style='color:red;'>2018</b> and <b style='color:blue;'>2019</b>", size = 6, 
             label.color = NA, fill = NA) +
-  geom_richtext(data = text4, aes(x = 3.6, y = 320), 
-            label = "30 years <br>monthly sum <br>precipitation", size = 2, 
+  geom_richtext(data = text4, aes(x = 3.6, y = 320), lineheight = 0.4, 
+            label = "30 years <br>monthly sum <br>precipitation", size = 6, 
              label.color = NA, fill = NA, color = "gray90") +
@@
              x = 2.6, y = 310, 
-             xend = 2, yend = 320,
+             xend = 2.1, yend = 320,
              arrow = arrow(length = unit(0.07, "inch")), 
              size = 0.4, curvature = -0.3, color = "gray90") + 
-  facet_grid(~ site) +
   labs(x = NULL, y = "Precipitation (mm)") +
   theme_test() +
-  theme(legend.position = "none") -> ppp
+  theme(legend.position = "none",
+        strip.text = element_markdown(size = 28),
+        axis.text.y = element_markdown(size = 25),
+        axis.title = element_markdown(size = 27),
+        axis.text.x = element_markdown(size = 25),
+        legend.text = element_markdown(size = 25)) -> ppp
 
-ggsave("ppp.png", height = 3, width = 6)
 ```

### 2021-08-16 Manuscript/Manuscript.Rmd
hgh

@@
  #   index = c("Bareground" = 2, "Corn" = 4, "Soybean" = 8)) %>% 
-  kable_styling(font_size = 12, 
+  kable_styling(font_size = 10, 
                 full_width = FALSE, 
                 position = "center",
-                latex_options = c("scale_down", "striped", "hold_position")) %>% 
+                latex_options = c("scale_down", "hold_position")) %>% 
   collapse_rows(columns = 1, latex_hline = "none", valign = "middle") 
@@
 
-```{r table1, echo = FALSE, message=FALSE, fig.show = "hide"}
+```{r fig1, echo = FALSE, message=FALSE, fig.show = "hide"}
 knitr::include_graphics("../data analysis/weather/Figure 1.png")
@@
 
-$$Y(x) = 0 + (d-0) exp (-exp(b(log(x)-e)))$$
+$Y(x) = 0 + (d-0) exp (-exp(b(log(x)-e)))$
+
 In this model, *Y* is the Palmer amaranth cumulative flowering, *d* is the upper limit (set to 100), and *e* is the XXX, and *x* day of year (doy).
@@
 Palmer amaranth gender was fitted to a binary logistic regression [@bangdiwala2018].
-Binary logistic regression is used for predicting binary classes, such gender of Palmer amaranth plants.
+Binary logistic regression is used for predicting binary classes, such as the probability of female in dioecious plants.
 Prior to the analysis, all missing values were removed from the dataset.
@@
 
-The model fit was 0.11, 0.14, 0.18 with using pseudo R squared test from McFadden, Cox and Snell, and Cragg and Uhler, respectively.
+The model fit was 0.23, 0.32, 0.40 with using pseudo R squared test from McFadden, Cox and Snell, and Cragg and Uhler, respectively.
 The likelihood ratio test showed a p-value of < 0.00.
-The average marginal effects showed that increasing a weight unit by 1 g increases the probability of having a female plant by 0.08% (P = 0.01).
-Similar trend is observed to height as well as doy, whereas the probability of being female increase by 0.2% (P = 0.01) and 0.4% (P = 0.00) when a unit of height (cm) and day increases, respectively.
+The average marginal effects showed that Palmer amaranth growing in corn resulted in 14.5% less females plants (Table 2).
+Moreover, increasing a unit doy  increases the probability of having a female plant by 0.4% (Table 2 and Figure4A).
+Similar trend is observed to weight as well as height, whereas the probability of being female increase by 0.2% (Figure 4B) and 0.1% (Figure 4C) when a unit of weight (g) and height (cm) increases, respectively.
 
-The model accuracy evaluation accuracy in the unseen 20% dataset was 0.61 with a cutoff value for female and male plants of 0.49.
-The model classification showed a precision of 0.59, recall of 0.76, and a F1 score of 0.66.
-In addition, the area under the curve was 0.61.
+```{r echo = FALSE, warning = FALSE, message=FALSE}
+read_csv("../data/table_a.csv") -> table_a
+
+#names(table_a)[2] <- paste0(names(table_a)[2], 
+#                                footnote_marker_symbol(1))
+
+table_a %>% 
+  kable(booktabs = TRUE, escape = TRUE, caption = "Average marginal means of gender logistic model. Factor pararemter values (e.g. crop) is shown related to soybean.") %>% 
+  kable_styling(font_size = 10, 
+                full_width = FALSE, 
+                position = "center",
+                latex_options = c("hold_position")) %>% 
+  collapse_rows(columns = 1, latex_hline = "none", valign = "middle") %>%
+  footnote(
+           alphabet = c("Average Marginal Effects.", "Standard Error."),
+           footnote_as_chunk = TRUE)
+```
+
+The model accuracy evaluation accuracy in the unseen 20% dataset was 0.62 with a cutoff value for female and male plants of 0.43.
+The model classification showed a precision of 0.64, recall of 0.66, and a F means score of 0.65.
+In addition, the area under the curve was 0.64.
 

### 2021-08-16 Manuscript/Manuscript.tex
hgh

@@
 \resizebox{\linewidth}{!}{
-\fontsize{12}{14}\selectfont
+\fontsize{10}{12}\selectfont
 \begin{tabular}[t]{lllllll}
@@
 \midrule
-\cellcolor{gray!6}{Bareground} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{}\\
+Bareground &  &  &  &  &  & \\
 
@@
 
-\cellcolor{gray!6}{Corn} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{}\\
+Corn &  &  &  &  &  & \\
 
@@
 
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Seeding rate} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{86487} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{}\\
-
- & Weed control & glyphosate / S-metolachor & S-metolachlor + trazine + mesotrione, + bicyclopyrone &  &  & \\
-
-\cellcolor{gray!6}{Soybean} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{}\\
+ & Seeding rate &  & 86487 &  &  & \\
+
+\multirow{-3}{*}{\raggedright\arraybackslash } & Weed control & glyphosate / S-metolachor & S-metolachlor + trazine + mesotrione, + bicyclopyrone &  &  & \\
+
+Soybean &  &  &  &  &  & \\
 
@@
 
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Seeding rate} & \cellcolor{gray!6}{296400} & \cellcolor{gray!6}{321237} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{}\\
-
- & Weed control & glyphosate / S-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
-
-\cellcolor{gray!6}{Planting day} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{May 10 to 20} & \cellcolor{gray!6}{May 10 to 14} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{}\\
+ & Seeding rate & 296400 & 321237 &  &  & \\
+
+\multirow{-3}{*}{\raggedright\arraybackslash } & Weed control & glyphosate / S-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
+
+Planting day &  & May 10 to 20 & May 10 to 14 &  &  & \\
 
@@
 
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Type} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{Crete Silt Loam} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{}\\
+ & Type &  & Crete Silt Loam &  &  & \\
 
@@
 
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{pH} & \cellcolor{gray!6}{6.6} & \cellcolor{gray!6}{6.5} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{}\\
-
- & Organic matter (\%) &  & 3 &  &  & \\
+ & pH & 6.6 & 6.5 &  &  & \\
+
+\multirow{-4}{*}{\raggedright\arraybackslash } & Organic matter (\%) &  & 3 &  &  & \\
 \bottomrule
@@
 
-\[Y(x) = 0 + (d-0) exp (-exp(b(log(x)-e)))\] In this model, \emph{Y} is
-the Palmer amaranth cumulative flowering, \emph{d} is the upper limit
-(set to 100), and \emph{e} is the XXX, and \emph{x} day of year (doy).
+\(Y(x) = 0 + (d-0) exp (-exp(b(log(x)-e)))\)
+
+In this model, \emph{Y} is the Palmer amaranth cumulative flowering,
+\emph{d} is the upper limit (set to 100), and \emph{e} is the XXX, and
+\emph{x} day of year (doy).
 
@@
 (Bangdiwala, 2018). Binary logistic regression is used for predicting
-binary classes, such gender of Palmer amaranth plants. Prior to the
-analysis, all missing values were removed from the dataset. Also, data
-from Grant was not used in this analysis due to the uniform plant
-harvesting at that location. The whole dataset was splitted into 80\%
-train and 20\% test data. The 80\% train is used for the model training
-and the rest 20\% is used for checking how the model generalized on
-unseen dataset. With 80\% dataset, a binary response variable, male (0)
-and female (1), was fitted to a generalized linear model (\emph{glm}
-function) including day of year harvest, height, weight, crop and month
-as independent variables. The model family was binomial with a logit
-function. The model fit was assessed through pseudo R squared values
-(McFadden, Cox and Snell, Cragg and Uhler) and likelihood ratio using
-\emph{nagelkerke} function (``rcompanion'' package). The marginal
-effects computation was performed with Average Marginal Effects at every
-observed value of x and average across the results (AMEs) (Leeper, 2017)
-using \emph{margins} function from ``margins'' package. The rest 20\%
-dataset was predicted using \emph{predict} function with a cutoff
-estimation for male or female using \emph{performance} function. The
-model quality prediction from the classification algorithm was measure
-with precision (\emph{precision} function), recall (\emph{recall}
-function) and F1 score (\emph{f\_meas} function) using ``yardstick''
-package. The precision determines the accuracy of positive predictions
-(female plants), recall determines the fraction of positives that were
-correctly identified, and F1 score is a weighted harmonic mean of
-precision and recall with the best score of 1 and the worst score of 0.
-F1 score conveys the balance between the precision and the recall.
+binary classes, such as the probability of female in dioecious plants.
+Prior to the analysis, all missing values were removed from the dataset.
+Also, data from Grant was not used in this analysis due to the uniform
+plant harvesting at that location. The whole dataset was splitted into
+80\% train and 20\% test data. The 80\% train is used for the model
+training and the rest 20\% is used for checking how the model
+generalized on unseen dataset. With 80\% dataset, a binary response
+variable, male (0) and female (1), was fitted to a generalized linear
+model (\emph{glm} function) including day of year harvest, height,
+weight, crop and month as independent variables. The model family was
+binomial with a logit function. The model fit was assessed through
+pseudo R squared values (McFadden, Cox and Snell, Cragg and Uhler) and
+likelihood ratio using \emph{nagelkerke} function (``rcompanion''
+package). The marginal effects computation was performed with Average
+Marginal Effects at every observed value of x and average across the
+results (AMEs) (Leeper, 2017) using \emph{margins} function from
+``margins'' package. The rest 20\% dataset was predicted using
+\emph{predict} function with a cutoff estimation for male or female
+using \emph{performance} function. The model quality prediction from the
+classification algorithm was measure with precision (\emph{precision}
+function), recall (\emph{recall} function) and F1 score (\emph{f\_meas}
+function) using ``yardstick'' package. The precision determines the
+accuracy of positive predictions (female plants), recall determines the
+fraction of positives that were correctly identified, and F1 score is a
+weighted harmonic mean of precision and recall with the best score of 1
+and the worst score of 0. F1 score conveys the balance between the
+precision and the recall.
 
@@
 
-The model fit was 0.11, 0.14, 0.18 with using pseudo R squared test from
+The model fit was 0.23, 0.32, 0.40 with using pseudo R squared test from
 McFadden, Cox and Snell, and Cragg and Uhler, respectively. The
 likelihood ratio test showed a p-value of \textless{} 0.00. The average
-marginal effects showed that increasing a weight unit by 1 g increases
-the probability of having a female plant by 0.08\% (P = 0.01). Similar
-trend is observed to height as well as doy, whereas the probability of
-being female increase by 0.2\% (P = 0.01) and 0.4\% (P = 0.00) when a
-unit of height (cm) and day increases, respectively.
+marginal effects showed that Palmer amaranth growing in corn resulted in
+14.5\% less females plants (Table 2). Moreover, increasing a unit doy
+increases the probability of having a female plant by 0.4\% (Table 2 and
+Figure4A). Similar trend is observed to weight as well as height,
+whereas the probability of being female increase by 0.2\% (Figure 4B)
+and 0.1\% (Figure 4C) when a unit of weight (g) and height (cm)
+increases, respectively.
+
+\begin{table}[!h]
+
+\caption{\label{tab:unnamed-chunk-3}Average marginal means of gender logistic model. Factor pararemter values (e.g. crop) is shown related to soybean.}
+\centering
+\fontsize{10}{12}\selectfont
+\begin{tabular}[t]{lrrrrrr}
+\toprule
+Term & AME & SE & Lower & Upper & Z-score & P-value\\
+\midrule
+crop\_bareground & -0.048 & 0.054 & -0.154 & 0.059 & -0.876 & 0.381\\
+
+crop\_corn & -0.148 & 0.052 & -0.250 & -0.046 & -2.842 & 0.004\\
+
+doyh & 0.004 & 0.001 & 0.003 & 0.006 & 4.959 & 0.000\\
+
+height & 0.002 & 0.001 & 0.001 & 0.003 & 2.953 & 0.003\\
+
+weight & 0.001 & 0.000 & 0.000 & 0.001 & 2.179 & 0.029\\
+\bottomrule
+\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a} Average Marginal Effects. \textsuperscript{b} Standard Error.}\\
+\end{tabular}
+\end{table}
 
 The model accuracy evaluation accuracy in the unseen 20\% dataset was
-0.61 with a cutoff value for female and male plants of 0.49. The model
-classification showed a precision of 0.59, recall of 0.76, and a F1
-score of 0.66. In addition, the area under the curve was 0.61.
+0.62 with a cutoff value for female and male plants of 0.43. The model
+classification showed a precision of 0.64, recall of 0.66, and a F means
+score of 0.65. In addition, the area under the curve was 0.64.
 
@@
 
+{\centering \includegraphics[width=160mm,height=100mm]{../data analysis/weather/Figure 1}
+
+}
+
+\caption{Mean average temperature (C) and montly sum precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-1}
+\end{figure}
+
+\begin{figure}
+
 {\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 3}

### 2021-08-16 data analysis/gender_analysis.Rmd
hgh

@@
   na.omit(data2) -> data2
+```
+
+```{r}
+data2 %>% 
+  distinct(location)
 ```
@@
          `P-value` = p,
+         `Z-score` = z,
          Lower = lower,
          Upper = upper) %>% 
-  dplyr::select(Term, AME, SE, Lower, Upper, `P-value`) %>% 
-  reactable()
+  dplyr::select(Term, AME, SE, Lower, Upper, `Z-score`, `P-value`) -> table_a
+
+write_csv(table_a, "../data/table_a.csv")
 ```
@@
 # predict the preproc_test dataset
+preproc_test <- na.omit(preproc_test)
 pred <- predict(model_logi, preproc_test, type="response") 
@@
 ```{r}
+#preproc_test <- na.omit(preproc_test)
 pred.rocr <- prediction(pred, preproc_test$fem)

### 2021-08-16 data analysis/weather/weather.Rmd
hgh

@@
 adaptation_weather %>% 
+  filter(month %in% c(6, 7, 8, 9)) %>% 
+  mutate(year = as.character(year),
+         highlight = ifelse(year %in% c("2018", "2019"), TRUE, FALSE),
+         variable = if_else(highlight == TRUE, year, "NA")) %>% 
   mutate(site = fct_recode(site, 
@@
                            "Lincoln, NE" = "Lincoln",
-                           "Macomb, IL" = "Macomb")) -> adaptation_weather
+                           "Macomb, IL" = "Macomb")) -> adaptation_weather1
+```
+
+```{r}
+adaptation_weather %>% 
+  filter(month %in% c(6, 7, 8, 9)) %>% 
+  group_by(month_lab, site) %>% 
+  summarise(tmean = mean(tmean))
 ```
@@
 ```{r}
-adaptation_weather %>% 
-  filter(month %in% c(6, 7, 8, 9)) %>% 
-  mutate(year = as.character(year),
-         highlight = ifelse(year %in% c("2018", "2019"), TRUE, FALSE),
-         variable = if_else(highlight == TRUE, year, "NA")) %>% 
+adaptation_weather1 %>% 
   ggplot(aes(x = month_lab, y = tmean)) +
@@
   scale_color_manual(values = c("red", "blue", "gray90")) +
-  geom_line(adaptation_weather %>% 
-               filter(month %in% c(6, 7, 8, 9) & year %in% c(2018)) %>% 
+  facet_grid(~ site) +
+  geom_line(adaptation_weather1 %>% 
+               filter(year %in% c(2018)) %>% 
                group_by(site, month_lab) %>% 
@@
              color = "red") +
-  geom_line(adaptation_weather %>% 
-               filter(month %in% c(6, 7, 8, 9) & year %in% c(2019)) %>% 
+  geom_line(adaptation_weather1 %>% 
+               filter(year %in% c(2019)) %>% 
                group_by(site, month_lab) %>% 
@@
              color = "blue") +
-  geom_point(adaptation_weather %>% 
-               filter(month %in% c(6, 7, 8, 9) & year %in% c(2018)) %>% 
+  geom_point(adaptation_weather1 %>% 
+               filter(year %in% c(2018)) %>% 
                group_by(site, month_lab) %>% 
@@
              fill = "red", shape = 22, size = 3) +
-  geom_point(adaptation_weather %>% 
-               filter(month %in% c(6, 7, 8, 9) & year %in% c(2019)) %>% 
+  geom_point(adaptation_weather1 %>% 
+               filter(year %in% c(2019)) %>% 
                group_by(site, month_lab) %>% 
@@
   scale_y_continuous(limits = c(0, 32)) +
-  facet_grid(~ site) +
   labs(x = NULL, y = "Temperature (C)") +
@@
   theme(legend.position = "none",
-        strip.text = element_markdown(size = 28),
-        axis.text.y = element_markdown(size = 25),
-        axis.title = element_markdown(size = 27),
-        axis.text.x = element_markdown(size = 25),
-        legend.text = element_markdown(size = 25))  -> temp
+#        strip.text = element_markdown(size = 15, hjust = 0.5),
+#        axis.text.y = element_markdown(size = 25),
+#        axis.title = element_markdown(size = 27),
+#        axis.text.x = element_markdown(size = 25)#,
+#        legend.text = element_markdown(size = 25)
+        )  -> temp
 
 ggsave("weather.png", height = 3, width = 6)
+```
+
+```{r}
+
 ```
@@
          highlight = ifelse(year %in% c("2018", "2019"), TRUE, FALSE),
-         variable = if_else(highlight == TRUE, year, "NA")) -> adaptation_pp
+         variable = if_else(highlight == TRUE, year, "NA")) %>% 
+  mutate(site = fct_recode(site, 
+                           "Arlington, WI" = "Arlington",
+                           "Clay Center, NE" = "Clay Center",
+                           "Grant, NE" = "Grant",
+                           "Lincoln, NE" = "Lincoln",
+                           "Macomb, IL" = "Macomb")) -> adaptation_pp
 ```
@@
   geom_richtext(data = text3, aes(x = 2.6, y = 330), lineheight = 0.4, 
-            label = "Monthly sum precipitation <br>in <b style='color:red;'>2018</b> and <b style='color:blue;'>2019</b>", size = 6, 
+            label = "Monthly sum precipitation <br>in <b style='color:red;'>2018</b> and <b style='color:blue;'>2019</b>", size = 2, 
             label.color = NA, fill = NA) +
   geom_richtext(data = text4, aes(x = 3.6, y = 320), lineheight = 0.4, 
-            label = "30 years <br>monthly sum <br>precipitation", size = 6, 
+            label = "30 years <br>monthly sum <br>precipitation", size = 2, 
              label.color = NA, fill = NA, color = "gray90") +
@@
   theme(legend.position = "none",
-        strip.text = element_markdown(size = 28),
-        axis.text.y = element_markdown(size = 25),
-        axis.title = element_markdown(size = 27),
-        axis.text.x = element_markdown(size = 25),
-        legend.text = element_markdown(size = 25)) -> ppp
+#        strip.text = element_markdown(size = 28),
+#        axis.text.y = element_markdown(size = 25),
+#        axis.title = element_markdown(size = 27),
+#        axis.text.x = element_markdown(size = 25),
+#        legend.text = element_markdown(size = 25)
+  ) -> ppp
 

### 2021-08-16 data analysis/layout.Rmd
b

---
title: "layout"
author: "Maxwel Coura Oliveira"
date: "8/16/2021"
output: html_document
---

```{r}
library(tidyverse)
library(ggimage)
library(ggtext)
```

```{r}
read_csv("../data/layout.csv") %>% 
  janitor::row_to_names(row_number = 1) %>% 
  mutate(distance = as.double(distance)) %>% 
  mutate(emoji = case_when(
    crop == "corn" ~ "https://img.icons8.com/color/48/000000/corn.png",
    crop == "soybean" ~ "https://img.icons8.com/color/48/000000/soy.png",
    TRUE ~ NA_character_)) %>% 
  mutate(weed = "https://hotemoji.com/images/emoji/0/1gqjxj81p5mvj0.png") -> layout
```

```{r}
layout %>% 
  ggplot(aes(x = row, y = distance)) +
  facet_grid(~crop) +
#  geom_point() +
#  scale_x_discrete(limits = c(-0.5, 4.5)) +
  geom_image(data = layout,
               aes(image=weed)) +
  geom_image(data = layout %>% filter(crop == "corn"),
               aes(x = 0.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "corn"),
               aes(x = 1.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "corn"),
               aes(x = 2.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "corn"),
               aes(x = 3.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "corn"),
               aes(x = 4.5, image=emoji)) +
  labs(x= NULL, y = NULL) +
  theme_test() +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        strip.text = element_markdown(size = 18),
        panel.background = element_rect(fill = "#f7e4d7")) -> fig1
```

```{r}
library(ggimage)
fig1 +
  geom_image(data = layout %>% filter(crop == "soybean"),
               aes(x = 0.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "soybean"),
               aes(x = 1.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "soybean"),
               aes(x = 2.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "soybean"),
               aes(x = 3.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "soybean"),
               aes(x = 4.5, image=emoji))

ggsave("figures/layout.png", height = 4, width = 9)
```

### 2021-08-17 data analysis/layout.Rmd
gh

@@
   geom_image(data = layout %>% filter(crop == "soybean"),
-               aes(x = 4.5, image=emoji))
+               aes(x = 4.5, image=emoji)) -> fig2
+```
+
+```{r}
+palmer <- tibble(crop = "bareground")
+corn <- tibble(crop = "corn")
+soybean <- tibble(crop = "soybean")
+```
+
+```{r}
+fig2 +
+  geom_text(data = palmer, aes(x = 2.7, y = 270), size = 2, hjust = 0, 
+            fontface = "italic",
+            label = "24 Palmer amaranth plants \nwere placed 76.2 cm \nat the same distance") +
+  geom_text(data = palmer, aes(x = 2.3, y = 228), size = 2,
+            label = "76.2 cm", fontface = "bold") +
+  geom_text(data = palmer, aes(x = 2.5, y = 390), size = 2,
+            label = "Palmer amaranth \nplants", fontface = "bold") +
+  geom_text(data = corn, aes(x = 2.5, y = 400), size = 2,
+            label = "Corn plants", fontface = "bold") +
+  geom_text(data = soybean, aes(x = 2.5, y = 400), size = 2,
+            label = "Soybean plants", fontface = "bold") -> fig3
+
+#ggsave("figures/layout.png", height = 4, width = 9)
+```
+
+```{r}
+fig3 +
+  geom_segment(data = palmer, mapping = aes(x = 2.1, y = 264, 
+                                            xend = 2.85, yend = 230),
+               linetype = "dotted") +
+  geom_segment(data = palmer, mapping = aes(x = 2.1, y = 188, 
+                                            xend = 2.85, yend = 228),
+               linetype = "dotted") +
+  geom_segment(data = palmer, mapping = aes(x = 2, y = 250, 
+                                            xend = 2, yend = 198),
+               linetype = "dotted") -> fig4
+
+#ggsave("figures/layout.png", height = 4, width = 9)
+```
+
+```{r}
+fig4 +
+  geom_curve(data = palmer, aes(x = 2.5, y = 405, 
+                               xend = 2.1, yend = 420),
+             arrow = arrow(length = unit(0.07, "inch")), 
+               size = 0.4, curvature = 0.3) +
+  geom_curve(data = palmer, aes(x = 2.5, y = 375, 
+                               xend = 2.85, yend = 375),
+             arrow = arrow(length = unit(0.07, "inch")), 
+               size = 0.4, curvature = 0.3) +
+  geom_curve(data = corn, aes(x = 2.3, y = 405, 
+                               xend = 2.4, yend = 420),
+             arrow = arrow(length = unit(0.07, "inch")), 
+               size = 0.4, curvature = -0.3) +
+  geom_curve(data = corn, aes(x = 2.7, y = 395, 
+                               xend = 2.6, yend = 380),
+             arrow = arrow(length = unit(0.07, "inch")), 
+               size = 0.4, curvature = -0.3) + 
+  geom_curve(data = soybean, aes(x = 2.3, y = 405, 
+                               xend = 2.4, yend = 420),
+             arrow = arrow(length = unit(0.07, "inch")), 
+               size = 0.4, curvature = -0.3) +
+  geom_curve(data = soybean, aes(x = 2.7, y = 395, 
+                               xend = 2.6, yend = 380),
+             arrow = arrow(length = unit(0.07, "inch")), 
+               size = 0.4, curvature = -0.3)
 

### 2021-08-23 Manuscript/Manuscript.Rmd
vb

@@
 ---
-title: "Palmer amaranth (*Amaranthus palmeri*) adaptation to agroecossystems"
+title: "Palmer amaranth (*Amaranthus palmeri*) adaptation to US Midwest agroecosystems"
 runtitle: "Palmer amaranth adaptation"
@@
 Palmer amaranth (*Amaranthus palmeri* S.
-Watson) is currently considered one of the most economically damaged weed species to cropping systems in the United States.
-Unmanaged Palmer amaranth in competition for water, light and nutrients can drastically impact on crop yields [@berger2015].
-For example, Palmer amaranth is documented with potential to reduce 91%, 68%, and 54% of corn [@massinga2001], soybean [@klingaman1994], and cotton [@morgan2001] yields, respectively.
-Moreover, Palmer amaranth has showed a remarkable capacity to evolve resistance to herbicides.
+Watson) is currently ranked as one of the most economically detrimental weed species to cropping systems in the United States (ADD CITATION; WSSA Survey Perhaps).
+Unmanaged Palmer amaranth plants compete for water, light, and nutrients, which can drastically impact crop yields [@berger2015].
+For example, Palmer amaranth has been documented to reduce up to 91%, 68%, and 54% corn [@massinga2001], soybean [@klingaman1994], and cotton [@morgan2001] yields, respectively.
+Moreover, Palmer amaranth has shown a remarkable capacity to evolve resistance to herbicides.
 To date, Palmer amaranth has evolved resistance to eight herbicide sites of action [@heap2021], increasing the weed management complexity [@lindsay2017].
-Thus, Palmer amaranth poses an economical risk to sustainable agriculture.
+Thus, Palmer amaranth poses an economical and environmental risk to sustainable agriculture.
 
-Palmer amaranth is a fast growing summer annual forb indigenous to Sonoran Desert [@sauer1957].
+Palmer amaranth is a fast growing summer annual forb indigenous to the Sonoran Desert [@sauer1957].
 The species would eventually emerge as a threat to US agriculture in the 1990s.
-Palmer amaranth weediness is likely a result of human-assisted selection in combination with species biology.
-Farm mechanization, conservation agriculture (e.g., no-till), and reliance on herbicides for weed management are the main human-mediated selection of Palmer amaranth into cropping systems [@ward2013].
+Palmer amaranth weediness is likely a result of human-assisted selection in combination with plant biology.
+Farm mechanization, adoption of conservation agriculture (e.g., no-till), and reliance on herbicides for weed management are the main human-mediated selection of Palmer amaranth into cropping systems [@ward2013].
 On the other hand, Palmer amaranth is a prolific seed producer with a C4 photosynthetic apparatus [@wang1992].
 With a dioecy nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging adaptive traits among plants [@oliveira2018].
-Also, Palmer amaranth small seed size (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure [@hartzler2016], animals [@farmer2017], and plant propagules [@yu2021].
-The dispersal capacity of Palmer amaranth make the species one of the most successful cases of weed adaption to cropping systems.
+Also, Palmer amaranth small seed size (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure [@hartzler2016], animals [@farmer2017], etc..
+The dispersal capacity of Palmer amaranth makes the species one of the most successful cases of weed adaptation to current cropping systems.
 
-Light and temperature are likely the main environment requirements for Palmer amaranth successful grow and development.
+Light and temperature are likely the main environment requirements for Palmer amaranth successful grow and development (CITATION).
 Palmer amaranth is reported with an extended germination period [@jha2010].
 Germination of Palmer amaranth is triggered by 18 C soil temperature [@keeley1987], and optimal germination and biomass production occur at 35/30 C day and night temperatures [@guo2003].
-Palmer amaranth germination is also human-mediated by tillage timings and preemergence-applied herbicides [@chahal2021].
+In addition, Palmer amaranth establishment is human-mediated by tillage timings and preemergence-applied herbicides [@chahal2021], which can result in weed germination shifts [@sbatella2010].
 Water has not shown to limit Palmer amaranth fitness.
@@
 Growing conditions and management practices also influence Palmer amaranth sex dimorphism and flowering pattern [@korres2017; @rumpa2019].
-The continuous global temperature warming can impact agriculture and promote niches for Palmer amaranth invasion/adaptation into new environments.
+Therefore, Palmer amaranth has shown plasticity to evolve and adapt into different environments.
+The continuous global temperature warming is likely impact agriculture and promote niches for Palmer amaranth invasion/adaptation into new environments.
 Currently, it is estimated that the greatest climatic risk of Palmer amaranth establishment are agronomic crops in Australia and Sub-Sahara Africa [@kistner2018].
@@
 Palmer amaranth is already found in agronomic crops of South America [@larran2017; @kupper2017] and Southern Europe [@milani2021].
-In the US, Palmer amaranth is well established at cotton fields [@garetson2019] and non-crop areas [@bagavathiannan2016] in the southern United States but its range is expanding northward.
-For example, herbicide resistant Palmer amaranth is widespread in Nebraska  [@oliveira2021a],  Michigan [@kohrt2017], Connecticut [@aulakh2021], and potential damaged to soybean fields in Illinois is reported [@davis2015].
+In the United States, Palmer amaranth is well established in the Cotton Belt [@garetson2019,  [@bagavathiannan2016] in the southern United States but its range is expanding northward.
+For example, herbicide resistant Palmer amaranth is widespread in Nebraska  [@oliveira2021a].
+There are some reported cases of Palmer amaranth in Michigan [@kohrt2017] and Connecticut [@aulakh2021].
+Also, is estimated that Palmer amaranth can cause high damage to soybean fields in Illinois [@davis2015], which is concerning as soybean along with corn make most of US Midwest agronomic hectares.
 In Iowa, a study showed that Palmer amaranth is still not well adapted compared to waterhemp (*Amaranthus tuberculatus*) [@baker2021].
-Successful cases of Palmer amaranth invasion and near to eradication is documented in Minnesota [@yu2021].
-No Palmer amaranth actively growing was found in Canada; however, Palmer amaranth seeds was detected in sweet potato slips [@page2021].
-Nonetheless, it seems fated the need to manage Palmer amaranth in agronomic crops throughout multiple environments in the near future.
+Palmer amaranth invasion and successful eradication is documented in Minnesota [@yu2021].
+Palmer amaranth infestations have not been detected in Canada; however, Palmer amaranth seeds were detected in sweet potato slips in the country [@page2021].
+Palmer amaranth is still not as well adapted and established to Northern as it is in the Southern North America.
+Nonetheless, its range of expansion into new environment is increasing.
+It seems fated the need to manage Palmer amaranth in agronomic crops throughout multiple environments in the near future.
 Strategies on Palmer amaranth management should encompass the agroecosystem level but not only attempts to eradicate the weed.
@@
 
-In the southeastern US, early growing Palmer amaranth is well known to have higher impact on cotton yields than late established plants [@macrae2013].
-In the northern states, Palmer amaranth impact on the agroecossystem is recent.
-Studies investigating Palmer amaranth in those locations is limited due to the plant classification as noxious weed species [@yu2021].
-Nonetheless, the continuous Palmer amaranth dispersal and potential establishment across northern United States is concerning and warrant investigations on species morphology in such environments.
-Understanding Palmer amaranth biology and growing strategies under different agroecossystems can enhance our knowledge on species adaptation.
-It can also aid on designing proactive and ecological tactics to limit the species range expansion, reduce its negative impact, and design resilient and sustainable farming systems [@maclaren2020].
-Therefore, the objective of this study was to investigate the flowering pattern, gender, biomass production, and height of Palmer amaranth growing under corn, soybean and bareground at two timings across five locations in the United States Midwest.
+In the southeastern US, early growing Palmer amaranth is well known to have a higher impact on cotton yields compared to late established plants [@macrae2013].
+In the northern states, Palmer amaranth impact on the agroecosystem is recent.
+Studies investigating Palmer amaranth in those locations are limited due to the plant classification as noxious weed species in some northern states [@yu2021].
+Nonetheless, the continuous Palmer amaranth dispersal and potential establishment across the northern United States is concerning and warrants investigations on species morphology in such environments.
+Understanding Palmer amaranth biology and growing strategies under different agroecosystems can enhance our knowledge on species adaptation and management practices.
+It can also aid in designing proactive and ecological tactics to limit the species range expansion, reduce its negative impact, and developing resilient and sustainable farming systems [@maclaren2020].
+Therefore, the objective of this study was to investigate the flowering pattern, gender, biomass production, and height of Palmer amaranth cohorts growing under corn, soybean and bareground across five locations in the United States Midwest.
 
@@
 
-The study was performed with a Palmer amaranth accession (Per1) from Perkins County, Nebraska.
-Per1 accession collection is documented with no reported herbicide resistance [@oliveira2021a].
-Three weeks prior to the field experiment, seeds were planted in plastic trays containing potting-mix.
+A Palmer amaranth accession (Per1) from Perkins County, Nebraska.
+with no reported herbicide resistance according to @oliveira2021a was selected for this study.
+Three weeks prior to the onset of the field experiments, seeds were planted in plastic trays containing potting-mix.
 Emerged seedlings (1 cm) were transplanted into 200 cm<sup>-3</sup> plastic pots (a plant pot<sup>-1</sup>).
-Palmer amaranth seedlings were supplied with adequate water and kept under greenhouse conditions at Arlington, Clay Center, Lincoln, and Macomb; and kept outdoors in Grant.
-Palmer amaranth seedlings were kept under greenhouse/outdoors until the onset of the experiment (2-3 leaf stage/5 to 8 cm height).
+Palmer amaranth seedlings were supplied with adequate water and kept under greenhouse conditions at Arlington, Clay Center, Lincoln, and Macomb; and kept outdoors in Grant until the 2-3 leaf stage (5 to 8 cm height) when they were transported to the field.
 
@@
 
-Fields were conventionally tilled and disked prior to planting.
+Fields were conventionally tilled prior to crop planting.
 Corn hybrid and soybean varieties were planted in 76-cm row spacing (Table 1).
-Monthly mean air temperature and sum precipitation were obtained using Daymet weather data from June through September across the five locations in 2018 and 2019 [@correndo2021] (Figure 1)
+Monthly mean air temperature and total precipitation were obtained using Daymet weather data from June through September across the five locations in 2018 and 2019 [@correndo2021] (Figure 1)
+
+```{r fig1, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../data analysis/weather/Figure 1.png")
+```
 
@@
 
-The field experimental unit were six adjacent 9.1 m wide (12 rows at 72.2 cm row spacing) by 10.7 m long.
-Each experimental unit was planted with corn or soybean, or under bareground condition.
-Palmer amaranth seedlings (potting mix + two seedlings) were and gently transferring to the ground (6 cm deep and 8 cm wide).
-Twenty-four plants were equidistantly placed (0.76 m apart) between rows within each agroecossystems.
+The field experimental units were three adjacent 9.1 m wide (12 rows at 76.2 cm row spacing) by 10.7 m long.
+Each experimental unit was planted with corn, soybean, or kept under bareground.
+The experimental design were three crops, two transplanting times simulating two cohorts.
+
+Palmer amaranth seedlings (potting mix + two seedlings) were transplanted (6 cm deep and 8 cm wide).
+Twenty-four plants were equidistantly placed (0.76 m apart) between rows within each crop (Figure 2).
 After a week, one was eliminated and one was kept.
@@
 
-```{r fig1, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/weather/Figure 1.png")
+```{r fig2, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../data analysis/figures/Figure 2.png")
 ```
@@
 
-```{r fig2, echo = FALSE, message=FALSE, fig.show = "hide"}
+```{r fig3, echo = FALSE, message=FALSE, fig.show = "hide"}
 knitr::include_graphics("../data analysis/figures/Figure 3.png")
@@
 
-```{r fig3, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/image.jpeg")
-```
-
 ```{r fig4, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/Figure 2.png")
+knitr::include_graphics("../data analysis/figures/Figure 4.png")
 ```
@@
 ```{r fig5, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/Figure 4.png")
+knitr::include_graphics("../data analysis/figures/Figure 5.png")
+```
+
+```{r fig6, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../data analysis/figures/Figure 6.png")
 ```
@@
 
-```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL', out.width = "150mm", out.height = "90mm"}
+```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Mean average temperature (C) and montly sum precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL', out.width = "150mm", out.height = "70mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 
-```{r, Figure-3, ref.label = "fig3", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Harvest Palmer amaranth plants at 40 days after first transplant timing.
-From left to right, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin', out.width = "90mm", out.height = "90mm"}
+```{r, Figure-3, ref.label = "fig3", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL', out.width = "150mm", out.height = "90mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 
-```{r, Figure-4, ref.label = "fig4", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "150mm", out.height = "150mm"}
+```{r, Figure-4, ref.label = "fig4", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='MCO (180 cm) holding harvested Palmer amaranth plants at 40 days after first transplant (1st cohort, A) and 33 days after second transplant (2nd cohort, B).
+From left to right, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin', out.width = "150mm", out.height = "90mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 
-```{r, Figure-5, ref.label = "fig5", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "170mm", out.height = "70mm"}
+```{r, Figure-5, ref.label = "fig5", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "150mm", out.height = "150mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
 ```
+
+```{r, Figure-6, ref.label = "fig6", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "170mm", out.height = "70mm"}
+# You can also refer to code chunks from above to place figures at the bottom.
+```

### 2021-08-23 Manuscript/Manuscript.tex
vb

@@
 \title[Palmer amaranth adaptation]{Palmer amaranth (\emph{Amaranthus
-palmeri}) adaptation to agroecossystems}
+palmeri}) adaptation to US Midwest agroecosystems}
 \author[\firstAuthorLast]{\Authors}
@@
 
-{\centering \includegraphics[width=160mm,height=100mm]{../data analysis/weather/Figure 1}
+{\centering \includegraphics[width=150mm,height=70mm]{../data analysis/figures/Figure 1}
 
@@
 
+{\centering \includegraphics[width=160mm,height=100mm]{../data analysis/weather/Figure 2}
+
+}
+
+\caption{Mean average temperature (C) and montly sum precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-2}
+\end{figure}
+
+\begin{figure}
+
 {\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 3}
@@
 
-\caption{Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-2}
+\caption{Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-3}
 \end{figure}
@@
 
-{\centering \includegraphics[width=90mm,height=90mm]{../data analysis/figures/image}
+{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 4}
 
@@
 
-\caption{Harvest Palmer amaranth plants at 40 days after first transplant timing.
-From left to right, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin}\label{fig:Figure-3}
+\caption{MCO (180 cm) holding harvested Palmer amaranth plants at 40 days after first transplant (1st cohort, A) and 33 days after second transplant (2nd cohort, B).
+From left to right, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin}\label{fig:Figure-4}
 \end{figure}
@@
 
-{\centering \includegraphics[width=150mm,height=150mm]{../data analysis/figures/Figure 2}
+{\centering \includegraphics[width=150mm,height=150mm]{../data analysis/figures/Figure 5}
 
@@
 
-\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)}\label{fig:Figure-4}
+\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)}\label{fig:Figure-5}
 \end{figure}
@@
 
-{\centering \includegraphics[width=170mm,height=70mm]{../data analysis/figures/Figure 4}
+{\centering \includegraphics[width=170mm,height=70mm]{../data analysis/figures/Figure 6}
 
@@
 
-\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)}\label{fig:Figure-5}
+\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)}\label{fig:Figure-6}
 \end{figure}

### 2021-08-23 Manuscript/Manuscript.Rmd
as

@@
 With a dioecy nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging adaptive traits among plants [@oliveira2018].
-Also, Palmer amaranth small seed size (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure [@hartzler2016], animals [@farmer2017], etc..
+Also, Palmer amaranth small seed size (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure [@hartzler2016], animals [@farmer2017].
 The dispersal capacity of Palmer amaranth makes the species one of the most successful cases of weed adaptation to current cropping systems.
@@
 Growing conditions and management practices also influence Palmer amaranth sex dimorphism and flowering pattern [@korres2017; @rumpa2019].
-Therefore, Palmer amaranth has shown plasticity to evolve and adapt into different environments.
-The continuous global temperature warming is likely impact agriculture and promote niches for Palmer amaranth invasion/adaptation into new environments.
+Therefore, Palmer amaranth has shown plasticity to evolve and adapt under selection pressure.
+The continuous global temperature warming is likely impact agriculture and promote niches for Palmer amaranth invasion/adaptation into new habitats.
 Currently, it is estimated that the greatest climatic risk of Palmer amaranth establishment are agronomic crops in Australia and Sub-Sahara Africa [@kistner2018].
@@
 Palmer amaranth is already found in agronomic crops of South America [@larran2017; @kupper2017] and Southern Europe [@milani2021].
-In the United States, Palmer amaranth is well established in the Cotton Belt [@garetson2019,  [@bagavathiannan2016] in the southern United States but its range is expanding northward.
+In the United States, Palmer amaranth is well established in the Cotton Belt [@garetson2019; @bagavathiannan2016] in the southern United States but its range is expanding northward.
 For example, herbicide resistant Palmer amaranth is widespread in Nebraska  [@oliveira2021a].
 There are some reported cases of Palmer amaranth in Michigan [@kohrt2017] and Connecticut [@aulakh2021].
-Also, is estimated that Palmer amaranth can cause high damage to soybean fields in Illinois [@davis2015], which is concerning as soybean along with corn make most of US Midwest agronomic hectares.
+Also, it is estimated that Palmer amaranth can cause high damage to soybean fields in Illinois [@davis2015], which is concerning as soybean along with corn make most of US Midwest agronomic hectares.
 In Iowa, a study showed that Palmer amaranth is still not well adapted compared to waterhemp (*Amaranthus tuberculatus*) [@baker2021].
-Palmer amaranth invasion and successful eradication is documented in Minnesota [@yu2021].
+Invasion and successful eradication of Palmer amaranth is documented in Minnesota [@yu2021].
 Palmer amaranth infestations have not been detected in Canada; however, Palmer amaranth seeds were detected in sweet potato slips in the country [@page2021].
 Palmer amaranth is still not as well adapted and established to Northern as it is in the Southern North America.
-Nonetheless, its range of expansion into new environment is increasing.
+Therefore, its range of expansion into new habitats is increasing.
 It seems fated the need to manage Palmer amaranth in agronomic crops throughout multiple environments in the near future.
@@
 Emerged seedlings (1 cm) were transplanted into 200 cm<sup>-3</sup> plastic pots (a plant pot<sup>-1</sup>).
-Palmer amaranth seedlings were supplied with adequate water and kept under greenhouse conditions at Arlington, Clay Center, Lincoln, and Macomb; and kept outdoors in Grant until the 2-3 leaf stage (5 to 8 cm height) when they were transported to the field.
+Palmer amaranth seedlings were supplied with adequate water and kept under greenhouse conditions at the University of Wisconsin-Madison, University of Nebraska-Lincoln, and Western Illinois University; and kept outdoors in the Perkins extension office at Grant, NE until the 2-3 leaf stage (5 to 8 cm height) when they were transported to the field.
 
@@
 
-The experiment was conducted in 2018 and 2019 under field conditions at five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE ('N, 'W), Grant (('N, 'W)), Lincoln (('N, 'W)), and Macomb (('N, 'W)).
+The experiment was conducted in 2018 and 2019 under field conditions at five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE ('N, 'W), Grant, NE (('N, 'W)), Lincoln, NE (('N, 'W)), and Macomb, IL (('N, 'W)).
 
@@
 The field experimental units were three adjacent 9.1 m wide (12 rows at 76.2 cm row spacing) by 10.7 m long.
+The experimental design were arranged in factorial design with three crops, two transplanting times simulating two cohorts, repeated across five locations.
 Each experimental unit was planted with corn, soybean, or kept under bareground.
-The experimental design were three crops, two transplanting times simulating two cohorts.
-
+The two transplanting timings were June 1 (first cohort) and July 1 (second cohort).
 Palmer amaranth seedlings (potting mix + two seedlings) were transplanted (6 cm deep and 8 cm wide).
-Twenty-four plants were equidistantly placed (0.76 m apart) between rows within each crop (Figure 2).
-After a week, one was eliminated and one was kept.
-There were two transplant timing: first (June 1) and second (July 1).
-There were 24 Palmer amaranth plants in each experimental unit, with a total of 144 plants for each location.
-The study was repeated twice.
+Forty-eight plants were equidistantly placed (0.76 m apart) between rows within each crop (Figure 2).
+After a week, one plant was eliminated and one was kept, resulting in 24 plants per experimental unit and transplanting time (Figure 2).
+When needed, Palmer amaranth plants were supplied with water during the first week after transplanting.
 
@@
 After transplanting, Palmer amaranth flowering was monitored until the end of the study.
-When a plant started flowering, the day was recorded, plant sex was identified as male or female, and plant height was measured from soil surface to the plant top.
-Then, aboveground plant biomass was harvest near soil surface and oven dried at 65 C until reaching constant weight before the weight of biomass (g plant 1) was recorded.
+When a plant flowered, the day was recorded, plant sex was identified (male or female), plant height was measured from soil surface to the top of plant.
+Also, aboveground plant biomass was harvest near soil surface and oven dried at 65 C until reaching constant weight before weighing (g plant-1 was recorded).
+
+Plants had to be harvested at flowering because Palmer amaranth is neither endemic in Wisconsin nor in Illinois.
+In our study, all locations follow the methodology of plant harvest at flowering initiation, except in Grant, NE.
+In this location, all Palmer amaranth plants were harvest at once on July 06, 2018 and 2019 (first cohort), and on August 17, 2018 and on July 31, 2019 (second cohort).
 
@@
 
-The statistical analyses were performed using R statistical software version 4.0.1. Data across locations and year were combined.
+The statistical analyses were performed using R statistical software version 4.0.1.
 
-The cumulative Palmer amaranth flowering estimation was determined using a asymmetrical three parameter log logistic Weibull model of the drc package [@ritz2015].
+Analyses of Palmer amaranth height and biomass were performed with a linear mixed model using *lmer* function from "lme4" package [@bates2015].
+Plant height and biomass were log transformed to meet model assumption of normality.
+In the model, crop (bareground, corn, soybean) was the fixed effect and year nested with location the random effects.
+Analysis of variance was performed with *anova* function from "car" package [@fox2018].
+Marginal means and compact letter display were estimated with *emmeans* and *cld* from packages "emmeans" and "multcomp" [@hothorn2008].
 
-$Y(x) = 0 + (d-0) exp (-exp(b(log(x)-e)))$
+The cumulative Palmer amaranth flowering estimation was determined using an asymmetrical three parameter log logistic Weibull model of the drc package [@ritz2015].
+
+$$Y(x) = 0 + (d-0) exp (-exp(b(log(x)-e)))$$
 
@@
 The doy for 10, 50, and 90% Palmer amaranth cumulative flowering were determined using the *ED* function of drc package.
-Also, the 10, 50, and 90% Palmer amaranth cumulative flowering were compared among agroecossystems and timings using the *EDcomp* function of drc package.
+Also, the 10, 50, and 90% Palmer amaranth cumulative flowering were compared among crops and cohorts using the *EDcomp* function of drc package.
 The EDcomp function compares the ratio of cumulative flowering using t-statistics, where P-value < 0.05 indicates that we fail to reject the null hypothesis.
 
-Palmer amaranth gender was fitted to a binary logistic regression [@bangdiwala2018].
-Binary logistic regression is used for predicting binary classes, such as the probability of female in dioecious plants.
+Palmer amaranth sex was fitted to a binary logistic regression [@bangdiwala2018].
+Binary logistic regression is used for predicting binary classes, such as the probability of a plant being female in a dioecious species.
 Prior to the analysis, all missing values were removed from the dataset.
 Also, data from Grant was not used in this analysis due to the uniform plant harvesting at that location.
-The whole dataset was splitted into 80% train and 20% test data.
-The 80% train is used for the model training and the rest 20% is used for checking how the model generalized on unseen dataset.
-With 80% dataset, a binary response variable, male (0) and female (1), was fitted to a generalized linear model (*glm* function) including day of year harvest, height, weight, crop and month as independent variables.
+The complete dataset was splitted into 80% train and 20% test data.
+The 80% train is used for the model training and the 20% test is used for checking how the model generalized on unseen dataset.
+With 80% dataset, a binary response variable, male (0) and female (1), was fitted to a generalized linear model (*glm* function) including day of year harvest, height, weight, crop and month as independent variables (without interaction).
 The model family was binomial with a logit function.
-The model fit was assessed through pseudo R squared values (McFadden, Cox and Snell, Cragg and Uhler) and likelihood ratio using *nagelkerke* function ("rcompanion" package).
-The marginal effects computation was performed with Average Marginal Effects at every observed value of x and average across the results (AMEs) [@leeper2017] using *margins* function from "margins" package.
+The model fit was assessed through pseudo R-squared values (McFadden, Cox and Snell, Cragg and Uhler) and likelihood ratio using *nagelkerke* function ("rcompanion" package).
+The marginal effects computation was performed with Average Marginal Effects (AMEs) at every observed value of x and average across the results [@leeper2017] using *margins* function from "margins" package.
 The rest 20% dataset was predicted using *predict* function with a cutoff estimation for male or female using *performance* function.
-The model quality prediction from the classification algorithm was measure with precision (*precision* function), recall (*recall* function) and F1 score (*f_meas* function) using "yardstick" package.
+The model quality prediction from the classification algorithm was measured with precision (*precision* function), recall (*recall* function) and F1 score (*f_meas* function) using the "yardstick" package.
 The precision determines the accuracy of positive predictions (female plants), recall determines the fraction of positives that were correctly identified, and F1 score is a weighted harmonic mean of precision and recall with the best score of 1 and the worst score of 0.
-F1 score conveys the balance between the precision and the recall.
-
-Palmer amaranth height and biomass were performed with a linear mixed model using *lmer* function from "lme4" package [@bates2015].
-Plant height and biomass were transformed to meet model assumption of normality.
-In the model, agroecosystem (crop, soybean, bareground) was the fixed effect and year nested with location the random effects.
-Analysis of variance was performed with *anova* function from "car" package [@fox2018].
-Marginal means and compact letter display were estimated with *emmeans* and *cld* from packages "emmeans" and "multcomp" [@hothorn2008].
+F1 score conveys the balance between the precision and the recall (CITATION).
 
@@
 
-Palmer amaranth accumulated more biomass when growing in bareground compared to Palmer amaranth growing in soybean and corn (figure 2A).
-At first cohort time, Palmer amaranth biomass was 75.5 g plant-1, 28.3 g plant-1 and 16.3 g plant-1 in bareground, soybean and corn, respectively.
-At second cohort timing, Palmer amaranth produced 62.6 g plant in bareground, followed by 6.3 g plant in soybean, and 1.4 g plant in corn.
+Palmer amaranth plants accumulated more biomass when growing in bareground compared to plants growing in soybean and corn (figure 3A).
+Palmer amaranth plants in the first cohort produced 75.5, 28.3 g, and 16.3 g plant-1 in bareground, soybean and corn, respectively.
+Plants from the second cohort produced 62.6 g plant in bareground, followed by 6.3 g plant in soybean, and 1.4 g plant in corn.
 
-Palmer amaranth height was more uniform across cohort timings, except when growing in corn (figure 2B).
-Palmer amaranth achieved 69.2 cm tall when growing at first cohort timing in bareground, which was not different to 70.7 cm tall at second cohort timing (P = 0.74).
-In addition, no difference in Palmer amaranth height (69.3 cm) was found in soybean at first cohort compared to bareground (P > 0.75).
-At second cohort in soybean, Palmer amaranth was near to 10 cm lower compared to its first cohort time (P = 0.04).
-The tallest and smallest Palmer amaranth plants were found in corn.
-Palmer amaranth reached 85.2 cm tall at first cohort and 38.2 cm at second cohort timings.
+```{r fig3, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../data analysis/figures/Figure 3.png")
+```
+
+Palmer amaranth height was more uniform across cohort timings, except when growing in corn (figure 3B).
+Palmer amaranth plants from the first cohort were on average 69.2 cm tall in bareground, which was not different from the 70.7 cm tall plants from the second cohort timing (P = 0.74).
+In addition, no difference in Palmer amaranth height (69.3 cm) was detected from first cohort plants in soybean to first and second cohort plants in bareground (P > 0.75).
+Palmer amaranth plants from the second cohort were nearly 10 cm lower compared to the first cohort in soybeans (P = 0.04).
+The tallest (first cohort) and smallest (second cohort) Palmer amaranth plants were found in corn.
+Palmer amaranth reached 85.2 and 38.2 cm tall, respectively.
 
@@
 
-Palmer amaranth growing in corn resulted in a longer flowering pattern compared to bareground and soybean at first cohort (figure3A).
-Nonetheless, the 10% cumulative Palmer amaranth flowering in soybean, bareground and corn occurred at the end of June.
-Palmer amaranth reached 10% flowering in soybean at doy 180, which was slightly different from bareground (doy 180.9; *P* = 0.01) and corn (doy 181.7; *P* = 0.00).
+Palmer amaranth plants from the first cohort growing in corn resulted in a longer flowering window compared to plants growing in bareground and soybean (Figure 4A).
+The 10% cumulative Palmer amaranth flowering in soybean, bareground and corn occurred at the end of June.
+Palmer amaranth reached 10% flowering in soybean, bareground and corn at doy 180, 180.9 and 181.7, respectively.
 The 50% Palmer amaranth cumulative flowering occurred in July.
-For example, Palmer amaranth reached 50% flowering in bareground at doy 193.4, followed by soybean (doy 194.8), corn (doy 206.6).
+Palmer amaranth reached 50% flowering in bareground, soybean and corn at doy 193.4, 194.8, and 206.6, respectively.
 Similar trend was observed at 90% Palmer amaranth cumulative flowering.
-Palmer amaranth growing in corn reached 90% flowering at doy 252.6 (early September), which was 37.8 and 32.2 days after Palmer amaranth 90% flowering in bareground and soybean, respectively.
+Palmer amaranth reached 90% flowering at doy 252.6 in corn (early September), which was 38 and 32 days after reaching 90% flowering in bareground and soybean, respectively.
 
-Palmer amaranth cumulative flowering at second cohort ranged from mid July to mid September (figure3B).
-Palmer amaranth growing in bareground resulted in earlier flowering time compared to soybean and corn.
+```{r fig4, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../data analysis/figures/Figure 5.png")
+```
+
+Palmer amaranth cumulative flowering at the second cohort ranged from mid July to mid September (Figure 4B).
+Palmer amaranth growing in the bareground resulted in earlier flowering time compared to soybean and corn.
 Palmer amaranth growing in bareground reached 10%, 50%, and 90% flowering time at day 203.8, 214.4, and 232.2, respectively.
@@
 
-The model fit was 0.23, 0.32, 0.40 with using pseudo R squared test from McFadden, Cox and Snell, and Cragg and Uhler, respectively.
+The model fit was 0.23, 0.32, 0.40 using pseudo R-squared test from McFadden, Cox and Snell, and Cragg and Uhler, respectively.
 The likelihood ratio test showed a p-value of < 0.00.
-The average marginal effects showed that Palmer amaranth growing in corn resulted in 14.5% less females plants (Table 2).
+The average marginal effects showed that Palmer amaranth growing in corn resulted in 14.8% less females plants (Table 2).
 Moreover, increasing a unit doy  increases the probability of having a female plant by 0.4% (Table 2 and Figure4A).
-Similar trend is observed to weight as well as height, whereas the probability of being female increase by 0.2% (Figure 4B) and 0.1% (Figure 4C) when a unit of weight (g) and height (cm) increases, respectively.
+Similar trend is observed for weight as well as height, whereas the probability of being female increase by 0.2% (Figure 4B) and 0.1% (Figure 4C) when a unit of weight (g) and height (cm) increases, respectively.
 
@@
 
-The model accuracy evaluation accuracy in the unseen 20% dataset was 0.62 with a cutoff value for female and male plants of 0.43.
+```{r fig5, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../data analysis/figures/Figure 6.png")
+```
+
+The model accuracy evaluation accuracy in the 20% test dataset was 0.62 with a cutoff value for female and male plants of 0.43.
 The model classification showed a precision of 0.64, recall of 0.66, and a F means score of 0.65.
 In addition, the area under the curve was 0.64.
-
-```{r fig3, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/Figure 3.png")
-```
-
-```{r fig4, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/Figure 4.png")
-```
-
-```{r fig5, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/Figure 5.png")
-```
-
-```{r fig6, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/Figure 6.png")
-```
 
@@
 Our study showed that Palmer amaranth biomass, height, flowering pattern and gender varied within agroecossystems and cohort timings.
-In general, Palmer amaranth produced more biomass and taller plants when growing at first cohort rather than second cohort.
-At first cohort, resources (e.g., soil nutrients) and conditions (e.g., light) were more timely available for both species, crop and weed.
+In general, first cohort of Palmer amaranth plants were heavier and taller when compared to the second cohort.
+At first cohort, resources (e.g., soil nutrients) and conditions (e.g., light) were more timely available for the species.
 High biomass and taller Palmer amaranth plants are likely a weed strategy to compete for light in between crop rows in absence of canopy.
-In such condition, Palmer amaranth showed an extraordinary plasticity to adapt upon the agroecosystem.
+In such conditions, Palmer amaranth showed an extraordinary plasticity to adapt upon the agroecosystem.
 This is evident when comparing Palmer amaranth canopy shape, and its extended flowering pattern when growing into corn compared to soybean.
-The Palmer amaranth competition strategy was to mimic the crop grow and development (Figure 4).
-These results suggests that Palmer amaranth can fast evolve life-history traits to adapt  into agroecossystems and cultural practices, including nitrogen fertilization [@bravo2018].
-Our results highlight the Palmer amaranth as a threat to agroecossystems as breeding more competitive crop varieties is likely to select more competitive biotypes [@bravo2017].
+The Palmer amaranth competition strategy was to mimic the crop grow and development (Figure 6).
+These results suggests that Palmer amaranth can fast evolve life-history traits to adapt into cropping systems and cultural practices, which was also showed in a study varying nitrogen fertilization [@bravo2018].
+Our results highlight the Palmer amaranth as a threat to field crops as breeding more competitive crop varieties is likely to select more competitive biotypes [@bravo2017].
 
-Palmer amaranth grow and development at second cohort was limited due to the crop competitive ability at advanced development stages.
+```{r fig6, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../data analysis/figures/Figure 4.png")
+```
+
+Palmer amaranth grow and development in second cohort was limited due to the crop competitive ability at advanced development stages.
 Palmer amaranth was transplanted when corn canopy was nearly closed, which reduced Palmer amaranth competitiveness.
 As a result, Palmer amaranth height and biomass was lower compared to its first cohort.
-Under crop canopy (e.g., second cohort), Palmer amaranth flowering pattern was near to similar in corn and soybean.
-Palmer amaranth growing without crop competition produced highest amounts of biomass and less extended flowering pattern.
+Under crop canopy (e.g., second cohort), Palmer amaranth flowering window was near to similar in corn and soybean.
+Palmer amaranth growing without crop competition produced the highest amounts of biomass and less extended flowering window.
 The Palmer amaranth strategy in bareground was to invest biomass in growing plant width and height.
@@
 In a bareground study, early emerged Palmer amaranth without competition was 50% taller than late emerged plants [@webster2015].
-These results suggests that crop competition is not the only factor limiting late Palmer amaranth establishment.
-The limited growth of Palmer amaranth at second cohort is likely a reduced plant response to thermal units (e.g, growing degree days).
+These results suggest that crop competition is not the only factor limiting late Palmer amaranth establishment.
+The limited growth of Palmer amaranth at second cohort is likely a reduced plant response to day length, light availability and thermal units (e.g, growing degree days).
 It is hypothesize that reduced day length contributed to smaller plants at second cohort as well as shorter flowering period.
@@
 I our study, Palmer amaranth growing at first cohort accumulated an overall 36% more biomass when compared to second cohort.
-Therefore, Palmer amaranth plants growing in the second cohort is likely to produce less seeds regardless the agroecossystem.
+Therefore, Palmer amaranth plants growing in the second cohort is likely to produce less seeds regardless the cropping system.
 Our observation is consistent with the findings that first Palmer amaranth cohort produced 50% more seeds per plant than Palmer amaranth plants established six weeks later in bareground [@webster2015].
@@
 When growing in soybean, Palmer amaranth flowering pattern was similar to bareground at first cohort but similar to corn at second cohort timing.
-Plant flowering initiation is complex and it depends on biological and ecological factors [@lang1965].
+Plant flowering initiation is complex and depends on biological and ecological factors [@lang1965].
 We hypothesize that when growing in high competition (e.g., second cohort), Palmer amaranth plants tend to initiate flowering early, as well as having an extended flowering pattern.
@@
 For example, when growing under water stress, early flowering in Palmer amaranth resulted in a mismatch between female and male plants by seven days [@mesgaran2021].
-A mismatch in Palmer amaranth male and female flowering period can minimize plant outcross, and thus reduce plant seed production.
+A mismatch in Palmer amaranth male and female flowering period can minimize plant outcross, and thus reduce plant seed production and exchange of resistant alleles [@jhala2021].
 Sex dimorphism manipulation is considered a potential ecological pest control [@mcfarlane2018; @schliekelman2005].
 
-The mechanisms of sex-determination in plant species is intriguing and arouse the curiosity of many scientists, including Darwin [@darwin1888].
-In our study, the gender model performance was decent considering the biology of plant flowering.
+The mechanisms of sex-determination in plant species is intriguing and arouse the curiosity of many scientists, including Charles Darwin [@darwin1888].
+In our study, the gender model performance was decent (AIC 0.64) considering the plant flowering biology.
 A 1:1 male and female sex ratio is a general evolutionary stable strategy for plant species perpetuation [@fisher1930].
@@
 
-RW: designed the experiments; AJ, CP, MB, MO, and SS: conducted the experiments; MO: analyzed the data and wrote the manuscript; AJ, CP, MB, MO, SS, and RW: conceptualized the research.
+RW and MO: designed the experiments; AJ, CP, MB, MO, and SS: conducted the experiments; MO: analyzed the data and wrote the manuscript; AJ, CP, MB, MO, SS, and RW: conceptualized the research.
 All authors reviewed the manuscript.
@@
 
-```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Mean average temperature (C) and montly sum precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL', out.width = "150mm", out.height = "70mm"}
+```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Palmer amaranth adaptation study layout of a transplanting time.
+Twenty-four Palmer amaranth plants were place 76.2 cm apart in each crop', out.width = "150mm", out.height = "70mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 
-```{r, Figure-4, ref.label = "fig4", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='MCO (180 cm) holding harvested Palmer amaranth plants at 40 days after first transplant (1st cohort, A) and 33 days after second transplant (2nd cohort, B).
+```{r, Figure-4, ref.label = "fig4", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "150mm", out.height = "150mm"}
+# You can also refer to code chunks from above to place figures at the bottom.
+```
+
+```{r, Figure-5, ref.label = "fig5", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "170mm", out.height = "70mm"}
+# You can also refer to code chunks from above to place figures at the bottom.
+```
+
+```{r, Figure-6, ref.label = "fig6", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='MCO (180 cm) holding harvested Palmer amaranth plants at 40 days after first transplant (1st cohort, A) and 33 days after second transplant (2nd cohort, B).
 From left to right, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin', out.width = "150mm", out.height = "90mm"}
@@
 ```
-
-```{r, Figure-5, ref.label = "fig5", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "150mm", out.height = "150mm"}
-# You can also refer to code chunks from above to place figures at the bottom.
-```
-
-```{r, Figure-6, ref.label = "fig6", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "170mm", out.height = "70mm"}
-# You can also refer to code chunks from above to place figures at the bottom.
-```

### 2021-08-23 Manuscript/Manuscript.tex
as

@@
 Palmer amaranth (\emph{Amaranthus palmeri} S. Watson) is currently
-considered one of the most economically damaged weed species to cropping
-systems in the United States. Unmanaged Palmer amaranth in competition
-for water, light and nutrients can drastically impact on crop yields
-(Berger et al., 2015). For example, Palmer amaranth is documented with
-potential to reduce 91\%, 68\%, and 54\% of corn (Massinga et al.,
-2001), soybean (Klingaman and Oliver, 1994), and cotton (Morgan et al.,
-2001) yields, respectively. Moreover, Palmer amaranth has showed a
-remarkable capacity to evolve resistance to herbicides. To date, Palmer
-amaranth has evolved resistance to eight herbicide sites of action
-(Heap, 2021), increasing the weed management complexity (Lindsay et al.,
-2017). Thus, Palmer amaranth poses an economical risk to sustainable
+ranked as one of the most economically detrimental weed species to
+cropping systems in the United States (ADD CITATION; WSSA Survey
+Perhaps). Unmanaged Palmer amaranth plants compete for water, light, and
+nutrients, which can drastically impact crop yields (Berger et al.,
+2015). For example, Palmer amaranth has been documented to reduce up to
+91\%, 68\%, and 54\% corn (Massinga et al., 2001), soybean (Klingaman
+and Oliver, 1994), and cotton (Morgan et al., 2001) yields,
+respectively. Moreover, Palmer amaranth has shown a remarkable capacity
+to evolve resistance to herbicides. To date, Palmer amaranth has evolved
+resistance to eight herbicide sites of action (Heap, 2021), increasing
+the weed management complexity (Lindsay et al., 2017). Thus, Palmer
+amaranth poses an economical and environmental risk to sustainable
 agriculture.
 
-Palmer amaranth is a fast growing summer annual forb indigenous to
+Palmer amaranth is a fast growing summer annual forb indigenous to the
 Sonoran Desert (Sauer, 1957). The species would eventually emerge as a
 threat to US agriculture in the 1990s. Palmer amaranth weediness is
-likely a result of human-assisted selection in combination with species
-biology. Farm mechanization, conservation agriculture (e.g., no-till),
-and reliance on herbicides for weed management are the main
+likely a result of human-assisted selection in combination with plant
+biology. Farm mechanization, adoption of conservation agriculture (e.g.,
+no-till), and reliance on herbicides for weed management are the main
 human-mediated selection of Palmer amaranth into cropping systems (Ward
@@
 2011), and spread across locations through farm equipment (Sauer, 1972),
-manure (Hartzler and Anderson, 2016), animals (Farmer et al., 2017), and
-plant propagules (Yu et al., 2021). The dispersal capacity of Palmer
-amaranth make the species one of the most successful cases of weed
-adaption to cropping systems.
+manure (Hartzler and Anderson, 2016), animals (Farmer et al., 2017). The
+dispersal capacity of Palmer amaranth makes the species one of the most
+successful cases of weed adaptation to current cropping systems.
 
 Light and temperature are likely the main environment requirements for
-Palmer amaranth successful grow and development. Palmer amaranth is
-reported with an extended germination period (Jha et al., 2010).
-Germination of Palmer amaranth is triggered by 18 C soil temperature
-(Keeley et al., 1987), and optimal germination and biomass production
-occur at 35/30 C day and night temperatures (Guo and Al-Khatib, 2003).
-Palmer amaranth germination is also human-mediated by tillage timings
-and preemergence-applied herbicides (Chahal et al., 2021). Water has not
-shown to limit Palmer amaranth fitness. Under continuous water stress,
-Palmer amaranth survived and produced at least 14000 seeds plant-1
-(Chahal et al., 2018). Seeds from Palmer amaranth growing with limited
-water conditions were heavier, less dormant, and prompt for germination
-(Matzrafi et al., 2021). Growing conditions and management practices
-also influence Palmer amaranth sex dimorphism and flowering pattern
-(Korres et al., 2017; Rumpa et al., 2019). The continuous global
-temperature warming can impact agriculture and promote niches for Palmer
-amaranth invasion/adaptation into new environments. Currently, it is
+Palmer amaranth successful grow and development (CITATION). Palmer
+amaranth is reported with an extended germination period (Jha et al.,
+2010). Germination of Palmer amaranth is triggered by 18 C soil
+temperature (Keeley et al., 1987), and optimal germination and biomass
+production occur at 35/30 C day and night temperatures (Guo and
+Al-Khatib, 2003). In addition, Palmer amaranth establishment is
+human-mediated by tillage timings and preemergence-applied herbicides
+(Chahal et al., 2021), which can result in weed germination shifts
+(Sbatella and Wilson, 2010). Water has not shown to limit Palmer
+amaranth fitness. Under continuous water stress, Palmer amaranth
+survived and produced at least 14000 seeds plant-1 (Chahal et al.,
+2018). Seeds from Palmer amaranth growing with limited water conditions
+were heavier, less dormant, and prompt for germination (Matzrafi et al.,
+2021). Growing conditions and management practices also influence Palmer
+amaranth sex dimorphism and flowering pattern (Korres et al., 2017;
+Rumpa et al., 2019). Therefore, Palmer amaranth has shown plasticity to
+evolve and adapt under selection pressure. The continuous global
+temperature warming is likely impact agriculture and promote niches for
+Palmer amaranth invasion/adaptation into new habitats. Currently, it is
 estimated that the greatest climatic risk of Palmer amaranth
@@
 (Larran et al., 2017; Küpper et al., 2017) and Southern Europe (Milani
-et al., 2021). In the US, Palmer amaranth is well established at cotton
-fields (Garetson et al., 2019) and non-crop areas (Bagavathiannan and
+et al., 2021). In the United States, Palmer amaranth is well established
+in the Cotton Belt (Garetson et al., 2019; Bagavathiannan and
 Norsworthy, 2016) in the southern United States but its range is
 expanding northward. For example, herbicide resistant Palmer amaranth is
-widespread in Nebraska (Oliveira et al., 2021), Michigan (Kohrt et al.,
-2017), Connecticut (Aulakh et al., 2021), and potential damaged to
-soybean fields in Illinois is reported (Davis et al., 2015). In Iowa, a
-study showed that Palmer amaranth is still not well adapted compared to
-waterhemp (\emph{Amaranthus tuberculatus}) (Baker, 2021). Successful
-cases of Palmer amaranth invasion and near to eradication is documented
-in Minnesota (Yu et al., 2021). No Palmer amaranth actively growing was
-found in Canada; however, Palmer amaranth seeds was detected in sweet
-potato slips (Page et al., 2021). Nonetheless, it seems fated the need
-to manage Palmer amaranth in agronomic crops throughout multiple
-environments in the near future. Strategies on Palmer amaranth
-management should encompass the agroecosystem level but not only
-attempts to eradicate the weed. Most tactics to manage Palmer amaranth
-are based on technology fixes (Scott, 2011), which are short-term (e.g.,
-herbicide and/or tillage) rather than long-term weed management. Palmer
-amaranth management should be built on minimizing the species ability to
-adapt, grow and develop into agroecossystems.
+widespread in Nebraska (Oliveira et al., 2021). There are some reported
+cases of Palmer amaranth in Michigan (Kohrt et al., 2017) and
+Connecticut (Aulakh et al., 2021). Also, it is estimated that Palmer
+amaranth can cause high damage to soybean fields in Illinois (Davis et
+al., 2015), which is concerning as soybean along with corn make most of
+US Midwest agronomic hectares. In Iowa, a study showed that Palmer
+amaranth is still not well adapted compared to waterhemp
+(\emph{Amaranthus tuberculatus}) (Baker, 2021). Invasion and successful
+eradication of Palmer amaranth is documented in Minnesota (Yu et al.,
+2021). Palmer amaranth infestations have not been detected in Canada;
+however, Palmer amaranth seeds were detected in sweet potato slips in
+the country (Page et al., 2021). Palmer amaranth is still not as well
+adapted and established to Northern as it is in the Southern North
+America. Therefore, its range of expansion into new habitats is
+increasing. It seems fated the need to manage Palmer amaranth in
+agronomic crops throughout multiple environments in the near future.
+Strategies on Palmer amaranth management should encompass the
+agroecosystem level but not only attempts to eradicate the weed. Most
+tactics to manage Palmer amaranth are based on technology fixes (Scott,
+2011), which are short-term (e.g., herbicide and/or tillage) rather than
+long-term weed management. Palmer amaranth management should be built on
+minimizing the species ability to adapt, grow and develop into
+agroecossystems.
 
 In the southeastern US, early growing Palmer amaranth is well known to
-have higher impact on cotton yields than late established plants (MacRae
-et al., 2013). In the northern states, Palmer amaranth impact on the
-agroecossystem is recent. Studies investigating Palmer amaranth in those
-locations is limited due to the plant classification as noxious weed
-species (Yu et al., 2021). Nonetheless, the continuous Palmer amaranth
-dispersal and potential establishment across northern United States is
-concerning and warrant investigations on species morphology in such
-environments. Understanding Palmer amaranth biology and growing
-strategies under different agroecossystems can enhance our knowledge on
-species adaptation. It can also aid on designing proactive and
+have a higher impact on cotton yields compared to late established
+plants (MacRae et al., 2013). In the northern states, Palmer amaranth
+impact on the agroecosystem is recent. Studies investigating Palmer
+amaranth in those locations are limited due to the plant classification
+as noxious weed species in some northern states (Yu et al., 2021).
+Nonetheless, the continuous Palmer amaranth dispersal and potential
+establishment across the northern United States is concerning and
+warrants investigations on species morphology in such environments.
+Understanding Palmer amaranth biology and growing strategies under
+different agroecosystems can enhance our knowledge on species adaptation
+and management practices. It can also aid in designing proactive and
 ecological tactics to limit the species range expansion, reduce its
-negative impact, and design resilient and sustainable farming systems
-(MacLaren et al., 2020). Therefore, the objective of this study was to
-investigate the flowering pattern, gender, biomass production, and
-height of Palmer amaranth growing under corn, soybean and bareground at
-two timings across five locations in the United States Midwest.
+negative impact, and developing resilient and sustainable farming
+systems (MacLaren et al., 2020). Therefore, the objective of this study
+was to investigate the flowering pattern, gender, biomass production,
+and height of Palmer amaranth cohorts growing under corn, soybean and
+bareground across five locations in the United States Midwest.
 
@@
 
-The study was performed with a Palmer amaranth accession (Per1) from
-Perkins County, Nebraska. Per1 accession collection is documented with
-no reported herbicide resistance (Oliveira et al., 2021). Three weeks
-prior to the field experiment, seeds were planted in plastic trays
-containing potting-mix. Emerged seedlings (1 cm) were transplanted into
-200 cm-3 plastic pots (a plant pot-1). Palmer amaranth seedlings were
-supplied with adequate water and kept under greenhouse conditions at
-Arlington, Clay Center, Lincoln, and Macomb; and kept outdoors in Grant.
-Palmer amaranth seedlings were kept under greenhouse/outdoors until the
-onset of the experiment (2-3 leaf stage/5 to 8 cm height).
+A Palmer amaranth accession (Per1) from Perkins County, Nebraska. with
+no reported herbicide resistance according to Oliveira et al. (2021) was
+selected for this study. Three weeks prior to the onset of the field
+experiments, seeds were planted in plastic trays containing potting-mix.
+Emerged seedlings (1 cm) were transplanted into 200 cm-3 plastic pots (a
+plant pot-1). Palmer amaranth seedlings were supplied with adequate
+water and kept under greenhouse conditions at Arlington, Clay Center,
+Lincoln, and Macomb; and kept outdoors in Grant until the 2-3 leaf stage
+(5 to 8 cm height) when they were transported to the field.
 
@@
 five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE ('N,
-'W), Grant (('N, 'W)), Lincoln (('N, 'W)), and Macomb (('N, 'W)).
-
-Fields were conventionally tilled and disked prior to planting. Corn
-hybrid and soybean varieties were planted in 76-cm row spacing (Table
-1). Monthly mean air temperature and sum precipitation were obtained
-using Daymet weather data from June through September across the five
+'W), Grant, NE (('N, 'W)), Lincoln, NE (('N, 'W)), and Macomb, IL (('N,
+'W)).
+
+Fields were conventionally tilled prior to crop planting. Corn hybrid
+and soybean varieties were planted in 76-cm row spacing (Table 1).
+Monthly mean air temperature and total precipitation were obtained using
+Daymet weather data from June through September across the five
 locations in 2018 and 2019 (Correndo et al., 2021) (Figure 1)
@@
 
-\multirow{-3}{*}{\raggedright\arraybackslash } & Weed control & glyphosate / S-metolachor & S-metolachlor + trazine + mesotrione, + bicyclopyrone &  &  & \\
+ & Weed control & glyphosate / S-metolachor & S-metolachlor + trazine + mesotrione, + bicyclopyrone &  &  & \\
+
+ & Stage 1 cohort & V2-3 &  &  &  & \\
+
+\multirow{-5}{*}{\raggedright\arraybackslash } & Stage 2 cohort & V6-7 &  &  &  & \\
 
@@
 
-\multirow{-3}{*}{\raggedright\arraybackslash } & Weed control & glyphosate / S-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
+ & Stage 1 cohort & V1-2 &  &  &  & \\
+
+ & Stage 2 cohort & V5-6 &  &  &  & \\
+
+\multirow{-5}{*}{\raggedright\arraybackslash } & Weed control & glyphosate / S-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
 
@@
 
-The field experimental unit were six adjacent 9.1 m wide (12 rows at
-72.2 cm row spacing) by 10.7 m long. Each experimental unit was planted
-with corn or soybean, or under bareground condition. Palmer amaranth
-seedlings (potting mix + two seedlings) were and gently transferring to
-the ground (6 cm deep and 8 cm wide). Twenty-four plants were
-equidistantly placed (0.76 m apart) between rows within each
-agroecossystems. After a week, one was eliminated and one was kept.
-There were two transplant timing: first (June 1) and second (July 1).
-There were 24 Palmer amaranth plants in each experimental unit, with a
-total of 144 plants for each location. The study was repeated twice.
+The field experimental units were three adjacent 9.1 m wide (12 rows at
+76.2 cm row spacing) by 10.7 m long. The experimental design were
+arranged in factorial design with three crops, two transplanting times
+simulating two cohorts, repeated across five locations. Each
+experimental unit was planted with corn, soybean, or kept under
+bareground. The two transplanting timings were June 1 (first cohort) and
+July 1 (second cohort). Palmer amaranth seedlings (potting mix + two
+seedlings) were transplanted (6 cm deep and 8 cm wide). Forty-eight
+plants were equidistantly placed (0.76 m apart) between rows within each
+crop (Figure 2). After a week, one plant was eliminated and one was
+kept, resulting in 24 plants per experimental unit and transplanting
+time (Figure 2). When needed, Palmer amaranth plants were supplied with
+water during the first week after transplanting.
 
 After transplanting, Palmer amaranth flowering was monitored until the
-end of the study. When a plant started flowering, the day was recorded,
-plant sex was identified as male or female, and plant height was
-measured from soil surface to the plant top. Then, aboveground plant
-biomass was harvest near soil surface and oven dried at 65 C until
-reaching constant weight before the weight of biomass (g plant 1) was
-recorded.
+end of the study. When a plant flowered, the day was recorded, plant sex
+was identified (male or female), plant height was measured from soil
+surface to the top of plant. Also, aboveground plant biomass was harvest
+near soil surface and oven dried at 65 C until reaching constant weight
+before weighing (g plant-1 was recorded).
+
+Plants had to be harvested at flowering because Palmer amaranth is
+neither endemic in Wisconsin nor in Illinois. In our study, all
+locations follow the methodology of plant harvest at flowering
+initiation, except in Grant, NE. In this location, all Palmer amaranth
+plants were harvest at once on July 06, 2018 and 2019 (first cohort),
+and on August 17, 2018 and on July 31, 2019 (second cohort).
 
@@
 The statistical analyses were performed using R statistical software
-version 4.0.1. Data across locations and year were combined.
+version 4.0.1.
+
+Analyses of Palmer amaranth height and biomass were performed with a
+linear mixed model using \emph{lmer} function from ``lme4'' package
+(Bates et al., 2015). Plant height and biomass were log transformed to
+meet model assumption of normality. In the model, crop (bareground,
+corn, soybean) was the fixed effect and year nested with location the
+random effects. Analysis of variance was performed with \emph{anova}
+function from ``car'' package (Fox and Weisberg, 2018). Marginal means
+and compact letter display were estimated with \emph{emmeans} and
+\emph{cld} from packages ``emmeans'' and ``multcomp'' (Hothorn et al.,
+2008).
 
 The cumulative Palmer amaranth flowering estimation was determined using
-a asymmetrical three parameter log logistic Weibull model of the drc
+an asymmetrical three parameter log logistic Weibull model of the drc
 package (Ritz et al., 2015).
 
-\(Y(x) = 0 + (d-0) exp (-exp(b(log(x)-e)))\)
+\[Y(x) = 0 + (d-0) exp (-exp(b(log(x)-e)))\]
 
@@
 50, and 90\% Palmer amaranth cumulative flowering were compared among
-agroecossystems and timings using the \emph{EDcomp} function of drc
-package. The EDcomp function compares the ratio of cumulative flowering
-using t-statistics, where P-value \textless{} 0.05 indicates that we
-fail to reject the null hypothesis.
-
-Palmer amaranth gender was fitted to a binary logistic regression
+crops and cohorts using the \emph{EDcomp} function of drc package. The
+EDcomp function compares the ratio of cumulative flowering using
+t-statistics, where P-value \textless{} 0.05 indicates that we fail to
+reject the null hypothesis.
+
+Palmer amaranth sex was fitted to a binary logistic regression
 (Bangdiwala, 2018). Binary logistic regression is used for predicting
-binary classes, such as the probability of female in dioecious plants.
-Prior to the analysis, all missing values were removed from the dataset.
-Also, data from Grant was not used in this analysis due to the uniform
-plant harvesting at that location. The whole dataset was splitted into
-80\% train and 20\% test data. The 80\% train is used for the model
-training and the rest 20\% is used for checking how the model
-generalized on unseen dataset. With 80\% dataset, a binary response
-variable, male (0) and female (1), was fitted to a generalized linear
-model (\emph{glm} function) including day of year harvest, height,
-weight, crop and month as independent variables. The model family was
-binomial with a logit function. The model fit was assessed through
-pseudo R squared values (McFadden, Cox and Snell, Cragg and Uhler) and
-likelihood ratio using \emph{nagelkerke} function (``rcompanion''
-package). The marginal effects computation was performed with Average
-Marginal Effects at every observed value of x and average across the
-results (AMEs) (Leeper, 2017) using \emph{margins} function from
-``margins'' package. The rest 20\% dataset was predicted using
+binary classes, such as the probability of a plant being female in a
+dioecious species. Prior to the analysis, all missing values were
+removed from the dataset. Also, data from Grant was not used in this
+analysis due to the uniform plant harvesting at that location. The
+complete dataset was splitted into 80\% train and 20\% test data. The
+80\% train is used for the model training and the 20\% test is used for
+checking how the model generalized on unseen dataset. With 80\% dataset,
+a binary response variable, male (0) and female (1), was fitted to a
+generalized linear model (\emph{glm} function) including day of year
+harvest, height, weight, crop and month as independent variables. The
+model family was binomial with a logit function. The model fit was
+assessed through pseudo R-squared values (McFadden, Cox and Snell, Cragg
+and Uhler) and likelihood ratio using \emph{nagelkerke} function
+(``rcompanion'' package). The marginal effects computation was performed
+with Average Marginal Effects at every observed value of x and average
+across the results (AMEs) (Leeper, 2017) using \emph{margins} function
+from ``margins'' package. The rest 20\% dataset was predicted using
 \emph{predict} function with a cutoff estimation for male or female
 using \emph{performance} function. The model quality prediction from the
-classification algorithm was measure with precision (\emph{precision}
+classification algorithm was measured with precision (\emph{precision}
 function), recall (\emph{recall} function) and F1 score (\emph{f\_meas}
-function) using ``yardstick'' package. The precision determines the
+function) using the ``yardstick'' package. The precision determines the
 accuracy of positive predictions (female plants), recall determines the
@@
 
-Palmer amaranth height and biomass were performed with a linear mixed
-model using \emph{lmer} function from ``lme4'' package (Bates et al.,
-2015). Plant height and biomass were transformed to meet model
-assumption of normality. In the model, agroecosystem (crop, soybean,
-bareground) was the fixed effect and year nested with location the
-random effects. Analysis of variance was performed with \emph{anova}
-function from ``car'' package (Fox and Weisberg, 2018). Marginal means
-and compact letter display were estimated with \emph{emmeans} and
-\emph{cld} from packages ``emmeans'' and ``multcomp'' (Hothorn et al.,
-2008).
-
 \hypertarget{results}{%
@@
 
-Palmer amaranth accumulated more biomass when growing in bareground
-compared to Palmer amaranth growing in soybean and corn (figure 2A). At
-first cohort time, Palmer amaranth biomass was 75.5 g plant-1, 28.3 g
-plant-1 and 16.3 g plant-1 in bareground, soybean and corn,
-respectively. At second cohort timing, Palmer amaranth produced 62.6 g
-plant in bareground, followed by 6.3 g plant in soybean, and 1.4 g plant
-in corn.
+Palmer amaranth plants accumulated more biomass when growing in
+bareground compared to plants growing in soybean and corn (figure 3A).
+Palmer amaranth plants in the first cohort produced 75.5, 28.3 g, and
+16.3 g plant-1 in bareground, soybean and corn, respectively. Plants
+from the second cohort produced 62.6 g plant in bareground, followed by
+6.3 g plant in soybean, and 1.4 g plant in corn.
 
 Palmer amaranth height was more uniform across cohort timings, except
-when growing in corn (figure 2B). Palmer amaranth achieved 69.2 cm tall
-when growing at first cohort timing in bareground, which was not
-different to 70.7 cm tall at second cohort timing (P = 0.74). In
-addition, no difference in Palmer amaranth height (69.3 cm) was found in
-soybean at first cohort compared to bareground (P \textgreater{} 0.75).
-At second cohort in soybean, Palmer amaranth was near to 10 cm lower
-compared to its first cohort time (P = 0.04). The tallest and smallest
-Palmer amaranth plants were found in corn. Palmer amaranth reached 85.2
-cm tall at first cohort and 38.2 cm at second cohort timings.
+when growing in corn (figure 3B). Palmer amaranth plants from the first
+cohort were on average 69.2 cm tall in bareground, which was not
+different from the 70.7 cm tall plants from the second cohort timing (P
+= 0.74). In addition, no difference in Palmer amaranth height (69.3 cm)
+was detected from first cohort plants in soybean to first and second
+cohort plants in bareground (P \textgreater{} 0.75). Palmer amaranth
+plants from the second cohort were nearly 10 cm lower compared to the
+first cohort in soybeans (P = 0.04). The tallest (first cohort) and
+smallest (second cohort) Palmer amaranth plants were found in corn.
+Palmer amaranth reached 85.2 and 38.2 cm tall, respectively.
 
@@
 
-Palmer amaranth growing in corn resulted in a longer flowering pattern
-compared to bareground and soybean at first cohort (figure3A).
-Nonetheless, the 10\% cumulative Palmer amaranth flowering in soybean,
-bareground and corn occurred at the end of June. Palmer amaranth reached
-10\% flowering in soybean at doy 180, which was slightly different from
-bareground (doy 180.9; \emph{P} = 0.01) and corn (doy 181.7; \emph{P} =
-0.00). The 50\% Palmer amaranth cumulative flowering occurred in July.
-For example, Palmer amaranth reached 50\% flowering in bareground at doy
-193.4, followed by soybean (doy 194.8), corn (doy 206.6). Similar trend
-was observed at 90\% Palmer amaranth cumulative flowering. Palmer
-amaranth growing in corn reached 90\% flowering at doy 252.6 (early
-September), which was 37.8 and 32.2 days after Palmer amaranth 90\%
-flowering in bareground and soybean, respectively.
-
-Palmer amaranth cumulative flowering at second cohort ranged from mid
-July to mid September (figure3B). Palmer amaranth growing in bareground
-resulted in earlier flowering time compared to soybean and corn. Palmer
-amaranth growing in bareground reached 10\%, 50\%, and 90\% flowering
-time at day 203.8, 214.4, and 232.2, respectively. Palmer amaranth
-growing in soybean reached 10\% flowering at doy 210.9, which was 6 days
-prior to corn (\emph{P}-value = 0.00). Similar trend was observed at
-50\% flowering, whereas Palmer amaranth reached 50\% flowering in corn
-(doy 233.0) 4 days after soybeans (doy 228.9; \emph{P} = 0.00). The 90\%
-Palmer amaranth cumulative flowering occurred at same day in corn
-(260.9) and soybean (260.5; \emph{P} = 0.66).
+Palmer amaranth plants from the first cohort growing in corn resulted in
+a longer flowering window compared to plants growing in bareground and
+soybean (Figure 4A). The 10\% cumulative Palmer amaranth flowering in
+soybean, bareground and corn occurred at the end of June. Palmer
+amaranth reached 10\% flowering in soybean, bareground and corn at doy
+180, 180.9 and 181.7, respectively. The 50\% Palmer amaranth cumulative
+flowering occurred in July. Palmer amaranth reached 50\% flowering in
+bareground, soybean and corn at doy 193.4, 194.8, and 206.6,
+respectively. Similar trend was observed at 90\% Palmer amaranth
+cumulative flowering. Palmer amaranth reached 90\% flowering at doy
+252.6 in corn (early September), which was 38 and 32 days after reaching
+90\% flowering in bareground and soybean, respectively.
+
+Palmer amaranth cumulative flowering at the second cohort ranged from
+mid July to mid September (Figure 4B). Palmer amaranth growing in the
+bareground resulted in earlier flowering time compared to soybean and
+corn. Palmer amaranth growing in bareground reached 10\%, 50\%, and 90\%
+flowering time at day 203.8, 214.4, and 232.2, respectively. Palmer
+amaranth growing in soybean reached 10\% flowering at doy 210.9, which
+was 6 days prior to corn (\emph{P}-value = 0.00). Similar trend was
+observed at 50\% flowering, whereas Palmer amaranth reached 50\%
+flowering in corn (doy 233.0) 4 days after soybeans (doy 228.9; \emph{P}
+= 0.00). The 90\% Palmer amaranth cumulative flowering occurred at same
+day in corn (260.9) and soybean (260.5; \emph{P} = 0.66).
 
@@
 
-The model fit was 0.23, 0.32, 0.40 with using pseudo R squared test from
+The model fit was 0.23, 0.32, 0.40 using pseudo R-squared test from
 McFadden, Cox and Snell, and Cragg and Uhler, respectively. The
@@
 marginal effects showed that Palmer amaranth growing in corn resulted in
-14.5\% less females plants (Table 2). Moreover, increasing a unit doy
+14.8\% less females plants (Table 2). Moreover, increasing a unit doy
 increases the probability of having a female plant by 0.4\% (Table 2 and
-Figure4A). Similar trend is observed to weight as well as height,
+Figure4A). Similar trend is observed for weight as well as height,
 whereas the probability of being female increase by 0.2\% (Figure 4B)
@@
 
-The model accuracy evaluation accuracy in the unseen 20\% dataset was
-0.62 with a cutoff value for female and male plants of 0.43. The model
+The model accuracy evaluation accuracy in the 20\% test dataset was 0.62
+with a cutoff value for female and male plants of 0.43. The model
 classification showed a precision of 0.64, recall of 0.66, and a F means
@@
 and gender varied within agroecossystems and cohort timings. In general,
-Palmer amaranth produced more biomass and taller plants when growing at
-first cohort rather than second cohort. At first cohort, resources
-(e.g., soil nutrients) and conditions (e.g., light) were more timely
-available for both species, crop and weed. High biomass and taller
-Palmer amaranth plants are likely a weed strategy to compete for light
-in between crop rows in absence of canopy. In such condition, Palmer
-amaranth showed an extraordinary plasticity to adapt upon the
-agroecosystem. This is evident when comparing Palmer amaranth canopy
-shape, and its extended flowering pattern when growing into corn
-compared to soybean. The Palmer amaranth competition strategy was to
-mimic the crop grow and development (Figure 4). These results suggests
-that Palmer amaranth can fast evolve life-history traits to adapt into
-agroecossystems and cultural practices, including nitrogen fertilization
-(Bravo et al., 2018). Our results highlight the Palmer amaranth as a
-threat to agroecossystems as breeding more competitive crop varieties is
-likely to select more competitive biotypes (Bravo et al., 2017).
-
-Palmer amaranth grow and development at second cohort was limited due to
+first cohort of Palmer amaranth plants were heavier and taller when
+compared to the second cohort. At first cohort, resources (e.g., soil
+nutrients) and conditions (e.g., light) were more timely available for
+the species. High biomass and taller Palmer amaranth plants are likely a
+weed strategy to compete for light in between crop rows in absence of
+canopy. In such conditions, Palmer amaranth showed an extraordinary
+plasticity to adapt upon the agroecosystem. This is evident when
+comparing Palmer amaranth canopy shape, and its extended flowering
+pattern when growing into corn compared to soybean. The Palmer amaranth
+competition strategy was to mimic the crop grow and development (Figure
+6). These results suggests that Palmer amaranth can fast evolve
+life-history traits to adapt into cropping systems and cultural
+practices, which was also showed in a study varying nitrogen
+fertilization (Bravo et al., 2018). Our results highlight the Palmer
+amaranth as a threat to field crops as breeding more competitive crop
+varieties is likely to select more competitive biotypes (Bravo et al.,
+2017).
+
+Palmer amaranth grow and development in second cohort was limited due to
 the crop competitive ability at advanced development stages. Palmer
@@
 height and biomass was lower compared to its first cohort. Under crop
-canopy (e.g., second cohort), Palmer amaranth flowering pattern was near
+canopy (e.g., second cohort), Palmer amaranth flowering window was near
 to similar in corn and soybean. Palmer amaranth growing without crop
-competition produced highest amounts of biomass and less extended
-flowering pattern. The Palmer amaranth strategy in bareground was to
+competition produced the highest amounts of biomass and less extended
+flowering window. The Palmer amaranth strategy in bareground was to
 invest biomass in growing plant width and height. Nonetheless, Palmer
@@
 without competition was 50\% taller than late emerged plants (Webster
-and Grey, 2015). These results suggests that crop competition is not the
+and Grey, 2015). These results suggest that crop competition is not the
 only factor limiting late Palmer amaranth establishment. The limited
 growth of Palmer amaranth at second cohort is likely a reduced plant
-response to thermal units (e.g, growing degree days). It is hypothesize
-that reduced day length contributed to smaller plants at second cohort
-as well as shorter flowering period. A study in North Carolina and
-Illinois predicted that less than 10\% Palmer amaranth seedlings
-emergence occurred after June (Piskackova et al., 2021). In addition,
-Palmer amaranth negative impact on soybean (Korres et al., 2020) and
-cotton (Webster and Grey, 2015) yields was higher when plants were
-established near to crop planting.
+response to day length, light availability and thermal units (e.g,
+growing degree days). It is hypothesize that reduced day length
+contributed to smaller plants at second cohort as well as shorter
+flowering period. A study in North Carolina and Illinois predicted that
+less than 10\% Palmer amaranth seedlings emergence occurred after June
+(Piskackova et al., 2021). In addition, Palmer amaranth negative impact
+on soybean (Korres et al., 2020) and cotton (Webster and Grey, 2015)
+yields was higher when plants were established near to crop planting.
 
@@
 plants growing in the second cohort is likely to produce less seeds
-regardless the agroecossystem. Our observation is consistent with the
+regardless the cropping system. Our observation is consistent with the
 findings that first Palmer amaranth cohort produced 50\% more seeds per
@@
 was similar to bareground at first cohort but similar to corn at second
-cohort timing. Plant flowering initiation is complex and it depends on
+cohort timing. Plant flowering initiation is complex and depends on
 biological and ecological factors (Lang, 1965). We hypothesize that when
@@
 The mechanisms of sex-determination in plant species is intriguing and
-arouse the curiosity of many scientists, including Darwin (Darwin,
-1888). In our study, the gender model performance was decent considering
-the biology of plant flowering. A 1:1 male and female sex ratio is a
-general evolutionary stable strategy for plant species perpetuation
-(Fisher, 1930). However, a slight deviation from 1:1 sex ratio might
-occur in some dioecious species. For example, the dioecious
-\emph{Halophila stipulacea} is a female-biased plant in its native
-habitat, but the naturalized \emph{H. stipulacea} have a 1:1 ratio
-(Nguyen et al., 2018). Naturalized of \emph{H. stipulacea} reduced
+arouse the curiosity of many scientists, including Charles Darwin
+(Darwin, 1888). In our study, the gender model performance was decent
+(AIC 0.64) considering the plant flowering biology. A 1:1 male and
+female sex ratio is a general evolutionary stable strategy for plant
+species perpetuation (Fisher, 1930). However, a slight deviation from
+1:1 sex ratio might occur in some dioecious species. For example, the
+dioecious \emph{Halophila stipulacea} is a female-biased plant in its
+native habitat, but the naturalized \emph{H. stipulacea} have a 1:1
+ratio (Nguyen et al., 2018). Naturalized of \emph{H. stipulacea} reduced
 female-male ratio to expand into its non-native habitat (Nguyen et al.,
@@
 
-RW: designed the experiments; AJ, CP, MB, MO, and SS: conducted the
-experiments; MO: analyzed the data and wrote the manuscript; AJ, CP, MB,
-MO, SS, and RW: conceptualized the research. All authors reviewed the
-manuscript.
+RW and MO: designed the experiments; AJ, CP, MB, MO, and SS: conducted
+the experiments; MO: analyzed the data and wrote the manuscript; AJ, CP,
+MB, MO, SS, and RW: conceptualized the research. All authors reviewed
+the manuscript.
 
@@
 
-{\centering \includegraphics[width=150mm,height=70mm]{../data analysis/figures/Figure 1}
+{\centering \includegraphics[width=160mm,height=100mm]{../data analysis/weather/Figure 1}
 
@@
 
-{\centering \includegraphics[width=160mm,height=100mm]{../data analysis/weather/Figure 2}
+{\centering \includegraphics[width=150mm,height=70mm]{../data analysis/figures/Figure 2}
 
@@
 
-\caption{Mean average temperature (C) and montly sum precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-2}
+\caption{Palmer amaranth adaptation study layout of a transplanting time. Twenty-four Palmer amaranth plants were place 76.2 cm apart in each crop}\label{fig:Figure-2}
 \end{figure}
@@
 
-{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 4}
+{\centering \includegraphics[width=150mm,height=150mm]{../data analysis/figures/Figure 5}
 
@@
 
-\caption{MCO (180 cm) holding harvested Palmer amaranth plants at 40 days after first transplant (1st cohort, A) and 33 days after second transplant (2nd cohort, B).
-From left to right, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin}\label{fig:Figure-4}
+\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)}\label{fig:Figure-4}
 \end{figure}
@@
 
-{\centering \includegraphics[width=150mm,height=150mm]{../data analysis/figures/Figure 5}
+{\centering \includegraphics[width=170mm,height=70mm]{../data analysis/figures/Figure 6}
 
@@
 
-{\centering \includegraphics[width=170mm,height=70mm]{../data analysis/figures/Figure 6}
+{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 4}
 
@@
 
-\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)}\label{fig:Figure-6}
+\caption{MCO (180 cm) holding harvested Palmer amaranth plants at 40 days after first transplant (1st cohort, A) and 33 days after second transplant (2nd cohort, B).
+From left to right, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin}\label{fig:Figure-6}
 \end{figure}
@@
 
+\leavevmode\hypertarget{ref-sbatella2010}{}%
+Sbatella, G. M., and Wilson, R. G. (2010). Isoxaflutole {Shifts Kochia}
+({Kochia} scoparia) {Populations} in {Continuous Corn}. \emph{Weed
+Technology} 24, 392--396.
+doi:\href{https://doi.org/10.1614/WT-D-09-00023.1}{10.1614/WT-D-09-00023.1}.
+
 \leavevmode\hypertarget{ref-schliekelman2005}{}%

### 2021-10-20 Manuscript/Manuscript.Rmd
as

@@
 Palmer amaranth (*Amaranthus palmeri* S.
-Watson) is currently ranked as one of the most economically detrimental weed species to cropping systems in the United States (ADD CITATION; WSSA Survey Perhaps).
+Watson) is currently ranked as one of the most economically detrimental weed species to cropping systems in the United States [@vanwychen2020].
 Unmanaged Palmer amaranth plants compete for water, light, and nutrients, which can drastically impact crop yields [@berger2015].
@@
 
-Light and temperature are likely the main environment requirements for Palmer amaranth successful grow and development (CITATION).
-Palmer amaranth is reported with an extended germination period [@jha2010].
-Germination of Palmer amaranth is triggered by 18 C soil temperature [@keeley1987], and optimal germination and biomass production occur at 35/30 C day and night temperatures [@guo2003].
+Light and temperature are the main environment requirements for Palmer amaranth successful grow and development [@jha2010].
+Palmer amaranth is reported with an extended germination period [@ward2013].
+Germination of Palmer amaranth was triggered by 18 C soil temperature at 5 cm depth [@keeley1987], and optimal germination and biomass production occurred at 35/30 C day and night temperatures [@guo2003].
 In addition, Palmer amaranth establishment is human-mediated by tillage timings and preemergence-applied herbicides [@chahal2021], which can result in weed germination shifts [@sbatella2010].
@@
 Growing conditions and management practices also influence Palmer amaranth sex dimorphism and flowering pattern [@korres2017; @rumpa2019].
-Therefore, Palmer amaranth has shown plasticity to evolve and adapt under selection pressure.
-The continuous global temperature warming is likely impact agriculture and promote niches for Palmer amaranth invasion/adaptation into new habitats.
+Therefore, Palmer amaranth has shown plasticity to evolve and fast adapt under selection pressure.
+Palmer amaranth invasion/adaptation into new habitats is likely to increase with the global temperature warming.
 Currently, it is estimated that the greatest climatic risk of Palmer amaranth establishment are agronomic crops in Australia and Sub-Sahara Africa [@kistner2018].
@@
 
-The cumulative Palmer amaranth flowering estimation was determined using an asymmetrical three parameter log logistic Weibull model of the drc package [@ritz2015].
+The Palmer amaranth flowering timing was estimated as cumulative flowering across all location.
+Dataset was nested but id did not include Grant, NE.
+Palmer amaranth cumulative flowering estimation was determined using an asymmetrical three parameter log logistic Weibull model of the drc package [@ritz2015].
 
@@
 The rest 20% dataset was predicted using *predict* function with a cutoff estimation for male or female using *performance* function.
-The model quality prediction from the classification algorithm was measured with precision (*precision* function), recall (*recall* function) and F1 score (*f_meas* function) using the "yardstick" package.
-The precision determines the accuracy of positive predictions (female plants), recall determines the fraction of positives that were correctly identified, and F1 score is a weighted harmonic mean of precision and recall with the best score of 1 and the worst score of 0.
-F1 score conveys the balance between the precision and the recall (CITATION).
+The model quality prediction from the classification algorithm was measured with precision (*precision* function), recall (*recall* function) and F1-score (*f_meas* function) using the "yardstick" package.
+The precision determines the accuracy of positive predictions (female plants), recall determines the fraction of positives that were correctly identified, and F1-score is a weighted harmonic mean of precision and recall with the best score of 1 and the worst score of 0.
+F1-score conveys the balance between the precision and the recall [@yacouby2020].
 
@@
 ```{r fig4, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/Figure 5.png")
+knitr::include_graphics("../data analysis/figures/Figure 4.png")
 ```
@@
 ```{r fig5, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/Figure 6.png")
+knitr::include_graphics("../data analysis/figures/Figure 5.png")
 ```
@@
 ```{r fig6, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/Figure 4.png")
+knitr::include_graphics("../data analysis/figures/Figure 6.png")
 ```
@@
 
-An ecological approach to reduce seed production in Palmer amaranth is understanding its flowering pattern.
-Our study suggests that Palmer amaranth flowering pattern was slightly influenced by cropping systems and cohort timings.
-Palmer amaranth growing in bareground and corn resulted in the overall shortest and longest flowering pattern, respectively.
-When growing in soybean, Palmer amaranth flowering pattern was similar to bareground at first cohort but similar to corn at second cohort timing.
+An ecological approach to reduce seed production in Palmer amaranth is understanding its flowering window.
+Our study suggests that Palmer amaranth flowering was slightly influenced by cropping systems and cohort timings.
+Palmer amaranth growing in bareground and corn resulted in the overall shortest and longest flowering window, respectively.
+When growing in soybean, Palmer amaranth flowering window was similar to bareground at first cohort but similar to corn at second cohort timing.
 Plant flowering initiation is complex and depends on biological and ecological factors [@lang1965].
-We hypothesize that when growing in high competition (e.g., second cohort), Palmer amaranth plants tend to initiate flowering early, as well as having an extended flowering pattern.
+We hypothesize that when growing in high competition (e.g., second cohort), Palmer amaranth plants tend to initiate flowering early, as well as having an extended flowering window.
 Early flower initiation is plant strategy when growing in stress conditions.
@@
 The mechanisms of sex-determination in plant species is intriguing and arouse the curiosity of many scientists, including Charles Darwin [@darwin1888].
-In our study, the gender model performance was decent (AIC 0.64) considering the plant flowering biology.
+In our study, the gender model performance was decent (AIC 0.64) considering the biology of plant flowering.
 A 1:1 male and female sex ratio is a general evolutionary stable strategy for plant species perpetuation [@fisher1930].
@@
 
-```{r, Figure-1, ref.label = "fig1", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Mean average temperature (C) and montly sum precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL', out.width = "160mm", out.height = "100mm"}
+```{r, Figure-1, ref.label = "fig1", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Mean average temperature (C) and total montly precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL', out.width = "160mm", out.height = "100mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 
-```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Palmer amaranth adaptation study layout of a transplanting time.
-Twenty-four Palmer amaranth plants were place 76.2 cm apart in each crop', out.width = "150mm", out.height = "70mm"}
+```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Palmer amaranth adaptation study layout of a plant cohort timing.
+Twenty-four Palmer amaranth plants were place 76.2 cm apart in each experimental unit', out.width = "150mm", out.height = "70mm"}
 # You can also refer to code chunks from above to place figures at the bottom.

### 2021-10-20 Manuscript/Manuscript.tex
as

@@
 ranked as one of the most economically detrimental weed species to
-cropping systems in the United States (ADD CITATION; WSSA Survey
-Perhaps). Unmanaged Palmer amaranth plants compete for water, light, and
-nutrients, which can drastically impact crop yields (Berger et al.,
-2015). For example, Palmer amaranth has been documented to reduce up to
-91\%, 68\%, and 54\% corn (Massinga et al., 2001), soybean (Klingaman
-and Oliver, 1994), and cotton (Morgan et al., 2001) yields,
-respectively. Moreover, Palmer amaranth has shown a remarkable capacity
-to evolve resistance to herbicides. To date, Palmer amaranth has evolved
-resistance to eight herbicide sites of action (Heap, 2021), increasing
-the weed management complexity (Lindsay et al., 2017). Thus, Palmer
-amaranth poses an economical and environmental risk to sustainable
-agriculture.
+cropping systems in the United States (Van Wychen, 2020). Unmanaged
+Palmer amaranth plants compete for water, light, and nutrients, which
+can drastically impact crop yields (Berger et al., 2015). For example,
+Palmer amaranth has been documented to reduce up to 91\%, 68\%, and 54\%
+corn (Massinga et al., 2001), soybean (Klingaman and Oliver, 1994), and
+cotton (Morgan et al., 2001) yields, respectively. Moreover, Palmer
+amaranth has shown a remarkable capacity to evolve resistance to
+herbicides. To date, Palmer amaranth has evolved resistance to eight
+herbicide sites of action (Heap, 2021), increasing the weed management
+complexity (Lindsay et al., 2017). Thus, Palmer amaranth poses an
+economical and environmental risk to sustainable agriculture.
 
@@
 
-Light and temperature are likely the main environment requirements for
-Palmer amaranth successful grow and development (CITATION). Palmer
-amaranth is reported with an extended germination period (Jha et al.,
-2010). Germination of Palmer amaranth is triggered by 18 C soil
-temperature (Keeley et al., 1987), and optimal germination and biomass
-production occur at 35/30 C day and night temperatures (Guo and
-Al-Khatib, 2003). In addition, Palmer amaranth establishment is
+Light and temperature are the main environment requirements for Palmer
+amaranth successful grow and development (Jha et al., 2010). Palmer
+amaranth is reported with an extended germination period (Ward et al.,
+2013). Germination of Palmer amaranth was triggered by 18 C soil
+temperature at 5 cm depth (Keeley et al., 1987), and optimal germination
+and biomass production occurred at 35/30 C day and night temperatures
+(Guo and Al-Khatib, 2003). In addition, Palmer amaranth establishment is
 human-mediated by tillage timings and preemergence-applied herbicides
@@
 Rumpa et al., 2019). Therefore, Palmer amaranth has shown plasticity to
-evolve and adapt under selection pressure. The continuous global
-temperature warming is likely impact agriculture and promote niches for
-Palmer amaranth invasion/adaptation into new habitats. Currently, it is
-estimated that the greatest climatic risk of Palmer amaranth
-establishment are agronomic crops in Australia and Sub-Sahara Africa
-(Kistner and Hatfield, 2018). Temperature is a key factor limiting
-Palmer amaranth expansion to cooler geographies (Briscoe Runquist et
-al., 2019); however, under future climate change Palmer amaranth is
-likely to expand northward into Canada and Northern Europe (Kistner and
-Hatfield, 2018; Briscoe Runquist et al., 2019).
+evolve and fast adapt under selection pressure. Palmer amaranth
+invasion/adaptation into new habitats is likely to increase with the
+global temperature warming. Currently, it is estimated that the greatest
+climatic risk of Palmer amaranth establishment are agronomic crops in
+Australia and Sub-Sahara Africa (Kistner and Hatfield, 2018).
+Temperature is a key factor limiting Palmer amaranth expansion to cooler
+geographies (Briscoe Runquist et al., 2019); however, under future
+climate change Palmer amaranth is likely to expand northward into Canada
+and Northern Europe (Kistner and Hatfield, 2018; Briscoe Runquist et
+al., 2019).
 
@@
 plant pot-1). Palmer amaranth seedlings were supplied with adequate
-water and kept under greenhouse conditions at Arlington, Clay Center,
-Lincoln, and Macomb; and kept outdoors in Grant until the 2-3 leaf stage
-(5 to 8 cm height) when they were transported to the field.
+water and kept under greenhouse conditions at the University of
+Wisconsin-Madison, University of Nebraska-Lincoln, and Western Illinois
+University; and kept outdoors in the Perkins extension office at Grant,
+NE until the 2-3 leaf stage (5 to 8 cm height) when they were
+transported to the field.
 
@@
 
-The cumulative Palmer amaranth flowering estimation was determined using
-an asymmetrical three parameter log logistic Weibull model of the drc
-package (Ritz et al., 2015).
+The Palmer amaranth flowering timing was estimated as cumulative
+flowering across all location. Dataset was nested but id did not include
+Grant, NE. Palmer amaranth cumulative flowering estimation was
+determined using an asymmetrical three parameter log logistic Weibull
+model of the drc package (Ritz et al., 2015).
 
@@
 generalized linear model (\emph{glm} function) including day of year
-harvest, height, weight, crop and month as independent variables. The
-model family was binomial with a logit function. The model fit was
-assessed through pseudo R-squared values (McFadden, Cox and Snell, Cragg
-and Uhler) and likelihood ratio using \emph{nagelkerke} function
-(``rcompanion'' package). The marginal effects computation was performed
-with Average Marginal Effects at every observed value of x and average
-across the results (AMEs) (Leeper, 2017) using \emph{margins} function
-from ``margins'' package. The rest 20\% dataset was predicted using
-\emph{predict} function with a cutoff estimation for male or female
-using \emph{performance} function. The model quality prediction from the
-classification algorithm was measured with precision (\emph{precision}
-function), recall (\emph{recall} function) and F1 score (\emph{f\_meas}
-function) using the ``yardstick'' package. The precision determines the
-accuracy of positive predictions (female plants), recall determines the
-fraction of positives that were correctly identified, and F1 score is a
-weighted harmonic mean of precision and recall with the best score of 1
-and the worst score of 0. F1 score conveys the balance between the
-precision and the recall.
+harvest, height, weight, crop and month as independent variables
+(without interaction). The model family was binomial with a logit
+function. The model fit was assessed through pseudo R-squared values
+(McFadden, Cox and Snell, Cragg and Uhler) and likelihood ratio using
+\emph{nagelkerke} function (``rcompanion'' package). The marginal
+effects computation was performed with Average Marginal Effects (AMEs)
+at every observed value of x and average across the results (Leeper,
+2017) using \emph{margins} function from ``margins'' package. The rest
+20\% dataset was predicted using \emph{predict} function with a cutoff
+estimation for male or female using \emph{performance} function. The
+model quality prediction from the classification algorithm was measured
+with precision (\emph{precision} function), recall (\emph{recall}
+function) and F1-score (\emph{f\_meas} function) using the ``yardstick''
+package. The precision determines the accuracy of positive predictions
+(female plants), recall determines the fraction of positives that were
+correctly identified, and F1-score is a weighted harmonic mean of
+precision and recall with the best score of 1 and the worst score of 0.
+F1-score conveys the balance between the precision and the recall
+(Yacouby and Axman, 2020).
 
@@
 An ecological approach to reduce seed production in Palmer amaranth is
-understanding its flowering pattern. Our study suggests that Palmer
-amaranth flowering pattern was slightly influenced by cropping systems
-and cohort timings. Palmer amaranth growing in bareground and corn
-resulted in the overall shortest and longest flowering pattern,
-respectively. When growing in soybean, Palmer amaranth flowering pattern
-was similar to bareground at first cohort but similar to corn at second
-cohort timing. Plant flowering initiation is complex and depends on
-biological and ecological factors (Lang, 1965). We hypothesize that when
-growing in high competition (e.g., second cohort), Palmer amaranth
-plants tend to initiate flowering early, as well as having an extended
-flowering pattern. Early flower initiation is plant strategy when
-growing in stress conditions. For example, when growing under water
-stress, early flowering in Palmer amaranth resulted in a mismatch
-between female and male plants by seven days (Mesgaran et al., 2021). A
-mismatch in Palmer amaranth male and female flowering period can
-minimize plant outcross, and thus reduce plant seed production. Sex
-dimorphism manipulation is considered a potential ecological pest
-control (McFarlane et al., 2018; Schliekelman et al., 2005).
+understanding its flowering window. Our study suggests that Palmer
+amaranth flowering was slightly influenced by cropping systems and
+cohort timings. Palmer amaranth growing in bareground and corn resulted
+in the overall shortest and longest flowering window, respectively. When
+growing in soybean, Palmer amaranth flowering window was similar to
+bareground at first cohort but similar to corn at second cohort timing.
+Plant flowering initiation is complex and depends on biological and
+ecological factors (Lang, 1965). We hypothesize that when growing in
+high competition (e.g., second cohort), Palmer amaranth plants tend to
+initiate flowering early, as well as having an extended flowering
+window. Early flower initiation is plant strategy when growing in stress
+conditions. For example, when growing under water stress, early
+flowering in Palmer amaranth resulted in a mismatch between female and
+male plants by seven days (Mesgaran et al., 2021). A mismatch in Palmer
+amaranth male and female flowering period can minimize plant outcross,
+and thus reduce plant seed production and exchange of resistant alleles
+(Jhala et al., 2021). Sex dimorphism manipulation is considered a
+potential ecological pest control (McFarlane et al., 2018; Schliekelman
+et al., 2005).
 
@@
 (Darwin, 1888). In our study, the gender model performance was decent
-(AIC 0.64) considering the plant flowering biology. A 1:1 male and
+(AIC 0.64) considering the biology of plant flowering. A 1:1 male and
 female sex ratio is a general evolutionary stable strategy for plant
@@
 
-\caption{Mean average temperature (C) and montly sum precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-1}
+\caption{Mean average temperature (C) and total montly precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-1}
 \end{figure}
@@
 
-\caption{Palmer amaranth adaptation study layout of a transplanting time. Twenty-four Palmer amaranth plants were place 76.2 cm apart in each crop}\label{fig:Figure-2}
+\caption{Palmer amaranth adaptation study layout of a plant cohort timing. Twenty-four Palmer amaranth plants were place 76.2 cm apart in each experimental unit}\label{fig:Figure-2}
 \end{figure}
@@
 
-{\centering \includegraphics[width=150mm,height=150mm]{../data analysis/figures/Figure 5}
+{\centering \includegraphics[width=150mm,height=150mm]{../data analysis/figures/Figure 4}
 
@@
 
-{\centering \includegraphics[width=170mm,height=70mm]{../data analysis/figures/Figure 6}
+{\centering \includegraphics[width=170mm,height=70mm]{../data analysis/figures/Figure 5}
 
@@
 
-{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 4}
+{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 6}
 
@@
 
+\leavevmode\hypertarget{ref-jhala2021}{}%
+Jhala, A. J., Norsworthy, J. K., Ganie, Z. A., Sosnoskie, L. M., Beckie,
+H. J., Mallory-Smith, C. A., Liu, J., Wei, W., Wang, J., and
+Stoltenberg, D. E. (2021). Pollen-mediated gene flow and transfer of
+resistance alleles from herbicide-resistant broadleaf weeds. \emph{Weed
+Technology} 35, 173--187.
+doi:\href{https://doi.org/10.1017/wet.2020.101}{10.1017/wet.2020.101}.
+
 \leavevmode\hypertarget{ref-keeley1987}{}%
@@
 
+\leavevmode\hypertarget{ref-vanwychen2020}{}%
+Van Wychen, L. (2020). 2020 {Survey} of the most common and troublesome
+weeds in grass crops, pasture, and turf in the {United States} and
+{Canada}. Available at:
+\url{https://wssa.net/wp-content/uploads/2020-Weed-Survey_grass-crops.xlsx}.
+
 \leavevmode\hypertarget{ref-wang1992}{}%
@@
 doi:\href{https://doi.org/10.1614/WS-D-14-00051.1}{10.1614/WS-D-14-00051.1}.
+
+\leavevmode\hypertarget{ref-yacouby2020}{}%
+Yacouby, R., and Axman, D. (2020). Probabilistic {Extension} of
+{Precision}, {Recall}, and {F1 Score} for {More Thorough Evaluation} of
+{Classification Models}. in \emph{Proceedings of the {First Workshop} on
+{Evaluation} and {Comparison} of {NLP Systems}} ({Online}: {Association
+for Computational Linguistics}), 79--91.
+doi:\href{https://doi.org/10.18653/v1/2020.eval4nlp-1.9}{10.18653/v1/2020.eval4nlp-1.9}.
 

### 2021-10-20 data analysis/gender_analysis.Rmd
as

@@
     TRUE ~ harvest
-  )) -> data1
+  )) %>% 
+  janitor::clean_names() %>% 
+  rename(planting = planting_4,
+         planting_1 = planting_9) -> data1
 ```
@@
        ylab="P(female)",
-       line = list(col="1"),
+       line = list(col="#333333"),
        fill=list(fill= "green", alpha = 0.2),
-       points=list(size=1, col = 1, pch=1),
+       points=list(size=1, col = "#333333", pch=1),
        ylim = c(0,1)) -> doy_f
 ```
+
 ```{r}
@@
 
-ggsave("figures/fig2.png")
+#ggsave("figures/fig2.png")
 ```
@@
        ylab="P(female)",
-       line = list(col="1"),
+       line = list(col="#333333"),
        fill=list(fill= "green", alpha = 0.2),
-       points=list(size=1, col = 1, pch=1),
+       points=list(size=1, col = "#333333", pch=1),
        ylim = c(0,1)) -> weight_f
@@
 
-ggsave("figures/fig1.png")
+#ggsave("figures/fig1.png")
 ```
@@
        ylab="P(female)",
-       line = list(col="1"),
+       line = list(col="#333333"),
        fill=list(fill= "green", alpha = 0.2),
-       points=list(size=1, col = 1, pch=1),
+       points=list(size=1, col = "#333333", pch=1),
        ylim = c(0,1)) -> height_f
@@
 
-ggsave("figures/fig.png")
+#ggsave("figures/fig.png")
 ```
@@
 
-ggsave("figures/Figure 4.png", width = 9, height = 3)
+ggsave("figures/Figure 5.png", width = 9, height = 3)
 ```

### 2021-10-20 data analysis/weight.Rmd
as

@@
     TRUE ~ harvest
-  )) -> data1
+  )) %>% 
+  janitor::clean_names() %>% 
+  rename(month = planting_4) -> data1
 ```
@@
 data1 %>% 
-  rename(month = planting) %>% 
+#  rename(month = planting) %>% 
   mutate_at(c("year", "location"), as_factor) %>% 
@@
 weight %>% 
+  mutate(crop = fct_relevel(crop, 
+                           levels = c("bareground", "corn", "soybean"))) %>% 
+  ggplot(aes(y = crop, 
+             x = response,
+             shape = month,
+             color = month))
+```
+
+```{r}
+weight %>% 
+  mutate(crop = fct_relevel(crop, 
+                           levels = c("bareground", "corn", "soybean"))) %>% 
   mutate(crop = fct_recode(crop,
@@
   scale_y_discrete(expand = c(0.1, 0.1)) +
+  geom_linerange(aes(xmin = lower.CL, xmax = upper.CL),
+                 position = position_dodge2(width = 0.4, reverse = TRUE)) +
   geom_jitter(data = data2 %>% 
@@
               position = position_dodge2(width = 0.4, reverse = TRUE)) +
-  geom_linerange(aes(xmin = lower.CL, xmax = upper.CL),
-                 position = position_dodge2(width = 0.4, reverse = TRUE)) +
-  scale_color_manual(values = c(2,1)) +
+  scale_color_manual(values = c(2, "#333333")) +
   geom_text(aes(label = round(response,1), x = upper.CL + 130, color = month),
@@
            label = "First cohort",
-           size = 7, color = 1,
+           size = 7, color = "#333333",
            fontface = "bold") + 
@@
 height %>% 
+  mutate(crop = fct_relevel(crop, 
+                           levels = c("bareground", "corn", "soybean"))) %>% 
   mutate(crop = fct_recode(crop,
@@
               position = position_dodge2(width = 0.4, reverse = TRUE)) +
-  geom_linerange(aes(xmin = lower.CL, xmax = upper.CL),
-                 position = position_dodge2(width = 0.4, reverse = TRUE)) +
-  scale_color_manual(values = c(2,1)) +
+  scale_color_manual(values = c(2,"#333333")) +
   geom_text(aes(label = round(response,1), x = upper.CL + 20, color = month),
@@
   facet_grid(~crop) +
-  scale_color_manual(values = c(2,1)) +
-  scale_fill_manual(values = c(2,1)) +
+  scale_color_manual(values = c(2,"#333333")) +
+  scale_fill_manual(values = c(2,"#333333")) +
   scale_x_continuous(limits = c(170, 260), 
@@
   geom_text(data = first, mapping = aes(x = x, y = y), label = first$label,
-            size = 10, color = 1) +
+            size = 10, color = "#333333") +
   geom_text(data = second, mapping = aes(x = x, y = y), label = second$label,
@@
   facet_grid(~crop, scales = "free") +
-  scale_color_manual(values = c(2,1)) +
-  scale_fill_manual(values = c(2,1)) +
+  scale_color_manual(values = c(2,"#333333")) +
+  scale_fill_manual(values = c(2,"#333333")) +
   scale_y_continuous( breaks = seq(0, 1, by = 0.2)) +
@@
   facet_grid(~crop) +
-  scale_color_manual(values = c(2,1)) +
-  scale_fill_manual(values = c(2,1)) +
+  scale_color_manual(values = c(2,"#333333")) +
+  scale_fill_manual(values = c(2,"#333333")) +
   scale_y_continuous( breaks = seq(0, 1, by = 0.2)) +
@@
 
-ggsave("figures/Figure 4.png", width = 6, height = 8)
+ggsave("figures/Figure 3.png", width = 6, height = 8)
 ```
